\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 2 \\ Functions on $\R^n$}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Fall 2018 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.}

\textbf{a.} Yes $f$ is injective because if $(x_0, y_0) \not= (x_1, y_1)$
then $(x_0 +1 , y_0 +1) \not= (x_1 +1 , y_1 +1)$ and so
 $(x_0 +1 , y_0 +1, 5) \not= (x_1 +1 , y_1 +1, 5)$.


\textbf{b.} No $f$ is not surjective. For example the origin of $\R^3$ is not in
the range. Only points with $z$-coordinate equal to 5 are in the range.



\textbf{c.} $\singleton{(4, -6)}$



\textbf{d.}$\emptyset$



\textbf{e.} The line: $y=1$, $z=5$. In other words $\setof{(x,y,z)}{y=1, z=5}$

\medskip

\textbf{Problem 2.}

\textbf{a.} No $g$ is not injective. For example
$g(0,0,0) = g(0, 0,1) = (0,0)$.


\textbf{b.} No $g$ is not onto. For example $(0, -1)$ is not in the range of
$g$ because only points with non-negative $y$-coordinates are in the range.


\textbf{c.} The inverse image of the origin in $\R^2$ is the $z$-axis in $\R^3$.


\textbf{d.} The inverse image of the $x$-axis in $\R^2$ is the $x$- $z$- plane
in $\R^3$

\textbf{e.} The inverse image of the horizontal line $y=1$ in $\R^2$ is
the union of two planes in $\R^3$ parallel to the $x$- $z$- plane: The
plane $y=1$ and the plane $y=-1$.

\textbf{f.} The image of the $x$-axis in $\R^3$ is the $x$-axis in $\R^2$.

\textbf{g.} The  image of the line $y=x$ in the $x$- $y$- plane
of $\R^3$ is the parabola $y=x^2$ in $\R^2$.


\medskip

\textbf{Problem 3.}

\textbf{a.}  The domain and target are $\R^2$. The range is

\medskip

\textbf{b.} Is $h$ one-to-one?

\medskip

\textbf{c.} Is $h$ surjective?

\medskip

\textbf{c.} What is $h^{-1}[\singleton{(0,0)}]$?

\medskip

\textbf{d.} What is the image of the $x$-axis under $h$?

\medskip

\textbf{e.} What is the image of the $y$-axis under $h$?

\medskip

\textbf{f.} What is the inverse image of the $x$-axis under $h$?

\medskip

\textbf{g.} What is the inverse image of the $y$-axis under $h$?

\medskip

\textbf{h.} Is $f \circ g$ defined? Explain?

\bigskip

\textbf{Problem 4.}
\smallskip

\textbf{a.} Give an example of a function $f:\R^2\map\R^3$ that is
one-to-one but not onto.

\bigskip

\textbf{b.} Give an example of a function $f:\R^2\map\R^3$ that is neither
one-to-one nor onto.

\bigskip

\textbf{c.} Give an example of a function $f:\R^3\map\R^2$ that is
onto but not one-to-one.

\bigskip

\textbf{d.} Give an example of a function $f:\R^3\map\R^2$ that is neither
one-to-one nor onto.


\bigskip

\textbf{Problem 5.} These questions are difficult.
\smallskip

\textbf{a.} Is it possible for $f:\R^2\map\R^2$ to be one-to-one but not onto?
If so give an example. If not explain why.

\bigskip

\textbf{b.} Is it possible for $f:\R^2\map\R^2$ to be onto but not one-to-one?
If so give an example. If not explain why.

\bigskip

\textbf{c.} Is it possible for $f:\R^3\map\R^2$ to be one-to-one?

\end{document}
