% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 18 \\ Fermat's Little Theorem}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$a^{p-1}$}

\begin{itemize}
  \item Let $p$ be a prime. We are going to consider integers of the form $a^{p-1}$.
  \item For example, let $p=5$. We are going to consider integers of the form $a^4$.
  \item $2^4=16$, $3^4=81$, $4^4=256$, $5^4=625$
  \item $6^4=1296$,  $7^4=2401$, $8^4=4096$.
  \item What do you notice about the above values?
  \item For all of the $a$ other than $a=5$, $a^4$ is a multiple of 5 plus one.
  \item In other words $a^4 \equiv 1 \pmod 5$.
\end{itemize}

\end{frame}

\begin{frame}{Fermat's Little Theorem}

\begin{itemize}
  \item \textbf{Fermat's Little Theorem.} Let $p$ be prime. Let $a$ be relatively prime to $p$.
  \item Then $a^{p-1} \equiv 1 \pmod p$.
  \item Example. Let $p=11$. Suppose $11\ndivides a$. Then $a^{10} \equiv 1 \pmod {11}$.
  \item $2^{10} = 1024$ and $1023 = 11 \cdot 93$.
  \item $3^{10} = 59049$ and $59048 = 11 \cdot 5368$.
  \item $4^{10} = 1048576$ and $1048575 = 11 \cdot 95325$.
  \item $15^{10} = 576650390625$ and  $576650390624 = 11 \cdot 52422762784$.
  \item What happens if $a$ is not relatively prime to $p$?
  \item $22^{10} = 26559922791424$.
  \item Since $11\divides 22$, $11\divides (22^{10})$.
  \item If $p\divides a$ then $a^{p-1} \equiv 0 \pmod p$.
\end{itemize}

\end{frame}

\begin{frame}{Equivalent in terms of $\Z_n^*$}

\begin{itemize}
  \item We know that $a^n \equiv (a \bmod p)^n \pmod p$.
  \item i.e. the value of $a^n \bmod p$ only depends on $a\bmod p$.
  \item So we may as well only consider $a = 1,2,3,\cdots p-1$.
  \item So then we may as well work in $\Z_p^*$.
  \item \textbf{Theorem.} Let $p$ be prime. Let $[a]\in\Z_p^*$.
  \item Then $[a]^{p-1} = [1]$.
  \item Notice that $p-1$ is the cardinality of $Z_p^*$.
  \item This theorem may be stated:
  \item ``Any element of the group raised to the power of the cardinality of the group equals one.''
  \item For example, work in $\Z_{11}^*$.
  \item Then $[a]^{10} = 1$, for all $[a]\in\Z_{11}^*$.
\end{itemize}

\end{frame}

\begin{frame}{Proof}

\begin{itemize}
  \item \textbf{proof.} Let $[a] \in \Z_p^*$. Consider the set of all multiples of $[a]$ in $\Z_p^*$.
  \item $[1][a], [2][a], [3][a], \cdots [p-1][a]$.
  \item \textbf{Claim.} The set of multiples of $[a]$ runs through all of the elements of $\Z_p^*$.
  \item For example, take $p=5$, $[a]=[3]$. The set of multiples of $[3]$ in $\Z_5^*$ is:
  \item $[1][3]=[3], [2][3]=[1], [3][3]=[4], [4][3] = [2]$.
  \item \textbf{proof of Claim.} Notice that there are $p-1$ values. So we just need to show they are all different.
  \item Suppose $[i][a]=[j][a]$. Multiplying both sides by $[a]^{-1}$ we get $[i]=[j]$.
  \item In other words, if $[i]\not=[j]$, then $[i][a]\not=[j][a]$.
  \item So all $p-1$ multiples of $[a]$ in $\Z_p^*$ are different.
  \item So the set of multiples of $[a]$ runs through all of the elements of $\Z_p^*$. $\diamondsuit$.
\end{itemize}

\end{frame}

\begin{frame}{Proof continued}

\begin{itemize}
  \item Now consider what happens when we multiply all of the elements of $\Z^*_p$ together.
  \item Let $[x] = [1]\cdot[2]\cdot[3]\cdots[p-1]$.
  \item We know from the proof of Wilson's Theorem that $[x] = -[1]$.
  \item But actually we are not going to need this fact.
  \item By the claim
  \item $[x] = [1][a]\cdot[2][a]\cdot[3][a]\cdots[p-1][a]$
  \item  $=[a]^{p-1}[1]\cdot[2]\cdot[3]\cdots[p-1]$
  \item $=[a]^{p-1}[x]$.
  \item Canceling $[x]$, we get $[a]^{p-1}=[1]$.
  \item $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Corollary}

\begin{itemize}
  \item \textbf{Theorem.} Suppose $p$ is prime and $a$ is any integer.
  \item Then $a^p \equiv a \pmod p$.
  \item Example: $971^{31} \equiv 971 \pmod{31}$.
  \item \textbf{proof.} If $p\ndivides a$ then Fermat's Little theorem says
  \item $a^{p-1} \equiv 1 \pmod p$.
  \item Multiplying both sides by $a$ we get $a^p \equiv a \pmod p$.
  \item If $p\divides a$ then $a^p \equiv a \equiv 0 \pmod p$. $\qed$.
  \item Example. Let $p=3$.
  \item $4^3 = 64 \equiv 4 \pmod 3$.
  \item $3^3 = 27 \equiv 3 \pmod 3$.
\end{itemize}

\end{frame}

\begin{frame}{Computing powers mod a prime}

\begin{itemize}
  \item Example. Compute $7^{39412} \bmod 11$.
  \item Solution. $11$ is prime and $7$ is relatively prime to $11$.
  \item So Fermat's Little Theorem tells us that $7^{10} \bmod 11 = 1$.
  \item $7^{39412} = 7^{10(3941) + 2}$
  \item $=\left(7^{10}\right)^{3941}\cdot 7^2$
  \item So $7^{39412} \bmod 11 = 1^{3941}\cdot 7^2 \bmod 11$
  \item $=49 \bmod 11 = 5$.
\end{itemize}

\end{frame}



\end{document}
