\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Course Info}
\author{Math 325, Linear Algebra \\ Section 2 \\ Fall 2018 \\ SFSU }
\date{}

\maketitle

\textbf{Basics:}
\begin{itemize}
\item \textbf{Instructor:} Mitch Rudominer
\item \textbf{e-mail:} rudominer@sfsu.edu
\item \textbf{Lecture:}  T,Th 8:00 - 9:15 am in TH 325
\item \textbf{Office Hours:} After class, 9:15 - 10:00 in TH 934
\end{itemize}

\bigskip

\textbf{How to get help:}
\begin{itemize}
\item Ask a question in the class discussion forum on iLearn.
\item Send me an e-mail.
\item Come to office hours.
\item Ask a question during class. The first 5-10 minutes of every lecture will be devoted to questions about homework and previous lectures.
\end{itemize}

\bigskip

\textbf{iLearn:}
\begin{itemize}
\item There is lots of information on the class iLearn page. Use it frequently.
\item This document
\item Announcements
\item Lecture Notes
\item Homework Assignments
\item Homework Solutions
\item Calendar
\end{itemize}

\bigskip

\textbf{Text:} \emph{Applied Linear Algebra: Edition 2} by Olver and Shakiban.
\begin{itemize}
\item This is a very large, heavy book. Sorry.
\item You do \textbf{not} not to bring this book to class.
\item The lecture notes will serve as the primary text. The textbook will be a secondary text.
\item But I will sometimes assign reading and homework from the book so you need to have access to a copy.
\item If you buy the book online (Amazon or Google Play) you can get an electronic copy if you would prefer it.
\end{itemize}

\bigskip

\textbf{Lecture Notes:}
\begin{itemize}
\item The bulk of the lecture material will be presented as slides projected using the overhead projector.
\item I will also write auxiliary information on the board.
\item You do \textbf{not} need to copy down the slides because I have already given you a copy of the slides in iLearn. I would rather you pay attention
and try to understand in real time than spend your mental energy taking notes.
\item Feel free to copy down what I write on the board.
\end{itemize}

\bigskip

\textbf{Exams:}
\begin{itemize}
\item Regular in-class \textbf{quizzes} during the last 15 minutes of most classes.
\item \textbf{Midterm:} on Thursday October 26, in class.
\item \textbf{Final Exam:} Sat, Dec 15, 8:00 am- 10:15 am
\end{itemize}

\bigskip

\textbf{Homework:}
\begin{itemize}
\item Homework will be assigned every class.
\item The homework assignments are available in iLearn
\item The homework solutions are available in iLearn
\item You should do all of the problems and then use the solutions to check all of your answers.
\item You should make sure you understand every problem and solution and ask questions if you do not.
\item The homework will not be graded, but you will want to do the homework because the quizzes are based directly on the homework and the quizzes are graded.
\end{itemize}

\bigskip

\textbf{Grades:} Percentages are flexible so as to give you the highest possible grade.
\begin{itemize}
\item Quizzes 50\% ($\pm5\%$) (I will drop the lowest 4 quizzes.)
\item Midterm 20\% ($\pm5\%$)
\item Final: 30\% ($\pm5\%$)
\item Homework: Self-corrected. Not graded but the quizzes are based directly on the homework.
\item Letter grades will be assigned based on straight numerical grades
without curving and using the usual mapping of number grades to letter grades:

\begin{center}
  \begin{tabular}{ | l | l |}
    \hline
    Number Grade & Letter Grade  \\ \hline \hline
    97 -- 100 & A+  \\ \hline
    93 -- 96 & A\\ \hline
    90 -- 92 & A- \\ \hline
    87 -- 89 & B+ \\ \hline
    83 -- 86 & B \\ \hline
    etc & etc \\ \hline
  \end{tabular}
\end{center}

\end{itemize}

\textbf{Add/Drop:}
\begin{itemize}
\item Last day to add/drop is Monday September 17. After this day, in order to withdraw you need my permission, and there needs to be a serious and compelling reason to do so.
\item Last day to withdraw is Saturday November 17. After this day withdrawal is permitted only in emergencies and you must provide written documentation.
\item Absolutely  no withdrawals will be permitted after December 13, the last day of class.
\item  The deadline for deciding to audit the course Monday September 17.
\item The deadline for the CR/NC option is Monday October 22.
\end{itemize}

\end{document}
