% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 3 \\ Quotients and Remainders}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{The Quotient-Remainder Theorem}


\begin{itemize}
  \item \textbf{Theorem:} Let $a$ and $b$ be integers with $b>0$.
  \item Then there are unique integers $q$ and $r$ such that
  \item $a=bq+r$ and
  \item $0\leq r < b$.
  \item Example: Let $a=20$ and $b=3$.
  \item $20 = 3 \times 6 + 2$
  \item So $q=6$ and $r=2$.
  \item ``3 goes into 20 6 times with 2 left over.''
  \item $q$ is the \emph{quotient}, $r$ is the \emph{remainder}.
  \item Sometimes people call this theorem \emph{The Division Algorithm}.
  But this is a misnomer because we are not describing any algorithm. Only
  a theorem.
\end{itemize}

\end{frame}

\begin{frame}{Negatives in the Quotient-Remainder Theorem}


\begin{itemize}
  \item The theorem says: Let $a$ and $b$ be integers with $b>0$.
  \item Then there are unique integers $q$ and $r$ such that
  $a=bq+r$ and $0\leq r < b$.
  \item So $b$ is positive and $r$ is non-negative. But $a$ and $q$ can be negative.
  \item Example: Find $q$ and $r$ if $a=-20$ and $b=3$.
  \item $-20 = 3 \times -7 + 1$.
  \item So $q=-7$ and $r=1$.
\end{itemize}

\end{frame}

\begin{frame}{Remainder can be zero}


\begin{itemize}
  \item Let $a$ and $b$ be integers with $b>0$.
  \item Let $q$ and $r$ be the quotient and remainder from the theorem.
  \item Then $b\divides a$ iff $r=0$.
  \item Example: Let $a=20$ and $b=3$.
  \item Then $20 = 3 \times 6 + 2$.
  \item Since $r=2$, $3\ndivides 20$.
  \item Example: Let $a=35$ and $b=5$.
  \item $35 = 5 \times 7 + 0$.
  \item So $r=0$ and $5\divides 35$.
\end{itemize}

\end{frame}

\begin{frame}{The Floor Function}

\begin{itemize}
  \item \textbf{Definition.} Let $x$ be any real number. Then $\floor{x}$ is the greatest integer less than or equal to $x$.
  \item $\floor{x}$ is pronounced ``floor of $x$''.
  \item Some authors use the notation $[x]$ instead of $\floor{x}$. You can use either notation.
  \item Example: What is $\floor{3.14159}$?
  \item $\floor{3.14159} = 3$.
  \item Example: What is $\floor{-3.14159}$?
  \item $\floor{-3.14159} = -4$.
  \item Example: What is $\floor{411}$?
  \item $\floor{411} = 411$.
\end{itemize}

\end{frame}

\begin{frame}{How do you find $q$?}

\begin{itemize}
  \item Let $a$ and $b$ be integers with $b>0$.
  \item Let $q$ and $r$ be the quotient
  and remainder when $a$ is divided by $b$.
  \item So $a = bq + r$ and $0\leq r < b$.
  \item Then $q = \floor{\frac{a}{b}}$.
  \item Example: Let $a = 17$ and $b=5$.
  \item $17 = 5 \times 3 + 2$ so $q=3, r=2$.
  \item $\frac{17}{5} = 3  + \frac{2}{5} = 3.4$
  \item $3 = \floor{3.4} = \floor{\frac{17}{5}}$.
\end{itemize}

\end{frame}

\begin{frame}{How do you find negative $q$?}

\begin{itemize}
  \item Example: Let $a = -17$ and $b=5$.
  \item $-17 = 5 \times (-4) + 3$ so $q=-4, r=3$.
  \item $\frac{-17}{5} = -4  + \frac{3}{5} = -3.4$
  \item $-4 = \floor{-3.4} = \floor{\frac{-17}{5}}$.
\end{itemize}

\end{frame}

\begin{frame}{The mod function}

\begin{itemize}
 \item \textbf{Definition.} Let $a$ and $b$ be integers with $b>0$.
 \item Then $a\bmod b$ is the remainder when $a$ is divided by $b$
 \item That is $a\bmod b =r$
 \item where $a = bq + r$ and $0\leq r < b$.
 \item Example: Find $17 \bmod 5$.
 \item $17 = 5 \times 3 + 2$. So $17 \bmod 5 = 2$
 \item Example: Find $-17 \bmod 5$.
 \item $-17 = 5 \times -4 + 3$. So $-17 \bmod 5 = 3$
\end{itemize}

\end{frame}

\begin{frame}{How do you find $a\bmod b$?}

\begin{itemize}
\item Let $a$ and $b$ be integers with $b>0$.
\item Let $q$ and $r$ be the quotient
  and remainder when $a$ is divided by $b$.
 \item So $a = bq + r$ and $0\leq r < b$.
 \item Then $r = a - bq$.
 \item So $a\bmod b =  a - b \floor{\frac{a}{b}}$.
 \item Example: Find $987654321 \bmod 123456789$.
 \item Want $q,r$ such that $987654321 = 123456789 q + r$.
 \item $q = \floor{\frac{987654321}{123456789}} = 8$.
 \item $r = 987654321 - 123456789 \cdot 8 = 9$.
 \item $987654321 = 123456789 \times 8 + 9$.
 \item So $987654321 \bmod 123456789 = 9$.
\end{itemize}

\end{frame}


\begin{frame}{Proof of the Quotient-Remainder Theorem}


\begin{itemize}
  \item \textbf{Proof:} Let $a$ and $b$ be integers with $b>0$.
  \item Let $S = \setof{a-tb}{t\in\Z\text{ and }a-tb\geq 0}$.
  \item By definition $S$ is a set of non-negative integers. We want to use
  the wellordering property get the least element of $S$.
  \item First we need to see that $S\not=\emptyset$.
  \item If $a\geq 0$ then letting $t=0$ we have $a-tb = a \geq 0$ so $a-tb\in S$.
  \item If $a < 0$ then letting $t=a$ we have $a-tb=a-ab = a (1-b) \geq 0$. So $a-tb\in S$.
  \item So $S$ is not empty. Let $r$ be the least element of $S$ and let $q$ be such that $r=a-qb$.
  \item So $a=bq+r$ as we are aiming for. We must see that $0\leq r < b$. We have $r\geq 0$ by definition since $r\in S$.
\end{itemize}

\end{frame}

\begin{frame}{Existence proof continued}


\begin{itemize}
  \item Summary: $S = \setof{a-tb}{t\in\Z\text{ and }a-tb\geq 0}$.
  \item $r$ is the least element of $S$. $r=a-qb$.
  \item To see that $r < b$, suppose $r \geq b$ and let $\rprime = r -b \geq 0$.
  \item Now $\rprime  = (a-qb) -b = a-(q+1)b$. So $\rprime \in S$.
  \item This contradicts the choice of $r$ as the least element of $S$.
  \item So $r < b$. This proves existence.
\end{itemize}

\end{frame}

\begin{frame}{Uniqueness proof}


\begin{itemize}
  \item To prove uniqueness, suppose we had $q_1, q_2, r_1, r_2$ such that
  \item $a = b q_1 + r_1$ and $a = b q_2 + r_2$ and
  \item $0\leq r_1 < b$ and $0 \leq r_2 < b$.
  \item Then $b q_1 + r_1 = b q_2 + r_2$.
  \item WLOG assume $r_1 \leq r_2$.
  \item Now $r_2 - r_1 = b (q_1 - q_2)$.
  \item So $b \divides (r_2 - r_1)$.
  \item But $0\leq (r_2 - r_1) < b$. So $r_2 - r_1 = 0$, i.e. $r_2 = r-1$.
  \item Since $b (q_1 - q_2) = (r_2 - r_1) = 0$, by cancellation $q_1 - q_2 = 0$ so $q_1 = q_2$.
  \item This proves uniqueness. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Existence and Uniqueness Proofs}


\begin{itemize}
  \item The previous proof was an example of an \emph{existence and uniqueness} proof.
  \item This is a proof of a statement of the form \emph{there exists a unique x such that ...}.
  \item In order to prove such a statement there are always two parts to the proof.
  \item In the existence part of the proof we must show that there is \emph{at least one x such that ...}
  \item In the uniqueness part of the proof we showed that there is \emph{not more than one}.
\end{itemize}

\end{frame}

\end{document}
