\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for 11 \\ Linear Independence and Span}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Fall 2018 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.} No $T_A$ is not one-to-one. You can see this because
the columns of $A$ are not linearly independent. The second column is
-4 times the first column.

\bigskip

\textbf{Problem 2.} Yes $T_A$ is one-to-one. You can see this because
the columns of $A$ are linearly independent. The second column is
not a multiple of the first column.

\bigskip

\textbf{Problem 3.} Yes the three vectors are linearly independent. You can
see this because if you arrange them in a matrix as
$A = (\bv_2 \vert \bv_3 \vert \bv_1)$ then $A$ is upper-triangular with
nonzeros on the diagonal.

\bigskip

\textbf{Problem 4.} No, the three vectors are linearly dependent.
$\bv_3 = 2\bv_1 + \bv_2$.

\bigskip

\textbf{Problem 5.} Let
$A = \left(\bv_3 \vert \bv_2 \vert \bv_1 \right) =
\begin{pmatrix}
1 & -1 & 2\\
0 &  2 & 1 \\
0 &  0 & 1  \\
\end{pmatrix}
$.
Then the problem is equivalent to finding an $\bx$ such that $A\bx=\bb$.
But $A$ is upper triangular with nonzeros on the diagonal, so we can
solve the equation $A\bx=\bb$ using the method of back-substitution.

$$
\begin{pmatrix}
1 & -1 & 2 \\
0 &  2 & 1 \\
0 &  0 & 1  \\
\end{pmatrix}
\begin{pmatrix}
x \\
y \\
z \\
\end{pmatrix}
=
\begin{pmatrix}
-2 \\
12 \\
2 \\
\end{pmatrix}
$$


Working our way up from the bottom row we get:
\begin{itemize}
\item $z = 2$
\item $2y + z = 12$ so $2y + 2 = 12$ so $y=5$.
\item $x - y + 2z = -2$ so $x - 5 + 4 = -2$ so $x= -1$.
\item The solution is $x=-1, y=5, z= 2$.
\end{itemize}

We check the solution by multiplying

$$
\begin{pmatrix}
1 & -1 & 2 \\
0 &  2 & 1 \\
0 &  0 & 1  \\
\end{pmatrix}
\begin{pmatrix}
-1 \\
5 \\
2 \\
\end{pmatrix}
=
\begin{pmatrix}
-2 \\
12 \\
2 \\
\end{pmatrix}
$$

Remember that we reordered the vectors of the original problem when we
formed the matrix $A$. The solution to the original version of the problem is
that $\bb = 2\bv_1  + 5 \bv_2 -\bv_3$.

\bigskip

\textbf{Problem 6.}  No. You can see this because the second components
of $\bv_1$ and $\bv_2$ are both zero so that would be true also for any linear
combination of $\bv_1$ and $\bv_2$. But the second coordinate of $\bb$ is 5,
not 0, so therefore $\bb$ cannot be written as a linear combination of $\bv_1$
and $\bv_2$.

\bigskip


\textbf{Problem 7.}

\bigskip

$
\begin{pmatrix}
-1 \\ 2 \\ 3
\end{pmatrix}
=
2
\begin{pmatrix}
2\\ -1 \\ 2
\end{pmatrix}
-
\begin{pmatrix}
5 \\ -4 \\ 1
\end{pmatrix}
$
\bigskip


\textbf{Problem 8.}
Yes it is in the span because

\bigskip

$
\begin{pmatrix}
1 \\ -2 \\ -3
\end{pmatrix}
=
\begin{pmatrix}
1 \\ 1 \\ 0
\end{pmatrix}
-
3
\begin{pmatrix}
0 \\ 1 \\ 1
\end{pmatrix}
$

\bigskip


\textbf{Problem 9.} True. Because $\bzero = 0 \bv$ for
any vector $\bv$.

\bigskip

\textbf{Problem 10.} This is false. For example in $\R^3$
let $\bz = \bzero$ and let $\bu,\bv,\bw$ be the three
standard basis vectors. Then by the previous problem
$\bzero$ is a linear combination of $\be_1,\be_2,\be_3$, but it is not true the $\be_3$ is a linear combination of $\be_1$, $\be_2$ and $\bzero$.


\bigskip


\textbf{Problem 11.} Do exercise 2.3.21 parts (a), (b), (d)

\textbf{(a)} Linearly independent.

\bigskip

\textbf{(b)} Linearly dependent.
$
-2
\begin{pmatrix}
1 \\ 3
\end{pmatrix}
=
\begin{pmatrix}
-2 \\ -6
\end{pmatrix}
$
\bigskip

\textbf(d) Linearly independent.


\bigskip


\textbf{Problem 12.} False. The zero vector belongs to the span of every
set of vectors, whether or not those vectors are linearly independent.

\bigskip

\textbf{Problem 13.} The only way that a set consisting of a single vector
can be linearly dependent is if the single vector is the zero vector.

\bigskip

\textbf{Problem 14.} This is false. For example
$
\begin{pmatrix}
1 \\ 0
\end{pmatrix}
$
and
$
\begin{pmatrix}
2 \\ 0
\end{pmatrix}
$
do not span $\R^2$ and they are not linearly independent.



\end{document}
