\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 13 \\ Rank and Row Echelon Form}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Spring 2019 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.}
\textbf{(a)} Columns 2 and 4

\medskip

\textbf{(b)} 2

\medskip

\textbf{(c)} Any of the following:
(i) Columns 3 and 4
(ii) Columns 2 and 5
(iii) Columns 3 and 5
(iv) Columns 4 and 5

\bigskip

\textbf{Problem 2.}

\medskip


\medskip
\textbf{(a)} Not in row-echelon form

\medskip

\textbf{(b)} In row-echelon form. Column 3 is the only pivot column.
$\rank(A)=1$.

\medskip

\textbf{(c)} Not in row-echelon form.

\medskip

\textbf{(d)} In row-echelon form. The pivot columns are 1,3,6.
$\rank(A)=3$.

\medskip

\textbf{(e)} In row-echelon form. The pivot columns are 2,3,6.
$\rank(A)=3$.

\medskip

\textbf{(f)} In row-echelon form. The pivot columns are 2,3,6.
$\rank(A)=3$.

\medskip

\textbf{(g)} In row-echelon form. The pivot columns are 2,3,6.
$\rank(A)=3$.

\medskip

\textbf{(h)} In row-echelon form. The pivot columns are 1,2.
$\rank(A)=2$.

\bigskip
\bigskip

\textbf{Problem 3.} Yes. All of the columns are pivot columns and the rank is 10.

\bigskip
\bigskip

\textbf{Problem 4.} False. $\diag(1,0,1)$ is a counterexample.

\bigskip
\bigskip

\textbf{Problem 5.} True. If the row vector is the zero vector than it has no
pivot columns and its rank is zero. Otherwise the pivot column is the
index of the first nonzero entry and the rank is one.

\bigskip
\bigskip

\textbf{Problem 6.} False. Only column vectors with a nonzero in the first component
and zeros in every other component are in row-echelon form.


\end{document}
