% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,xcolor}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 15 \\ Applications of Gaussian Elimination}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{General Systems of Linear Equations}

\begin{itemize}
\item Suppose $A\bx=\bb$ is a general system of linear equations.
\item Recall that the possibilities for the solution set are:
\item (i) There are no solutions. The system is incompatible (or inconsistent.) $\bb\notin\ran(T_A)$.
\item (ii) There is exactly one solution. This happens when $T_A$ is one-to-one. ( $\ker(A)$ is trivial.)
\item (iii) There are infinitely many solutions. The system is underdetermined.
\item If $A$ is in row-echelon form then it is easy to see which of these three cases occur.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Incompatible systems}

\begin{itemize}
\item Suppose $A\bx=\bb$ is a general system of linear equations, and $A$ is in row-echelon form.
\item Then the system is incompatible iff
\item there is a row of $A$ that is all zeros but the corresponding element of $\bb$ is not zero.
\item Example:
$$
A =
\begin{pmatrix}
1 & 2 & 3 & 4 \\
0 & 0 & 5 & 6 \\
0 & 0 & 6 & 8 \\
0 & 0 & 0 & 0 \\
\end{pmatrix}
\quad
\bb=
\begin{pmatrix}
2 \\ 3 \\ 4 \\ 5
\end{pmatrix}
$$

\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Unique solutions}

\begin{itemize}
\item Suppose $A\bx=\bb$ is a general system of linear equations, and $A$ is in row-echelon form.
\item Then the system has a unique solution iff
\item $A$ is actually upper triangular with nonzeros on the diagonal.
\item Example:
$$
A =
\begin{pmatrix}
1 & 2 & 3 & 4 \\
0 & 5 & 6 & 7 \\
0 & 0 & 9 & 8 \\
0 & 0 & 0 & 7 \\
\end{pmatrix}
\quad
\bb=
\begin{pmatrix}
2 \\ 3 \\ 4 \\ 5
\end{pmatrix}
$$
\item We can find the unique solution using back-substitution.

\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Underdetermined systems}

\begin{itemize}
\item Suppose $A\bx=\bb$ is a general system of linear equations, and $A$ is in row-echelon form.
\item Then the system is underdetermined iff
\item neither of the previous two cases occur
\item Example:
$$
A =
\begin{pmatrix}
1 & 2 & 3 & 4 \\
0 & 0 & 5 & 6 \\
0 & 0 & 6 & 8 \\
0 & 0 & 0 & 9 \\
\end{pmatrix}
\quad
\bb=
\begin{pmatrix}
2 \\ 3 \\ 4 \\ 5
\end{pmatrix}
$$
\item We can find the infinitely many solutions through generalized back-substitution.

\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Generalized back-substitution}

\begin{itemize}
\item
\end{itemize}
\end{frame}


\end{document}


