\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Course Info}
\author{Math 310, Elementary Number Theory  \\ Fall 2019 \\ SFSU }
\date{}

\maketitle

\textbf{Basics:}
\begin{itemize}
\item \textbf{Instructor:} Mitch Rudominer
\item \textbf{e-mail:} rudominer@sfsu.edu
\item \textbf{Lecture:}  T,Th 8:00 - 9:15 am in TH 326
\item \textbf{Office Hours:} After class, 9:15 - 10:00 in TH 927. But you must show up at the beginning of office hours or else tell me to expect you later because if nobody shows up for my office hours I don't stick around--I leave campus.
\end{itemize}

\bigskip

\textbf{How to get help:}
\begin{itemize}
\item Ask a question in the class discussion forum on iLearn.
\item Send me an e-mail.
\item Come to office hours.
\item Ask a question during class. 
\end{itemize}

\bigskip

\textbf{iLearn:}
\begin{itemize}
\item There is lots of information on the class iLearn page. Use it frequently.
\item This document
\item Announcements
\item The class discussion forum. Register to receive e-mail notifications.
\item Lecture Notes
\item Homework Assignments
\item Homework Solutions
\end{itemize}

\bigskip

\textbf{Lecture Notes} 
\begin{itemize}
\item I will provide lecture notes on our iLearn site. This will be the primary source for the course.
\item But you will also need to have access to the textbooks.
\end{itemize}


\bigskip
\textbf{Text 1:} \emph{A Computational Introduction to Number Theory and Algebra, 2nd edition} by Victor Shoup.
\begin{itemize}
\item This book is available for free online at https://shoup.net/ntb/ntb-v2.pdf.
\item Optionally you can purchase it from amazon or your favorite online bookseller.
\item I will assign reading and exercises from this book so you must have access to it.
\end{itemize}


\bigskip

\textbf{Text 2:} \emph{Number Theory} by George E. Andrews.
\begin{itemize}
\item This book is available for purchase from the bookstore. It is also very cheap on Amazon.
\item It will be our secondary text but you must have a copy of it because I will also assign reading and exercises from it.
\end{itemize}

\bigskip

\textbf{Exams:}
\begin{itemize}
\item \textbf{Midterm:} Thursday October 24, in class.
\item \textbf{Final Exam:} Sat, Dec 14, 8:00 am- 10:15 am
\end{itemize}

\bigskip

\textbf{Homework:}
\begin{itemize}
\item Homework will be assigned every class.
\item It is due the following week. You may turn it in either Tuesday or Thursday at the beginning of class.
\item In unusual circumstances you may e-mail me your homework if you cannot come to class. But this must not be a common occurrence.
\item I will pick some of the exercises from the homework set to grade carefully.
\item You will get additional credit for finishing the other problems.
\item The homework assignments will be posted in iLearn.
\item No late homework will be accepted. After the due date, homework solutions will be posted in iLearn.
\item You may collaborate with your classmates on the homework, but you must write up your own homework to turn in.
\end{itemize}

\bigskip

\textbf{Grades:} Percentages are flexible so as to give you the highest possible grade.
\begin{itemize}
\item Homework 40\% ($\pm5\%$) (I will drop the 3 homeworks.)
\item Midterm 25\% ($\pm5\%$)
\item Final: 35\% ($\pm5\%$)
\item Letter grades will be assigned based on straight numerical grades
without curving and using the usual mapping of number grades to letter grades:

\begin{center}
  \begin{tabular}{ | l | l |}
    \hline
    Number Grade & Letter Grade  \\ \hline \hline
    93 -- 100 & A\\ \hline
    90 -- 92 & A- \\ \hline
    87 -- 89 & B+ \\ \hline
    83 -- 86 & B \\ \hline
    80 -- 82 & B- \\ \hline
    etc & etc \\ \hline
  \end{tabular}
\end{center}

\end{itemize}

\end{document}
