\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Syllabus}
\author{Math 310, Elementary Number Theory  \\ Fall 2019 \\ SFSU }
\date{}

\maketitle

\textbf{Basics:}
\begin{itemize}
\item \textbf{Instructor:} Mitch Rudominer
\item \textbf{e-mail:} rudominer@sfsu.edu
\item \textbf{Lecture:}  T,Th 8:00 - 9:15 am in TH 326
\item \textbf{Office Hours:} After class, 9:15 - 10:00 in TH 927. But you must show up at the beginning of office hours or else tell me to expect you later because if nobody shows up for my office hours I don't stick around--I leave campus.
\end{itemize}

\bigskip

\textbf{Main topics of this course:}
The theory of the integers: divisibility, congruence,
primitive roots, quadratic residues, quadratic reciprocity.


\bigskip

\textbf{Learning outcomes:}
When a student has successfully completed this course they will be able to:
\begin{itemize}
\item do number-theoretic proofs
\item explain and be able to use the theorems of Fermat, Wilson and Euler
\item given an integer $n$, determine whether or not there are any primitive
roots modulo $n$ and if so, how many
\item perform the extended Euclidean algorithm and the Chinese remainder theorem algorithm
\item express the greatest common divisor of two integers $a$ and $b$ as an integer
linear combination of $a$ and $b$
\item solve systems of simultaneous linear congruences
\item determine  whether  a  quadratic congruence has solutions, and if so, find them
\end{itemize}

\bigskip

\textbf{How to get help:}
\begin{itemize}
\item Ask a question in the class discussion forum on iLearn.
\item Send me an e-mail.
\item Come to office hours.
\item Ask a question during class.
\end{itemize}

\bigskip

\textbf{iLearn:}
\begin{itemize}
\item There is lots of information on the class iLearn page. Use it frequently.
\item This document
\item Announcements
\item The class discussion forum. Register to receive e-mail notifications.
\item Lecture Notes
\item Homework Assignments
\item Homework Solutions
\end{itemize}

\bigskip

\textbf{Lectures}
\begin{itemize}
\item This will be a traditional lecture format course.
\item I will provide \emph{lecture notes} on our iLearn site. This will be the primary source for the course.
\item But you will also need to have access to the textbooks.
\end{itemize}

\bigskip

\textbf{Text 1:} \emph{Number Theory} by George E. Andrews.
\begin{itemize}
\item This book is available for purchase from the bookstore. It is also very cheap on Amazon.
\end{itemize}

\bigskip

\textbf{Text 2:} \emph{A Computational Introduction to Number Theory and Algebra, 2nd edition} by Victor Shoup.
\begin{itemize}
\item This book is available for free online at https://shoup.net/ntb/ntb-v2.pdf.
\item Optionally you can purchase it from amazon or your favorite online bookseller.
\item I will assign reading and exercises from this book so you must have access to it.
\end{itemize}

\bigskip

\textbf{Exams:}
\begin{itemize}
\item \textbf{Midterm:} Thursday October 24, in class.
\item \textbf{Final Exam:} Sat, Dec 14, 8:00 am- 10:15 am
\end{itemize}

\bigskip

\textbf{Homework:}
\begin{itemize}
\item Homework will be assigned every class.
\item It is due the following week. You may turn it in either Tuesday or Thursday at the beginning of class.
\item In unusual circumstances you may e-mail me your homework if you cannot come to class. But this must not be a common occurrence.
\item I will pick some of the exercises from the homework set to grade carefully.
\item You will get additional credit for finishing the other problems.
\item The homework assignments will be posted in iLearn.
\item No late homework will be accepted. After the due date, homework solutions will be posted in iLearn.
\item You may collaborate with your classmates on the homework, but you must write up your own homework to turn in.
\end{itemize}

\bigskip

\textbf{Grades:} Percentages are flexible so as to give you the highest possible grade.
\begin{itemize}
\item Homework 40\% ($\pm5\%$) (I will drop the lowest 3 homework grades.)
\item Midterm 25\% ($\pm5\%$)
\item Final: 35\% ($\pm5\%$)
\item Letter grades will be assigned based on straight numerical grades
without curving and using the usual mapping of number grades to letter grades:

\begin{center}
  \begin{tabular}{ | l | l |}
    \hline
    Number Grade & Letter Grade  \\ \hline \hline
    93 -- 100 & A\\ \hline
    90 -- 92 & A- \\ \hline
    87 -- 89 & B+ \\ \hline
    83 -- 86 & B \\ \hline
    80 -- 82 & B- \\ \hline
    etc & etc \\ \hline
  \end{tabular}
\end{center}

\end{itemize}

\textbf{Policy on WU grades and International Students:}

WU (Withdrawal Unauthorized) indicates that an enrolled student did not withdraw from the course and also failed to complete course requirements. It is used when, in the opinion of the instructor, completed assignments or course activities or both were insufficient to make normal evaluation of academic performance possible. For purposes of grade point average and progress point computation, this symbol is equivalent to an F.

International students who have enrollment unit requirements for their stay in the US need to be aware that a course with a WU grade is not viewed as a course attempted whereas a course with an F grade is considered to be a course attempted. The Mathematics Department will not approve petitions to retroactively change a grade of WU to a grade of F.

\textbf{Disability access:}
Students with disabilities who need reasonable accommodations are encouraged to contact
the instructor.  The Disability Programs and Resource Center (DPRC) is available to facilitate
the reasonable accommodations process. The DPRC is located in the Student Service Building
and can be reached by telephone (voice/415-338-2472, video phone/415-335-7210) or by
email (dprc@sfsu.edu).

\textbf{Student disclosures of sexual violence}
SF State fosters a campus free of sexual violence including sexual harassment, domestic
violence, dating violence, stalking, and/or any form of sex or gender discrimination.  If you
disclose a personal experience as an SF State student, the course instructor is required to
notify the Title IX Coordinator by completing the report form available
at http://titleix.sfsu.edu, emailing vpsaem@sfsu.edu or calling 338-2032.

\textbf{To disclose any such violence confidentially, contact:}
\begin{itemize}
\item The SAFE Place - (415) 338-2208; http://www.sfsu.edu/\~safe\_plc/
\item Counseling and Psychological Services Center - (415) 338-2208; http://psyservs.sfsu.edu/
\item For more information on your rights and available resources:  http://titleix.sfsu.edu
\end{itemize}

\end{document}
