% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 12 \\ Bases and Dimension}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Basis}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item We say that $\bv_1,\bv_2,\cdots,\bv_n$ is a \emph{basis} for $V$ if
\item (i) $\bv_1,\bv_2,\cdots,\bv_n$ are linearly independent
\item (ii) $\bv_1,\bv_2,\cdots,\bv_n$ span $V$.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example of a basis}

\begin{itemize}
\item The standard basis vectors of $\R^n$ are a basis for $\R^n$.
\item Let $V = \R^3$.
\item Let $\be_1 =
\begin{pmatrix}
1 \\ 0 \\ 0
\end{pmatrix}
$,
$\be_2 =
\begin{pmatrix}
0 \\ 1 \\ 0
\end{pmatrix}
$,
$\be_3=
\begin{pmatrix}
0 \\ 0 \\ 1
\end{pmatrix}
$.

\item Then $\be_1,\be_2,\be_3$ are a basis for $\R^3$.
\item To see this we need to recall that they are
\item (i) linearly independent, and
\item (ii) span $\R^3$.
\item So the standard basis vectors form a basis.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Dimension}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space
\item Suppose $V$ has a basis consisting of $n$ vectors.
\item Then we say that the \emph{dimension} of $V$ is $n$.
\item Or that $V$ is $n$-dimensional.
\item \textbf{Example.} $\R^n$ is $n$-dimensional.
\item Because $\R^n$ has a basis consisting of the $n$ standard basis vectors.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Is dimension well-defined?}

\begin{itemize}
\item Let $V$ be an $n$-dimensional vector space.
\item So $V$ has a basis consisting of $n$ vectors.
\item Question: Is it possible that $V$ also has a different basis
consisting of $m$ vectors, for some $m\not=n$?
\item If this is possible, then $V$ is also $m$-dimensional.
\item And so the dimension of $V$ is not well-defined.
\item We will show that this cannot happen.
\item We will show that every basis for $V$ has the same size.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Dimension is well-defined}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be a basis for $V$.
\item Let $\bw_1,\bw_2,\cdots,\bw_n$  be any $n$ linearly independent vectors of $V$.
\item Then $\bw_1,\bw_2,\cdots,\bw_n$ are also a basis.
\item \textbf{Proof.} Let's just deal with the two dimensional case.  Let $V$ be a two dimensional
vector space with basis $\bv_1,\bv_2$.
\item Let $\bw_1,\bw_2$ be linearly
independent. We will show they also form a basis.
\item We need to show that the span of $\bw_1,\bw_2$ is all of $V$.
\item Notice that it suffices to show that $\bv_1,\bv_2\in\spn(\bw_1,\bw_2)$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof continued.}

\begin{itemize}
\item $\bw_1 = c_1\bv_1 + c_2\bv_2$, for some scalars $c_1, c_2$.
\item Since $\bw_1\not=\bzero$, either $c_1$ or $c_2$ is nonzero. Say $c_1 \not= 0$.
\item So then $\bv_1 = (1/c_1)\bw_1 - (c_2/c_1)\bv_2$.
\item So $\bv_1\in\spn(\bw_1,\bv_2)$.
\item But then $\bv_1,\bv_2\in\spn(\bw_1,\bv_2)$.
\item So $\spn(\bw_1,\bv_2)=V$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof continued.}

\begin{itemize}
\item $\bw_2 = c_1\bw_1 + c_2\bv_2$, for some scalars $c_1, c_2$.
\item Since $\bw_1, \bw_2$, are linearly independent, $c_2 \not= 0$.
\item So then $\bv_2 = (1/c_2)\bw_2 - (c_1/c_2)\bw_1$.
\item So $\bv_2$ is in the span of $\bw_1$ and $\bw_2$.
\item So $V=\spn(\bw_1,\bv_2)\subseteq\spn(\bw_1,\bw_2)$.
\item So $\spn(\bw_1,\bw_2) = V$.
\item So $\bw_1,\bw_2$ span all of $V$ and so they form a basis. $\qed$
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{More vectors than the dimension}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be a basis for $V$.
\item Let $\bw_1,\bw_2,\cdots\bw_m$ be any other vectors with $m > n$.
\item Then $\bw_1,\bw_2,\cdots\bw_m$ are linearly dependent.
\item \textbf{proof.} Suppose $\bw_1,\bw_2,\cdots\bw_m$ are linearly independent. We will derive a contradiction.
\item $\bw_1,\bw_2,\cdots,\bw_n,\cdots,\bw_m$ are linearly independent.
\item In particular $\bw_1,\bw_2,\cdots,\bw_n$ are linearly independent.
\item So by the previous lemma $\bw_1,\bw_2,\cdots,\bw_n$ are a basis
\item so then $\bw_{n+1}$ can be written as a linear combination of $\bw_1,\bw_2,\cdots,\bw_n$
\item but then $\bw_1,\bw_2,\cdots\bw_m$ are not linearly independent. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example}

\begin{itemize}
\item Let $A=
\begin{pmatrix}
1 & 2 & 3 & 4 \\
5 & 6 & 7 & 8 \\
9 & 10 & 11 & 12\\
\end{pmatrix}
$
\item Is $T_A$ one-to-one?
\item The columns of $A$ consist of four vectors in $\R^3$.
\item $\R^3$ is 3-dimensional.
\item By the previous lemma, a set of four vectors in a three dimensional space is necessarily linearly dependent.
\item So the columns of $A$ are linearly dependent.
\item So there is a nonzero $\bx$ such that $A\bx=\bzero$.
\item So $\ker(T_A)$ is not-trivial.
\item So $T_A$ is not one-to-one.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Underdetermined systems}

\begin{itemize}
\item Let $A$ be a rectangular matrix with more columns than rows.
\item Then for any $\bb$ the system $A\bx=\bb$ is always underdetermined: There is not a unique solution and in fact there are infinitely many solutions.
\item If there are more unknowns than equations there is never a unique solution.
\item $T_A$ is always not one-to-one.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Dimension is well-defined}

\begin{itemize}
\item \textbf{Theorem.} Let $V$ be a vector space and suppose $V$ has two different bases,
\item one with $n$ elements and one with $m$ elements.
\item Then $n=m$.
\item i.e. all bases for $V$ have the same number of elements. So the dimension of $V$ is well-defined.
\item \textbf{Proof.} This follows from the previous lemma
\item because if one basis had fewer
elements than the other one, then the one with more elements would be linearly dependent. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Other bases for $R^n$}

\begin{itemize}
\item The standard basis vectors form a basis for $\R^n$.
\item But $\R^n$ has other bases besides the standard basis.
\item But every basis for $\R^n$ has $n$ elements.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: The columns of an upper-triangular matrix.}

\begin{itemize}
\item The columns of an upper-triangular matrix with nonzeros on the
diagonal always form a basis.
\item Let
$
\bv_1 =
\begin{pmatrix}
-4 \\ 0 \\ 0
\end{pmatrix}
$,
$
\bv_2 =
\begin{pmatrix}
3 \\ -8 \\ 0
\end{pmatrix}
$,
$
\bv_3 =
\begin{pmatrix}
-6 \\ 11 \\ -37
\end{pmatrix}
$
\item Then $\bv_1,\bv_2,\bv_3$ form a basis for $\R^3$. Why?
\item They are linearly independent, because the matrix is upper-triangular with nonzero diagonal entries.
\item But then by an earlier lemma, in a three dimensional space, any set of three linearly independent
vectors form a basis.
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Dimension of subspaces}

\begin{itemize}
\item \textbf{Theorem.} Let $V$ be an $n$-dimensional vector space.
\item Let $W$ be a subspace of $V$.
\item Then $\dim(W) \leq \dim(V)$.
\item \textbf{proof}
\item By a previous lemma, any collection of more than $n$ vectors in $V$ is linearly dependent.
\item So then any collection of more than $n$ vectors of $W$ is linearly dependent and so does not form a basis.
\item So we know that the dimension of $W$ cannot be greater than the dimension of $V$.
\item So why aren't we done with the proof yet?
\item How do we know $W$ has a basis at all? That's what we still need to show.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{proof, continued}

\begin{itemize}
\item If $W$ is the zero space then it has dimension zero so we are done.
\item So now assume that $W$ is not the zero space. So let $\bw_1\in W$, $\bw_1\not=\bzero$.
\item If $\bw_1$ spans $W$ then $\singleton{\bw_1}$ is a basis for $W$ and so then $W$ is one dimensional and we are done.
\item So suppose $\bw_1$ does not span $W$. Let $\bw_2\in W$, $\bw_2\notin\spn(\bw_1)$.
\item So then $\singleton{\bw_1,\bw_2}$ are linearly independent.
\item If $\singleton{\bw_1,\bw_2}$ is a spanning set for $W$ then  $\singleton{\bw_1,\bw_2}$ is a basis and
so $W$ is two dimensional and we are done.
\item So suppose $\bw_3\in W$, $\bw_3\notin\spn(\bw_1,\bw_2)$.
\end{itemize}

\end{frame}


\end{document}


