% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 5 \\ GCD}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{GCD}

\begin{itemize}
  \item \textbf{Definition.} Let $a,b$ be any integers.
  \item Then $x$ is a \emph{common divisor} of $a$ and $b$ iff $x\divides a$ and $x\divides b$.
  \item Example: $4$ is a common divisor of 16 and 12 because $4\divides 16$ and $4\divides 12$.
\end{itemize}

\begin{itemize}
  \item \textbf{Definition.} Let $a,b$ be any integers, not both zero.
  \item Then the \emph{greatest common divisor} of $a$ and $b$ is the largest integer that divides both $a$ and $b$.
  \item We will always say ``GCD'' instead of ``greatest common divisor".
  \item This is written $\gcd(a,b)$.
  \item Because $\gcd$ is so important in number theory this is often abbreviated to $(a,b)$.
\end{itemize}

\end{frame}

\begin{frame}{GCD Examples}

\begin{itemize}
  \item Example: Find $\gcd(12,18)$.
  \item Example: Find $(5, 30)$.
  \item Example: Find $(-4, 6)$.
  \item Example: Find $\gcd(9, 10)$.
  \item Example: Find $\gcd(1, 30)$.
  \item Example: Find $\gcd(0,30)$.
  \item Example: Find $\gcd(0,0)$.
\end{itemize}

\end{frame}

\begin{frame}{$\gcd(0,0)$}

\begin{itemize}
  \item \textbf{Definition.} We define $\gcd(0,0)=0$.
  \item This may seem arbitrary. But later we will give a more
  sophisticated definition of GCD where this will naturally fall out.
\end{itemize}

\end{frame}

\begin{frame}{Relatively Prime}

\begin{itemize}
  \item \textbf{Definition.} Two non-zero integers $a$ and $b$ are called
  \emph{relatively prime} iff $\gcd(a,b)=1$.
  \item Example: $\gcd(9,25) = 1$ so 9 and 25 are relatively prime.
\end{itemize}

\end{frame}

\begin{frame}{Sneak Peak}

\begin{itemize}
  \item We will study GCD more in later lessons. Here is a sneak peak at two
  things we will learn about it.
  \item \textbf{FACT:} Suppose $a$ and $b$ are relatively prime. Then there
  is an integer linear combination of $a$ and $b$ that equals 1.
  \item  Example: We saw that 9 and 25 are relatively prime. So there is an
  integer linear combination of 9 and 25 that equals 1.
  \item There are integers $x$ and $y$ such that $9x + 25y = 1$.
  \item $x=-11, y=4$.
  \item In fact it is an iff:
  \item \textbf{FACT:} $a$ and $b$ are relatively prime iff 1 can be expressed
  as an integer linear combination of $a$ and $b$.
  \item Example: Are 789,567,345 and 54321 relatively prime?
  \item How would you go about computing this?
  \item We will learn an algorithm that allows us to compute this efficiently.
\end{itemize}

\end{frame}


\end{document}



\end{document}
