% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 14 \\ Span}
\subtitle{Math 325, Linear Algebra \\ Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Linear Combinations of Vectors in a Vector Space}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item Let $c_1, c_2, \cdots c_k$ be $k$ scalars.
\item Then $$c_1 \bv_1 + c_2 \bv_2 + \cdots + c_k \bv_k$$ is
called a \emph{linear combination} of $\bv_1, \bv_2, \cdots \bv_k$.
The $c_i$ are called the \emph{coefficients} of the linear combination.
\item Using Sigma notation we would write this as
$$\sum_{i=1}^{k} c_i \bv_i$$
\item Example. Let $\bv = (2, 1, 0)^{\text{T}}$ and $\bw = (0, 1, 3)^{\text{T}}$ be vectors in $\R^3$.
Express $\bu = (6, 8, 15)^{\text{T}}$ as a linear combination of $\bv$ and $\bw$.
\item Answer: $\bu = 3 \bv + 5 \bw$.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Linear combination of the columns of a matrix}

\begin{itemize}
\item Let $\bv_1,\bv_2,\cdots,\bv_n \in \R^m$.
\item Let $A =\left( \bv_1 \vert \bv_2 \vert \cdots \vert \bv_n \right)$
be the $m\times n$ matrix whose columns are those vectors.
\item Let $\bc = \left(c_1, c_2,\cdots c_n\right)^{\text{T}}$ be a column vector.
\item Then $A\bc = c_1\bv_1 + c_2\bv_2 + \cdots + c_n\bv_n$.
\item In words: A matrix times a column vector yields the linear
combination of the columns of the matrix with coefficients given by
the vector.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Linear combination of the columns of a matrix}

\begin{itemize}
\item Example: Let
$A =
\begin{pmatrix}
1 & 2 \\
3 & 4 \\
\end{pmatrix}
$,
$\bc =
\begin{pmatrix}
2 \\
-3 \\
\end{pmatrix}
$
\item Then $A\bc$=
\item
$
\begin{pmatrix}
1 & 2 \\
3 & 4 \\
\end{pmatrix}
\begin{pmatrix}
2 \\
-3 \\
\end{pmatrix}
=
\begin{pmatrix}
-4 \\
-6 \\
\end{pmatrix}
$
\item
$ =
2
\begin{pmatrix}
1 \\
3 \\
\end{pmatrix}
-3
\begin{pmatrix}
2 \\
4 \\
\end{pmatrix}
$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Span}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item Then the \emph{span} of the vectors is the set of all linear combinations of them.
\item $\spn(\bv_1,\bv_2,\cdots\bv_n)=\setof{c_1\bv_1+c_2\bv_2+\cdots+c_n\bv_n}{c_1,c_2,\cdots,c_n\in\R}$.
\item \textbf{Example.} In $\R^3$, consider the span of the first two standard basis vectors.
\item $\spn(\be_1,\be_2) = \setof{c_1\be_1+c_2\be_2}{c_1,c_2,\in\R}$.
\item This is the $x$- $y$-plane in $\R^3$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Span is a subspace}

\begin{itemize}
\item \textbf{Lemma.} $\spn(\bv_1,\bv_2,\cdots\bv_n)$ is a subspace of $V$.
\item \textbf{Proof.} We must show that the span contains zero and is closed under vector addition
and scalar multiplication.
\item The span contains zero because $\bzero$ can always be expressed as the linear combination
of the $\bv_i$ in which all coefficients are zero.
\item Now suppose that $\bx$ and $\by$ are in the span and $a$ is a scalar.
\item we will show that $\bx+\by$ and $a \bx$ are in the span.
\item Write $\bx = c_1\bv_1 + c_2\bv_2 + \cdots c_n \bv_n$
\item and $\by = d_1\bv_1 + d_2\bv_2 + \cdots d_n \bv_n$.
\item Then $\bx + \by = (c_1 + d_1) \bv_1 + (c_2 +d_2)\bv)2 + \cdots + (c_n+d_n)\bv_n$.
\item And $a\bx = (ac_1)\bv_1 + (ac_2)\bv_2 + \cdots + (ac_n)\bv_n$.
\item So both $\bx+\by$ and $a\bx$ are in the span of $\bv_1,\bv_2,\cdots,\bv_n$.
\item So the span is a subspace of $V$. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Span of the columns of a matrix}

\begin{itemize}
\item Let $\bv_1,\bv_2,\cdots,\bv_n \in \R^m$.
\item Let $A =\left( \bv_1 \vert \bv_2 \vert \cdots \vert \bv_n \right)$
be the $m\times n$ matrix whose columns are those vectors.
\item Let $\bb\in\R^m$.
\item Then $\bb\in\spn(\bv_1,\bv_2,\cdots,\bv_n)$
\item iff the system of equations $A\bx=\bb$ is compatible
\item iff $\bb\in\ran(T_A)$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The image of $T_A$}

\begin{itemize}
\item \textbf{Theorem.} Let $A = \left(\ba_1 \vert \ba_2 \vert \cdots \vert \ba_n \right)$
be an $m\times n$ matrix.
\item Then $\ran(T_A)$ is equal to the span of $\ba_1,\ba_2, \cdots,\ba_n$.
\item \textbf{proof.} $\bb\in\ran(T_A)$ iff
\item there is a $\bc\in\R^n$ such that $A\bc=\bb$ iff
\item there is an $\bc=(c_1, c_2,\cdots,c_n)^{\text{T}}$ such that
$$c_1\ba_1 + c_2\ba_2+\cdots+c_n\ba_n = \bb.$$
\item iff $\bb\in\spn\left( \ba_1,\ba_2, \cdots,\ba_n\right) \qed$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example}

\begin{itemize}
\item Let
$A=
\begin{pmatrix}
2 & 1 &  3 \\
0 & -1 & 0 \\
0 & 0 &  4 \\
\end{pmatrix}
$,
$\bb=
\begin{pmatrix}
 -5 \\
 -2 \\
 -12 \\
\end{pmatrix}
$.
\item Is $\bb\in\ran(T_A)$?
\item Equivalently is $\bb$ in the span of the columns of $A$?
\item Equivalently, can you solve the system $A\bx=\bb$?
\item Notice that $A$ is upper-triangular with non-zeroes on the
diagonal.
\item So we can use the method of back substitution.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example continued}

\begin{itemize}
\item Solving
$
\begin{pmatrix}
2 & 1 &  3 \\
0 & -1 & 0 \\
0 & 0 &  4 \\
\end{pmatrix}
\begin{pmatrix}
x \\ y \\ z
\end{pmatrix}
=
\begin{pmatrix}
 -5 \\
 -2 \\
 -12 \\
\end{pmatrix}
$
\item $4z=-12$ so $z=-3$.
\item $-y = -2$ so $y=2$.
\item $2x +y +3z = -5$
\item $2x + 2  -9 = -5$
\item $x=1$
\item
$
\begin{pmatrix}
2 & 1 &  3 \\
0 & -1 & 0 \\
0 & 0 &  4 \\
\end{pmatrix}
\begin{pmatrix}
1 \\ 2 \\ -3
\end{pmatrix}
=
\begin{pmatrix}
 -5 \\
 -2 \\
 -12 \\
\end{pmatrix}
$
\item So the answer is yes, $\bb\in\ran(T_A)$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example}

\begin{itemize}
\item Let
$\bv_1 =
\begin{pmatrix}
-3 \\ 0 \\ 5
\end{pmatrix}
$,
$\bv_2 =
\begin{pmatrix}
1 \\ 0 \\ 0
\end{pmatrix}
$,
$\bv_3 =
\begin{pmatrix}
0 \\ 0 \\ -8
\end{pmatrix}
$.
\item Let
$\bb=
\begin{pmatrix}
5 \\ -3 \\0
\end{pmatrix}
$.
\item Is $\bb$ in the subspace of $\R^3$ spanned by $\bv_1,\bv_2,\bv_3$?
\item Equivalently, is there a solution to $A\bx=\bb$ where
$A= \left(\bv_1 \vert \bv_2 \vert \bv_3 \right)$?
\item Note that $A$ is not in upper-triangular form. So we can't just use the technique of back substitution.
\item But notice that all three of $\bv_1,\bv_2,\bv_3$ have a zero in the
second position.
\item So every linear combination of $\bv_1,\bv_2,\bv_3$ also has a zero in the
second position.
\item But $\bb$ does not have a zero in the second position.
\item So $\bb\notin\spn\left( \bv_1,\bv_2,\bv_3\right)$

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Right-hand side of a system of equations}

\begin{itemize}
\item The system of equations $A\bx=\bb$ has a solution iff $b\in\ran(T_A)$.
\item iff $b$ is in the subspace spanned by the columns of $A$.
\item If $\bb\notin\ran(T_A)$ then $A\bx=\bb$ is inconsistent (incompatible.)
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Spanning set of vectors}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item We say that $\bv_1,\bv_2,\cdots,\bv_n$ \emph{span} $V$
\item or that $\singleton{\bv_1,\bv_2,\cdots,\bv_n}$ is a \emph{spanning set} for $V$
\item if $V=\spn(\bv_1,\bv_2,\cdots,\bv_n)$.
\item In other words, if every vector of $V$ can be written as a linear combination of
 $\bv_1,\bv_2,\cdots,\bv_n$.
 \item \textbf{Example.} In $\R^n$, $\singleton{\be_1,\be_2,\cdots,\be_n}$ is a
 spanning set.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Facts about span}

\begin{itemize}
\item Let $W$ be a subspace of $V$ and suppose $\bv_1,\bv_2,\cdots,\bv_n\in W$.\
\item Then $\spn(\bv_1,\bv_2,\cdots,\bv_n)\subseteq W$.
\item \textbf{proof.} Because the subspace $W$ is closed under linear combinations. $\qed$
\item Suppose that $\bv_1,\bv_2,\cdots,\bv_n$ span all of $V$.
\item Then $W=V$.
\item i.e. If a subspace contains a set of vectors that span the whole space then
that subspace is the whole space.
\end{itemize}

\end{frame}


\end{document}


