% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 1 \\ The Integers}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Number Systems}

  \begin{itemize}
  \item $\Z = \singleton{\cdots -3,-2,-1,0,1,2,3,\cdots} = $ the integers
  \item $\Q = \setof{\frac{n}{m}}{n,m\in \Z, m\not= 0} = $ the rationals
  \item $\R = $ the reals
  \item $\Q = \setof{x\in\R}{\text{decimal expansion of $x$ is terminating or repeating}}$
  \item $\C = $ the complex numbers
  \item $\Zpos = \singleton{1,2,3,\cdots} =$ the positive integers
  \item $\Znonneg = \singleton{0,1,2,3,\cdots} =$ the non-negative integers
  \item  $\Zpos \subset \Znonneg \subset \Z \subset \Q \subset \R \subset \C$
  \end{itemize}


\end{frame}

\begin{frame}{Number Systems: Example}
\begin{itemize}
  \item  Give an example of an $x\in \Znonneg \setminus \Zpos$.
  \item $x=0$.
  \item  Give an example of an $x\in \Z \setminus \Z^{\geq 0}$.
  \item $x=-1$.
  \item  Give an example of an $x\in \Q \setminus \Z$.
  \item $x=1/2$.
  \item  Give an example of an $x\in \R \setminus \Q$.
  \item $x=\sqrt{2}$.
  \item How do you know $\sqrt{2}\notin \Q$?
  \item Give an example of $x\in\C \setminus \R$.
  \item $x=i$.
  \end{itemize}
\end{frame}

\begin{frame}{This class is about $\Z$}

\begin{itemize}
\item Elementary number theory studies the integers $\Z$.
\item Mostly we will be studying the positive integers $\Zpos$.
\item At first blush you might think that $\Zpos$ is not interesting.
\item After all when you visualize it, it looks very simple: just an infinite sequence of points.
How interesting could the theory be?
\item But it turns out there is a lot of complexity, even in the set of positive integers less than 100.
\item If we consider very large integers it becomes even more interesting and has applications
to computer cryptography.
\end{itemize}
\end{frame}

\begin{frame}{Axioms of the integers}

\begin{itemize}
  \item The integers are familiar objects you have known about for most of your life.
  \item Early in this course we will prove some facts about the integers
  that you have "known to be true" for many years:
  \begin{itemize}
    \item When you divide one integer by another you get a unique quotient and remainder.
    \item Every integer has a unique prime factorization.
  \end{itemize}
  \item The question comes up: What assumptions about the integers are we allowing
  ourselves to make when proving these ``elementar facts''.
  \item In the following few slides we will give the list of all assumptions
  about the integers that we will make in this class.
  \item Everything else we prove in this class may be derived from this list of assumptions.
\end{itemize}
\end{frame}

\begin{frame}{Axioms about addition}

\begin{itemize}
  \item $\left(\Z, +, 0\right)$ is an Abelian group. This means:
  \item  $+$ is a binary operation on $\Z$: For every $x,y\in\Z$, $x+y\in\Z$.
  \item  $+$ is commutative: $x+y=y+x$, for all $x,y\in\Z$.
  \item  $+$ is associative: $x + (y+z) = (x+y) +z$, for all $x,y,z\in\Z$.
  \item  $0$ is an additive identity element: $x+0 = x$, for all $x\in\Z$.
  \item  Additive inverses exist: For every $x\in\Z$ there is an integer called $-x$ such that $x+(-x) = 0$.
  \item  (Note: We write $x-y$ to mean $x+(-y)$.)
\end{itemize}

\end{frame}

\begin{frame}{Axioms about multiplication}

\begin{itemize}
  \item $\left(\Z, +, \times, 0, 1\right)$ is a commutative ring. This means (in addition to the previous axioms):
  \item  $\times$ is a binary operation on $\Z$: For every $x,y\in\Z$, $x\times y\in\Z$.
  \item (Note that $x\times y$ is also written $x\cdot y$ and $x y$.)
  \item  $\times$ is commutative: $x\cdot y= y \cdot x$, for all $x,y\in\Z$.
  \item  $\times $ is associative: $x \cdot (y\cdot z) = (x\cdot y) \cdot z$, for all $x,y,z\in\Z$.
  \item  $1$ is a multiplicative identity element: $x\cdot 1 = x$, for all $x\in\Z$.
  \item $1\not=0$.
  \item Distributive law: $x\cdot (y + z) = x \cdot y + x \cdot z$.
  \item Cancellation law: If $x\cdot z = y \cdot z$ and $z\not=0$ then $x=y$, for all $x,y,z\in\Z$.
  \item  (Note that we do not have multiplicative inverses.)
\end{itemize}

\end{frame}

\begin{frame}{The ordering on the integers}

\begin{itemize}
  \item We want to now introduce some definitions and axioms that describe the
  ordering on the integers.
  \item Attempt: For $x,y\in\Z$ define $x<y$ iff $y-x\in\Zpos$.
  \item Wait, there is a problem with this definition.
  \item $\Zpos = \setof{x\in\Z}{0<x}$.
  \item So $\Zpos$ is defined in terms of $<$, so we cannot define $<$ in terms of $\Zpos$.
  \item We need to introduce more axioms to make this work. We can either introduce
  axioms about $<$ directly, or we can introduce axioms about $\Zpos$ directly.
\end{itemize}
\end{frame}

\begin{frame}{Axioms about the positive integers}

\begin{itemize}
  \item There is a subset $\Zpos\subset\Z$ called the set of \emph{positive integers} such that:
  \item For every $x\in\Z$ exactly one of the following holds:
    \begin{itemize}
      \item $x=0$
      \item $x\in\Zpos$
      \item $-x\in\Zpos$
    \end{itemize}
  \item If $x,y\in\Zpos$ then $x+y\in\Zpos$ and $xy\in\Zpos$.
  \item (Now it makes sense to define $x<y$ iff $y-x\in\Zpos$.)
\end{itemize}

\end{frame}

\begin{frame}{The well-ordering axiom}

\begin{itemize}
  \item  $\Zpos$ is well-ordered by $<$. This means:
  \item Let $S$ be any non-empty subset of $\Z^+$. Then $S$ contains
  a least element.
  \item (Note this is not true of $\Z,\Q,\R,\C$.)
  \item (It is not true of $\R^+$.)
  \item (What about $\Znonneg$?)
\end{itemize}

\end{frame}

\begin{frame}{Simple facts about $+$, $-$, $\times$ and $<$}

\begin{itemize}
  \item Going forward, the only axiom we will ever mention explicitly is the last one: the well-ordering axiom.
  \item We will often use ``common, simple facts'' about integer arithmetic that follow
  from the axioms without bothering to explicitly mention them.
  \item Example: If $a<b$ and $b<c$ then $a < c$.
  \item Example: If $a<b$ and $c>0$ then $ac < bc$.
  \item Example: $-1\times a = -a$ for all $a\in\Z$.
  \item Example: 1 is the least positive integer.
\end{itemize}

\end{frame}

\begin{frame}{Example application of the well-ordering property}

\begin{itemize}
  \item Theorem: $\sqrt{2}$ is irrational.
  \item proof: Suppose that $\sqrt{2}$ were rational.
  \item Let $a,b$ be positive integers such that $\sqrt{2}=a/b$.
  \item Notice that $a=b\sqrt{2}$ is a positive integer.
  \item Let $S=\setof{k\sqrt{2}}{k,k\sqrt{2}\in\Zpos}.$
  \item $S$ is a set of positive integers. $S\not=\emptyset$ because $a\in S$.
  \item By the well-ordering property there is a least element of $S$, say $s$.
  \item So $s\in\Zpos$ and there is a $t\in\Zpos$ such that $s=t\sqrt{2}$.
  \item Let $\sprime = s(\sqrt{2} - 1)$. Notice that $0< \sprime<s$.
  \item We will get a contradiction by showing that $\sprime\in S$.
\end{itemize}

\end{frame}

\begin{frame}{Proof continued}

\begin{itemize}
  \item Remember $S=\setof{k\sqrt{2}}{k,k\sqrt{2}\in\Zpos}.$
  \item $s$ is the least element of $S$. $s=t\sqrt{2}$
  \item $\sprime = s(\sqrt{2} - 1)$.
  \item $0< \sprime<s$.
  \item We will get a contradiction by showing that $\sprime\in S$.
  \item $\sprime = s\sqrt{2} -s = t\sqrt{2}\sqrt{2} - s = 2t -s$.
  \item So $\sprime$ is a  positive integer.
  \item $\sprime = s\sqrt{2} -s = s\sqrt{2} - t\sqrt{2} = (s-t)\sqrt{2}$.
  \item So $\sprime\in S$.
  \item Contradiction. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Reading assignment}

\begin{itemize}
  \item \emph{What is Number Theory?} pages 1-3.
  \item Beginning of Chapter 1. page 5 - top of page 7.
  \item You can stop before the definition of \emph{algebraic} number.
  \item Appendix A: \emph{Axioms for the Set of Integers.} pages 605-606.
\end{itemize}

\end{frame}

\end{document}
