% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 17 \\ Euler's Theorem and Fermat's Little Theorem}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Euler's Theorem}

\begin{itemize}
  \item \textbf{Euler's Theorem.} Let $n>1$ be an integer. Let $a$ be relatively prime to $n$.
  \item Then $a^{\phi(n)} \equiv 1 \pmod n$.
  \item \textbf{Example.} Let $n=10$.
  \item $\phi(10)=4$.
  \item Suppose $\gcd(a,10)=1$.
  \item Euler's Thereom says that $a^4 \equiv 1 \pmod{10}$.
  \item \textbf{Examples.} 3,7, 9 are all relatively prime to 10.
  \item $3^4=81$.
  \item $7^4=201$
  \item $9^4 = 6561$.
  \item All of these are congruent to 1 mod 10.
  \item $2^4 = 16 \not\equiv 1 \pmod{10}$.
  \item But 2 is not relatively prime to 10 so this does not violate Euler's Theorem.
\end{itemize}

\end{frame}

\begin{frame}{More Euler's Theorem Examples}

\begin{itemize}
  \item \textbf{Example.} Let $n=9$.
  \item $\phi(9)=6$.
  \item Suppose $a$ is relatively prime to $9$.
  \item Euler's Thereom says that $a^6 \equiv 1 \pmod{9}$.
  \item \textbf{Examples.} 2 and 5 are relatively prime to 9.
  \item $2^6 = 64$
  \item $9\cdot 7 = 63$.
  \item So  $2^6 \equiv 1 \pmod 9$.
  \item $5^6=15,625$
  \item $9\cdot 1736 = 15,624$
  \item So $5^6\equiv 1\pmod 9$.
  \item $3^6 = 729$.
  \item $9\ndivides 728$.
  \item But $\gcd(3,9)\not=1$ so this does not violate Euler's Theorem.
\end{itemize}

\end{frame}

\begin{frame}{Example}
\begin{itemize}
\item Example: Find the last digit in the decimal expansion of $7^{4000000002}$.
\item Solution: If $n$ is any positive integer, the last digit in the decimal
expansion of $n$ is $n\bmod 10$.
\item So we must find $7^{4000000002} \bmod 10$.
\item 7 is relatively prime to 10 so we can apply Euler's Theorem.
\item $\phi(10) = 4$, so $7^4 \bmod 10 = 1$.
\item $7^{4000000000} = (7^4)^{1000000000}$.
\item $7^{4000000002} = 7^{400000000} \cdot 7^2$.
\item So $7^{4000000002} \equiv 7^2 \pmod{10}$.
\item $7^{4000000002} \bmod 10 = 7^2 \bmod 10 = 49 \bmod 10 = 9$.
\item So the last digit in the decimal expansion of $7^{4000000002}$ is 9.
\end{itemize}
\end{frame}


\begin{frame}{Fermat's Little Theorem}

\begin{itemize}
  \item When $p$ is prime, $\phi(p)=p-1$. As a special case of Euler's Theorem we have:
  \item \textbf{Fermat's Little Theorem.} Let $p$ be prime. Let $a$ be an integer not divisible by $p$.
  \item Then $a^{p-1}\equiv 1 \pmod{p}$.
  \item \textbf{Example.} Let $p=11$. Let $a$ be an integer such that $11\ndivides a$. Then $a^{10} \equiv 1 \pmod{11}$.
  \item $2^{10} = 1024$ and $1023 = 11 \cdot 93$.
  \item $3^{10} = 59049$ and $59048 = 11 \cdot 5368$.
  \item $4^{10} = 1048576$ and $1048575 = 11 \cdot 95325$.
  \item $15^{10} = 576650390625$ and  $576650390624 = 11 \cdot 52422762784$.
  \item What happens if $p\divides a$?
  \item $22^{10} = 26559922791424$.
  \item Since $11\divides 22$, $11\divides (22^{10})$.
  \item If $p\divides a$ then $a^{p-1} \equiv 0 \pmod p$.
\end{itemize}

\end{frame}

\begin{frame}{More Fermat's Little Theorem}

\begin{itemize}
  \item Let $p=5$. Consider integers of the form $a^4$.
  \item $2^4=16$, $3^4=81$, $4^4=256$, $5^4=625$
  \item $6^4=1296$,  $7^4=2401$, $8^4=4096$.
  \item What do you notice about the above values?
  \item For all of the $a$ other than $a=5$, $a^4$ is a multiple of 5 plus one.
  \item In other words $a^4 \equiv 1 \pmod 5$.
  \item $5^4 \equiv 0 \pmod {5}$.
\end{itemize}

\end{frame}

\begin{frame}{Euler's Theorem in terms of $\Z_n^*$.}
\begin{itemize}
  \item \textbf{Theorem.} Let $n>1$ be an integer. Let $[a]\in\Z_n^*$.
  \item Then $[a]^{\phi(n)} = [1]$.
  \item \textbf{Example.} Let $[a]\in\Z_{10}^*$.
  \item Then $[a]^4 = [1]$.
  \item $\Z^*_{10} = \singleton{[1],[3],[7],[9]}$. Working in $\Z^*_{10}$...
  \item $[1]^4=[1], [3]^4=[1], [7]^4 = [1], [9]^4 = [1]$.
  \item ``Any element of the group raised to the power of the cardinality of the group equals one.''
\end{itemize}

\end{frame}

\begin{frame}{Fermat's Little Theorem in terms of $\Z_n^*$.}
\begin{itemize}
  \item \textbf{Theorem.} Let $p$ be prime. Let $[a]\in\Z_p^*$.
  \item Then $[a]^{p-1} = [1]$.
  \item \textbf{Example.} Let $[a]\in\Z_{11}^*$.
  \item Then $[a]^{10} = [1]$.
  \item Work in $\Z^*_{11}$. Then...
  \item $[1]^{10}=[2]^{10}=[3]^{10} = [4]^{10} = [5]^{10}=[6]^{10}=[7]^{10}=[8]^{10} = [9]^{10}=[10]^{10} = [1]$.
  \item ``Any element of the group raised to the power of the cardinality of the group equals one.''
\end{itemize}

\end{frame}


\begin{frame}{Proof of Euler's Theorem}

\begin{itemize}
  \item \textbf{proof.} Let $[a] \in \Z_n^*$. Consider the set of all multiples of $[a]$ in $\Z_n^*$.
  \item $S=\setof{[x][a]}{[x]\in\Z^*_n}$.
  \item $S=\singleton{[1][a], \cdots, [x][a],[y][a], \cdots, [n-1][a]}$.
  \item For example, consider $\Z^*_{10} = \singleton{[1],[3],[7],[9]}$.
  \item Consider the set $S$ of all multiples of $[3]$ in $\Z_{10}^*$:
  \item $S = \singleton{[1][3],[3][3],[7][3],[9][3]}$.
  \item $[1][3]=[3], [3][3]=[9], [7][3]=[1], [9][3] = [7]$.
  \item So $S = \singleton{[3],[9],[1],[7]} = \Z^*_{10}$.
  \item That is, the elements of $S$ are all of the elements of $Z^*_{10}$.
  \item They are just presented in a different order.
\end{itemize}

\end{frame}

\begin{frame}{Proof of Euler's Theorem, Continued}

\begin{itemize}
  \item \textbf{Claim.}  Let $[a] \in \Z_n^*$.
  \item Let $S=\setof{[x][a]}{[x]\in\Z^*_n}$ be the set of multiples of $[a]$ in $\Z^*_n$.
  \item Then $S=\Z^*_n$.
  \item \textbf{proof of Claim.} Notice that there are $\phi(n)$ multiples of $[a]$, one for each
  $[x]\in\Z^*_n$.
  \item So we just need to show they are all different.
  \item Suppose $[i][a]=[j][a]$. Multiplying both sides by $[a]^{-1}$ we get $[i]=[j]$.
  \item In other words, if $[i]\not=[j]$, then $[i][a]\not=[j][a]$.
  \item So all $\phi(n)$ multiples of $[a]$ in $\Z_n^*$ are different.
  \item So the set of multiples of $[a]$ runs through all of the elements of $\Z_n^*$. $\diamondsuit$.
\end{itemize}

\end{frame}


\begin{frame}{Proof of Euler's Theorem continued}

\begin{itemize}
  \item Now consider what happens when we multiply all of the elements of $\Z^*_n$ together.
  \item Let $[b]$ be the product of all of the elements of $\Z^*_n$.
  \item For example, the product of all of the elements of $\Z^*_{10}$ is
  \item $[b] = [1][3][7][9]$.
  \item In general, $[b]=[1]\cdots[x]\cdot[y] \cdots [n-1]$.
  \item By the claim, $[b]$ is also equal to the product of all of the multiples of $[a]$:
  \item $[b]=[1][a]\cdot \cdots[x][a]\cdot[y][a] \cdots [n-1][a]$
  \item  $=[a]^{\phi(n)}([1]\cdots[x]\cdot[y] \cdots [n-1])$
  \item $=[a]^{\phi(n)}[b]$.
  \item That is, $[b] = [a]^{\phi(n)} [b]$.
  \item Canceling $[b]$, we get $[a]^{\phi(n)}=[1]$. $\qed$.
\end{itemize}

\end{frame}

\end{document}
