% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 10 \\ Linear transformations on vector spaces}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Linear transformations}

\begin{itemize}
\item The definition of linear transformation makes sense on an arbitrary vector space.
\item \textbf{Definition.} Let $V$ and $W$ be two vector spaces.
\item Let $T:V\map W$.
\item Then $T$ is \emph{linear} iff
\item (a) $T$ preserves vector addition: $T(\bv+\bw) = T(\bv) + T(\bw)$.
\item (b) $T$ preserves scalar multiplication: $T(c\bv) = c T(\bv)$.
\item Linear Algebra is the study of linear transformations between vector spaces.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Examples of linear transformations}

\begin{itemize}
\item If $V=\R^n$ and $W=\R^m$ and $A$ is an $m\times n$ matrix then
$T_A$ is a linear transformation from $V$ to $w$.
\item Let $\cD$ be the space of all differentiable functions $f:\R\map\R$.
\item Let $\cF$ be the space of all functions $f:\R\map\R$.
\item Let $T:\cD\map\cF$ be defined by $T(f) = \fprime$.
\item Then $T$ is linear.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{More examples of linear transformations}

\begin{itemize}
\item Let $C^0$ be the space of all continuous functions $f:\R\map\R$.
\item Let $T:C^0\map C^0$ be defined by $T(f) = F$ where $F$ is the
anti-derivative of $F$ such that $F(0) = 0$.
\item i.e. $F(x) = \int_{t=0}^x f(t) dt$.
\item Then $T$ is linear.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The image of zero}

\begin{itemize}
\item \textbf{Lemma.} Let $T:V\map W$ be a linear transformation.
\item Then $T(\bzero) = \bzero$.
\item \textbf{proof.} Let $\bw = T(\bzero)$. We must show that $\bw=\bzero$.
\item $\bw = T(\bzero)$
\item $= T(0 \bzero)$
\item $= 0 T(\bzero)$
\item $= 0 \bw$
\item $= \bzero$. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Kernel of a linear transformation}

\begin{itemize}
\item \textbf{Definition.} Let $T:V\map W$ be a linear transformation.
\item The \textbf{kernel} of $T$, $\ker(T)$ is
\item $\setof{v\in V}{T(\bv) = \bzero}$.
\item $\ker(T) = T^{-1}[\singleton{\bzero}]$.
\item The kernel of $T$ is also called the \textbf{nullspace} of $T$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Kernel is a subspace}

\begin{itemize}
\item \textbf{Lemma.} Let $T:V\map W$ be a linear transformation.
\item Then $\ker(T)$ is a subspace of $V$.
\item \textbf{proof.} We must show $\ker(T)$ contains $\bzero$ and is closed
under addition and scalar multiplication.
\item $T(\bzero) = \bzero$. So $\bzero\in\ker(T)$.
\item Let $\bv_1,\bv_2\in\ker(T)$. Let $c\in\R$. We will show $\bv_1+\bv_2\in\ker(T)$ and $c\bv_1\in\ker(T)$.
\item $T(\bv_1+\bv_2)=T(\bv_1) + T(\bv_2) = \bzero + \bzero = \bzero$.
\item So $\bv_1+\bv_2\in\ker(T)$.
\item $T(c\bv_1)=cT(\bv_1)=c\bzero=\bzero$.
\item So $c\bv_1\in\ker(T)$.
\item So $\ker(T)$ is a subspace of $V$. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Homogeneous system of equations}

\begin{itemize}
\item \textbf{Definition.} A system of linear equations is called
\emph{homogeneous} if the right had side of the system consists of all zeroes.
\item As a matrix equation this means $\bb=0$ so the system looks like
$$A\bx=\bzero.$$
\item Notice: $\bx$ is a solution iff $\bx\in\ker(T_A)$.
\item Therefore the set of all solutions is a subspace.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Homogeneous system of equations}

\begin{itemize}
\item \textbf{Facts.} Let $A\bx=\bzero$ be a homogeneous system of equations. Then
\item $\bzero$ is always a solution.
\item So a homogeneous system is never inconsistent. It always has at least one solution.
\item $\bzero$ is called the \emph{trivial} solution.
\item If $\bx_1$ and $\bx_2$ are solutions then so is $\bx_1+\bx_2$.
\item Also so is $c\bx_1$ for any $c\in\R$.
\item If there is any non-trivial solution then there are infinitely many solutions.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Range of a linear transformation}

\begin{itemize}
\item \textbf{Lemma.} Let $T:V\map W$ be a linear transformation.
\item Then $\ran(T)$ is a subspace of $W$.
\item The range of $T$ is also called the \emph{image} of $T$.
\item \textbf{proof.} We must show $\ran(T)$ contains $\bzero$ and is closed
under addition and scalar multiplication.
\item $T(\bzero) = \bzero$. So $\bzero\in\ran(T)$.
\item Let $\bw_1,\bw_2\in\ran(T)$. Let $c\in\R$. We will show $\bw_1+\bw_2\in\ran(T)$ and $c\bw_1\in\ran(T)$.
\item Let $\bv_1$ and $\bv_2$ be such that $T(\bv_1)=\bw_1$ and $T(\bv_2)=\bw_2$.
\item $T(\bv_1+\bv_2)=T(\bv_1) + T(\bv_2) = \bw_1+\bw_2$.
\item So $\bw_1+\bw_2\in\ran(T)$.
\item $T(c\bv_1)=cT(\bv_1)=c\bw_1$.
\item So $c\bw_1\in\ran(T)$.
\item So $\ran(T)$ is a subspace of $W$. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Right-hand side of a system of equations}

\begin{itemize}
\item The system of equations $A\bx=\bb$ has a solution iff $b\in\ran(T_A)$.
\item If $\bb\notin\ran(T_A)$ then $A\bx=\bb$ is inconsistent (incompatible.)
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Inhomogeneous system of equations}

\begin{itemize}
\item Let $A$ be an $m\times n$ matrix. Let $\bb\in\R^m$. $\bb\not=\bzero$.
\item The system $A\bx=\bb$ is called \emph{inhomogeneous.}
\item Let $S=\setof{\bx \in \R^n}{A \bx = \bb}$ be the \emph{solution set}.
\item What are the possibilities for $S$?
\item If $\bb\notin\ran(T_A)$ then $S=\emptyset$.
\item Suppose $\bb\in\ran(T_A)$.
\item Let $\bx_0\in\R^n$ be a particular solution. $A\bx_0=\bb$. So $\bx_0\in S$.
\item If $T_A$ is one-to-one then $S=\singleton{\bx_0}$.
\item $S$ is \emph{not} a subspace:
\item $\bzero\notin S$. $T(\bzero) = \bzero \not= \bb$.
\item $S$ is not closed under vector addition. If $\bx_0\in S$ and $\bx_1\in S$
then $\bx_0+\bx_1\notin S$:
\item $A(\bx_0+\bx_1) = A\bx_0+A\bx_1 = \bb+\bb = 2\bb \not= \bb$.
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Solution set}

\begin{itemize}
\item \textbf{Lemma.} Let $A\bx=\bb$ be an inhomogeneous set of linear equations.
\item Let $S=\setof{\bx \in \R^n}{A \bx = \bb}$ be the solution set.
\item Let $\bx_0\in S$ be a particular solution. $A\bx_0=\bb$.
\item Then the general solution is given by
$$S=\setof{\bx_0 + \bx}{\bx\in\ker(T_A)}.$$
\item That is, the general solution is the kernel plus any particular solution.
\end{itemize}

\end{frame}



\end{document}


