% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref,changepage}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 23 \\ Proof of the Primitve Root Theorem}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Primitive Root Theorem}

\begin{itemize}
  \item \textbf{Theorem.} Let $n>1$ be an integer. Then
  \item $n$ has a primitive root iff one of the following is true:
  \item (a) $n$ is a power of an odd prime. $n=p^k$, $k\geq 1$.
  \item (b) $n$ is twice a power of an odd prime. $n=2p^k$, $k\geq 1$.
  \item (c) $n=2$ or $n=4$.
  \item In other words, $n$ does \emph{not} have a primitive root iff one of the following is true:
  \item (d) $n$ is divisible by two distinct odd primes.
  \item (e) $n$ is divisible by $4$ but not equal to $4$.
  \item In other words, $n$ does \emph{not} have a primitive root iff one of the following is true:
  \item (f) $n=ab$ where $\gcd(a,b)=1$ and $a,b>2$.
  \item (g) $n=2^k$ with $k\geq 3$.
\end{itemize}

\end{frame}

\begin{frame}{Powers of 2}

\begin{itemize}
  \item We will prove some parts of the theorem. Beginning with...
  \item \textbf{Theorem.} Let $n=2^k$ with $k\geq 3$.
  \item Then $n$ does not have any primitive roots.
  \item \textbf{proof.} We will show that there are more than two incongruent square-roots of 1 modulo $n$.
  \item In the previous lesson we learned that if $n$ has a primitive root, then it has only two square-roots of 1, so this will show that $n$ does not have any primitive roots.
  \item $1$ and $n-1 \equiv -1 \pmod n$ are always two square-roots of 1 modulo $n$, for any $n$,
  so it suffices for us to find any other square-root of 1 modulo $n$.
  \item We claim that $2^{k-1}+1$ is a square-root of 1 modulo $n$.
  \item $(2^{k-1}+1)^2 = 2^{2k-2} + 2^k + 1$.
  \item Since $k\geq 3$, $2k-2 \geq k$, so $2^k \divides (2^{2k-2} + 2^k)$.
  \item So $(2^{k-1}+1)^2 \equiv 1 \pmod n$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Divisible by two odd primes}

\begin{itemize}
  \item Next we prove another of the negative cases...
  \item \textbf{Theorem.} Let $n=ab$ where $\gcd(a,b)=1$ and $a,b>2$.
  \item Then $n$ does not have any primitive roots.
  \item \textbf{proof.} Again we will find a third square-root of 1 modulo $n$.
  \item Consider this system of congruences:
  \item $x\equiv 1 \pmod a$.
  \item $x\equiv -1 \pmod b$
  \item By the Chinese Remainder Theorem there is a unique $x$ with $0\leq x < n$
  that satisfies both of these congruences.
  \item Clearly $x\not=1$. Also $x\not=n-1$ because then $x\equiv -1 \pmod n$ and
  so $x\equiv -1 \pmod a$.
  \item But $a>2$ so $1\not\equiv -1 \pmod a$.
  \item For this $x$, $x^2 \equiv 1 \pmod a$ and $x^2 \equiv 1 \pmod b$.
  \item So $x^2\equiv 1 \mod n$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{The proof for primes.}

\begin{itemize}
  \item \textbf{Theorem.} Let $p$ be prime. Then $p$ has a primitive root.
  \item \textbf{proof.} Trivially 1 is a primitive root of 2. So assume $p>2$. We will work in $\Z_p^*$.
  \item We will show that $\Z_p^*$ is cyclic.
  \item We will show that there is an $[a]\in\Z_p^*$ of order $\phi(p)=p-1$.
  \item Write the prime decomposition $p-1 = q_1^{m_1}q_2^{m_2}\cdots q_k^{m_k}$.
  \item For each $j$ with $1\leq j \leq k$, we will find an $[a_j]$ of order $q_j^{m_j}$.
  \item Then we will let $[a]=[a_1][a_2]\cdots[a_k]$.
  \item $[a]$ will have order $q_1^{m_1}q_2^{m_2}\cdots q_k^{m_k}=p-1$.
  \item This is true because of the following...
  \item \textbf{Lemma.} If the order of $[a]$ and the order of $[b]$ are relatively prime, then
  the order of $[a][b]$ is equal to the order of $[a]$ times the order of $[b]$.

\end{itemize}

\end{frame}

\begin{frame}{proof of lemma}

\begin{itemize}
  \item \textbf{proof of lemma.} Let $s$ be the order of $[a]$ and $t$ the order of $[b]$.
  \item Let $r$ be the order of $[a][b]$. We want to see that $r=st$.
  \item Since $([a][b])^{st}=[1]$, $r\divides (st)$.
  \item So we want to show that $st\divides r$.
  \item Now $([a][b])^{rs}= (([a][b])^r)^s = [1]$.
  \item So $[1] = [a]^{rs}[b]^{rs}=[1][b]^{rs} = [b]^{rs}$.
  \item So $t\divides (rs)$.
  \item Since $\gcd(s,t)=1$, we have $t\divides r$.
  \item Similarly $s\divides r$.
  \item So $st=\gcd(s,t) \divides r$. $\diamondsuit$.
\end{itemize}

\end{frame}

\begin{frame}{The proof for primes, continued}

\begin{itemize}
  \item So fix $j$ with $1\leq j \leq k$.
  \item To simplify notation, let $q=q_j$ and $m={m_j}$. We will find an element of order $q^m$.
  \item Let $z=\frac{p-1}{q^m}$. So $p-1 = q^mz$ and $\gcd(q,z)=1$.
  \item If $[a]$ is any element of $\Z_p^*$ then the order of $[a]$ divides $q^mz$.
  \item \textbf{Notice:} Suppose the order of $[a]$ is a multiple of $q^m$.
  \item Then the order of $[a]$ is $q^mx$ with $x\divides z$ and so $\gcd(q,x)=1$.
  \item Then the order of $[a]^x=q^m$.
  \item So it suffices for us to find an $[a]$ whose order is divisible by $q^m$.
  \item Equivalently, it suffices for us to find an $[a]$ whose order does not divide $q^{m-1}z$.
  \item Therefore it suffices for us to prove the following theorem...
\end{itemize}
\end{frame}

\begin{frame}{The proof for primes, continued}

\begin{itemize}
  \item \textbf{Theorem.} Suppose $1\leq c \leq p-1$.
  \item Then there are at most $c$ elements $[a]$ of $\Z_p^*$ such that $[a]^c = [1]$.
  \item Equivalently, there are at most $c$ elements $[a]$ of $\Z_p^*$ whose order divides $c$.
  \item This theorem is enough to complete our argument from the previous slide.
  \item By the Theorem there are only $q^{m-1}z$ elements whose order divides $q^{m-1}z$.
  \item So there must be at least one element whose order does not divide $q^{m-1}z$.
  \item This is what we said we wanted on the previous slide.
\end{itemize}

\end{frame}

\begin{frame}{Polynomials over $\Z_p$}

\begin{itemize}
  \item We can rephrase the Theorem from the previous slide in terms of polynomials.
  \item Let $1\leq c \leq p-1$. We want to see that there are at most $c$ solutions
  \item to the polynomial equation $x^c-[1] = [0]$.
  \item Notice that $x^c-[1]$ is a polynomial of degree $c$.
  \item We will actually prove a a more general theorem...
  \item \textbf{Theorem.} Let $p$ be prime.
  \item Let $1\leq c \leq p-1$.
  \item Let $f(x)$ be a polynomial of degree $c$ with coefficients in $\Z_p^*$.
  \item Then there are at most $c$ elements $[a]$ of $\Z_p$ that are solutions to
  the equation $f([a])=[0]$.
\end{itemize}

\end{frame}


\end{document}
