% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 1 \\ The Integers}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Number Systems}

  \begin{itemize}
  \item Number theory is the theory of numbers.
  \item But what is a number?
  \item $\Z = \singleton{\cdots -3,-2,-1,0,1,2,3,\cdots} = $ the integers
  \item $\Q = \setof{\frac{n}{m}}{n,m\in \Z, m\not= 0} = $ the rationals
  \item $\R = $ the reals
  \item $\Q = \setof{x\in\R}{\text{decimal expansion of $x$ is terminating or repeating}}$
  \item $\C = $ the complex numbers
  \item Number theory is the theory of the integers, $\Z$.
  \end{itemize}


\end{frame}

\begin{frame}{Number Systems: Example}
\begin{itemize}
  \item  Give an example of an $x\in \Q \setminus \Z$.
  \item $x=1/2$.
  \item  Give an example of an $x\in \R \setminus \Q$.
  \item $x=\sqrt{2}$, $x=\pi$.
  \item Give an example of $x\in\C \setminus \R$.
  \item $x=2i+3$.
  \end{itemize}
\end{frame}

\begin{frame}{The well-ordering property}

\begin{itemize}
  \item Let $S$ be any non-empty set of positive integers. Then $S$ contains
  a least element.
  \item This is not true of $\Q,\R$.
  \item This is not true if you drop either ``non-empty'' or ``positive''.
\end{itemize}

\end{frame}

\begin{frame}{Proof by Induction}
\begin{itemize}
  \item Suppose $P$ is some \emph{property} about positive integers.
  \item For example $P(n)$ might be the property ``$n$ has a unique prime decomposition.''
  \item Or $P(n)$ might be the property ``either $n$ is even or $n+1$ is even.``
  \item Suppose we want to prove that property $P$ is true of all positive integers.
  \item We can do this by proving two things:
  \item \textbf{The base step:} Prove that $P(1)$ holds.
  \item \textbf{The Inductive step:} Prove that for all $n\geq 1$, $P(n) \Implies P(n+1)$.
  \item Today we are not going to do any proofs by induction.
  \item Instead we are going to justify the technique of proof by induction.
\end{itemize}
\end{frame}

\begin{frame}{Proof of  Proof by Induction}
\begin{itemize}
  \item \textbf{Proof that proof by induction works:} Suppose we have proven the base step and the inductive step.
  \item We want to see that $P(n)$ holds for all positive integers $n$.
  \item Suppose $P(n)$ doesn't hold for some positive integers, and we will derive a contradiction.
  \item Let $S$ be the set of positive integers $n$ for which $P(n)$ does not hold.
  \item We are assuming that $S$ is not empty.
  \item $S$ is a non-empty set of positive integers.
  \item By the well-ordering property, $S$ has a least element, $m$.
\end{itemize}
\end{frame}

\begin{frame}{Proof of  Proof by Induction, continued}
\begin{itemize}
  \item $m$ is the least element of $S$.
  \item $m$ cannot be $1$ because by the base step $P(1)$ holds so $1\notin S$.
  \item Since $m>1$, $m=n+1$ for some positive integer $n$.
  \item Since $n<m$, $n\notin S$. So $P(n)$ holds.
  \item By the inductive step, $P(n+1)$ holds, i.e. $P(m)$ holds.
  \item This contradicts our assumption that $m$ is the least element of $S$.
  \item Since we derived a contradiction, our assumption that $S\not=\emptyset$ must be wrong.
  \item So $S=\emptyset$, i.e. $P(n)$ holds for all positive integers $n$. $\qed$.
\end{itemize}
\end{frame}

\begin{frame}{Proof by Contradiction}
\begin{itemize}
\item The preceding proof was an example of \emph{proof by contradiction}.
\item Here is how such a proof works.
\item Suppose we want to prove a statement $\sigma$.
\item We start by assuming the \emph{negation} of $\sigma$, i.e. we assume
that $\sigma$ is not true.
\item Then we derive a \emph{contradiction}. This means we prove that some
sentence is both true and false.
\item Of course its impossible for a sentence to be both true and false.
\item We derived this impossible thing by assuming that $\sigma$ was false.
\item Thus our assumption that $\sigma$ was false must have been wrong.
\item So $\sigma$ is true.
\end{itemize}
\end{frame}


\end{document}
