% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 11 \\ Bases and Dimension}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Linear dependence}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item The vectors are called \emph{linearly dependent} if
\item there are scalars $c_1,c_2,\cdots,c_n$ such that
\item (i) not all of the $c_i = 0$, and
\item (ii) $c_1 \bv_1 + c_2 \bv_2 + \cdots + c_n\bv_n = \bzero$.
\item In words: A set of vectors is linearly dependent iff there is a non-trivial linear combination
of them that equals $\bzero$.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example of linear dependence}

\begin{itemize}
\item Let $V = \R^3$.
\item Let $\bv_1 =
\begin{pmatrix}
1 \\ 0 \\ 2
\end{pmatrix}
$,
$\bv_2 =
\begin{pmatrix}
0 \\ 1 \\ -1
\end{pmatrix}
$,
$\bv_3 =
\begin{pmatrix}
2 \\ 3 \\ 1
\end{pmatrix}
$.

\item Then $2\bv_1+3\bv_2 - \bv_3 = \bzero$.
\item So $\bv_1, \bv_2, \bv_3$ are linearly dependent.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Equivalent definition of linear dependence}

\begin{itemize}
\item \textbf{Lemma.} Let $\bv_1,\bv_2,\cdots\bv_n$ be $n$ vectors.
\item Then the following are equivalent:
\item (a) The vectors are linearly dependent.
\item (b) One of the velors can be expressed as a linear combination of the others.
\item \textbf{Proof.} First suppose that the vectors are linearly dependent.
\item Let $c_1,c_2,\cdots,c_n$ be scalars so that
\item $c_1\bv_1+c_2\bv_2\cdots c_n\bv_n = \bzero$, and at least one of the $c_i$ is not 0.
\item Suppose $c_1\not=0$. Then
\item $\bv_1 = (-c_2/c_1)\bv_2 + (-c_3/c_1)\bv_3 + \cdots + (-c_n/c_1)\bv_n$.
\item So we have expressed one of the vectors as a linear combination of the others.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{proof continued}

\begin{itemize}
\item Conversely, suppose we have expressed one of the vectors as a linear combination
of the others. Say it is $\bv_1$.
\item So $\bv_1 = c_2\bv_2 + c_3\bv_3 + \cdots + c_n \bv_n$.
\item Then $\bv_1 - c_2\bv_2 - c_3\bv_3 - \cdots - c_n\bv_n = \bzero$.
\item That is a non-trivial linear combination because the coefficient of $\bv_1$ is 1.
\item So the vectors are linearly dependent. $\qed$

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example of linear dependence}

\begin{itemize}
\item Let $V = \R^3$.
\item Let $\bv_1 =
\begin{pmatrix}
1 \\ 0 \\ 2
\end{pmatrix}
$,
$\bv_2 =
\begin{pmatrix}
0 \\ 1 \\ -1
\end{pmatrix}
$,
$\bv_3 =
\begin{pmatrix}
2 \\ 3 \\ 1
\end{pmatrix}
$.

\item We saw before that $2\bv_1+3\bv_2 - \bv_3 = \bzero$.
\item So $\bv_1, \bv_2, \bv_3$ are linearly dependent.
\item Now also notice that $\bv_3 = 2\bv_1 + \bv_2$.
\item So we have expressed one of the vectors as a linear combination of the others.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Linear independence}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item The vectors are called \emph{linearly independent} iff
\item they are not linearly dependent.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example of linear independence}

\begin{itemize}
\item The standard basis vectors of $\R^n$ are always linearly independent.
\item Let $V = \R^3$.
\item Let $\be_1 =
\begin{pmatrix}
1 \\ 0 \\ 0
\end{pmatrix}
$,
$\be_2 =
\begin{pmatrix}
0 \\ 1 \\ 0
\end{pmatrix}
$,
$\be_3=
\begin{pmatrix}
0 \\ 0 \\ 1
\end{pmatrix}
$.

\item Then $\be_1,\be_2,\be_3$ are linearly independent.
\item To see this, let $c_1,c_2,c_3$ be scalars and consider the linear combination
\item $c_1\be_1+c_2\be_2+c_3\be_3$.
\item How can this linear combination be $\bzero$?
\item That linear combination is equal to the vector
$
\begin{pmatrix}
c_1 \\ c_2 \\ c_3
\end{pmatrix}
$.
\item So it can be zero iff $c_1=c_2=c_3=0$.
\item In other words, no non-trivial linear combination of the $\be_i$ can be zero.
So the $\be_i$ are linearly independent.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Linear independence and the zero vector}

\begin{itemize}
\item \textbf{Lemma.} Any set of vectors that includes the zero vector is linearly dependent.
\item \textbf{proof:} Suppose $\bv_0=\bzero$.
\item Let $\bv_1,\bv_2,\cdots,\bv_n$ be any
other vectors.
\item Then
$$\bv_0 + 0\bv_1 + 0\bv_2 + \cdots + 0\bv_n = \bzero.$$
\item This is a non-trivial linear combination that equals zero.
\item So $\bv_0,\bv_1,\cdots,\bv_n$ is linearly dependent. $\qed$
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Span}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item Then the \emph{span} of the vectors is the set of all linear combinations of them.
\item $\spn(\bv_1,\bv_2,\cdots\bv_n)=\setof{c_1\bv_1+c_2\bv_2+\cdots+c_n\bv_n}{c_1,c_2,\cdots,c_n\in\R}$.
\item \textbf{Example.} In $\R^3$, consider the span of the first two standard basis vectors.
\item $\spn(\be_1,\be_2) = \setof{c_1\be_1+c_2\be_2}{c_1,c_2,\in\R}$.
\item This is the $x$- $y$-plane in $\R^3$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Span is a subspace}

\begin{itemize}
\item \textbf{Lemma.} $\spn(\bv_1,\bv_2,\cdots\bv_n)$ is a subspace of $V$.
\item \textbf{Proof.} We must show that the span contains zero and is closed under vector addition
and scalar multiplication.
\item The span contains zero because $\bzero$ can always be expressed as the linear combination
of the $\bv_i$ in which all coefficients are zero.
\item Now suppose that $\bx$ and $\by$ are in the span and $a$ is a scalar.
\item we will show that $\bx+\by$ and $a \bx$ are in the span.
\item Write $\bx = c_1\bv_1 + c_2\bv_2 + \cdots c_n \bv_n$
\item and $\by = d_1\bv_1 + d_2\bv_2 + \cdots d_n \bv_n$.
\item Then $\bx + \by = (c_1 + d_1) \bv_1 + (c_2 +d_2)\bv)2 + \cdots + (c_n+d_n)\bv_n$.
\item And $a\bx = (ac_1)\bv_1 + (ac_2)\bv_2 + \cdots + (ac_n)\bv_n$.
\item So both $\bx+\by$ and $a\bx$ are in the span of $\bv_1,\bv_2,\cdots,\bv_n$.
\item So the span is a subspace of $V$. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Spanning set of vectors}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item We say that $\bv_1,\bv_2,\cdots,\bv_n$ \emph{span} $V$
\item or that $\singleton{\bv_1,\bv_2,\cdots,\bv_n}$ is a \emph{spanning set} for $V$
\item if $V=\spn(\bv_1,\bv_2,\cdots,\bv_n)$.
\item In other words, if every vector of $V$ can be written as a linear combination of
 $\bv_1,\bv_2,\cdots,\bv_n$.
 \item \textbf{Example.} In $\R^n$, $\singleton{\be_1,\be_2,\cdots,\be_n}$ is a
 spanning set.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Containing a spanning set}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space. Suppose $\bv_1,\bv_2,\cdots,\bv_n$ span $V$.
\item Let $\bw_1,\bw_2,\cdots\bw_m$ be another set of vectors of $V$.
\item Suppose $\bv_1,\bv_2,\cdots\bv_n\in\spn(\bw_1,\bw_2,\cdots,\bw_m)$.
\item Then $\bw_1,\bw_2,\cdots,\bw_m$ also span $V$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example}

\begin{itemize}
\item Let $V=\R^2$
\item Let $\bw_1 =
\begin{pmatrix}
1 \\ 1
\end{pmatrix}
$, $\bw_2 =
\begin{pmatrix}
1 \\ 2
\end{pmatrix}
$
\item Suppose we want to show that $\bw_1,\bw_2$ spans all of $\R^2$.
\item Because we already know that $\be_1,\be_2$ is a spanning set for $\R^2$,
\item it suffices to show $\be_1,\be_2\in\spn(\bw_1,\bw_2)$.
\item $\be_1 = 2\bw_1 - \bw_2$.
\item $\be_2 = \bw_2 - \bw_1$.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Basis}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be $n$ vectors in $V$.
\item We say that $\bv_1,\bv_2,\cdots,\bv_n$ is a \emph{basis} for $V$ if
\item (i) $\bv_1,\bv_2,\cdots,\bv_n$ are linearly independent
\item (ii) $\bv_1,\bv_2,\cdots,\bv_n$ span $V$.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example of a basis}

\begin{itemize}
\item The standard basis vectors of $\R^n$ are a basis for $\R^n$.
\item Let $V = \R^3$.
\item Let $\be_1 =
\begin{pmatrix}
1 \\ 0 \\ 0
\end{pmatrix}
$,
$\be_2 =
\begin{pmatrix}
0 \\ 1 \\ 0
\end{pmatrix}
$,
$\be_3=
\begin{pmatrix}
0 \\ 0 \\ 1
\end{pmatrix}
$.

\item Then $\be_1,\be_2,\be_3$ are a basis for $\R^3$.
\item To see this we need to recall that they are
\item (i) linearly independent, and
\item (ii) span $\R^3$.
\item So the standard basis vectors form a basis.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Dimension}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space
\item Suppose $V$ has a basis consisting of $n$ vectors.
\item Then we say that the \emph{dimension} of $V$ is $n$.
\item Or that $V$ is $n$-dimensional.
\item \textbf{Example.} $\R^n$ is $n$-dimensional.
\item Because $\R^n$ has a basis consisting of the $n$ standard basis vectors.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Is dimension well-defined?}

\begin{itemize}
\item Let $V$ be an $n$-dimensional vector space.
\item So $V$ has a basis consisting of $n$ vectors.
\item Question: Is it possible that $V$ also has a different basis
consisting of $m$ vectors, for some $m\not=n$?
\item If this is possible, then $V$ is also $m$-dimensional.
\item And so the dimension of $V$ is not well-defined.
\item We will show that this cannot happen.
\item We will show that every basis for $V$ has the same size.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Dimension is well-defined}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be a basis for $V$.
\item Let $\bw_1,\bw_2,\cdots,\bw_n$  be any $n$ linearly independent vectors of $V$.
\item Then $\bw_1,\bw_2,\cdots,\bw_n$ are also a basis.
\item \textbf{Proof.} Let's just deal with the two dimensional case.  Let $V$ be a two dimensional
vector space with basis $\bv_1,\bv_2$.
\item Let $\bw_1,\bw_2$ be linearly
independent. We will show they also form a basis.
\item We need to show that the span of $\bw_1,\bw_2$ is all of $V$.
\item It suffices to show that $\bv_1,\bv_2\in\spn(\bw_1,\bw_2)$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof continued.}

\begin{itemize}
\item $\bw_1 = c_1\bv_1 + c_2\bv_2$, for some scalars $c_1, c_2$.
\item Since $\bw_1\not=\bzero$, either $c_1$ or $c_2$ is non-zero. Say $c_1 \not= 0$.
\item So then $\bv_1 = (1/c_1)\bw_1 - (c_2/c_1)\bv_2$.
\item So $\bv_1\in\spn(\bw_1,\bv_2)$.
\item But then $\bv_1,\bv_2\in\spn(\bw_1,\bv_2)$.
\item So $\spn(\bw_1,\bv_2)=V$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof continued.}

\begin{itemize}
\item $\bw_2 = c_1\bw_1 + c_2\bv_2$, for some scalars $c_1, c_2$.
\item Since $\bw_1, \bw_2$, are linearly independent, $c_2 \not= 0$.
\item So then $\bv_2 = (1/c_2)\bw_2 - (c_1/c_2)\bw_1$.
\item So $\bv_2$ is in the span of $\bw_1$ and $\bw_2$.
\item But then $\bv_1$ and $\bv_2$ are in the span of $\bw_1$ and $\bw_2$
and $\bv_1,\bv_2$ span all of $V$.
\item So $\bw_1,\bv_2$ span all of $V$. $\qed$
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Is dimension well-defined}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space and let $\bv_1,\bv_2,\cdots,\bv_n$ be a basis for $V$.
\item Let $\bw_1,\bw_2,\cdots,\bw_{n+1}$  be any $n+1$ vectors of $V$.
\item Then $\bw_1,\bw_2,\cdots,\bw_{n+1}$ are linearly dependent.
\item For example, in $\R^2$ and $3$ vectors are linearly dependent.
\item And in $\R^3$ any $4$ vectors are linearly dependent.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}


