% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 13 \\ Linear Congruences}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Solve for $x$}

\begin{itemize}
  \item Solve for $x$:
  \item $2x  \equiv 4 \mod 6$.
  \item $x=2$ works.
  \item So does $x=8$ and $x=-4$
  \item because these are congruent to 2 mod 6.
  \item Any element of the congruence class of 2 modulo 6 is a solution.
  \item But we consider all of these to be the same solution.
  \item Are there any other solutions besides these?
\end{itemize}

\end{frame}

\begin{frame}{Find all incongruent solutions.}

\begin{itemize}
  \item When we consider problems where we solve for $x$ in a congruence
  \item if $x$ and $\xprime$ are two solutions and $x$ and $\xprime$ are congruent
  \item then we consider them to be the same solution.
  \item In general our goal is to find all incongruent solutions.
  \item Equivalently, find all least non-negative residues that are solutions.
\end{itemize}

\end{frame}


\begin{frame}{Find all incongruent solutions}
\begin{itemize}
  \item Find all incongruent solutions to
  \item $2x  \equiv 4 \mod 6$.
  \item Equivalently, find all $x$ in the range $[0,5]$.
  \item $x=2$ is a solution. Are there any others?
  \item Compute $2x$ for $x=0,1,2,3,4,5$.
  \item i.e. find the multiples of 2 mod 6
  \item Answer: 0,2,4,0,2,4
  \item So $2\cdot 5 \equiv 4 \pmod 6$ and $x=5$ is another solution.
  \item \textbf{Answer:} There are two incongruent solutions: $2,5$.
\end{itemize}
\end{frame}

\begin{frame}{Representatives of the solutions}
\begin{itemize}
  \item $x=-4,8,14$ are also solutions. But they are congruent to 2.
  \item $x=-1,11,17$ are also solutions. But they are congruent to 5.
  \item This would also have been correct:
  \item \textbf{Answer:} There are two incongruent solutions: $-4,17$.
  \item You need to say how many incongruent solutions there are and
  give one representative from each congruence class of the solution.
  \item It doesn't matter which representative you give. You might as well give
  the least non-negative residue.
\end{itemize}
\end{frame}

\begin{frame}{Linear congruence}
\begin{itemize}
  \item A \emph{linear congruence} problem is a problem of the following kind:
  \item Find all incongruent solutions to $ax \equiv b \pmod n$
  \item where $a,b,n$ are integers,  $n>1$.
  \item We will learn how to solve all problems of this kind.
  \item This involves:
  \item How many incongruent solutions are there? (There might be none.)
  \item Finding all of the incongruent solutions.
  \item Equivalently, finding all solutions in the range $0,1,\cdots, n-1$.
\end{itemize}
\end{frame}

\begin{frame}{Modular Inverses}
\begin{itemize}
  \item We start by looking at the special case: $ax\equiv 1 \pmod n$.
  \item \textbf{Definition.} $x$ is called a \emph{multiplicative inverse}  of $a$ mod $n$
  \item iff $ax \equiv 1 \pmod n$.
  \item \textbf{Theorem.} $a$ has a multiplicative inverse mod $n$ iff $\gcd(a,n)=1$.
  \item If $\gcd(a,n) = 1$ then $a$ has a unique multiplicative inverse, up to congruence.
  \item That means there is exactly one incongruent solution to $ax\equiv 1 \pmod n$.
\end{itemize}
\end{frame}

\begin{frame}{Examples of Modular Inverses}
\begin{itemize}
  \item Let $n=9$.
  \item $\gcd(4,9)=1$. So $4$ has a unique multiplicative inverse mod 9. Find it.
  \item Compute $4x \bmod 9$ for $x=0,1,2,3,4,5,6,7,8$.
  \item The multiples of 4 mod 9 are: 0,4,8,3,7,2,6,1,5.
  \item So $4\cdot 7 \equiv 1 \pmod 9$ and $x=7$ is the unique solution, up to congruence.
  \item 7 is the multiplicative inverse of 4 mod 9.
  \item $\gcd(6,9) \not= 1$.
  \item So 6 should not have a multiplicative inverse mod 9. Check this.
  \item Compute $6x\bmod 9$ for $x=0,1,2,3,4,5,6,7,8$.
  \item The multiples of 6 mod 9 are: 0,6,3,0,6,3,0,6,3.
  \item So 6 does not have a multiplicative inverse mod 9.
\end{itemize}
\end{frame}

\begin{frame}{Proof of Modular Inverses Theorem}
\begin{itemize}
  \item \textbf{proof.} $\gcd(a,n)=1$ iff
  \item there are $x,y$ such that $ax + ny = 1$ iff
  \item there are $x,y$ such that $ax - 1 = -ny$ iff
  \item there is an $x$ such that $n\divides (ax -1)$ iff
  \item there is an $x$ such that $ax \equiv 1 \pmod n$ iff
  \item $a$ has a multiplicative inverse mod n.
  \item For uniqueness, suppose $ax \equiv a\xprime \pmod n$.
  \item Since $\gcd(a, n)=1$, we can cancel the $a$ to conclude
  \item $x\equiv \xprime \pmod n$.
  \item So there is only one solution, up to congruence. $\qed$.
\end{itemize}
\end{frame}

\begin{frame}{Solutions when $\gcd(a,n)=1$}
\begin{itemize}
  \item \textbf{Theorem.} Consider the congruence $ax\equiv b \pmod n$.
  \item Suppose $\gcd(a,n) = 1$.
  \item Then there is a unique solution, up to congruence.
  \item \textbf{proof.} Let $\aprime$ be the unique multiplicative inverse
  of $a$ mod $n $.
  \item Then $ax\equiv b \pmod n$ iff
  \item $a\aprime x\equiv b\aprime \pmod n$ iff
  \item $x \equiv b\aprime \pmod n$.
  \item So $x=b\aprime$ is the unique solution, up to congruence. $\qed$.
\end{itemize}
\end{frame}

\begin{frame}{Example when $\gcd(a,n)=1$}
\begin{itemize}
  \item Solve $4x \equiv 5 \pmod 9$.
  \item Solution: Recall that the multiplicative inverse of 4 mod 9 is 7.
  \item $4x \equiv 5 \pmod 9$ iff $7\cdot 4 x  \equiv 7\cdot 5  \pmod 9$
  \item iff $x \equiv 35 \pmod 9$.
  \item So we could say that the unique solution is $x=35$, up to congruence.
  \item But $35\bmod 9 = 8$.
  \item So using least non-negative residues we would say
  \item the unique solution is $x=8$, up to congruence.
  \item Check $4\cdot 8 = 32 \equiv 5 \pmod 9$.
\end{itemize}
\end{frame}


\begin{frame}{General Theorem}
\begin{itemize}
  \item \textbf{Theorem} Consider the congruence $ax\equiv b \pmod n$.
  \item Let $d=\gcd(a,n)$.
  \item If $d\not\divides b$ then there are no solutions.
  \item If $d\divides b$ then there are $d$ incongruent solutions.
  \item If $x_0$ is any solution then the $d$ incongruent solutions are
  \item $x_0 + \frac{kn}{d}$ for $k=0,1,2,\cdots d-1$.
  \item To find the first solution, $x_0$,  we can solve
  \item $\frac{a}{d}x \equiv \frac{b}{d} \pmod {\frac{n}{d}}$.
  \item This congruence has a unique solution $x_0$ with $0\leq x_0 < \frac{n}{d}$.
  \item $x_0$ is also a solution to the original congruence.
\end{itemize}
\end{frame}

\begin{frame}{Example}
\begin{itemize}
  \item Solve $14x \equiv 15 \pmod {35}$.
  \item Let $d=\gcd(14,35) = 7$.
  \item $7\not\divides 15$. So there are no solutions.
  \item Solve $10x \equiv 15 \pmod {35}$.
  \item Let $d=\gcd(10,35) = 5$.
  \item $5\divides 15$. So there are 5 incongruent solutions.
  \item To find one solution, solve $2x \equiv 3 \pmod 7$.
  \item $\gcd(2,7) = 1$ so there is a unique solution in $\singleton{0,1,2,3,4,5,6}$.
  \item The multiples of 2 mod 7 are: 0,2,4,1,3,5.
  \item So $2\cdot 4 \equiv 3 \pmod 7$.
  \item So $x=4$ is one solution to the original congruence.
  \item Check: $10\cdot 4 = 40 \equiv 5 \pmod {35}$.
\end{itemize}
\end{frame}

\begin{frame}{Example Continued}
\begin{itemize}
  \item We are solving $10x \equiv 15 \pmod {35}$.
  \item We found $x_0=4$ is one solution.
  \item We have $d=\gcd(10,35) = 5$.
  \item $5\divides 15$. So there are 5 incongruent solutions:
  \item $x_0 + \frac{kn}{d}$ for $k=0,1,2,\cdots d-1$, i.e.
  \item $4 + 7k$, for $k=0,1,2,3,4$.
  \item The solutions are $x=4,11,18,25,32$.
  \item Check.
\end{itemize}
\end{frame}

\begin{frame}{Proof of General Theorem}
\begin{itemize}
  \item \textbf{proof.} Let $d=\gcd(a,n)$.
  \item There is a solution to $ax\equiv b \pmod n$ iff
  \item there is an $x$ such that $n\divides (ax - b)$ iff
  \item there are $x,y$ such that $ax-b = ny$ iff
  \item there are $x,y$ such that $ax - ny = b$ iff
  \item $b$ is an integer linear combination of $a$ and $n$ iff
  \item $d\divides b$.
  \item So if $d\not\divides b$ there are no solutions.
  \item Now suppose that $d\divides b$.
  \item $ax\equiv b \pmod n$ iff
  \item $\frac{a}{d}x \equiv \frac{b}{d} \pmod {\frac{n}{d}}$.
  \item The two congruences have the same integer solutions.
  \item Now $\gcd(\frac{a}{d},\frac{n}{d}) = 1$.
  \item So there is a unique solution $x_0$ to the two congruences with
  $0\leq x_0 < \frac{n}{d}$.
\end{itemize}
\end{frame}

\begin{frame}{Proof of General Theorem, continued}
\begin{itemize}
  \item Let $x_0$ be the unique solution to the two congruences with
 with $0\leq x_0 < \frac{n}{d}$.
  \item All of the other integer solutions to the two congruences are of the form
  \item $x=x_0+k\frac{n}{d}$ for some integer $k$.
  \item The solutions with $0\leq x < n$ are those with $0\leq k < d$. $\qed$
\end{itemize}
\end{frame}


\end{document}
