% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 8\\ Gaussian Elimination, Part 1}
\subtitle{Math 325, Linear Algebra \\ Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Solving a matrix equation}

\begin{itemize}
\item Solve the equation $A\bx=\bb$.
\item Here $A$ is an $m\times n$ matrix, $\bb\in\R^{m}$.
\item How should we think about this?
\item Idea: What would we do if $A$, $\bb$ and $\bx$ were numbers instead of matrices?
\item $\bx = \frac{\bb}{A}$.
\item This doesn't work with matrices because matrix division is not defined. Is it?
\item Suppose $A$ were an invertible matrix.
\item Then multiply both sides of the equation by $A^{-1}$. We get
\item $\bx = A^{-1}\bb$. We have solved the equation!
\item But this is not a good general solution.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Solving a matrix equation with the inverse}

\begin{itemize}
\item Solve the equation $A\bx=\bb$.
\item $\bx = A^{-1}\bb$ is a solution if $A$ is invertible.
\item But there are two problems:
\item (i) $A$ might not be invertible.
\item (ii) If $A$ is invertible but it is a large matrix, then computing
$A^{-1}$ is very expensive--more expensive than solving the original equation.
\item Besides, we haven't yet learned how to find the inverse of a large matrix.
\item So while $\bx = A^{-1}\bb$ is interesting theoretically, that is
not the techniqe we will be using for solving $A\bx=\bb$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{A different idea}

\begin{itemize}
\item Solve the equation $A\bx=\bb$.
\item Here $A$ is an $m\times n$ matrix, $\bb\in\R^{m}$.
\item Suppose $G$ is an invertible $m\times m$ matrix.
\item Then for all $\bx$,
\item $A\bx=\bb$ iff $GA\bx = G\bb$.
\item Why?
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{A different idea, continued}

\begin{itemize}
\item $A\bx=\bb$ iff $GA\bx = G\bb$.
\item What good is this? Is the second equation easier to solve?
\item Let $A^{\prime} = GA$, $\bb^{\prime} = G\bb$.
\item So the second equatin is $A^{\prime}\bx=\bb^{\prime}$.
\item This looks just like the original equation, so have we gotten anywhere?
\item Idea: Pick the invertible matrix $G$ so that $A^{\prime}=GA$ is in a simple form.
\item If $A^{\prime}$ is in a simple form then it is easy to solve $A^{\prime}\bx=\bb^{\prime}$.
\item Plan: Find an invertible matrix $G$ so that $GA$ is in a simple form.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Plan}

\begin{itemize}
\item Plan: Find an invertible matrix $G$ so that $GA$ is in a simple form.
\item So what is $G$?
\item $G$ will be the product of elementary matrices of types I and II.
\item $G=G_1\cdot G_2 \cdot \cdots \cdot G_k$, where each $G_k$ is an elementary matrix.
\item So we will be performing a series of elementary row operations on $A$
\item until we have put $A$ in a simple form.
\item And we will be performing the same series of elementary row operations
on the right-hand-side vector $\bb$.
\item So what is the simple form of the matrix we are looking for?
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Upper-triangular form}

\begin{itemize}
\item \textbf{Definition} A square matrix $A$ is called \emph{upper-triangular}
\item if all entries of $A$ below the main diagonal are zero.
\item For example
$$A=
\begin{pmatrix}
3 & -1  & 3 \\
0 & 2  & 1 \\
0 & 0 & -2 \\
\end{pmatrix}
$$
\item We are going to learn how to solve the equation $A\bx=\bb$
\item in the case where $A$ is
\item upper triangular, with non-zeros on the main diagonal.
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The method of back-substitution}

\begin{itemize}
\item Consider the equation
$$
\begin{pmatrix}
3 & -1  & 3 \\
0 & 2  & 1 \\
0 & 0 & -2 \\
\end{pmatrix}
\begin{pmatrix}
x \\ y \\ z
\end{pmatrix}
=
\begin{pmatrix}
4 \\ 3 \\ 2
\end{pmatrix}
$$
\item We can solve this equation via the method of back-substitution.
\item The third equation is $-2z=2$.
\item So we can see that $z=-1$.
\item The second equation is $2y+z = 3$.
\item But we can substitue $z=-1$ and get $2y -1 = 3$.
\item So $y=2$.
\item The first equation is $3x -y +3z = 4$.
\item But we can substitue $z=-1, y=2$ and get $3x -2  - 3 = 4$
\item So $3x = 9$ or $x=3$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The method of back-substitution}

\begin{itemize}
\item So we have solved
$$
\begin{pmatrix}
3 & -1  & 3 \\
0 & 2  & 1 \\
0 & 0 & -2 \\
\end{pmatrix}
\begin{pmatrix}
3 \\ 2 \\ -1
\end{pmatrix}
=
\begin{pmatrix}
4 \\ 3 \\ 2
\end{pmatrix}
$$
\item Notice that at each stage we found the only possible value for $z$,
then for $y$, then for $x$.
\item So $\bx=(3,2,-1)^{\text{T}}$ is the only solution to $A\bx=\bb$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Upper-triangular matrices with non-zero diagonal entries}

\begin{itemize}
\item \textbf{Theorem} Let $A$ be an $n\times n$ matrix in upper-triangular
form with all non-zeros on the diagonal.
\item Let $\bb$ be any vector in $\R^n$.
\item Then there is a unique solution to the equation $A\bx=\bb$.
\item \textbf{proof} We can use the method of back substitution as in the
previous example to find the unique solution. $\qed$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Gaussian Elimination}

\begin{itemize}
\item Gaussian Elimination is an algorithm that
\item - takes a rectangular $m\times n$ matrix $A$ as input
\item - performs a series of row operations on the matrix
\item - resulting in a new $m\times n$ matrix $\Aprime$, such that
\item - $\Aprime$ is in a particular simple form.
\item What is this particularly simple form? It is called row-echelon form.
\item We will learn about row-echelon form later. For now we only consider the easy case.
\item In the easy case, $\Aprime$ will be upper-triangular with non-zeros on the diagonal.
\item This is the case we will study first.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Augmented Matrix}

\begin{itemize}
\item \textbf{Definition.} If $A\bx=\bb$ is a system of linear equations then the
\emph{augmented matrix} for the system is the matrix whose columns are the columns
of $A$ with one additional column on the right consisting of $\bb$. We draw
a vertical line between the columns of $A$ and $\bb$.
\item For the system we are currently working with
$A=
\begin{pmatrix}
 0 & 1 & -1 \\
-2 & 4 & -1 \\
-2 & 5 & -4 \\
\end{pmatrix}
$,
and
$
\bb =
\begin{pmatrix}
3 \\ 1 \\ -2
\end{pmatrix}
$.
\item The augmented matrix is
$
\begin{pmatrix}
 0 & 1 & -1  &  \aug & 3  \\
-2 & 4 & -1  &  \aug & 1  \\
-2 & 5 & -4  &  \aug & -2  \\
\end{pmatrix}
$
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example Gaussian Elimination with an Augmented Matrix}

\begin{itemize}
\item $
\begin{pmatrix}
 0 & 1 & -1  &  \aug & 3  \\
-2 & 4 & -1  &  \aug & 1  \\
-2 & 5 & -4  &  \aug & -2  \\
\end{pmatrix}
$
\item Swap rows 1 and 2
\item $
\begin{pmatrix}
\textbf{-2} & 4 & -1  &  \aug & 1  \\
        0   & 1 & -1  &  \aug & 3  \\
        -2  & 5 & -4  &  \aug & -2  \\
\end{pmatrix}
$
\item Replace row 3 with row 3 minus row 1
\item $
\begin{pmatrix}
\textbf{-2} & 4 & -1  &  \aug & 1  \\
        0   & 1 & -1  &  \aug & 3  \\
        0   & 1 & -3  &  \aug & -3  \\
\end{pmatrix}
$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example Gaussian Elimination with an Augmented Matrix}

\begin{itemize}
\item $
\begin{pmatrix}
\textbf{-2} &         4  & -1  &  \aug & 1  \\
        0   & \textbf{1} & -1  &  \aug & 3  \\
        0   &         1  & -3  &  \aug & -3  \\
\end{pmatrix}
$
\item Replace row 3 with row 3 minus row 2.
\item $
\begin{pmatrix}
\textbf{-2} &         4  &          -1  &  \aug & 1  \\
        0   & \textbf{1} &          -1  &  \aug & 3  \\
        0   &         0  & \textbf{-2}  &  \aug & -6  \\
\end{pmatrix}
$
\item That is now upper-triangular with nonzeros on the diagonal.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Back substitution.}

\begin{itemize}
\item $
\begin{pmatrix}
\textbf{-2} &         4  &          -1  &  \aug & 1  \\
        0   & \textbf{1} &          -1  &  \aug & 3  \\
        0   &         0  & \textbf{-2}  &  \aug & -6  \\
\end{pmatrix}
$
\item To solve the equation we use back-substitution.
\item $-2 w = -6 \quad \implies \quad w = 3$
\item $v - w = 3 \quad \implies \quad v - 3 = 3 \quad \implies \quad v=6$.
\item $-2 u + v4 -w  = 1 \quad \implies \quad -2u +24 - 3 = 1 \quad \implies \quad u = 10$.
\item This is the solution to the \emph{original} system too.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Gaussian Elimination to Solve Systems of Linear Equations}

We have solved the original system of linear equations.

\pause

\begin{align*}
 \quad v - w         &= 3 \\
-2u + 4v - w         &= 1 \\
-2u + 5v - 4w        &= -2 \\
\end{align*}

\pause

The solution is $u = 10, v=6, w=3$.

\end{frame}



\end{document}


