\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Notes on Differential Privacy}
\author{Mitch Rudominer}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Amplificatin by Shuffling}

In this section I am trying to understand the amplification by shuffling idea.

In order to avoid dealing with unnecessry exponentials, we define 
$\alpha$- and $(\alpha,\delta)$-differential-privacy to
be just like $\epsilon$- and  and $(\epsilon,\delta)$-differential-privacy except that
we replace $e^{\epsilon}$ with $\alpha$.

Consider the experiment in which a user has a true bit value (0 or 1) and the user
emits the true value with probability $3/4$ and the opposite value with
probability $1/4$. Cleary this mechanism satisfies $\alpha$-differential-privacy
with $\alpha=3$. 

Let $\delta=(0.75)^{101} \approx 2.4\times 10^{-13}$. Then this mechanism also satisfies
$(\alpha,\delta)$-differential-privacy with 
$$\alpha=\frac{0.75 - \delta}{0.25} = 3-4\delta \approx 3-9.6\times 10^{-13}.$$
This is the local model.

Now consider the the experiment in which there are 100 users each of which perform the
above local experiment and emit a bit value. Call the true bit values $b_i$ and the
emitted values $x_i$ for $i=0,\cdots,99$.

Let $Y$ be the sum of the $x_i$, and consider the mechanism that emits $Y$. So this
is the central model.

In fact this is the shuffled model because the shuffled model is equivalent to emitting
the multi-set of locally-emitted values, or equivalently the histogram of locally-emitted
values. So we can use this example to analyze amplification by Shuffling.


Consider the case that $b_i=1$ for $i=1,\cdots,99$. Then

$$P(Y=100 \mid b_0 = 1) = \left(\frac{3}{4}\right)^{100}$$
and
$$P(Y=100 \mid b_0 = 0) = \left(\frac{3}{4}\right)^{99}\frac{1}{4}$$

Thus this mechanism again satisfies $\alpha$-differential-privacy with 

$$\alpha=\frac{(\frac{3}{4})^{100}}{(\frac{3}{4})^{99}\frac{1}{4}}=3.$$

This indicates that there is no amplification by Shuffling in the pure $\alpha$-differential-privacy
model. One needs to include $\delta$ to see the amplification.

Again letting $\delta=(0.75)^{101}$ this mechanism satisfies $(\alpha,\delta)$ differential privacy
where $\alpha$ satisifes
$$\left(\frac{3}{4}\right)^{100} = \alpha \left(\frac{3}{4}\right)^{99}\frac{1}{4} + \delta$$
or
\begin{align*}
\alpha &=\frac{(\frac{3}{4})^{100} - \delta}{(\frac{3}{4})^{99}\frac{1}{4}} \\
&= 4\left(\frac{3}{4}  - \frac{\delta}{\left(\frac{3}{4}\right)^{99}} \right) \\
&= 3 - 4\delta\left(\frac{4}{3}\right)^{99} \\
&= 3 - 4(\frac{3}{4})^{101}\left(\frac{4}{3}\right)^{99} \\
&= 3 - 4\left(\frac{3}{4}\right)^2 \\
&= 3 - 3\frac{3}{4} = 3 \times (1/4) = 3/4.
\end{align*}

\bibliographystyle{amsalpha}

\bibliography{math}

\end{document}
