% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 9 \\ Congruence}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Congruence}

\begin{itemize}
  \item \textbf{Definition.} Let $a,b,m$ be integers with $m>1$.
  \item Then we say that $a$ is congruent to $b$ modulo $m$ iff $a\bmod m = b\bmod m$.
  \item We write this as $a\equiv b \pmod m$.
  \item If $a$ is not congruent to $b$ modulo $m$ we write $a\not\equiv b \pmod m$.
  \item \textbf{Example.} $13\bmod 5 = 3$.
  \item $33 \bmod 5 = 3$.
  \item So $13 \equiv 33 \pmod 5$.
  \item \textbf{Example.} $13 \bmod 7 = 6$
  \item $33 \bmod 7 = 5$.
  \item So $13 \not\equiv 33 \bmod 7$.
\end{itemize}

\end{frame}

\begin{frame}{Two different uses of ``mod''}

\begin{itemize}
  \item Notice the two different ways of using the symbol ``mod''.
  \item When we write $13\bmod 5 = 3$ the symbol ``mod'' is being used as a function.
  \item The expression ``$13\bmod 5$'' has a value that is an integer. (The value is 3.)
  \item When we write $13 \equiv 33 \pmod 5$ the pair of symbols $\equiv$ and ``mod'' are being
  used together to represent a relation.
  \item The expression ``$13 \equiv 33 \pmod 5$'' has a value that is either true or false. (The value is true.)
  \item You can always tell which meaning of ``mod'' is intended by checking for the $\equiv$ and
  also checking for parentheses around  ``mod'' next to a number.
\end{itemize}

\end{frame}

\begin{frame}{More examples ``mod''}

\begin{itemize}
  \item Give the value of each expression:
  \item \textbf{(a)} $16 \bmod 6$.
  \item \textbf{Answer:} 4
  \item \textbf{(b)} $26 \bmod 6$
  \item \textbf{Answer:} 2
  \item \textbf{(c)} $8 \bmod 6$.
  \item \textbf{Answer:} 2
  \item \textbf{(d)} $16 \equiv 26 \pmod 6$.
  \item \textbf{Answer:} False
  \item \textbf{(e)} $8 \equiv 26 \pmod 6$.
  \item \textbf{Answer:} True
\end{itemize}

\end{frame}

\begin{frame}{Slots in Blocks}

\begin{itemize}
  \item  You can visualize congruence in terms of ``slots in blocks.''
  \item Take congruence mod 3 for example.
  \item We can divide the integers into infinitely many blocks of size 3.
  \item $\singleton{0,1,2}$, $\singleton{3,4,5}$, $\singleton{6,7,8}$, $\cdots$
  \item Every integer is in exactly one block.
  \item For example 7 is in the block $\singleton{6,7,8}$.
  \item Each block has three slots: slot 0, slot 1, slot 2.
  \item 7 is in slot 1 of it's block. (i.e. $7\bmod 3 = 1$)
  \item 4 is also in slot 1 of it's block.
  \item So $4\equiv 7 \pmod 3$.
\end{itemize}

\end{frame}


\begin{frame}{Another characterization of congruence}

\begin{itemize}
  \item  Let $a,b,m$ be integers with $m>1$.
  \item Then $a\equiv b \pmod m$ iff $m \divides (a-b)$.
  \item Example: Suppose we want to know whether or not $26 \equiv 8 \pmod 6$.
  \item $26 - 8 = 18$.
  \item $6\divides 18$.
  \item So $26 \equiv 8 \pmod 6$.
  \item Proof of the new characterization:
  \item Write $a=q_1 m + r_1$, $b=q_2 m + r_2$, with $0\leq r_1,r_2 < m$.
  \item $a -b = (q_1-q_2) m + (r_1 - r_2)$.
  \item $a\equiv b \pmod m$ iff $r_1 = r_2$ iff $a-b$ is a multiple of $m$.
\end{itemize}

\end{frame}

\begin{frame}{Finding other things congruent to $a$.}

\begin{itemize}
  \item  Question: Suppose you have an integer $a$ and you want
  a different integer $b$ such that $a\equiv b \pmod m$.
  \item How would you go about finding $b$?
  \item Answer: You would add a multiple of $m$.
  \item Example: Find several integers $b$ such that $7\equiv b \pmod 3$.
  \item Solution: Try 10, 13, 16, 19.
  \item Why? Because $a\equiv b \pmod m$ iff $b=a + qm$ for some $q$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Congruence classes}

\begin{itemize}
  \item  Let $a,m$ be integers with $m>1$.
  \item The congruence class of $a$ mod $m$ is the set of all integers $b$ such that $a\equiv b \pmod m$.
  \item $[a]_m = \setof{b\in\Z}{a\equiv b \pmod m}$.
  \item Example. Find $[4]_7$.
  \item $[4]_7 = \singleton{\cdots -24, -17, -10, -3, 4, 11, 18, 25, 32, \cdots}$.
  \item Example. Find $[11]_7$.
  \item $[11]_7 = \singleton{\cdots -24, -17, -10, -3, 4, 11, 18, 25, 32, \cdots}$.
\end{itemize}

\end{frame}

\begin{frame}{Congruence is transitive}

\begin{itemize}
  \item  Let $a,b,c,m$ be integers with $m>1$.
  \item Suppose $a\equiv b \pmod m$ and $b \equiv c \pmod m$.
  \item Then $a\equiv c \pmod m$.
  \item Example: $1 \equiv 4 \pmod 3$.
  \item $4 \equiv 7 \pmod 3$.
  \item So $1 \equiv 7 \pmod 3$.
  \item Proof: If $a \bmod m = b \bmod m$
  \item and $b \bmod m = c \bmod m$,
  \item then $a\bmod m = c \bmod m$.
\end{itemize}

\end{frame}


\begin{frame}{Congruence is symmetric}

\begin{itemize}
  \item  Let $a,b,m$ be integers with $m>1$.
  \item Suppose $a\equiv b \pmod m$.
  \item Then $b\equiv a \pmod m$.
  \item Proof: If $a \bmod m = b \bmod m$,
  \item then $b\bmod m = a \bmod m$.
\end{itemize}

\end{frame}

\begin{frame}{Congruence is reflexive}

\begin{itemize}
  \item  Let $a, m$ be integers with $m>1$.
  \item Then $a\equiv a \pmod m$.
  \item Proof: $a \bmod m = a \bmod m$.
\end{itemize}

\end{frame}


\begin{frame}{Congruence is an equivalence relation}

\begin{itemize}
  \item  Congruence is transitive, symmetric and reflexive.
  \item This makes it an equivalence relation.
  \item This means that the congruence classes mod $m$ partition the integers into $m$ pieces.
  \item Example: There are three congruence classes mod 3.
  \item They partition the integers into three disjoint infinite sets.
  \item $[0]_3 = \singleton{\cdots -9, -6, -3, 0, 3, 6, 9,\cdots}$
  \item $[1]_3 = \singleton{\cdots -8, -5, -2, 1, 4, 7, 10,\cdots}$
  \item $[2]_3 = \singleton{\cdots -7, -4, -1, 2, 5, 8, 11,\cdots}$
  \item Every integer is in exactly one of those sets.
  \item In terms of slots in blocks, we have partitioned the integers into slot 0 integers,
  slot 1 integers, and slot 2 integers.
\end{itemize}

\end{frame}


\begin{frame}{Congruence vs Congruence Classes}

\begin{itemize}
  \item  $a\equiv b \bmod m$ iff $[a]_m = [b]_m$
  \item iff $a\in[b]_m$ iff $b\in[a]_m$.
  \item Example:
  \item $[0]_3 = \singleton{\cdots -9, -6, -3, 0, 3, 6, 9,\cdots} = [9]_3$
  \item $[1]_3 = \singleton{\cdots -8, -5, -2, 1, 4, 7, 10,\cdots} =[10]_3$
  \item $[2]_3 = \singleton{\cdots -7, -4, -1, 2, 5, 8, 11,\cdots} = [11]_3$
  \item $1 \equiv 10 \bmod 3$
  \item $[1]_3 = [10]_3$
  \item $10\in [1]_3$
  \item $1\in[10]_3$.
\end{itemize}

\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A complete system of residues}

\begin{itemize}
  \item  A \emph{complete system of residues}  mod $m$ is a set of integers that
  chooses exactly one integer from each of the congruence classes mod $m$.
  \item Example. Give a complete system of residues mod $3$.
  \item $[0]_3 = \singleton{\cdots -9, -6, -3, 0, 3, 6, 9,\cdots}$
  \item $[1]_3 = \singleton{\cdots -8, -5, -2, 1, 4, 7, 10,\cdots}$
  \item $[2]_3 = \singleton{\cdots -7, -4, -1, 2, 5, 8, 11,\cdots}$
  \item Therefore each of the following are a complete system of residues mod 3.
  \item (a) $\singleton{0,1,2}$.
  \item (b) $\singleton{-3,-2,-1}$.
  \item (c) $\singleton{0,1,-1}$.
  \item (d) $\singleton{-9,7,5}$.
  \item In all cases the system consists of 3 mutually non-congruent integers.
\end{itemize}

\end{frame}

\begin{frame}{Least non-negative residues}

\begin{itemize}
  \item  The \emph{least non-negative residues} modulo $m$ is the complete set of residues
  consisting of the least non-negative integer from each congruence class.
  \item Example. Find the least non-negative residues mod $3$.
  \item $[0]_3 = \singleton{\cdots -9, -6, -3, 0, 3, 6, 9,\cdots}$
  \item $[1]_3 = \singleton{\cdots -8, -5, -2, 1, 4, 7, 10,\cdots}$
  \item $[2]_3 = \singleton{\cdots -7, -4, -1, 2, 5, 8, 11,\cdots}$
  \item Therefore the least non-negative residues modulo 3 are:
  \item $\singleton{0,1,2}$.
\end{itemize}

\end{frame}

\begin{frame}{Absolute least residues}

\begin{itemize}
  \item  The \emph{absolute least residues} modulo $m$ is the complete set of residues
  consisting of the element of each congruence class of smallest absolute value.
  \item Example. Find the absolute least residues mod $3$.
 \item $[0]_3 = \singleton{\cdots -9, -6, -3, 0, 3, 6, 9,\cdots}$
  \item $[1]_3 = \singleton{\cdots -8, -5, -2, 1, 4, 7, 10,\cdots}$
  \item $[2](2)_3 = \singleton{\cdots -7, -4, -1, 2, 5, 8, 11,\cdots}$
  \item Therefore the absolute residues modulo 3 are:
  \item $\singleton{0,1,-1}$.
  \item Usually we will write this as $\singleton{-1,0,1}$.
\end{itemize}

\end{frame}

\begin{frame}{Example}

\begin{itemize}
  \item  Give the set of least non-negative residues mod 5.
  \item Answer: $\singleton{0,1,2,3,4}$.
  \item Every integer is congruent mod 5 to exactly one of these.
  \item Give the set of absolute least residues mod 5.
  \item Answer: $\singleton{-2,-1,0,1,2}$.
  \item Every integer is congruent mod 5 to exactly one of these.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Characterization of a complete system of residues}

\begin{itemize}
  \item  \textbf{Lemma} Let $m>1$ be an integer. Let $S$ be a finite set of integers.
  \item Then $S$ is a complete system of residues modulo $m$ iff
  \item (a) $S$ contains $m$ elements.
  \item (b) The elements of $S$ are mutually incongruent.
  \item \textbf{proof} If $S$ is a complete system of residues then it contains exactly
  one element from each of the $m$ congruence classes.
  \item So it contains $m$ mutually incongruent elements.
  \item Conversely, suppose $S$ contains $m$ mutually incongruent elements.
  \item Since the elements are mutually incongruent, they come from different congruence classes.
  \item Since there are $m$ of them they must represent all $m$ of the congruence classes. $\qed$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example problem: Is a complete system of residues?}

\begin{itemize}
  \item  Example: Is $S=\singleton{0, -10, 31, 15}$ a complete system of residues modulo 6?
  \item Answer: No, it contains only 4 elements.
  \item  Example: Is $S=\singleton{0,-10,31,  15, 3, 11}$ a complete system of residues modulo 6?
  \item Answer: No, 15 and 3 are congruent modulo 6.
  \item Example: Is $S=\singleton{0, -10,31, 15, 22, 11}$ a complete system of residues modulo 6?
  \item Answer: Yes
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Consecutive Complete Systems}

\begin{itemize}
  \item  \textbf{Lemma} Let $m>1$ be an integer. Then every set of $m$ consecutive integers is
  a complete system of residues modulo $m$.
  \item \textbf{proof} Let $S=\singleton{x, x+1, x+2, \cdots ,x+(m-1)}$.
  \item Then every two elements of $S$ are non-congruent mod $m$ because their difference is less than $m$. $\qed$.
  \item Example: The following are all complete systems of residues mod 7
  \item $\singleton{14,15,16,17,18,19, 20}$
  \item $\singleton{-103, -102, -101, -100, -99, -98, -97}$.
  \item $\singleton{0,1,2,3,4,5, 6}$. (The least non-negative residues.)
  \item $\singleton{-3,-2,-1,0,1,2,3}$. (The absolute least residues.)
  \item It is useful to understand this in terms of slots and blocks.
  \item The \emph{standard} blocks are lined up so the slots yield the least non-negative residues.
  \item But we can consider non-standard blocks that are shifted relative to the standard blocks.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Residue Of an Integer}

\begin{itemize}
  \item  \textbf{Definition} Let $a, m \in\Z$ with $m>1$.
  \item The least non-negative residue of $a$ mod $m$ is just $a\mod m$.
  \item The absolute least residue of $a$ mod $m$ is the unique element $x$ of
  the set of absolute least residues mod $m$ such that $x\equiv a \pmod m$.
  \item Example: Find the least non-negative residue of 39 mod 7.
  \item Answer: 4
  \item Example: Find the absolute least residue of 39 mod 7.
  \item Answer: $-3$.
  \item Why: Because $-3\equiv 4 \equiv 39 \pmod 7$ and $-3$ is an absolute least residue mod 7.
  Note that 4 is not an absolute least residue mod 7.
\end{itemize}

\end{frame}



\end{document}
