% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 8 \\ More about Primes}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Non-Standard Prime Decomposition}

\begin{itemize}
  \item The prime decomposition of 20 is $20 = 2^2 \times 5$.
  \item But we could also write $20 = 2^2 \times 3^0 \times 5$.
  \item The \emph{standard} prime decomposition only uses positive exponents.
  But a non-standard decomposition may include the exponent 0.
  \item The advantage of a non-standard decomposition is it allows us to express
  two different integers using the same set of primes.
  \item Example: Give (non-standard) prime decompositions of 12 and 20 that use the same set of primes.
  \item $12 = 2^2 \cdot 3 \cdot 5^0$.
  \item $20 = 2^2 \cdot 3^0 \cdot 5$.
\end{itemize}

\end{frame}

\begin{frame}{Divisibility In Terms of Prime Powers}

\begin{itemize}
  \item \textbf{Lemma.} Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $a\divides b$ iff $e_i \leq f_i$ for $i=1,2,\cdots k$.
  \item Examples
\end{itemize}

\end{frame}


\begin{frame}{GCD In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $\gcd(a,b) = p_1^{\min(e_1,f_1)} \cdot p_2^{\min(e_2,f_2)} \cdots p_k^{\min(e_k,f_k)}$
\end{itemize}

\end{frame}

\begin{frame}{Example}

\begin{itemize}
  \item \textbf{Example} Let $a=108$, $b=120$. Find $\gcd(a,b)$.
  \item $108 = 2 \cdot 54 = 2 \cdot 2 \cdot 27 = 2^2 \cdot 3^3 = 2^2 \cdot 3^3 \cdot 5^0$.
  \item $120 = 2 \cdot 60 = 2 \cdot 2 \cdot 3 = 2\cdot 2 \cdot 2 \cdot 15 = 2^3 \cdot 3 \cdot 5$.
  \item $\gcd(108,120) = 2^2\cdot 3^1 \cdot 5^0 = 12$.
  \item \textbf{Example} Two positive integers are relatively prime, iff their two standard prime
  decompositions do not share any common primes.
\end{itemize}

\end{frame}

\begin{frame}{LCM}

\begin{itemize}
  \item \textbf{Definition.} Let $a$ and $b$ be positive integers.
  Then their \emph{least common multiple} is the least integer that is
  a multiple of both $a$ and $b$.
  \item This is written $\lcm(a,b)$.
  \item Notice that $ab$ is a common multiple of $a$ and $b$.
  \item So $a,b\leq \lcm(a,b) \leq ab$.
  \item Example: Find $\lcm(10,12)$.
\end{itemize}

\end{frame}

\begin{frame}{LCM In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $\lcm(a,b) = p_1^{\max(e_1,f_1)} \cdot p_2^{\max(e_2,f_2)} \cdots p_k^{\max(e_k,f_k)}$
\end{itemize}

\end{frame}

\begin{frame}{Products In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $ab = p_1^{e_1+f_1} \cdot p_2^{e_2+f_2} \cdots p_k^{e_k+f_k}$
\end{itemize}

\end{frame}

\begin{frame}{GCD, LCM and Product}

\begin{itemize}
  \item \textbf{Lemma.} Let $a,b$ be positive integers.
  \item Then $ab = \gcd(a,b) \cdot \lcm(a,b)$.
  \item \textbf{proof.} Write the prime decompositions of $a$ and $b$ as
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item $\gcd(a,b)\times \lcm(a,b) =$
  \item $p_1^{\max(e_1,f_1)+\min(e_1,f_1)} \cdot p_2^{\max(e_2,f_2)+\min(e_2,f_2)} \cdots p_k^{\max(e_k,f_k)+\min(e_k,f_k)}$
  \item $=p_1^{e_1+f_1} \cdot p_2^{e_2+f_2} \cdots p_k^{e_k+f_k}$
  \item $=ab$ $\qed$
\end{itemize}

\end{frame}

\begin{frame}{Factor out the GCD}

\begin{itemize}
  \item Let $a,b$ be positive integers. Let $d=\gcd(a,b)$.
  \item Then there are positive integers $\abar,\bbar$ such that
  \item $a=d\abar$, $b=d\bbar$
  \item $\abar$ and $\bbar$ are relatively prime.
  \item Example: $a=2^2\cdot 3 \cdot 5^2 = 300$.
  \item $b=2 \cdot 5^3 \cdot 7 =1750$.
  \item Then $d=2\cdot 5^2 = 50$.
  \item $\abar = 2\cdot 3 = 6$
  \item $\bbar = 5 \cdot 7 = 35$.
  \item $300 = 50 \cdot 6$
  \item $1750 = 50 \cdot 35$
  \item $\gcd(6,35) = 1$.
\end{itemize}

\end{frame}

\begin{frame}{Distribution of Primes}

\begin{itemize}
  \item Draw a number line and put a red dot wherever there is a prime number.
  \item What does the resulting picture look like? Some questions:
  \item Does the red go on forever or does it eventually stop?
  \item What happens to the \emph{density} of red as we move to the right?
  \item Does the density increase or decrease?
\end{itemize}

\end{frame}

\begin{frame}{Distribution of Primes}

\begin{itemize}
  \item \textbf{Theorem.} There are infinitely many primes.
  \item \textbf{Theorem.} Given any positive integer $k$ there exists $k$
  consecutive composite integers.
  \item \textbf{proof.} Consider $(k+1)! + 2, (k+1)! + 3, \cdots (k+1)! + k + 1$.
  \item \textbf{Definition.} For $n$ a positive integer $\pi(n)$ is the number
  of primes less than $n$.
  \item The \emph{density} of primes less than $n$ is $\pi(n)/n$.
  \item \textbf{The Prime Number Theorem.} As $n$ grows large the density of
  primes approaches $1/\ln(n)$.
  \item https://primes.utm.edu/howmany.html
  \item https://en.wikipedia.org/wiki/Prime-counting\_function
\end{itemize}

\end{frame}


\end{document}
