\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 9 \\ Vector Spaces}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Fall 2018 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.}

\smallskip

\textbf{(a)} (i) no, (ii) yes, (iii) yes, (iv) yes, (v) no, (vi) no

\smallskip

\textbf{(b)}
$$f(x,y) =
-5
\begin{pmatrix}
x - y + e^x  + 1 \\
xy + \cos y + 3 \\
\end{pmatrix}
$$

\smallskip

\textbf{(b)}
$$f(x,y) =
\begin{pmatrix}
0 \\ 0
\end{pmatrix}
$$


\textbf{Problem 2.}  This is false. There can be only one addiditve identity. Here
is one way to see this. Consider the value of $\bv + \bzero$. On the one
hand since $\bzero$ is an addiditve identity $\bv + \bzero = \bv$.
On the other hand since $\bv$ is an addiditve identity $\bv + \bzero = \bzero$.
So then $\bv=\bzero$, so we never really had a second addiditve identity.

\medskip

\textbf{Problem 3.}

\smallskip

\textbf{(a)} Not a subspace. It does not contain the origin because
$0+0+0+1\not=0$.

\smallskip

\textbf({b}) Is a subspace. It contains the origin and is closed under
vector addition and scalar mulitplication.

\smallskip

\textbf{(c)} Is a subspace. It contains the origin and is closed under
vector addition and scalar mulitplication.

\smallskip

\textbf{(d)} Is a subspace. It contains the origin and is closed under
vector addition and scalar mulitplication.

\smallskip

\textbf{(e)} Not a subspace. It does not contain the origin.

\smallskip

\textbf{(f)} Not a subspace. It is not closed under scalar multiplication.
For example.
$\transpose{(2,1,0)}$ is in this set but $(-1)\transpose{(2,1,0)}=\transpose{(-2,-1,0)}$
is not.

\smallskip

\textbf{(g)} Is a subspace. It contains the origin and is closed under
vector addition and scalar mulitplication.

\smallskip

\textbf{(h)} Not a subspace. It is not closed under scalar multiplication.
For example $\transpose{(1,1,1)}$ is in the set because $1=1\cdot1$ but
$\transpose{(-1,-1,-1)}$ is not in the set because $-1\not=-1\cdot -1$.

\smallskip

\textbf{(i)} This is the zero subspace.

\smallskip

\textbf{(j)} Not a subspace. It is not closed under vector addition. For
example $\transpose{(1,0,0)}$ is in the set
because $1\cdot 0 =  0 \cdot 0 = 1\cdot 0$ and
and $\transpose{(0,1,0)}$ is in the set because $0\cdot 1 = 1\cdot 0 = 0\cdot 0$
but $\transpose{(1,1,0)}$ is not in the set because $1\cdot1 \not= 1\cdot 0$.

\smallskip

\textbf{Problem 4.} Let $I\subset\R^1$ be an interval. In general $I$ is not
a subspace of $\R^1$ because it is not closed under scalar multiplication.
For example if $I=(0,1)$ Then $\frac{1}{2}\in I$ but $1\notin I$. The one exception
is the interval $(-\infty, \infty)$. This is a subspace of $\R^1$ because it
is equal to $\R^1$.

\smallskip

\textbf{Problem 5.} Do exercise 2.2.6 from the textbook.

\smallskip

\textbf{(a)} The union of any two lines through the origin is an example.
For example the union of the $x$- and $y$-axes.

\smallskip

\textbf{(b)} The positive $x$-axis is an example.

\smallskip

Additionally the empty set is an example that works for both (a) and (b).

\smallskip

\textbf{Problem 6.}

Let $S_{(a,b)} = \setof{f\in C^0(\R)}{f(a)=b}$. Then $S$ is a subspace of
$C^0(\R)$ iff $b=0$.


\end{document}
