% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 7 \\ Primes}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Prime Numbers}

\begin{itemize}
  \item \textbf{Definition.} Let $n>1$ be an integer. Then $n$ is \emph{prime} iff its only positive divisors are 1 and itself.
  \item If $n>1$ is not prime it is called \emph{composite}.
  \item Example: 11 is prime. 12 is composite.
  \item Notice that 1 and 0 are special cases. They are not considered prime or composite.
  \item We could define prime and composite for negative integers less than -1 but we will not.
\end{itemize}

\end{frame}


\begin{frame}{The Fundamental Theorem of Arithmetic}

\begin{itemize}
  \item \textbf{Theorem.} Every integer greater than 1 can be expressed as a product of primes.
  \item This \emph{prime decomposition} or \emph{prime factorization} is unique up to the order of the primes.
  \item Example: $12 = 2\cdot 2 \cdot 3$ is the unique prime factorization of 12 up to the order of the primes.
  \item What does up to the order of the primes mean?
  \item $12 = 2 \cdot 3 \cdot 2$.
\end{itemize}

\end{frame}

\begin{frame}{The Fundamental Theorem of Arithmetic}

\begin{itemize}
  \item We can extend the theorem to all non-zero integers by allowing an initial factor of $\pm 1$.
  \item If we say the primes have to be in increasing ordering we can get rid of ``up to the order of the primes.''
  \item \textbf{Theorem.} Let $n$ be a non-zero integer. Then $n$ has a unique prime decomposition:
  \item $n=\pm1 \cdot p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item where each $p_i$ is prime and $p_1<p_2<\cdots < p_k$
  \item and each $e_i$ is a positive integer.
  \item Example: $12 = 1 \times 2^2 \times 3^1$.
  \item We will prove the Fundamental Theorem of Arithmetic over the course of the next few slides.
  \item This is an existence and uniqueness proof so there will be two directions. First we
  will prove that every integer has some prime decomposition.
\end{itemize}

\end{frame}

\begin{frame}{Proof By Induction}

\begin{itemize}
  \item Suppose $\sigma$ is a statement that we want to prove about all positive integers $\geq N$.
  \item Then it suffices to do the following two steps:
  \item (i) Show that $\sigma(N)$ holds.
  \item (ii) Fix $n > N$. Assume that $\sigma(k)$ holds for all integers $k$ with $N\leq k < n$.  Show that $\sigma(n)$ holds.
  \item For example, we want to show that for all integers $n\geq 2$, $n$ has a prime decomposition.
  \item Let $\sigma(n)$ be the statement ``$n$ has a prime decomposition''.
  \item Then it suffices to do the following:
  \item (i) Show that 2 has a prime decomposition.
  \item (ii) Fix $n>2$. Assume that for all $k$ with $2\leq k < n$, $k$ has a prime decomposition. Then prove that $n$ has a
  prime decomposition.
\end{itemize}
\end{frame}

\begin{frame}{Proof By Induction}

\begin{itemize}
  \item Proof that proof by induction is valid:
  \item Suppose that
  \item (i) $\sigma(N)$ holds and
  \item (ii) For all $n>N$, if $\sigma(k)$ holds for all $k$ with $N\leq k < n$ then $\sigma(n)$ holds.
  \item We must show that $\sigma(n)$ holds for all $n\geq N$.
  \item Suppose otherwise.
  \item Let $S=\setof{n\geq N}{\sigma(n)\text{ does not hold}}$. $S\not=\emptyset$.
  \item Let $n$ be the least element of $S$. (Using the wellordering property.)
  \item By (i) $\sigma(N)$ holds so $N\notin S$ so $n>N$.
  \item For all $k$ with $N\leq k < n$, $k\notin S$, so $\sigma(k)$ holds.
  \item By (ii), $\sigma(n)$ holds. Contradiction. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Fundamental Theorem of Algebra: Existence}

\begin{itemize}
  \item Proof that proof by induction is valid:
  \item Suppose that
  \item (i) $\sigma(N)$ holds and
  \item (ii) For all $n>N$, if $\sigma(k)$ holds for all $k$ with $N\leq k < n$ then $\sigma(n)$ holds.
  \item We must show that $\sigma(n)$ holds for all $n\geq N$.
  \item Suppose otherwise.
  \item Let $S=\setof{n\geq N}{\sigma(n)\text{ does not hold}}$. $S\not=\emptyset$.
  \item Let $n$ be the least element of $S$. (Using the wellordering property.)
  \item By (i) $\sigma(N)$ holds so $N\notin S$ so $n>N$.
  \item For all $k$ with $N\leq k < n$, $k\notin S$, so $\sigma(k)$ holds.
  \item By (ii), $\sigma(n)$ holds. Contradiction. $\qed$.
\end{itemize}

\end{frame}



\end{document}
