\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Notes On Quaternions}
\author{Mitch Rudominer}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In these notes we will study the quaternions and their relationship to the double covering $\pi:\SU(2)\to \SO(3)$.

\section{Quaternions}
\emph{The quaternions} is an algebra obtained from $\R^4$ by adding an associative multiplication operation. There are several natural ways to define the multiplcation. We give three and show they are equivalent.

\subsection{The Quaternions as a Matrix Algebra}

\begin{definition}
Let $\H_M \subset M(2; \C)$ be defined by 
$$
\H_M = \Bigsetof{ \begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix} 
}{\alpha,\beta \in \C}.
$$
\end{definition}

\begin{lemma}
\label{lemma:det_of_A}
Let
$$A= \begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix}
\in\H_M.$$
Then $\det(A) = |\alpha|^2 + |\beta|^2$ and $\det(A)>=0$ and $\det(A)=0$ iff $A=0$.
\end{lemma} 
\begin{proof}
Immediate.
\end{proof}


\begin{note}
$\setof{A\in \H_M}{\det(A) = 1} = \SU(2).$
\end{note}


\begin{lemma} $\H_M$  is a real sub-algebra of $M(2; \C)$.
\end{lemma}
\begin{proof}
Because complex conjugation is linear over $\R$, $\H_M$ is closed under addition and real scalar multiplication.
To see that $\H_M$ is closed under multiplication we compute 

\begin{equation}
\label{eq:matrixmult}
\begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix}
\times
\begin{pmatrix}
\gamma & \delta \\
-\deltabar & \gammabar
\end{pmatrix} 
= 
\begin{pmatrix}
\alpha\gamma - \beta\deltabar & \alpha\delta + \beta\gammabar \\
-\betabar\gamma -\alphabar\deltabar & -\betabar\delta + \alphabar\gammabar
\end{pmatrix} .
\end{equation}
\end{proof}

\begin{lemma}  $\H_M$ is a real division algebra.
\end{lemma}
\begin{proof}
Let $A\in \H_M$, $A\not=0$. We need to show that $A$ is invertible and $A^{-1}\in \H_M$.
Write $A = \begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix}$. Then

$\det(A) = |\alpha|^2 + |\beta|^2 \not = 0$ so $A$ is invertible.
Also $\det(A)\in\R$ so

$$A^{-1} = \det(A)^{-1} \begin{pmatrix}
\alphabar & -\beta \\
\betabar & \alpha
\end{pmatrix} \in \H_M.$$

\end{proof}

\begin{lemma}
$H_M$ is closed under Hermitian adjoint.
\end{lemma}
\begin{proof}
\begin{equation}
\label{eq:quaternion_matrix_adjoint}
\begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix}^{*}=
\begin{pmatrix}
\alphabar & -\beta \\
\betabar & \alpha
\end{pmatrix}\in \H_M.
\end{equation}
\end{proof}

\begin{lemma}
\label{lemma:a_times_a_star}
For $A\in \H_M$, $AA^* = \det(A) I$.
\end{lemma}
\begin{proof}
Let $A= \begin{pmatrix}\alpha & \beta \\-\betabar & \alphabar\end{pmatrix}$.
Let $d=\det(A) = |\alpha|^2+|\beta|^2$. then
$$
AA^* = \begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix}
\begin{pmatrix}
\alphabar & -\beta \\
\betabar & \alpha
\end{pmatrix}
=\begin{pmatrix}d & 0\\ 0 & d\end{pmatrix}.
$$
\end{proof}


For $\alpha,\beta \in \C$ let 
$$Q(\alpha,\beta) = \begin{pmatrix}
\alpha & \beta \\
-\bar{\beta} & \bar{\alpha}
\end{pmatrix}.$$

For $x,y\in\R$ let $\alpha(x,y) = x + iy\in\C$.

For $x,y,z,w\in\R$ let $Q(x,y,z,w) = Q(\alpha(x,y), \alpha(z,w))$.

For $v\in\R^4$ let $Q(v) = Q(v^0, v^1, v^2, v^3)$.

This last $Q$ is clearly a bijection from $\R^4$ to the matrix algebra $\H_M$ defined above. 

\begin{definition}
We
define quaternionic multiplication on $\R^4$ using this mapping.
For $v,w\in\R^4$, define
$$vw = Q^{-1}(Q(v)Q(w)).$$

With this multiplication, in addition to the
standard vector space structure,
$\R^4$ becomes a real associative algebra we call the quaternions.
We use the symbol $\H$ to refer to the quaternions.
\end{definition}

\begin{remark}
$Q:\H\to \H_M$ is an algebra isomorphism. This is trivial
as we defined multiplication in $\H$ so as to make this so.
\end{remark}

\begin{definition}
$0,1\in\H$ are defined by
$$0=\begin{pmatrix}0\\0\\0\\0\end{pmatrix}\text{   }
1=\begin{pmatrix}1\\0\\0\\0\end{pmatrix}.$$

Since $Q(0)$ is the zero matrix and $Q(1)$ is the identity matrix,
$0$ is the zero element of $\H$ and $1$ is the identity element
of $\H$.

We can tell from context when the symbols 0 and 1 refer to
quaternions instead of numbers.
\end{definition}

\begin{definition}
We define quaternionic inverses using the mapping $Q$.
For $v\in\H$, $v\not=0$,
$$v^{-1} = Q^{-1}(Q(v)^{-1}).$$

With this inverse $\H$ is a division algebra.
\end{definition}

\begin{definition}
A \emph{scalar} quaternion is a $q\in\H$ of the form
$q=\begin{pmatrix}c\\0\\0\\0\end{pmatrix}$ for $c\in\R$,
and in this case we say that $c$ is the \emph{scalar part}
of $q$.
Equivalently $q$ is a scalar quaternion iff $Q(q)$ is
a scalar matrix. $0,1\in\H$ are examples of scalar quaternions.
\end{definition}

\begin{lemma}
Let $c\in\R$ and $q\in\H$.
Then $cq = \begin{pmatrix}c\\0\\0\\0\end{pmatrix} q$. That is,
scalar multiplication by $c$ is the same as quaternionic
multiplication by the scalar quaternion with scalar part $c$.
\end{lemma}
\begin{proof}
This is true becuase the analogous fact is true about scalar matrices.
\end{proof}

\begin{lemma}
\label{corollary:scalars_commute}
A scalar quaternion commutes with all quaternions.
Also real scalar multiplication commutes with quaternionic multiplication.
\end{lemma}
\begin{proof}
This is true because the analogous fact is true about
scalar matrices and matrix multiplication.
\end{proof}

\begin{remark}
If $c\in \R$ and $1$ represents the unit quaternion, then $c 1$ is the scalar quaternion
with scalar part $c$. We will henceforth simplify this and write $c 1$ as $c$.
Thus we are identifying the real numbers and the scalar quaternions. One can tell from
context which kind of object a symbol refers to.
\end{remark}


\begin{definition}
For $q\in\H$ let $|q|$ denote the Euclidean norm of $q$ in $\R^4$.
\end{definition}

\begin{lemma}
\label{lemma:norm_is_determinant}
For $q\in\H$, $|q|^2 = \det(Q(q))$.
\end{lemma}
\begin{proof}
By Lemma \ref{lemma:det_of_A},
$\det(Q(q)) = |q^0 + iq^1|^2 + |q^2 + iq^3|^2 = |q|^2$.
\end{proof}

\begin{lemma}
For all $q_1,q_2 \in \H$, $|q_1 q_2| = |q_1| |q_2|$.
\end{lemma}
\begin{proof}
$|q_1 q_2|^2 = \det(Q(q_1 q_2)) = \det(Q(q_1) Q(q_2)) = \det(Q(q_1))\det(Q(q_2)) = |q_1|^2 |q_2|^2.$
The result follows since norms are non-negative.
\end{proof}

\begin{definition}
For $q\in\H$ we define its \emph{conjugate} to be $\bar{q} = Q^{-1}(Q(q)^{*})$.
That is, the conjugate of $q$ corresponds to the Hermitian adjoint of $Q(q)$.
\end{definition}

\begin{corollary}
\label{corollary:q_times_qbar}
 For $q\in\H$,  $q \qbar = |q|^2 $. (We are using our convention of conflating a real number with its scalar quaternion.)
 \end{corollary}
 \begin{proof}  $Q(q \qbar) = Q(q) Q(q)^* = \det(Q(q))^2 I$ (by Lemma \ref{lemma:a_times_a_star})
 $= |q|^2 I$ (by Lemma \ref{lemma:norm_is_determinant}) $ = Q(|q|^2 1)$.
 \end{proof}

 \begin{corollary}
 \label{corollary:q_inverse_by_qbar}
 For $q\in\H$, 
 $$q^{-1} = \frac{1}{|q|^2} \qbar.$$
 \end{corollary}
 \begin{proof}
 By Corollary \ref{corollary:q_times_qbar}
 $$\frac{1}{|q|^2} q \qbar = 1.$$ 
 \end{proof}

 \begin{lemma}
 For $q\in\H$, $\bar{\qbar} = q$.
 \end{lemma}
 \begin{proof}
 Immediate because for $A$ a matrix $(A^{*})^{*} = A$.
 \end{proof}

\subsection{The Quaternions in terms of vector operations}

We can write an element of $\R^4$ as
$\begin{pmatrix}
c\\
\vec{v}
\end{pmatrix}$ with $c\in \R$ and $\vec{v}\in\R^3$. When
we do this we call $c$ the \emph{real} or \emph{scalar}
part of the quaternion and $\vec{v}$ the \emph{pure} part. 

\begin{definition}
A \emph{pure} quaternion is one in which the real part is zero.
A \emph{scalar} quaternion is one in which the pure part is zero.
(This is a restatement of the definition from the previous subsection.)
\end{definition}

\begin{lemma}
\label{lemma:in_terms_of_vector_products}
With quaternionic multiplication as defined in the previous subsection and the vector notation of the current subsection we have
$$
\begin{pmatrix}
c\\
\vec{v}
\end{pmatrix}
\begin{pmatrix}
d\\
\vec{w}
\end{pmatrix}
=
\begin{pmatrix}
cd - \vec{v} \cdot \vec{w} \\
c\vec{w} + d\vec{v} + \vec{v}\times \vec{w}
\end{pmatrix}.
$$
\end{lemma}
\begin{proof}
Let $\alpha = c + iv_0$, $\beta = v_1 + i v_2$,
$\gamma = d + iw_0$, $\delta = w_1 + iw_2$.

Let 

$$
\begin{pmatrix}
e\\
\vec{u}
\end{pmatrix}
=
\begin{pmatrix}
c\\
\vec{v}
\end{pmatrix}
\begin{pmatrix}
d\\
\vec{w}
\end{pmatrix}.
$$

By equation \ref{eq:matrixmult} we have that
\begin{itemize}
\item $e=\re(\alpha\gamma - \beta\deltabar) = (cd -v_0 w_0) - (v_1 w_1 + v_2 w_2)$
\item $u_0 = \im(\alpha\gamma - \beta\deltabar) = (c w_0 + d v_0) - (- v_1 w2 + w_1 v_2)$
\item $u_1 = \re(\alpha\delta + \beta\gammabar) = (c w_1 - v_0 w_2) + (d v_1 + v_2 w_0)$
\item $u_2 = \im(\alpha\delta + \beta\gammabar) = (c w_2 + v_0 w_1) + (-v_1 w_0 + d v_2)$
\end{itemize}

So

$$
\begin{pmatrix}
e\\
\vec{u}
\end{pmatrix}
=
\begin{pmatrix}
cd - \vec{v} \cdot \vec{w} \\
c w_0 + d v_0 + ( v_1 w2 -  v_2 w_1) \\
c w_1 + d v_1 + (v_2 w_0 - v_0 w_2))\\
c w_2 + d v_2 + (v_0 w_1 - v_1 w_0)
\end{pmatrix}
=
\begin{pmatrix}
cd - \vec{v} \cdot \vec{w} \\
c\vec{w} + d\vec{v} + \vec{v}\times \vec{w}
\end{pmatrix}.
$$
\end{proof}

\begin{note}
We could have made Lemma \ref{lemma:in_terms_of_vector_products} the definition of quaternion
multiplication. But then we would have needed to prove that it is associative and
a division algebra.
\end{note}

\begin{corollary}
\label{corollary:pure_mult}
Multiplication of pure quaternions is given by the vector dot and cross products:
$$
\begin{pmatrix}
0\\
\vec{v}
\end{pmatrix}
\begin{pmatrix}
0\\
\vec{w}
\end{pmatrix}
=
\begin{pmatrix}
\vec{v} \cdot \vec{w} \\
\vec{v}\times \vec{w}
\end{pmatrix}.
$$
\end{corollary}

\begin{lemma}
\label{lemma:commute_criterion}
Let $q_1, q_2 \in \H$. Then $q_1 q_2 = q_2 q_1$ iff 
the pure parts of $q_1$ and $q_2$ are colinear.
\end{lemma}
\begin{proof}
Let $\vec{v}$ and $\vec{w}$ be the pure parts of $q_1$ and $q_2$ respectively.
By Lemma \ref{lemma:in_terms_of_vector_products}, $q_1$ and $q_2$ commute
iff $v_1 \times v_2$ = $v_2 \times v_1$ and this is so iff
$v_1 \times v_2 = 0$ iff $v_1$ and $v_2$ are colinear.
\end{proof}

\begin{lemma}
\label{lemma:qbar_for_vectors}
 For $q\in\H$, with $q=\begin{pmatrix}c \\ \vec{v}\end{pmatrix}$,
 $$\qbar = \begin{pmatrix}c \\ -\vec{v}\end{pmatrix}.$$
 \end{lemma}
 \begin{proof}
 Write $q=\begin{pmatrix}x \\ y \\ z \\ w\end{pmatrix}$.
 Let $\alpha = x + iy$, $\beta=z+iw$.
 Then $\alphabar = x - iy$ and $-\beta = -z - iw$.
 By equation \ref{eq:quaternion_matrix_adjoint},
 $$
 \qbar=Q^{-1}(Q(\alpha,\beta)^*)=Q^{-1}(Q(\alphabar,-\beta))=
 \begin{pmatrix}
 x \\ -y \\ -z \\ -w
 \end{pmatrix}.
 $$
 \end{proof}

 \begin{lemma}
Let $q\in\H$, $q\not=0$.
Write $q=\begin{pmatrix}c\\ \vec{v}\end{pmatrix}$. Then
$$q^{-1} = \frac{1}{|q|^2}\begin{pmatrix}c\\ -\vec{v}\end{pmatrix}.$$
\end{lemma}
\begin{proof}
By Corollary \ref{corollary:q_inverse_by_qbar}, 
$$q^{-1} = \frac{1}{|q|^2}\qbar$$ so this lemma
follws from Lemma \ref{lemma:qbar_for_vectors}.
\end{proof}


\subsection{The Quaternions in terms of canonical basis elements}

We define four standard basis quaternions:
$$
1 = \begin{pmatrix}1\\0\\0\\0\end{pmatrix},\text{  }
i =  \begin{pmatrix}0\\1\\0\\0\end{pmatrix},\text{  }
j =  \begin{pmatrix}0\\0\\1\\0\end{pmatrix},\text{  }
k =  \begin{pmatrix}0\\0\\0\\1\end{pmatrix},\text{.}
$$

\begin{note}
We have already defined the quaternion 1 in an earlier subsection.
As with 1, we are overloading the symbol $i$ to refer to a quaternion in addition
to a complex number. One can tell from context which object $i$ refers to.
\end{note}

\begin{remark}
Let $\vec{i}$, $\vec{j}$, $\vec{k}$ be the standard basis vectors of three-dimensional vector calculus.
Then
$i = \begin{pmatrix}0\\\vec{i}\end{pmatrix}$, 
$j = \begin{pmatrix}0\\\vec{j}\end{pmatrix}$, 
$k = \begin{pmatrix}0\\\vec{k}\end{pmatrix}$.
\end{remark}

\begin{note}
So in addition to $i$ as a complex number and $i$ as a quaternion we also have $\vec{i}$
as a vector in $\R^3$. We remark below that all three objects are connected.
\end{note}

\begin{lemma} We record the following basic equations about the basis elements.
\label{lemma:basic_equations}
\begin{enumerate}
\item $1$ is the multiplicative identity in the quaternions.
\item $i^2=j^2=k^2 = -1$.
\item $ij = -ji = k$.
\item $jk = -kj = i$.
\item $ki = -ik = j$.
\end{enumerate}
Here $-1$ is a scalar quaternion.
\end{lemma}
\begin{proof}
(1) $Q(1)$ is the identity matrix. We have already seen this
earlier in these notes.

(2) Note that in the equation below we use the letter $i$ to mean two different things
on the left and right side of the equation.

$$Q(i^2) = \begin{pmatrix}i & 0 \\ 0 & -i\end{pmatrix}^2 = \begin{pmatrix}-1 & 0 \\ 0 & -1\end{pmatrix} = Q(-1).$$

$$Q(j^2) = \begin{pmatrix}0 & 1 \\ -1 & 0\end{pmatrix}^2 = \begin{pmatrix}-1 & 0 \\ 0 & -1\end{pmatrix} = Q(-1).$$

$$Q(k^2) = \begin{pmatrix}0 & i \\ i & 0\end{pmatrix}^2 = \begin{pmatrix}-1 & 0 \\ 0 & -1\end{pmatrix} = Q(-1).$$

(3),(4),(5) follow from Lemma \ref{lemma:in_terms_of_vector_products} because the identities are true when we
 replace $i,j,k$ with $\vec{i},\vec{j},\vec{k}$, and replace quaternionic multiplication with the vector cross product.
 \end{proof}

 \begin{lemma}
 Quaternionic multiplication as we have defined it earlier is the unique multiplication on $\R^4$ 
that distributes over addition, commutes with scalar multiplication,
 and satisfies the equations in Lemma \ref{lemma:basic_equations}. Thus we could have defined the quaternions
 using the equations in Lemma \ref{lemma:basic_equations}.
 \end{lemma}
 \begin{proof}
 A binary operation on $\R^4$ that distributes over addition and commutes with scalar multiplication
 is the same thing as a bilinear transformation from $\R^4 \times \R^4$ to $\R^4$ and such a transformation
 is completely determined by its values on all ordered pairs from a fixed basis.
 \end{proof}

 \begin{remarks} $\R$, $\R^3$ and $\C$ all naturally embed into $\H$ and going forward we will often
 tacitly conflate all of these with their images in $\H$.
 \begin{enumerate}
 \item As previously mentioned, we will identify $\R$ with the scalar quaternions via the identification $x\mapsto x 1$ where $x$ is a
 real number and $1$ is the unit quaternion.
 \item Define the \emph{complex} quaternions to be the two-dimensional subspace of $\H$ spanned by 1 and $i$.
 It is easy to see that the complex quaternions are a sub-algebra of $\H$.
 We will identify $\C$ with this sub-algebra by identifying the number 1 with the quaternion 1 and the
 number $i$ with the quaternion $i$.
 \item We will identify $\R^3$ with the pure quaternions by identifying $\vec{i}$ with the quaternion $i$,
 $\vec{j}$ with the quaternion $j$, and $\vec{k}$ with the quaternion $k$.
 \item It is pleasing that these identifications unify all three uses of the letter $i$. In particular it is
 not obvious to the student who first encounters vector calculus that the vector $\vec{i}$ has anything
 to do with the imaginary unit $i$.
 \end{enumerate}
 \end{remarks}

 \section{Lie Group and Lie Algebra Structures for the Quaternions}

 Because $\H$ and $\H_M$ are associative algebras, they naturally have a commutator bracket
 making them into Lie algebras. For $\H_M$ this is the usual matrix commutator and so
 $\H_M$ is a Lie subalgebra of $M(2; \C)$. We have that $Q:\H\map\H_M$ is a Lie algebra isomorphism.

\begin{lemma}
\label{lemma:bracket_in_terms_of_vector_products}
The Lie bracket on $\H$ may be computed as
$$
\left[
\begin{pmatrix}
c\\
\vec{v}
\end{pmatrix},
\begin{pmatrix}
d\\
\vec{w}
\end{pmatrix}
\right]
=
\begin{pmatrix}
0 \\
2\vec{v}\times \vec{w}
\end{pmatrix}.
$$
\end{lemma}
\begin{proof}
Using Lemma \ref{lemma:in_terms_of_vector_products} we compute
$$
\begin{pmatrix}
cd - \vec{v} \cdot \vec{w} \\
c\vec{w} + d\vec{v} + \vec{v}\times \vec{w}
\end{pmatrix}
-
\begin{pmatrix}
dc - \vec{w} \cdot \vec{v} \\
d\vec{v} + c\vec{w} + \vec{w}\times \vec{v}
\end{pmatrix}
=
\begin{pmatrix}
0 \\
 \vec{v}\times \vec{w} - (-\vec{v}\times \vec{w})
\end{pmatrix}
$$
\end{proof}

\begin{definition}
Let $\H_0$ be the space of pure quaternions. For $\vec{v}\in\R^3$, let $q(\vec{v})\in\H_0$
be the pure quaternion whose pure part is $\vec{v}$.
\end{definition}

\begin{corollary}
$\H_0$ with the commutator bracket is a three-dimensional Lie algebra, isomorphic to $\R^3$ with
the cross product:
$$\left(\H_0,[\cdot]\right) \isomorphic \left(\R^3, \times \right).$$
\end{corollary}
\begin{proof}
$\H_0$ is a three dimensional subspace closed under the the bracket, so it is a Lie
subalgebra of $\H$.
Let $\pi:\H_0\map\R^3$ be defined by $\pi(q(\vec{v}))=2\vec{v}$. This is a linear isomorphism
and we claim that for $q_1,q_2\in\H_0$, $\pi([q_1,q_2]) = \pi(q_1) \times \pi(q_2)$. To see
this fix $q_1=q(\vec{v}_1),q_2=q(\vec{v}_2)$. Then
$\pi([q_1,q_2]) = \pi(q(2 \vec{v}_1\times\vec{v}_2)) = 4 \vec{v}_1\times\vec{v}_2
= 2\vec{v}_1\times 2\vec{v}_2 = \pi(q_1) \times \pi(q_2)$.
\end{proof}

\begin{remark}
We need to be careful because we now have two different natural isomorphisms between $\H_0$ and $\R^{3}$.
We have the isomorphism that maps $i$ to $\vec{i}$ which is useful when we want to consider $\R^3$ as a subset of
$\H_0$ and which identifies the multiple
uses of the letter $i$, and we have the isomorphism that maps $i$ to $2\vec{i}$ which is a Lie algebra isomrphism.
\end{remark}

\begin{lemma}
\label{lemma:quaternion_lie_group}
$\H -\singleton{0}$ is a four-dimensional Lie group.
\end{lemma}
\begin{proof}
$M(2; \C)$ is a real vector space and so an 8-dimensional real manifold. 

$\H_M$ is a 4-dimensional vector subspace and so a 4-dimensional embedded submanifold of $M(2; \C)$.

$\H_M- \singleton{0}$ is an open submanifold of $\H_M$ and so a a 4-dimensional embedded submanifold of $M(2; \C)$.


$\H_M- \singleton{0}\subset \GL(2;\C)$ so $\H_M- \singleton{0}$ is a 4-dimensional embedded submanifold of $\GL(2;\C)$.


$\H_M- \singleton{0}$ is also a subgroup of $\GL(2;\C)$ and so it is a Lie subgroup.

$Q:\H\map\H_M$ is a vector space isomorphism and so a diffeomorphism.

So $Q:\H - \singleton{0}\map\H_M- \singleton{0}$ a diffeomorphism.

$Q:\H- \singleton{0}\map\H_M- \singleton{0}$ is also a group isomorphism.

So $\H - \singleton{0}$ is also a four-dimensional Lie group and $Q:\H-\singleton{0}\map\H_M-\singleton{0}$
 is a Lie group isomorphism.
\end{proof}


\begin{definition}
Let $\H_1$ be the space of quaternions with norm 1. Topologically $\H_1$ is $S^3$, the three-sphere in $\R^4$.
\end{definition}

\begin{corollary}
$\H_1$ is a three-dimensional Lie sub-group of $\H-\singleton{0}$, and $\H_1$ isomorphic to $\SU(2)$.
\end{corollary}
\begin{proof}
The proof of Lemma \ref{lemma:quaternion_lie_group} shows that $Q:\H-\singleton{0}\map\H_M-\singleton{0}$
is a Lie group isomorphism. The current result follows since
$Q$ maps $\H_1$ one-to-one and onto $\SU(2)$.
\end{proof}

\begin{lemma}
\begin{enumerate}
\item[(a)] $\H$ with the commutator bracket is the Lie algebra of the Lie group $\H-\singleton{0}$.


\item[(b)] $\H_0$ is the Lie algebra of $\H_1$.
\end{enumerate}
\end{lemma}
\begin{proof}
Let's take as given that $M(2;\C)$ with the commutator bracket is the Lie algebra for
$\GL(2;\C)$. 

For (a) it suffices to show that if $A\in M(2;\C)$ then $A$ is tangent to $\H_M - {0}$ at the identity
iff $A\in\H_M$.


Let 
$$
B(t) = 
\begin{pmatrix}
\alpha(t) & \beta(t) \\
-\betabar(t) & \alphabar(t)
\end{pmatrix} 
$$
be a smooth path in $\H_M - {0}$ with $B(0) = I$.
Let $A = B^{\prime}(0)$. We want to show that $A\in\H_M$.

We have 
$$
B^{\prime}(t) = 
\begin{pmatrix}
\alphaprime(t) & \betaprime(t) \\
-\betabar^{\prime}(t) & \alphabar^{\prime}(t)
\end{pmatrix} 
\in\H_M
$$
since taking the derivative commutes with complex conjugation and negation.

Conversely let 
$$
A = 
\begin{pmatrix}
\alpha & \beta \\
-\betabar & \alphabar
\end{pmatrix} 
\in\H_M.
$$
Let
$$
B(t) = 
\begin{pmatrix}
1 + \alpha t & \beta t \\
-\betabar t & 1 + \alphabar t
\end{pmatrix} .
$$
Then $B(0) = I$, $B(t)\in\H_M - {0}$ (for small $t$), and $B^{\prime} = A$.

Given part (a), in order to prove part (b) it suffices to show that for $q\in\H$,
$q$ is tangent to $\H_1$ at 1 iff $q\in\H_0$.

Let $s(t)$ be a path in $\H_1$ with $s(0) = 1$. We want to show that the real part of $q = \sprime(0)$ is zero.
This is true because the real part of $s(0)$  is 1 and that is the maximum value of the real part of $s(t)$
for all $t$ since $|s(t)| = 1$ for all $t$.

Conversely let $q\in\H_0$. We want to show that $q$ is tangent to $\H_1$ at 1. 
If $q=0$ then it is clearly tangent to $\H_1$ so suppose $q\not=0$.
Let us write the pure part of $q$ as $c\vec{v}$ where $\vec{v}$ is a unit 3 vector and $c$ is a positive real number.
Let
$$
s(t) = 
\begin{pmatrix}
\cos(ct) \\ \sin(ct) \vec{v}
\end{pmatrix}
.$$
Then $s(0) = 1$, $s(t)\in \H_1$ for all $t$, and
$$
\sprime(t) = 
\begin{pmatrix}
-c\sin(ct) \\ c\cos(ct) \vec{v}
\end{pmatrix}
$$
so that 
$$
\sprime(0) = 
\begin{pmatrix}
0 \\ c\vec{v}
\end{pmatrix}
= q.
$$
\end{proof}

\begin{definition}
For $q\in\H$, let $L_q:\H\map\H$ be left multiplication by $q$,
and let $R_q$ be right multiplication by $q$.
\end{definition}

\begin{lemma}
\label{lemma:mult_deriv}
If $q\in\H-\singleton{0}$, then the derivative of $L_q:\H-\singleton{0}\map\H-\singleton{0}$ is
$L_q:\H\map\H$, and similarly for $R_q$.

If $q\in\H_1$, then the derivative of $L_q:\H_1\map\H_1$ is
$L_q:\H_0\map q\H_0$, and similarly for $R_q$. 
\end{lemma}
\begin{proof}
This follows from the fact that the analogous facts are true of matrix multiplication.
\end{proof}

\begin{lemma}
Let $\exp:\H_0\map\H_1$ be the exponential map, as defined in Lie group theory.
Then 
$$\exp
\begin{pmatrix}
0 \\ c\vec{v}
\end{pmatrix}
 = 
\begin{pmatrix}
\cos(c) \\ \sin(c) \vec{v}
\end{pmatrix}
$$
where $\vec{v}$ is a unit vector in $\R^3$ and $c\geq 0$.
\end{lemma}
\begin{proof}
Fix 
$$q=
\begin{pmatrix}
0 \\ c\vec{v}
\end{pmatrix}
\in\H_0$$
with  $\vec{v}$ a unit vector in $\R^3$ and $c\geq 0$.

Let $V_q$ be the left-invariant vector field on $\H_1$ whose value at 1 is $q$.

By Lemma \ref{lemma:mult_deriv}
for $p\in\H_1$,$V_q(p) = pq$. 

Let
$$s(t)=
\begin{pmatrix}
\cos(ct) \\ \sin(ct) \vec{v}
\end{pmatrix}
.
$$
Then
$$V_q(s(t)) = \begin{pmatrix}
\cos(ct) \\ \sin(ct) \vec{v}
\end{pmatrix}
\begin{pmatrix}
0 \\ c \vec{v}
\end{pmatrix} =
\begin{pmatrix}
-c\sin(ct) \\ c \cos(ct)\vec{v}
\end{pmatrix} = s^{\prime}(t).
$$
Thus $s(t)$ is an integral curve of $V_q$ and
$$
s(1)=\begin{pmatrix}
\cos(c) \\ \sin(c) \vec{v}
\end{pmatrix}.
$$
\end{proof}

\begin{remarks} Let's compare $\exp:\H_0 \map \H_1$ to 
$\exp:\R\map C^1$. Let $v\in\H_0\intersect\H_1$. That is,
let $v = x i + y j + z k$ with $|v| = 1$. Then we have
\begin{enumerate}
\item[(a)] $\exp(\theta i) = \cos(\theta) + \sin(\theta)i$
\item[(b)] $\exp(\theta v) = \cos(\theta) + \sin(\theta)v$
\end{enumerate}
so we can see that (b) is a generalization of (a), since we get (a) by setting $x=1,y=z=0$. 

Of course in (a) we have that $\exp$ is given by the complex exponential function
$$\exp(z) = e^z = 1 + z + \frac{z^2}{2} + \frac{z^3}{3!} + ...$$
It turns out that we can say the same for (b). The abstract exponential function $\exp:\H_0 \map \H_1$
is also given by a concrete exponentional function on the quaternions given by the same power series.
But we do not pursue this further here as we do not need this result.
\end{remarks}

\section{Conjugation}

Let $G$ be a group and $g\in G$. Conjugation by $g$ is the map
$\phi_g:G\map G$ given by $\phi_g(h) = g h g^{-1}$. $\phi_g$ is a group automorphism.
If $G$ is a Lie group then $\phi_g$ is also a diffeomorphism, so it is a Lig group automorphism
and so it induces a Lie algebra homomorphism $T_{\phi_g}$.

\begin{lemma}
Let $G$ be a Lie group and $A$ the associated Lie algebra. Let $\exp:A\map G$ be the exponential map.
Let $a\in A$, and let $g=\exp(a)$. Then $T_{\phi_g}(a) = a$. 
\end{lemma}

We illustrate the previous Lemma in the case of the Quaternions by computing the conjugation explicitly.
\begin{multline*}
\begin{pmatrix}
\cos(c) \\ \sin(c)v
\end{pmatrix}
\begin{pmatrix}
0 \\ cv
\end{pmatrix}
\begin{pmatrix}
\cos(c) \\ - \sin(c)v
\end{pmatrix}
= 
\begin{pmatrix}
-c \sin(c) \\ c \cos(c) v
\end{pmatrix}
\begin{pmatrix}
\cos(c) \\ - \sin(c)v
\end{pmatrix} = \\
\begin{pmatrix}
-c \sin(c) \cos(c) + c \sin(c) \cos(c)\\\left(\sin^2(c) + \cos^2(c)\right) c v
\end{pmatrix}
=
\begin{pmatrix}
0 \\ cv
\end{pmatrix}.
\end{multline*}


\bibliographystyle{amsalpha}

\bibliography{math}

\end{document}