% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 5 \\ Matrices}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}



\begin{frame}{Matrices}

\begin{itemize}
\item A (real) \emph{matrix} is a rectangular array of real numbers.
\item We draw matrices, as in the example below, inside of elongated, curved
braces.
\item An $m\times n$ matrix (pronounded ``$m$-by-$n$'') has $m$-rows and $n$-columns.
\item If $A$ is an $m$-by-$n$ matrix, we say its \emph{size} is $m$-by-$n$.
\item We always list the number of rows before the number of columns.
\item The matrix $A$ below is a $4\times 5$ matrix because it has 4 rows
and 5 columns.
\item The size of $A$ is $4\times 5$.
\end{itemize}

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Entries of a matrix}

\begin{itemize}
\item We usually use capital Roman letters (lightface font), like $A$ or $M$, to name matrices.
\item The number in row $i$, column $j$ is called the $(i,j)$-th \emph{entry}.
\item If the matrix is called $A$ then by convention the symbol $a_{ij}$ refers
to the $(i,j)$-th entry of $A$.
We always list the row index first and the column index second.
\item For the matrix $A$ displayed below, the $(2,3)$-entry is 1 and the
$(3,2)$-entry is -3.
\end{itemize}

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Square matrices}

\begin{itemize}
\item If the number of rows equals the number of columns then the matrix is
called a \emph{square} matrix.
\item The matrix $M$ below is a square $2\times 2$ matrix.
\item If a matrix is not square it is called \emph{rectangular.}
\item The $4\times 5$ matrix $A$ from the previous slide is rectangular.
\end{itemize}

$$
M =
\begin{pmatrix}
\frac{1}{3} & \frac{2}{3} \\
0 & -1
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Row vectors and column vectors}

\begin{itemize}
\item A $1\times n$ matrix, with $n>1$, is called a \emph{row vector}.
\item $\sigma$ below has size $1\times 4$ so it is a row vector.
\item An $m\times 1$ matrix, with $m>1$, is called a \emph{column vector}.
\item $\bv$ below has size $3\times 1$ so it is a column vector.
\item We often use a different naming convention for row and column vectors
than we do for other matrices, as below where we used $\sigma$ and $\bv$
instead of names like $A$ and $M$.
\end{itemize}

$$
\begin{matrix}
\sigma =
\begin{pmatrix}
2, 2, 0, 1
\end{pmatrix}

&

\bv =
\begin{pmatrix}
1 \\
0 \\
-5
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Vector indexing}

\begin{itemize}
\item For a row vector like $\sigma$ we will often drop the row
index 1 and write only the column index.
\item So instead of writing $\sigma_{1,3} = 0$
\item we may write $\sigma_3 = 0$.
\item Similarly for column vectors like $\bv$ below we will often
drop the column index 1 and write only the row index.
\item So instead of writing $v_{2,1} = 0$,
\item we may write $v_2 = 0$.
\end{itemize}

$$
\begin{matrix}
\sigma =
\begin{pmatrix}
2, 2, 0, 1
\end{pmatrix}

&

\bv =
\begin{pmatrix}
1 \\
0 \\
-5
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{What are vectors in $\R^n$?}

\begin{itemize}
\item Previously, before we started talking about matrices, we said that
a \emph{vector in $\R^n$} is an $n$-tuple of numbers.
\item But now we have row vectors and column vectors. So which one of these
things is what we were previously calling a vector in $\R^n$?
\item Answer: Column vectors.
\item Convention: A vector in $\R^n$, from now on, is an $n\times 1$
column vector.
\item So then what is a row vector?
\item We will use row vectors to represent covectors.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Transpose of a vector}

\begin{itemize}
\item If $\bv$ is an $n\times 1$ column vector, then $\bv^{\text{T}}$
(pronounced $\bv$-\emph{transpose}) is the corresponding $1\times n$ row vector.
\item If $\bv$ is a $1\times n$ row vector, then $\bv^{\text{T}}$ is the
corresponding $n\times 1$ column vector.
\item $$(1, 0, -5)^{\text{T}} = \begin{pmatrix} 1 \\ 0 \\ -5 \end{pmatrix}$$
\item $$\begin{pmatrix} 1 \\ 0 \\ -5 \end{pmatrix}^{\text{T}} =  (1, 0, -5)$$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Transpose of a vector (continued)}

\begin{itemize}
\item Previously, before we started talking about matrices, we might have
written: $(1, 0, -5)\in\R^3$.
\item But the problem with this is that because we write an $n$-tuple from
left-to-right on a single line, it looks like a row vector, and we just said
that elements of $\R^n$ are column vectors.
\item So from now on we will write instead, $(1, 0, -5)^{\text{T}}\in\R^3$.
\item Similarly, suppose we want $\bw$ to be a row vector. We might write
let $\bv\in\R^n$, let $\bw=\bv^{\text{T}}$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{$1\times 1$ matrices}

\begin{itemize}
\item A $1\times 1$ matrix is identified with a scalar.
\item For example the $1\times 1$ matrix $(42)$ is considered to
be just the number 42.
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix multiplication}

\begin{itemize}
\item The most important thing we do with matrices is multiply two of
them together.
\item Let $A$ be an $n\times m$ matrix. Let $B$ be a $m\times k$ matrix.
\item Then the product $A B$ is defined.
\item $A B$ is the $n\times k$ matrix defined as follow:
\item Let $C=A B$. Then for $1\leq i \leq n$, and $1\leq j \leq k$,
\item $$c_{ij} = \sum_{t=1}^m a_{it} b_{tj}$$
\item To understand this better, we will look at some special cases first.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Which products are defined?}

\begin{itemize}
\item Suppose $A$ and $B$ are matrices of the sizes given below. Say
whether the product $A B$ is defined and if so say what its size is.
\item $A$ is $3\times 4$ and $B$ is $3\times 5$.
\item  $A B$ is not defined.
\item $A$ is $3 \times 4$ and $B$ is $4 \times 5$.
\item $A B$ is $3 \times 5$.
\item $A$ is $5\times 1$ and $B$ is $6\times 1$.
\item $A B$ is not defined.
\item $A$ is $5\times 1$ and $B$ is $1\times 6$.
\item $A B$ is $5\times 6$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A row vector times a column vector}

\begin{itemize}
\item Let's look at the special case of the previous definition where $n=k=1$.
\item Let $\bw$ be a $1\times m$ row vector. Let $\bv$ be a $m\times 1$ column vector.
\item Then the product $\bw \bv$ is defined.
\item $\bw \bv$ is the $1\times 1$ matrix (i.e. the scalar) defined as follow:
\item $$\bw \bv  = \sum_{t=1}^m w_t v_t$$
\item Let's look at an example to try to understand this.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A row vector times a column vector: Example}

\begin{itemize}
\item Let $\bw$ be the $1\times 3$ row vector below.
\item Let $\bv$ be the $3\times 1$ column vector below.
\item $$\bw \bv  = \sum_{t=1}^3 w_t  v_t = $$
\item $ = 1\cdot 4 + (-1)\cdot (-2) + 2\cdot 5$
\item $=16$
\item The product of a row vector times a column vector
\emph{of the same length} is just the \emph{dot product} of the two vectors.
\end{itemize}

$$
\begin{matrix}

\bw =

\begin{pmatrix}
1 & -1 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
4 \\ -2 \\ 5
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A new way of writing dot product.}

\begin{itemize}
\item Let $\bv,\bw \in \R^n$.
\item Then $\bv \bw$ is not defined.
\item But $\bv^{\text{T}} \bw$ is defined and
\item  $\bv^{\text{T}} \bw = \bv \cdot \bw$
\item Is $\bv \bw^{\text{T}}$ defined?
\item Yes $\bv \bw^{\text{T}}$ is an $n\times n$ matrix. We will look
at what this is later.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector}

\begin{itemize}
\item Let's look at the special case of the  definition where $k=1$.
\item Let $A$ be an $n\times m$ matrix. Let $\bv$ be a $m\times 1$ column vector.
\item Then the product $A \bv$ is defined.
\item $A \bv$ is the $n\times 1$ column vector defined as follows:
\item Let $\bw=A \bv$. Then for $1\leq i \leq n$,
\item $$w_i = \sum_{t=1}^m a_{it} v_{t}$$
\item Let's look at an example to try to understand what this means.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector: Example}

\begin{itemize}
\item Let $A$ be a $4 \times 5$ matrix.
\item Let $\bv$ be a $5\times 1$ column vector.
\item Then $A \bv$ is a $4 \times 1$ column vector.
\item Let $\bw=A \bv$. Then for $1\leq i \leq 4$,
\item $$w_i = \sum_{t=1}^5 a_{it} v_{t}$$
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
2 \\ 0 \\ -1 \\ 3 \\ 0
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector: Example continued}

\begin{itemize}
\item $$w_i = \sum_{t=1}^5 a_{it} v_{t}$$
\item The $i$-th element of $A\bv$ is the $i$-th row of $A$ times $\bv$.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
2 \\ 0 \\ -1\\ 3 \\ 0
\end{pmatrix}

&

A \bv =

\begin{pmatrix}
-3 \\ 15 \\ 10 \\ -23
\end{pmatrix}

\end{matrix}
$$

\end{frame}


\end{document}


