% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 6 \\ More about Primes}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Non-Standard Prime Decomposition}

\begin{itemize}
  \item The prime decomposition of 20 is $20 = 2^2 \times 5$.
  \item But we could also write $20 = 2^2 \times 3^0 \times 5$.
  \item The \emph{standard} prime decomposition only uses positive exponents.
  But a non-standard decomposition may include the exponent 0.
  \item The advantage of a non-standard decomposition is it allows us to express
  two different integers using the same set of primes.
  \item Example: Give (non-standard) prime decompositions of 12 and 20 that use the same set of primes.
  \item $12 = 2^2 \cdot 3 \cdot 5^0$.
  \item $20 = 2^2 \cdot 3^0 \cdot 5$.
\end{itemize}

\end{frame}

\begin{frame}{Divisibility In Terms of Prime Powers}

\begin{itemize}
  \item \textbf{Lemma.} Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $a\divides b$ iff $e_i \leq f_i$ for $i=1,2,\cdots k$.
  \item Example. Does 20 divide 150? Use prime decompositions to decide.
  \item $20 = 2^2 \cdot 3^0 \cdot 5^1$
  \item $150 = 2^1 \cdot 3^1 \cdot 5^2$
  \item Since the exponent on 2 is greater for 20 than for 150, $2\ndivides 150$.
\end{itemize}

\end{frame}


\begin{frame}{GCD In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $\gcd(a,b) = p_1^{\min(e_1,f_1)} \cdot p_2^{\min(e_2,f_2)} \cdots p_k^{\min(e_k,f_k)}$
\end{itemize}

\end{frame}

\begin{frame}{Example}

\begin{itemize}
  \item \textbf{Example} Let $a=108$, $b=120$. Use prime factorizations to find $\gcd(a,b)$.
  \item $108 =  2^2 \cdot 3^3 \cdot 5^0$.
  \item $120 =  2^3 \cdot 3^1 \cdot 5^1$.
  \item $\gcd(108,120) = 2^2\cdot 3^1 \cdot 5^0 = 12$.
  \item \textbf{Fact} Two positive integers are relatively prime, iff their two standard prime
  decompositions do not share any common primes.
  \item \textbf{Example}. Use prime factorization to decide if 45 and 49 are relatively prime.
  \item $45 = 3^2 \cdot 5$ and $49 = 7^2$. The two don't have any primes in common.
  \item So $\gcd(45,49) = 1$.
\end{itemize}

\end{frame}

\begin{frame}{LCM}

\begin{itemize}
  \item \textbf{Definition.} Let $a$ and $b$ be positive integers.
  Then their \emph{least common multiple} is the least integer that is
  a multiple of both $a$ and $b$.
  \item This is written $\lcm(a,b)$.
  \item Notice that $ab$ is a common multiple of $a$ and $b$.
  \item So $a,b\leq \lcm(a,b) \leq ab$.
  \item Example: Find $\lcm(10,12)$.
  \item Look at multiples of 12 until you find one that is divisible by 10:
  \item 12, 24, 36, 48, 60
  \item $\lcm(10,12) = 60$.
\end{itemize}

\end{frame}

\begin{frame}{LCM In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $\lcm(a,b) = p_1^{\max(e_1,f_1)} \cdot p_2^{\max(e_2,f_2)} \cdots p_k^{\max(e_k,f_k)}$
  \item \textbf{Example} Use prime factorizations to find $\lcm(10, 75)$.
  \item $10=2\cdot 3^0 \cdot 5$ and $75 = 2^0\cdot 3 \cdot 5^2$
  \item $\lcm(10,75) = 2 \cdot 3 \cdot 5^2 = 150$.
\end{itemize}

\end{frame}

\begin{frame}{Products In Terms of Prime Powers}

\begin{itemize}
  \item Let $a$ and $b$ be integers greater than 1 with prime decompositions
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item where the $p_i$ are distinct and the $e_i$ and the $f_i$ are allowed to be zero.
  \item Then $ab = p_1^{e_1+f_1} \cdot p_2^{e_2+f_2} \cdots p_k^{e_k+f_k}$
  \item \textbf{Example.} What is the prime factorization of $75 \times 18$?
  \item $75 = 3\cdot 5^2$, $18 = 2 \cdot 3^2$.
  \item $75 \times 18 = 2 \cdot 3^3 \cdot 5^2$.
\end{itemize}

\end{frame}

\begin{frame}{GCD, LCM and Product}

\begin{itemize}
  \item \textbf{Lemma.} Let $a,b$ be positive integers.
  \item Then $ab = \gcd(a,b) \cdot \lcm(a,b)$.
  \item \textbf{proof.} Write the prime decompositions of $a$ and $b$ as
  \item $a=p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item $b=p_1^{f_1} \cdot p_2^{f_2} \cdots p_k^{f_k}$
  \item $\gcd(a,b)\times \lcm(a,b) =$
  \item $p_1^{\max(e_1,f_1)+\min(e_1,f_1)} \cdot p_2^{\max(e_2,f_2)+\min(e_2,f_2)} \cdots p_k^{\max(e_k,f_k)+\min(e_k,f_k)}$
  \item $=p_1^{e_1+f_1} \cdot p_2^{e_2+f_2} \cdots p_k^{e_k+f_k}$
  \item $=ab$ $\qed$
\end{itemize}

\end{frame}

\begin{frame}{Factor out the GCD}

\begin{itemize}
  \item Let $a,b$ be positive integers. Let $d=\gcd(a,b)$.
  \item Then there are positive integers $\abar,\bbar$ such that
  \item $a=d\abar$, $b=d\bbar$
  \item $\abar$ and $\bbar$ are relatively prime.
  \item \textbf{Example.} Suppose $a=300$ and $b=1750$.
  \item Find $\abar,\bbar, d$ such that $a=d\abar, b=d\bbar, \gcd(\abar,\bbar)=1$.
  \item $300=2^2\cdot 3 \cdot 5^2$.
  \item $1750=2 \cdot 5^3 \cdot 7$.
  \item Then $d=\gcd(300,1750) = 2\cdot 5^2 = 50$.
  \item $\abar = 2\cdot 3 = 6$
  \item $\bbar = 5 \cdot 7 = 35$.
  \item $300 = 50 \cdot 6$
  \item $1750 = 50 \cdot 35$
  \item $\gcd(6,35) = 1$.
\end{itemize}

\end{frame}

\begin{frame}{Divisibility Tests}
\begin{itemize}
  \item \textbf{Fact.} Let $n$ be a positive integer.
  \item $n$ is divisible by 2 iff the last digit if $n$ is even.
  \item \textbf{proof.} Write $n=10k + d$ with $0\leq d < 10$.
  \item Then $d$ is the last digit of $n$.
  \item Since $2\divides 10$, we have that $2\divides n$ iff $2\divides d$. $\qed$.
  \item \textbf{Fact.}  Let $n$ be a positive integer.
  \item $n$ is divisible by 5 iff the last digit if $n$ is 5 or 0.
  \item \textbf{proof.} Write $n=10k + d$ with $0\leq d < 10$.
  \item Then $d$ is the last digit of $n$.
  \item Since $5\divides 10$, we have that $5\divides n$ iff $5\divides d$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Test for divisibility by 3}
\begin{itemize}
  \item \textbf{Fact.} Let $n$ be a positive integer.
  \item $n$ is divisible by 3 iff the sum of the digits of $n$ is divisible by 3.
  \item \textbf{proof.} Write $n=d_0 + 10d_1 + 100 d_2 + \cdots + 10^k d_k$.
  \item where the digits of $n$ are $d_0,d_1,\cdots d_k$.
  \item $n\bmod 3 = (d_0 + 10d_1 + 100 d_2 + \cdots + 10^k d_k)\bmod 3$
  \item  $=\left(d_0\bmod 3 + 10d_1\bmod 3 + \cdots + 10^k d_k \bmod 3 \right) \bmod 3$
  \item  $=\big(d_0\bmod 3 + (10\bmod 3)(d_1\bmod 3) + \cdots $
  \item  $  \cdots + (10^k\bmod 3) (d_k \bmod 3) \big) \bmod 3$
  \item  $=\left(d_0\bmod 3 + d_1\bmod 3 + \cdots + d_k \bmod 3 \right) \bmod 3$
  \item  $=\left(d_0 + d_1 + \cdots + d_k  \right) \bmod 3$. $\qed$
  \item \textbf{Note:} You will have to wait until a later lesson on modular arithmetic
  to understand why the above steps are valid.
\end{itemize}
\end{frame}

\begin{frame}{Factoring Algorithms}

\begin{itemize}
  \item Given an integer $n$, how can you determine whether or not $n$ is prime?
  \item An algorithm that does this is called a \emph{primality test.}
  \item More generally, given an integer $n$ how can you find its prime factorization?
  \item An algorithm that does this is called a factoring algorithm.
  \item Example: Let $n=187,272,000$. Is $n$ prime? Give its prime factorization.
  \item How would you figure this out?
  \item The method of \emph{trial division}:
  \item Iterate through the primes $p$ and start trying to divide $n$ by $p$.
  \item First we notice that $187,272,000$ is divisible by 1000.
  \item $187,272,000 = 187,272 \times 1000 $
  \item $= 187,272 \times 10^3$
  \item $= 187,272 \times 2^3 \times 5^3$
\end{itemize}

\end{frame}

\begin{frame}{Example Trial Division}

\begin{itemize}
  \item $187,272,000 = 187,272 \times 2^3 \times 5^3$
  \item(Factor out 2) $ = 93636 \times 2^4 \times 5^3$
  \item(Factor out 2) $ = 46818 \times 2^5 \times 5^3$
  \item(Factor out 2) $ = 23409 \times 2^6 \times 5^3$
  \item(Factor out 3) $ = 7803 \times 2^6 \times 3  \times 5^3$
  \item(Factor out 3) $ = 2601 \times 2^6 \times 3^2  \times 5^3$
  \item(Factor out 3) $ = 867 \times 2^6 \times 3^3  \times 5^3$
  \item(Factor out 3) $ = 289 \times 2^6 \times 3^4  \times 5^3$
  \item(Not divisible by 5)
  \item(Try to factor out 7: nope)
  \item(Try to factor out 11: nope)
  \item(Try to factor out 13: nope)
  \item(Try to factor out 17) $ = 2^6 \times 3^4  \times 5^3 \times 17^2$
\end{itemize}

\end{frame}

\begin{frame}{Prime Factors and Square Root}

\begin{itemize}
  \item \textbf{Lemma.} Let $n$ be a composite integer. Then $n$ has a prime factor less than or equal to $\sqrt{n}$.
  \item \textbf{proof.} Since $n$ is composite, $n=ab$ with $1<a\leq b < n$.
  \item $a^2 \leq ab = n$ so $a \leq \sqrt{n}$.
  \item Let $p$ be a prime factor of $a$.
  \item Then $p$ is a prime factor of $n$ and $p\leq a \leq \sqrt{n}$. $\qed$.
  \item This lemma tells us that when using the method of trial division we only
  have to search through $\sqrt{n}$ to determine if $n$ is prime.
\end{itemize}
\end{frame}

\begin{frame}{Example}
\begin{itemize}
  \item \textbf{Example.} Is $149$ prime?
  \item $11^2 = 121 < 149 < 169 = 13^2$.
  \item So we only need to check primes through 11.
  \item By inspection, 149 is not divisible by 2,3 or 5.
  \item Try 7. No $7\ndivides 149$.
  \item Try 11. No $11\ndivides 149$.
  \item So 149 is prime.
\end{itemize}
\end{frame}

\begin{frame}{Primality test for large numbers}
\begin{itemize}
  \item Suppose you are given a 200 digit integer  and you want to determine if it is prime.
  \item How might you do it? How might a computer do it?
  \item Trial division: Try every prime number less than $\sqrt{n}$ to see if it divides $n$.
  \item How long will this take on a very fast computer?
  \item Let's play around with Colab to try factoring integers of different sizes.
\end{itemize}
\end{frame}

\begin{frame}{How Big is Googol?}
\begin{itemize}
  \item To get a sense of how long it would take to factor a 200 digit integer using trial division,
  let's muse about how ``big'' the number $10^{100}$ is.
  \item Question: Which is bigger: (a) $10^{100}$ or (b) The number of grains of sand on an average beach?
  \item Put these in order: (a) $10^{100}$ (b) Grains of sand on a beach
  (c) Atoms in the Earth.
  \item Put these in order: (a) $10^{100}$ (b) Grains of sand on a beach
  (c) Atoms in the Earth (d) Elementary particles in the universe.
  \item See \href{https://en.wikipedia.org/wiki/Elementary_particle\#Cosmic_abundance_of_elementary_particles}{Wikipedia article on Elementary Particles}
  \item Which is larger: $10^{100}$ or the number of seconds the universe has existed so far?
  \item Conclusion: Any algorithm that requires $10^{100}$ steps will not terminate soon enough to be useful. (understatement)
\end{itemize}
\end{frame}

\begin{frame}{Distribution of Primes}

\begin{itemize}
  \item Draw a number line and put a red dot wherever there is a prime number.
  \item What does the resulting picture look like? Some questions:
  \item Does the red go on forever or does it eventually stop?
  \item What happens to the \emph{density} of red as we move to the right?
  \item Does the density increase or decrease?
\end{itemize}

\end{frame}

\begin{frame}{Infinitely Many Primes}

\begin{itemize}
  \item \textbf{Theorem.} There are infinitely many primes.
  \item \textbf{proof} Suppose towards a contradiction that there were only finitely many primes.
  \item List all of the primes: $p_1, p_2,  \cdots p_n$.
  \item Let $a = p_1 \cdot p_2 \cdot \cdots \cdot p_n$ be the product of all of the primes.
  \item Let $b=a+1$.
  \item We know that every positive integer has a prime decomposition. In particular every positive
  integer is divisible by at least one prime. So $b$ is divisible by at least one prime.
  \item So there is some $i$ with $1\leq i\leq n$ such that $p_i \divides b$.
  \item But $p_i \divides a$ and $\gcd(a,b)=1$. This is a contradiction.
  \item So our assumption that there were finitely many primes must be wrong.
  \item So there are infinitely many primes. $\qed$
\end{itemize}

\end{frame}

\begin{frame}{Arbitrarily Long Gaps}

\begin{itemize}
  \item \textbf{Theorem.} Given any positive integer $k$ there exists $k$
  consecutive composite integers.
  \item \textbf{proof.} Consider $(k+1)! + 2, (k+1)! + 3, \cdots (k+1)! + k + 1$.
  \item $(k+1)! + 2$ is divisible by 2, so it is not prime.
  \item $(k+1)! + 3$ is divisible by 3, so it is not prime.
  \item etc. So we have exhibited $k$ consecutive composite integers. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Twin Prime Conjecture}

\begin{itemize}
  \item \textbf{Statement.} There are infinitely many integers $n$ such that
  both $n$ and $n+2$ are prime.
  \item Is this statement true or false?
  \item It is an open question. We don't know if it is true or false.
  \item We believe it is probably true.
\end{itemize}

\end{frame}


\begin{frame}{Distribution of Primes}

\begin{itemize}
  \item \textbf{Definition.} For $n$ a positive integer $\pi(n)$ is the number
  of primes less than $n$.
  \item The \emph{density} of primes less than $n$ is $\pi(n)/n$.
  \item \textbf{The Prime Number Theorem.} As $n$ grows large the density of
  primes approaches $1/\ln(n)$.
  \item https://primes.utm.edu/howmany.html\#3
\end{itemize}

\end{frame}


\end{document}
