% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 6 \\ Systems of Linear Equations over $\R$}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{A system with 4 equations and five unknowns}

Consider the following system of equations:

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

\begin{itemize}
\item There are 5 \emph{unknowns}: $x, y, z, u, v$.
\item There are 4 \emph{equations}.
\item The numbers 2, 8, 13, -13 constitute the \emph{right-hand-side}
\item The goal is to find values for the unknowns in order to make all of the
equations simultaneously true.
\item What does this have to do with what we have been learning?
\item Notice: LH sides are linear combinations of the unknowns.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The matrix of coefficients}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $A$ be the 4 x 5 \emph{matrix of coefficients}:

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The column vector of unknowns}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $\bx$ be the 5 x 1  \emph{column vector of unknowns}:

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The right-hand-side column vector}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $\bb$ be the 4 x 1  \emph{right-hand-side column vector}:

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Equivalent matrix equation}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Then the system says exactly the same thing as the matrix equation $A \bx = \bb$.

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Drop the original system}

Notice the matrix equation $A \bx = \bb$ contains exactly the same information
as the original system of equations. So we may as well drop the original
system and work only with the matrix equation.

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix Equations}

\begin{itemize}
\item The names of the unknown variables in the original system of equations
are irrelevant.
\item So we may as well drop the definition of $\bx$.
\item And instead phrase the problem as:
\item Find all vectors $x \in \R^5$ such that $A \bx = \bb$.
\item Solving this matrix equation is the same thing as solving the
original system of equations.

\end{itemize}

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix Equations}

\begin{itemize}
\item Notice following are equivalent:
\item (a) $\bx$ is a solution to the original system of linear equations.
\item (b) $A \bx = \bb$.
\item (c) $T_A(\bx) = \bb$.
\item (d) $\bx \in T_A^{-1}[\singleton{\bb}]$
\item Put another way, the following are equivalent:
\item (i) Solve the original system of linear equations
\item (ii) Solve the matrix equation $A \bx = \bb$.
\item (iii) Solve the equation $T_A(\bx) = \bb$.
\item (iv) Find the inverse image under $T_A$ of $\singleton{\bb}$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{What kinds of questions do we want to ask?}

\begin{itemize}
\item What kinds of questions might we want to ask about our original system of linear equations?
\item (1) Are there any solutions?
If there are no solutions the system is called \emph{inconsistent}. Otherwise it is called \emph{consistent}.
\item (2) Is there more than one solution? If so the system is called
\emph{underdetermined}.
\item (3) If there is a unique solution, find it.
\item (4) If there is more than one solution, describe the structure of the
solution set.
\item Each of these questions can be translated into corresponding questions
about the abstract problems.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Translation of the questions:}

\begin{itemize}
\item (1) Is the original system of equations consistent?
\item Are there any vectors $\bx\in\R^5$ such that $A \bx = \bb$?
\item Is $\bb\in\ran(T_A)$?
\item (2) Is the original system of equations underdetermined?
\item Is there more than one vector $\bx\in\R^5$ such at $A \bx = \bb$?
\item Is the size of $T_A^{-1}[\singleton{\bb}]$ greater than 1?
\item (3) If there is a unique solution find it.
\item Find the unique vector $\bx$ such that $A\bx = \bb$.
\item Find the unique element of $T_A^{-1}[\singleton{\bb}]$.
\item (4) Describe the structure of the solution set of the original system.
\item Describe the set of all vectors $\bx$ such that $A\bx = \bb$.
\item Describe the inverse image under $T_A$ of $\singleton{\bb}$.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}


