% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 7 \\ Systems of Linear Equations over $\R$}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{A system with 4 equations and five unknowns}

Consider the following system of equations:

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

\begin{itemize}
\item There are 5 \emph{unknowns}: $x, y, z, u, v$.
\item There are 4 \emph{equations}.
\item The numbers 2, 8, 13, -13 constitute the \emph{right-hand-side}
\item The goal is to find values for the unknowns in order to make all of the
equations simultaneously true.
\item What does this have to do with what we have been learning?
\item Notice: LH sides are linear combinations of the unknowns.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The matrix of coefficients}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $A$ be the $4 \times 5$ \emph{matrix of coefficients}:

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The column vector of unknowns}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $\bx$ be the $5 \times 1$  \emph{column vector of unknowns}:

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The right-hand-side column vector}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Let $\bb$ be the $4 \times 1$  \emph{right-hand-side column vector}:

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Equivalent matrix equation}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

Then the system says exactly the same thing as the matrix equation $A \bx = \bb$.

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Drop the original system}

Notice the matrix equation $A \bx = \bb$ contains exactly the same information
as the original system of equations. So we may as well drop the original
system and work only with the matrix equation.

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix Equations}

\begin{itemize}
\item The names of the unknown variables in the original system of equations
are irrelevant.
\item So we may as well drop the definition of $\bx$.
\item And instead phrase the problem as:
\item Find all vectors $\bx \in \R^5$ such that $A \bx = \bb$.
\item Solving this matrix equation is the same thing as solving the
original system of equations.

\end{itemize}

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix Equations}

\begin{itemize}
\item Notice following are equivalent:
\item (a) $\bx$ is a solution to the original system of linear equations.
\item (b) $A \bx = \bb$.
\item (c) $T_A(\bx) = \bb$.
\item (d) $\bx \in T_A^{-1}[\singleton{\bb}]$
\item Put another way, the following are equivalent:
\item (i) Solve the original system of linear equations
\item (ii) Solve the matrix equation $A \bx = \bb$.
\item (iii) Solve the equation $T_A(\bx) = \bb$.
\item (iv) Find the inverse image under $T_A$ of $\singleton{\bb}$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{What kinds of questions do we want to ask?}

\begin{itemize}
\item What kinds of questions might we want to ask about our original system of linear equations?
\item (1) Are there any solutions?
If there are no solutions the system is called \emph{inconsistent}. Otherwise it is called \emph{consistent}.
\item (2) Is there more than one solution? If so the system is called
\emph{underdetermined}.
\item (3) If there is a unique solution, find it.
\item (4) If there is more than one solution, describe the structure of the
solution set.
\item Each of these questions can be translated into corresponding questions
about the abstract problems.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Translation of the questions:}

\begin{itemize}
\item (1) Is the original system of equations consistent?
\item Are there any vectors $\bx\in\R^5$ such that $A \bx = \bb$?
\item Is $\bb\in\ran(T_A)$?
\item (2) Is the original system of equations underdetermined?
\item Is there more than one vector $\bx\in\R^5$ such at $A \bx = \bb$?
\item Is the size of $T_A^{-1}[\singleton{\bb}]$ greater than 1?
\item (3) If there is a unique solution find it.
\item Find the unique vector $\bx$ such that $A\bx = \bb$.
\item Find the unique element of $T_A^{-1}[\singleton{\bb}]$.
\item (4) Describe the structure of the solution set of the original system.
\item Describe the set of all vectors $\bx$ such that $A\bx = \bb$.
\item Describe the inverse image under $T_A$ of $\singleton{\bb}$.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{General system of linear equations}

A general linear system of $m$ equations in $n$ unknowns has the form:

\begin{align*}
 a_{11}x_1 &+ a_{12}x_2 + &\cdots + a_{1n} x_n &= b_1\\
 a_{21}x_1 &+ a_{22}x_2 + &\cdots + a_{2n} x_n &= b_2\\
 \quad &{ \vdots \quad} \quad &{ \vdots \quad} \quad  &{ \vdots \quad} \quad\\
 a_{m1}x_1 &+ a_{m2}x_2 + &\cdots + a_{1mn} x_n &= b_m
\end{align*}


\end{frame}

\begin{frame}{Associated Matrix Equation}

Associated to a general linear system of equations is the matrix of coefficients
$A$ and the right-hand-side vector $\bb$. The
matrix equation $A\bx = \bb$ is equivalent to the original system.

$$
\begin{matrix}
A =
\begin{pmatrix}
a_{11} & a_{12} & \cdots & a_{1n} \\
a_{21} & a_{22} & \cdots & a_{2n} \\
\vdots & \vdots & \ddots & \vdots \\
a_{m1} & a_{m2} & \cdots & a_{mn}
\end{pmatrix}
&
\bb =
\begin{pmatrix}
b_1 \\ b_2 \\ \vdots \\ b_n
\end{pmatrix}
\end{matrix}
$$


\end{frame}

\begin{frame}{Start with a matrix equation}

Conversely, given an arbitrary $m \times n$ matrix $A$ and an arbitrary
$n \times 1$ column vector $\bb$ we can consider the matrix equation
$A\bx = \bb$. The system of linear equations displayed two slides back
is equivalent to this matrix equation.

$$
\begin{matrix}
A =
\begin{pmatrix}
a_{11} & a_{12} & \cdots & a_{1n} \\
a_{21} & a_{22} & \cdots & a_{2n} \\
\vdots & \vdots & \ddots & \vdots \\
a_{m1} & a_{m2} & \cdots & a_{mn}
\end{pmatrix}
&
\bb =
\begin{pmatrix}
b_1 \\ b_2 \\ \vdots \\ b_n
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Start with a linear transformation}

\begin{itemize}
\item Let $T$ be a linear transformation from $\R^n$ to $\R^m$.
\item Let $\bb\in \R^m$.
\item Then we may consider the problem of finding all vectors $\bx\in\R^n$
such that $T(\bx) = \bb$.
\item In other words the problem of finding $T^{-1}[\singleton{\bb}]$.
\item Let $A$ be the unique $m \times n$ matrix such that $T = T_A$.
\item Then this linear system on $\R^n$ is equivalent to the problem of
solving the matrix equation $A \bx = \bb$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Three equivalent things}

\begin{itemize}
\item In summary, the following three things are equivalent.
\item They are different ways of thinking of the same thing.
\item (1) A system of $m$ linear equations in $n$ unknowns.
\item (2) A matrix equation of the form $A\bx=\bb$ where $A$ is
$m\times n$ and $\bb$ is $m\times 1$.
\item (3) The problem of finding $T^{-1}[\singleton{\bb}]$ where
$T:\R^n\map\R^m$ is linear and $\bb\in\R^m$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Square Matrices}

\begin{itemize}
\item Under the equivalence described above, the following things correspond:
\item (1) The system has the same number of equations as unknowns
\item (2) The matrix $A$ is square.
\item (2) $n=m$ so that $T$ maps $\R^n$ to itself.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Inconsistent systems}

\begin{itemize}
\item Under the equivalence described above, the following things correspond:
\item (1) The system of linear equations is inconsistent.
\item (2) There is no vector $\bx \in \R^n$ such that $A \bx = \bb$.
\item (2) $\bb \notin \ran T$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Underdetermined systems}

\begin{itemize}
\item Under the equivalence described above, the following things correspond:
\item (1) The system of linear equations is underdetermined.
\item (2) There is more than one vector $\bx \in \R^n$ such that $A \bx = \bb$.
\item (2) $T^{-1}[\singleton{\bb}]$ has size greater than 1.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Structure of the solution set}

\begin{itemize}
\item Under the equivalence described above, the following things correspond:
\item (1) The structure of the set of all solutions to the system of linear equations.
\item (2) The structure of the set of all vectors $\bx \in \R^n$ such that $A \bx = \bb$.
\item (2) The structure of $T^{-1}[\singleton{\bb}]$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Unique solutions}

\begin{itemize}
\item Under the equivalence described above, the following things correspond:
\item (1) The system of linear equations has a unique solution.
\item (2) There is exactly one vector $\bx \in \R^n$ such that $A \bx = \bb$.
\item (2) $T^{-1}[\singleton{\bb}]$ has size exactly 1.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Study the abstract in order to inderstand the concrete}

\begin{itemize}
\item We are interested in finding ways to determine when a system of linear
equations:
\item (a) has a unique solution, or
\item (b) is inconsistent, or
\item (c) is underdetermined. And if it is underdetermined understanding
\item (d) the structure of the solution set.
\item In order to help us understand these things we will go on in the
following lessons to
study the corresponding issues about linear transformations on $\R^n$.
\end{itemize}

\end{frame}


\end{document}


