% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 14 \\ $\Z_n^*$}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$\Z^*_n$}

\begin{itemize}
  \item Let $n>1$ be an integer.
  \item $\Z^*_n$ is pronounced $\Z$-mod-n-star.
  \item It is also written $\Z^*/(n)$.
  \item It is a subset of $\Z_n$. $\Z_n^* \subset \Z_n$.
  \item $\Z^*_n$ is the set of units of $\Z_n$.
  \item Find $\Z_6^*$.
  \item $\Z_6^*=\singleton{[1],[5]}$.
  \item Find $\Z_7^*$.
  \item $\Z_7^*=\singleton{[1],[2],[3],[4],[5],[6]}$.
  \item Find $\Z_8*$.
  \item $\Z_8* = \singleton{[1],[3],[5],[7]}$.
\end{itemize}

\end{frame}

\begin{frame}{Properties of $\Z^*_n$}

\begin{itemize}
  \item If $[a]\in\Z_n$ then $[a]\in\Z_n^*$ iff ...
  \item iff $\gcd(a,n)=1$
  \item iff $[a]$ has a multiplicative inverse ($[a]$ is a unit)
  \item iff $a\not= 0$ and $[a]$ is not a zero-divisor.
  \item If $p$ is prime, then $\Z_p^*=\setof{[a]\in\Z_p}{a\not=0}$.
  \item $\Z_5^*=\singleton{[1],[2],[3],[4]}$
  \item $\Z_7^*=\singleton{[1],[2],[3],[4],[5],[6]}$.
  \item $\Z_{11}^*=\singleton{[1],[2],[3],[4],[5],[6],[7],[8],[9],[10]}$.
  \item If $n$ is not prime then this is not the case.
  \item $\Z_{10}^*=\singleton{[1],[3],[7],[9]}$.
  \item $\Z_n^*$ never includes $[0]$.
\end{itemize}

\end{frame}

\begin{frame}{Multiplication in $\Z^*_n$}

\begin{itemize}
  \item $\Z_n^*$ is a self-contained arithmetic universe inside of a self-contained
  arithmetic universe.
  \item $\Z_n^*$ is closed under the multiplication of $\Z_n$.
  \item If $[a],[b]\in\Z_n^*$ and $[a]\cdot[b]$ is computed in $\Z_n$, then
  $[a]\cdot[b]\in\Z_n^*$.
  \item For example work in $\Z_8$...
  \item $\Z_8* = \singleton{[1],[3],[5],[7]}$.
  \item $[3]\cdot [3] = [1]$.
  \item $[3]\cdot [5] = [7]$.
  \item $[3] \cdot [7] = [5]$.
\end{itemize}

\end{frame}

\begin{frame}{$\Z^*_n$ is closed under multiplication}

\begin{itemize}
  \item If $[a],[b] in\Z_n^*$ then so is $[a][b]$.
  \item \textbf{Proof} Suppose $[a],[b]\in\Z_n^*$.
  \item So $[a]^{-1}$ and $[b]^{-1}$ exist.
  \item To see $[a][b]\in\Z_n^*$ we must see that $([a][b])^{-1}$ exists.
  \item Notice that $([a][b])\cdot ([a]^{-1}[b]^{-1}) = [1]$.
  \item This means that $([a][b])^{-1} = [a]^{-1}[b]^{-1}$.
  \item So $([a][b])^{-1}$ exists and $[a][b]\in\Z_n^*$. $\qed$.
  \item \textbf{Proof 2}. Suppose $a$ and $b$ are relatively prime to $n$.
  \item Then $a$ and $n$ don't share any prime factors and neither do $b$ and $n$.
  \item But then $ab$ and $n$ don't share any prime factors.
  \item So $ab$ and $n$ are relatively prime. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Cancellation in $\Z^*_n$}

\begin{itemize}
  \item If $[c]\in\Z_n^*$, then so is $[c]^{-1}$.
  \item For example work in  $\Z_{10}^* = \singleton{[1],[3],[7],[9]}$.
  \item $[1]^{-1} = [1], [3]^{-1}=[7], [7]^{-1} = [3], [9]^{-1} = [9]$.
  \item This is always true because if $[c]$ is a unit then so is $[c]^{-1}$.
  \item Suppose $[a],[b],[c] \in \Z_n^*$
  \item and $[a][c] = [b][c]$.
  \item Then $[a] = [b]$.
  \item \textbf{Proof.} Multiply both sides by $[c]^{-1}$.
  \item For example work in  $\Z_{10}^* = \singleton{[1],[3],[7],[9]}$.
  \item Suppose $[3][x] = [3][y]$.
  \item Then $[x] = [y]$.
\end{itemize}

\end{frame}

\begin{frame}{$\Z^*_n$ is a commutative group under multiplication}

\begin{itemize}
  \item Multiplication is commutative and associative.
  \item $[1]$ is a multiplicative identity.
  \item Every element has a multiplicative inverse.
  \item Note that $\Z^*_n$ is not closed under addition.
  \item Example: Work in $\Z_{10}$...
  \item $[1]$ and $[3]$ are units but $[4]$ is not.
  \item Note that $\Z^*_n$ is closed under additive inverse.
  \item If $[a]\in\Z^*n$ then $-[a]\in\Z^*_n$.
  \item Work in $\Z_{10}$...
  \item $[3]\in\Z^*_{10}$. $-[3]=[7]\in\Z^*_{10}$.
  \item This is because $-[1]$ is a unit.
\end{itemize}

\end{frame}


\begin{frame}{The Orbit of an element of $\Z_n^*$}

\begin{itemize}
  \item Let $[a]\in\Z^*_n$. The \emph{orbit} of $[a]$ is
  the set of powers of $[a]$.
  \item Careful not to confuse powers with multiples.
  \item Example: Work in $\Z^*_7$...
  \item Find the orbit of $[2]$.
  \item $[2]^1 = 2$, $[2]^2 = 4$, $[2]^3=[1]$
  \item $[2]^4 = 2$, $[2]^5 = 4$, $[2]^6=[1], \cdots$
  \item So the orbit of $[2]$ in $\Z^*_7$ is: $\singleton{[1],[2],[4]}$.
  \item Find the orbit of $[3]$ in $\Z^*_7$.
  \item $[3]^1=[3]$, $[3]^2=[2]$, $[3]^3=[6]$, $[3]^4=[4]$
  \item $[3]^5=[5]$, $[3]^6=[1]$,
  \item $[3]^7=[3]$, $[3]^8=[2]$, $[3]^9=[6]$, $[3]^{10}=[4],\cdots$
  \item So the orbit of $[3]$ in $\Z^*_7$ is $\singleton{[1],[2],[3],[4],[5],[6]}$.
  \item i.e. the orbit of $[3]$ in $\Z^*_7$ is all of $\Z^*_7$.
\end{itemize}

\end{frame}

\begin{frame}{The Order of $[a]$ in $\Z_n^*$}

\begin{itemize}
  \item Let $[a]\in\Z^*_n$. The \emph{order} of $[a]$ is
  the least integer $k>0$ such that $[a]^k=[1]$.
  \item Example: Work in $\Z^*_7$...
  \item Find the order of $[2]$.
  \item $[2]^1 = 2$, $[2]^2 = 4$, $[2]^3=[1]$
  \item So the order of $[2]$ in $\Z^*_7$ is 3.
  \item Find the order of $[3]$ in $\Z^*_7$.
  \item $[3]^1=[3]$, $[3]^2=[2]$, $[3]^3=[6]$, $[3]^4=[4]$
  \item $[3]^5=[5]$, $[3]^6=[1]$.
  \item So the order of $[3]$ in $\Z^*_7$ is 6.
\end{itemize}

\end{frame}

\begin{frame}{The Order of $[a]$ exists}

\begin{itemize}
  \item \textbf{Theorem.} Let $[a]\in\Z^*_n$.
  \item Then there is and integer $k>0$ such that $[a]^k=[1]$.
  \item \textbf{Proof.} $\Z^*_n$ is finite. So there must be
  \item two integers $s$ and $t$ such that $0<s<t$ and $[a]^s=[a]^t$.
  \item Let $k=t-s$.
  \item Then $[a]^s = [a]^t = [a]^s [a]^k$
  \item Cancel $[a]^s$.
  \item It follows that $[1] = [a]^k$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Square Roots}

\begin{itemize}
  \item Let $[a]\in\Z_n^*$.
  \item $[b]\in\Z^*_n$ is a square root of $[a]$ iff $[b]^2=[a]$.
  \item Example: Work in $\Z_7^*$.
  \item $[1]^2=[1], [2]^2=[4], [3]^2 = [2], [4]^2=[2], [5]^2=[4], [6]^2 = [1]$.
  \item Find all of the square roots of $[4]$ in $\Z_7^*$.
  \item Answer $[2]$ and $[5]$. Note that $[5] = -[2]$.
  \item Find all of the square roots of $[2]$ in $\Z_7^*$.
  \item Answer $[3]$ and $[4]$. Note that $[4] = -[3]$.
  \item In general if $[b]^2=[a]$ then $(-[b])^2=[a]$.
  \item Find all of the square roots of $[1]$.
  \item Answer $[1]$ and $-[1]=[6]$.
  \item Find all of the square-roots of $[3]$.
  \item Answer, there are none.
\end{itemize}

\end{frame}

\begin{frame}{More Square Roots}

\begin{itemize}
  \item Can there be a square root of $-[1]$?
  \item In $\Z_5$, $-[1] = [4]$, so $[2]$ is a square root of $-[1]$.
  \item In $\Z_7$, $-[1] = [6]$ and there is no square root of $-[1]$.
  \item Square roots of $[1]$:
  \item In any $\Z_n^*$, $[1]$ and $-[1]$ are always square roots of $[1]$.
  \item Can there be more than those two?
  \item Work in $Z_8^* = \singleton{1,3,5,7}$.
  \item $[1]^2=[1],[3]^2=[1],[5]^2=[1],[7]^2=1$.
  \item So in $Z_8^*$, all four elements are square roots of $[1]$.
\end{itemize}

\end{frame}




\end{document}
