\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 12 \\ Bases and Dimension}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Spring 2019 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.}
\textbf{(a)} Yes, a basis.
\textbf{(b)} No, linearly dependent.
\textbf{(c)} Yes, a basis.
\textbf{(d)} No, linearly dependent.
\textbf{(e)} No, linearly dependent.
\bigskip

\textbf{Problem 2.}
\textbf{(a)} No, doesn't span
\textbf{(d)} No, linearly dependent.

\bigskip


\textbf{Problem 3.} $\singleton{\bv_1,\bv_2}$ is linearly independent.
But $\bv_3 = \bv_1+\bv_2$ so $\singleton{\bv_1,\bv_2,\bv_3}$ is linearly
dependent. So we discard $\bv_3$. $\singleton{\bv_1,\bv_2,\bv_4}$ is linearly
independent. We can see this because the matrix
$\left(\bv_4 \vert \bv_2 \vert \bv_1 \right)$ is upper-triangular with
nonzeros on the diagonal. Thus our maximal linearly independent set is
$\singleton{\bv_1,\bv_2,\bv_4}$ and this is a basis for $\R^3$.

\bigskip

\textbf{Problem 4.}

\textbf{(a)} No $\bv_1,\bv_2,\bv_3$ do not span $\R^3$. We can see this because
all three vectors have a zero in the second position and so every linear combination
of these three vectors do also. So the span of these three vectors does not include
every vector in $\R^3$.

\bigskip



\textbf{(b)} No $\bv_1,\bv_2,\bv_3$ do not span $\R^3$. That's what we saw in part (a) above.

\bigskip


\textbf{(c)} The dimension of the subspace of $\R^3$ spanned by $\bv_1,\bv_2,\bv_3$ is 2.
We know it is $<3$ because the three vectors do not span $\R^3$. But we can see that,
for example $\bv_2$ and $\bv_3$ are linearly independent so the dimension of the
span is at least 2. So it must be exactly 2.

\bigskip


\textbf{(d)} For example $\singleton{\bv_2,\bv_3}$ is a maximal linearly independent subset.
(Actually any set of two of the three vectors works.)

\bigskip

\textbf{(e)} Let
$\bv =
\begin{pmatrix}
0 \\ 1 \\0
\end{pmatrix}.
$
Then $\bv_2,\bv_3,\bv$ is a basis fro $\R^3$. We know this, for example, because
the matrix $\left(\bv_2 \vert \bv \vert \bv_3 \right)$ is upper triangular with
nonzeros on the diagonal.

\bigskip

\newpage

\textbf{Problem 5.}
\textbf{(a)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=1$.
$\singleton{\bv_1}$ is a maximal linearly independent set. $\bv_2,\bv_3,\bv_4$
are all multiples of $\bv_1$.


\bigskip

\textbf{(b)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=3$.
$\singleton{\bv_1,\bv_2,\bv_4}$ is a maximal linearly independent set. To
see this first note that $\singleton{\bv_1,\bv_4}$ is linearly independent
and then note that $\bv_2\notin\spn(\bv_1,\bv_4)$ because $\bv_1$ and $\bv_4$
have a zero in the third component but $\bv_2$ does not. $\bv_3$ is a scalar
multiple of $\bv_1$.

\bigskip

\textbf{(c)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=2$.
$\singleton{\bv_1,\bv_4}$ is a maximal linearly independent set. (Actually
any set of two of the vectors works here.) To
see this first note that $\singleton{\bv_1,\bv_4}$ is linearly independent
and then note that $\bv_2,\bv_3\in\spn(\bv_1,\bv_4)$.

\bigskip

\textbf{Problem 6.} For which of the following matrices $A$ is it true that $T_A$ is one-to-one?


\textbf{(a)} $T_A$ is one-to-one. The columns of $A$ are linearly independent.
To see this note that we could add a fourth column to the matrix, for example,
$
\begin{pmatrix}
0  \\
0  \\
0  \\
1  \\
\end{pmatrix}.
$
and then the matrix would be upper-triangular with nonzeros on the diagonal.


\bigskip

\textbf{(b)} $T_A$ is not one-to-one. The columns of $A$ are not linearly
independent. To see this note that the matrix is $4\times 5$ so that the
columns represent 5 vectors in $\R^4$.

\bigskip

\textbf{(c)} $T_A$ is not one-to-one. The columns of $A$ are not linearly
independent. The second and third columns are scalar multiples of the first.

\bigskip

\textbf{Problem 7.} The problem is equivalent to solving the following
system of linear equations

$$
\begin{pmatrix}
4 & 1 & 2 \\
0 & 2 & 0 \\
0 & 0 & -1 \\
\end{pmatrix}
\begin{pmatrix}
x \\
y \\
z \\
\end{pmatrix}
=
\begin{pmatrix}
-5 \\
6 \\
2 \\
\end{pmatrix}
.
$$

Since the matrix is upper-triangular with nonzeros on the diagonal we can
solve the system using back-substitution.

\begin{itemize}
\item $-1z = 2 \quad\implies\quad z = -2$
\item $2y = 6 \quad\implies\quad y = 3$
\item $4x + y + 2z = -5 \quad\implies\quad 4x + 3 -4 = -5 \quad\implies\quad x=-1$
\end{itemize}
So
$$
\begin{pmatrix}
4 & 1 & 2 \\
0 & 2 & 0 \\
0 & 0 & -1 \\
\end{pmatrix}
\begin{pmatrix}
-1 \\
3 \\
-2 \\
\end{pmatrix}
=
\begin{pmatrix}
-5 \\
6 \\
2 \\
\end{pmatrix}
$$
so
$$
\begin{pmatrix}
-5 \\
6 \\
2 \\
\end{pmatrix}
=
-1
\begin{pmatrix}
4 \\
0\\
0\\
\end{pmatrix}
+
3
\begin{pmatrix}
1 \\
2\\
0\\
\end{pmatrix}
-2
\begin{pmatrix}
2  \\
0  \\
-1 \\
\end{pmatrix}
$$
so $\bb = -\bv_1 + 3\bv_2 -2\bv_3$.
So the coordinates of $\bb$ in the basis $\bv_1,\bv_2,\bv_3$ are
$-1,3,-2$.



\end{document}
