% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 4 \\ Floor and Mod}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{The Floor Function}

\begin{itemize}
  \item \textbf{Definition.} Let $x$ be any real number. Then $\floor{x}$ is the greatest integer less than or equal to $x$.
  \item $\floor{x}$ is pronounced ``floor of $x$''.
  \item Our textbook uses the notation $[x]$ instead of $\floor{x}$. You can use either notation.
  \item Example: What is $\floor{3.14159}$?
  \item $\floor{3.14159} = 3$.
  \item Example: What is $\floor{-3.14159}$?
  \item $\floor{-3.14159} = -4$.
  \item Example: What is $\floor{411}$?
  \item $\floor{411} = 411$.
\end{itemize}

\end{frame}

\begin{frame}{Quotient and Remainder}

\begin{itemize}
  \item Let $a$ and $b$ be integers with $b>0$. Let $q$ and $r$ be the quotient
  and remainder when $a$ is the dividend and $b$ is the divisor.
  \item So $a = bq + r$ and $0\leq r < b$.
  \item Then $q = \floor{\frac{a}{b}}$.
  \item Example: Let $a = 17$ and $b=5$.
  \item $17 = 5 \times 3 + 2$ so $q=3$.
  \item $3 = \floor{\frac{17}{5}}$.
  \item $\frac{17}{5} = 3.4$.
\end{itemize}

\end{frame}

\begin{frame}{More Quotient and Remainder Examples}

\begin{itemize}
  \item Find the quotient and remainder when the divided is $a=987654321$
  and the divisor is $b=123456789$.
  \item  987654321/123456789 = 8.0000000729
  \item $\floor{987654321/123456789} = 8$.
  \item So $q=8$.
  \item How do we find $r$?
  \item $987654321 = 123456789 \times 8 + r$.
  \item $123456789 \times 8 = 987654312$.
  \item So $r=987654321 - 987654312 = 9$.
\end{itemize}

\end{frame}

\begin{frame}{More Quotient and Remainder Examples}

\begin{itemize}
  \item Find the quotient and remainder when the divided is $a=-17$
  and the divisor is $b=5$.
  \item $\frac{-17}{5} = -3.4$.
  \item $\floor{-3.4} = -4$.
  \item So $q=-4$.
  \item How do we find $r$?
  \item $-17 = 5 \times (-4) + r$.
  \item $r=3$.
\end{itemize}

\end{frame}

\begin{frame}{The mod function}

\begin{itemize}
 \item \textbf{Definition.} Let $a$ and $b$ be integers with $b>0$.
 Then $a\bmod b$ is the remainder when $a$ is divided by $b$
 (as given by the quotient and remainder theorem.)
 \item Example: Find $17 \bmod 5$.
 \item $17 = 5 \times 3 + 2$. So $17 \bmod 5 = 2$
 \item Example: Find $987654321 \bmod 123456789$.
 \item $987654321 = 123456789 \times 8 + 9$.
 \item So $987654321 \bmod 123456789 = 9$.
 \item Example: Find $-17 \bmod 5$.
 \item $-17 = 5 \times -4 + 3$. So $-17 \bmod 5 = 3$
\end{itemize}

\end{frame}

\begin{frame}{Modulus}

\begin{itemize}
 \item It is common to fix some integer $n>1$ and then consider $x\bmod n$
 for various integers $x$.
 \item When we do this we call $n$ the \emph{modulus}.
 \item Example: Let's use $n=3$ as the modulus.
 \item Compute $x\bmod 3$ for $x = 0,1,2,3,4,5,-1,-2,-3,-4,-5$.
\end{itemize}

\end{frame}

\begin{frame}{$x\bmod n$}

\begin{itemize}
 \item Fix an integer $n>1$ to use as a modulus.
 \item Let $x$ be any integer. What are the possible values of $x\bmod n$?
 \item Answer: $0,1,2\cdots n-1$.
 \item What are the possible values of $x\bmod 3$?
 \item Answer: 0,1,2.
\end{itemize}

\end{frame}

\begin{frame}{Generalization of Even}

\begin{itemize}
 \item Let $x$ be any integer. What is $x\bmod 2$?
 \item Answer: $x\bmod 2 = 0$ if $x$ is even and $x\bmod 2 = 1$ if $x$ is odd.
 \item Let $n>1$ be an integer.
 \item Quotient and Remainder theorem gives $x=kn+r$.
 \item where $r=x\bmod n$.
 \item Example with $n=3$. Every integer can be written in exactly one of the following forms:
 \item $3k$
 \item $3k+1$
 \item $3k+2$.
\end{itemize}

\end{frame}

\begin{frame}{$x=kn + r$.}

\begin{itemize}
 \item Let $x,n$ be integers with $n>1$.
 \item Quotient and Remainder theorem gives $x=kn+r$.
 \item where $r=x\bmod n$.
 \item Example with $n=3$. Every integer can be written in exactly one of the following forms:
 \item $3k$
 \item $3k+1$
 \item $3k+2$.
 \item You can think of this as a generalization of odd and even.
\end{itemize}

\end{frame}

\begin{frame}{Example Exercise}

\begin{itemize}
  \item \textbf{Example:} Suppose I tell you the following:
  \item I have an integer $x$. I square $x$ to get $x^2$. Then
  I divide the result by 3 and get a remainder of 2.
  \item How do you know I am not telling the truth?
  \item Solution: $x$ must be of the form $3k$, $3k+1$ or $3k+2$.
  \item $(3k)^2=9k^2= 3(3k^2)$. So $(3k)^2 \bmod 3 = 0$.
  \item $(3k+1)^2 = 9k^2 + 6k + 1 = 3(3k^2 + 2k) + 1$.
  \item So $(3k+1)^2 \bmod 3 = 1$.
  \item $(3k+2)^2 = 9k^2 + 12k + 4 = 9k^2 + 12k + 3 + 1$.
  \item $=3(3k^2 + 4k + 1) + 1$. So $(3k+2)^2 \bmod 3 = 1$.
  \item Conclusion: For all integers $x$, $x^2\bmod 3 = 0,1$. $x^2\bmod 3 \neq 2$.
\end{itemize}

\end{frame}

\begin{frame}{Example Exercise}

\begin{itemize}
  \item \textbf{Example:} How many positive integers less than 1000 are divisible
  by 7?
  \item Solution: An integer is divisible by 7 if it is of the form $7k$.
  \item So how many positive integers of the form $7k$ are less than 1000?
  \item $7, 7\times 2, 7\times 3, 7\times 4, \cdots$.
  \item $1000/7 = 142.857142857$. So $\floor{\frac{100}{7}} = 142$.
  \item $7\times 142 = 994$. So $1000 = 7 \times 142 + 6$.
  \item So $7k < 1000$ for $k=1,2,\cdots 142$.
  \item So the answer is 142.
\end{itemize}

\end{frame}

\begin{frame}{Reading assignment}

\begin{itemize}
  \item More of \emph{Divisibility}, section 5.1 of Rosen. From the end
  of the proof of Theorem 1.10 in the middle of page 38
  through the top of page 39. Stop when you get to
  \emph{Greatest Common Divisors}.
\end{itemize}

\end{frame}

\end{document}



\end{document}
