\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 5 \\ Linear functions $\R^n$}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Fall 2018 \\ SFSU }
\date{}

\maketitle

\textbf{Problem 1.} Yes, if $f$ is linear then necessarily
$f(\mathbf{0}) = \mathbf{0}$. (Here the $\mathbf{0}$ on the left is the
zero vector from the domain and the $\mathbf{0}$ on the right is the zero vector
from the codomain.) To see this use the fact that $f$ preserves scalar multiplication
and set the scalar equal to 0. Let $\bv$ be any vector in the domain. Then
$$f(\mathbf{0}) = f(0 \bv) = 0 f(\bv) = \mathbf{0}$$

\smallskip

\textbf{Problem 2.} No $f$ is not linear. One way to see this is to use the
result of problem 1 and notice that $f(0) = 1 \not= 0$.

\smallskip

\textbf{Problem 3.}

\textbf{(a)} Yes the constantly zero function is linear. It preserves vector
addition because
$$f(\bv + \bw) = \mathbf{0} = \mathbf{0} + \mathbf{0} = f(\bv) + f(\bw).$$
It preserves scalar multiplication because
$$f(c \bv) = \mathbf{0} = c \mathbf{0} = c f(\bv).$$

\smallskip

\textbf{(b)} No this constant function is not linear. One way to see that is
to use the result of problem 1 and notice that
$$f(\mathbf{0}) = (1, 0)^{\textbf{T}} \not= \mathbf{0}.$$

\smallskip

\textbf{(c)} A constant function is linear if and only if the constant is zero.
We saw in (a) that the constantly zero function is zero. But as in (b), a constant
function for any other constant cannot be linear because it does not map
zero to zero.

\smallskip

\textbf{Problem 4.}


\textbf{(a)} Suppose that for all vectors $\bv, \bw$ and all scalars $c, d$,
$$f(c\bv + d\bw) = c f(\bv) + d f (\bw).$$

Then yes it does follow that $f$ is linear.

To see that $f$ preserves vector addition set $c=d=1$. To see that $f$
preserves scalar multiplication, set $d=0$.

\smallskip

\textbf{(b)} Suppose that $f$ is linear. Yes it does follow that $f$ preserves
linear combinations of length 2.
\begin{align*}
f(c\bv + d\bw) &= f(c\bv) + f(d\bw) \quad \text{ (because $f$ preserves vector addition.)} \\
               &= c f(\bv) + d f(\bw) \quad \text{ (because $f$ preserves scalar multiplication.)}
\end{align*}

\smallskip

for all vectors $\bv_1,\bv_2 \cdots \bv_k$, and all scalars
$c_1, c_2, \cdots c_k$,

$$f(c_1 \bv_1 + c_2 \bv_2 + \cdots + c_k \bv_k)
= c_1 f(\bv_1) + c_2 f(\bv_2) + \cdots + c_k f(\bv_k)$$


\textbf{(c)} Suppose that $f$ is linear. Yes it does follow that $f$ preserves
all linear combinations. We know that $f$ preserves linear combinations of
length 2 by part (b). To see that $f$ preserves longer linear combinations
we keep using the same argument repeatedly. For example to see that $f$
preserves linear combinations of length three notice that
\begin{align*}
&f(c_1 \bv_1 + c_2 \bv_2 + c_3 \bv3) \\
&= f\left( (c_1 \bv_1 + c_2 \bv_2 )+ c_3 \bv3 \right) \\
&= f(c_1 \bv_1 + c_2 \bv_2) + f(c_3 \bv3) \quad \text{ (because $f$ preserves vector addition.)} \\
&= f(c_1 \bv_1 + c_2 \bv_2) + c_3f(\bv3) \quad \text{ (because $f$ preserves scalar multiplication.)} \\
&= c_f(\bv_1) + c_2 f(\bv_2) + c_3f(\bv3) \quad \text{ (by part (b))} \\
\end{align*}

To prove this formally for all linear combinations we would use the technique
of proof by induction.

\smallskip

\textbf{Problem 5.} Suppose $f:\R^n\map\R^m$ and $g:\R^m\map\R^k$ are both linear.
Let $h=g\circ f$ so that $h:\R^n\map\R^k$. Then yes, $h$ is necessarily linear.

By part (a) of problem 4 above it suffices for us to show that $h$ preserves
linear combinations of length 2. Fix vectors $\bv, \bw \in \R^n$ and scalars
$c,d$. We must show that $h(c\bv +d \bw) = c \cdot h(\bv) + d \cdot h(\bw)$.

\begin{align*}
&h(c\bv +d \bw) \\
&= g\left( f(c\bv +d \bw) \right)  \quad \text{ (by definition of $h$)} \\
&= g\left( cf(\bv) +df(\bw) \right)  \quad \text{ (because $f$ is linear)} \\
&= c g\left( f(\bv) \right) +d g \left(f(\bw) \right)  \quad \text{ (because $g$ is linear)} \\
& = c \cdot h(\bv) + d \cdot h(\bw) \quad \text{ (by definition of $h$)}
\end{align*}


\end{document}
