\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for 16 \\ Bases and Dimension}
\author{Math 325, Linear Algebra \\ Section 6\\ Spring 2019 \\ SFSU }
\date{}

\maketitle

\textbf{Problem 1.} $\singleton{\bv_1,\bv_2}$ is linearly independent.
But $\bv_3 = \bv_1+\bv_2$ so $\singleton{\bv_1,\bv_2,\bv_3}$ is linearly
dependent. So we discard $\bv_3$. $\singleton{\bv_1,\bv_2,\bv_4}$ is linearly
independent. We can see this because the matrix
$\left(\bv_4 \vert \bv_2 \vert \bv_1 \right)$ is upper-triangular with
nonzeros on the diagonal. Thus our maximal linearly independent set is
$\singleton{\bv_1,\bv_2,\bv_4}$ and this is a basis for $\R^3$.

\bigskip

\textbf{Problem 2.}

\textbf{(a)} No $\bv_1,\bv_2,\bv_3$ do not span $\R^3$. We can see this because
all three vectors have a zero in the second position and so every linear combination
of these three vectors do also. So the span of these three vectors does not include
every vector in $\R^3$.

\bigskip



\textbf{(b)} No $\bv_1,\bv_2,\bv_3$ do not span $\R^3$. That's what we saw in part (a) above.

\bigskip


\textbf{(c)} The dimension of the subspace of $\R^3$ spanned by $\bv_1,\bv_2,\bv_3$ is 2.
We know it is $<3$ because the three vectors do not span $\R^3$. But we can see that,
for example $\bv_2$ and $\bv_3$ are linearly independent so the dimension of the
span is at least 2. So it must be exactly 2.

\bigskip


\textbf{(d)} For example $\singleton{\bv_2,\bv_3}$ is a maximal linearly independent subset.
(Actually any set of two of the three vectors works.)

\bigskip

\textbf{(e)} Let
$\bv =
\begin{pmatrix}
0 \\ 1 \\0
\end{pmatrix}.
$
Then $\bv_2,\bv_3,\bv$ is a basis fro $\R^3$. We know this, for example, because
the matrix $\left(\bv_2 \vert \bv \vert \bv_3 \right)$ is upper triangular with
nonzeros on the diagonal.

\bigskip

\newpage

\textbf{Problem 3.}
\textbf{(a)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=1$.
$\singleton{\bv_1}$ is a maximal linearly independent set. $\bv_2,\bv_3,\bv_4$
are all multiples of $\bv_1$.


\bigskip

\textbf{(b)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=3$.
$\singleton{\bv_1,\bv_2,\bv_4}$ is a maximal linearly independent set. To
see this first note that $\singleton{\bv_1,\bv_4}$ is linearly independent
and then note that $\bv_2\notin\spn(\bv_1,\bv_4)$ because $\bv_1$ and $\bv_4$
have a zero in the third component but $\bv_2$ does not. $\bv_3$ is a scalar
multiple of $\bv_1$.

\bigskip

\textbf{(c)} $\dim\left(\spn(\bv_1,\bv_2,\bv_3,\bv_4)\right)=2$.
$\singleton{\bv_1,\bv_4}$ is a maximal linearly independent set. (Actually
any set of two of the vectors works here.) To
see this first note that $\singleton{\bv_1,\bv_4}$ is linearly independent
and then note that $\bv_2,\bv_3\in\spn(\bv_1,\bv_4)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip

\textbf{Problem 4.} Do exercise 2.4.2 on page 103.

\textbf{(b)} The three vectors are linearly independent so they do form a basis.
The way you can tell that they are linearly independent is to form a matrix
out of them and use Gaussian elimination to determine that the rank is 3.

\bigskip

\textbf{(c)} The three vectors are not linearly independent so they do not form
a basis. The way you can tell that they are not linearly independent is to
form a matrix out of them and use Gaussian elimination to determine that the rank is 2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip

\textbf{Problem 5.} To solve 2.4.3, form a matrix out of the 4 vectors and
perform Gaussian elimination on it. This reveals that the rank of the matrix
is 2.

\bigskip

\textbf{(a)} The vectors do not span $\R^3$ since the dimension of their span is 2.

\textbf{(b)} The vectors are not linearly independent since  there are four of them in $\R^3$.

\textbf{(c)} They do not form a basis for $\R^3$ because they are neither linearly independent
nor span. It is not possible to choose a subset of them that form a basis since they do
not span.

\textbf{(d)} The dimension of the span is 2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip

\textbf{Problem 6.} To solve 2.4.4, form a matrix out of the 4 vectors and
perform Gaussian elimination on it. This reveals that the rank of the matrix
is 3 and that the first 3 columns are basic columns.

\bigskip

\textbf{(a)} The vectors do  span $\R^3$ since the dimension of their span is 3.

\textbf{(b)} The vectors are not linearly independent since  there are four of them in $\R^3$.

\textbf{(c)} They do not form a basis for $\R^3$ because they are not linearly independent.
It is possible to chose a subset of them that form a basis, namely the first three vectors
form a basis.

\textbf{(d)} The dimension of the span is 3.

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\end{document}
