\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Homework for Lesson 6 \\ Matrices and Linear Transformations $\R^n$}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Spring 2019 \\ SFSU }
\date{}

\maketitle

\textbf{Problem 1.} Suppose $T:\R^2\map\R^3$ is a linear transformation.
Suppose you know that $T$ sends the first standard basis vector of of $\R^2$
to the vector $(1, 2, 3)^{\text{T}}$ and the second standard basis vector
of $\R^2$ to the vector $(2,3,4)^{\text{T}}$. Write down the matrix of $T$.

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\textbf{Problem 2.}  Suppose $A$ is a matrix such that $T_A:\R^3\map\R^2$ and
$$
\begin{matrix}
T_A \begin{pmatrix} 0 \\ 0 \\ 1 \end{pmatrix} = \begin{pmatrix} 1 \\ 2 \end{pmatrix}\text{; and }
&
T_A \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} = \begin{pmatrix} 3 \\ 4 \end{pmatrix}\text{; and }
&
T_A \begin{pmatrix} 0 \\ 1 \\ 0 \end{pmatrix} = \begin{pmatrix} 5 \\ 6 \end{pmatrix}
\end{matrix}
$$

Give the matrix $A$.

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\textbf{Problem 3.}  Suppose $T:\R^2\map\R^2$ is a linear transformation
that swaps the two standard basis vectors. That is, it
maps $\be_1$ to $\be_2$ and maps $\be_2$ to $\be_1$. Write down the matrix
for $T$.

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\textbf{Problem 4.}  Let
$$
A = \begin{pmatrix}
9 & 8 & 7 \\
6 & 5 & 4 \\
3 & 2 & 1
\end{pmatrix}
$$

Where does $T_A$ send the third standard basis vector of $\R^3$?

\bigskip
\bigskip

\textbf{Problem 5.} Suppose $A$ and $B$ are both $2\times 3$ matrices and
$$
\begin{matrix}
A \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} = B \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix}\text{; and }
&
A \begin{pmatrix} 0 \\ 1 \\ 0 \end{pmatrix} = B \begin{pmatrix} 0 \\ 1 \\ 0 \end{pmatrix}\text{.}
\end{matrix}
$$
True or false? Then we must have that $A = B$. Explain.


\bigskip
\bigskip


\textbf{Problem 6.} Suppose $T:\R^2\map\R^2$ is a linear transformation and
$T(\be_1) = \begin{pmatrix} 1 \\ 2\end{pmatrix}$
and $T(\be_2) = \begin{pmatrix} 2 \\ 3\end{pmatrix}$. Then find
$T\begin{pmatrix} 3 \\ 4\end{pmatrix}$.


\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip


\textbf{Problem 7.} Let $\bv = (2, 0, 2)$ be a row vector. Where does
the covector for $\bv$ send $(4, 5, 6)^{\text{T}}$?


\bigskip

\textbf{Problem 8.} Let $T:\R^3\map\R^4$ be a linear transformation.

\smallskip

\textbf{a.} How many component covectors does $T$ have?

\bigskip

\textbf{b.} Let $f$ be the second component covector for $T$. Is $f$ linear?

\bigskip

\textbf{c.} Suppose you are told that
$$
\begin{matrix}
f \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} = 7 \text{; and }
&
f \begin{pmatrix} 0 \\ 1 \\ 0 \end{pmatrix} = 8 \text{; and }
&
f \begin{pmatrix} 0 \\ 0 \\ 1 \end{pmatrix} = 9 \text{.}
\end{matrix}
$$

Let $A$ be the matrix for $T$. What do you now know about $A$?

\bigskip
\bigskip
\bigskip

\textbf{d.} Suppose you are now told that all of the other component covectors
for $T$ are identically zero. Write down the matrix $A$.

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\textbf{Problem 9.}  Let
$$
A = \begin{pmatrix}
9 & 8 & 7 \\
6 & 5 & 4 \\
3 & 2 & 1
\end{pmatrix}
$$


\textbf{(a.)} Where does the second component covector of $T_A$ send the
third standard basis vector?

\bigskip

\textbf{(b.)} Where does the third component covector of $T_A$ send the
second standard basis vector?

\bigskip

\textbf{Problem 10} Suppose $T$ and $S$ are linear transformations with
the same component covectors. Does it follow that $T = S$? Explain.



\end{document}
