% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 19 \\ Euler's Theorem}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$\phi(n)$.}

\begin{itemize}
  \item \textbf{Definition.} Let $n>1$ be an integer.
  \item $\phi(n)$ is the number of positive integers less than $n$ and relatively prime to $n$.
  \item Equavlanetly, $\phi(n)$ is the number of elements of $\Z_n^*$.
  \item Example: Let $n=10$. The positive integers less than 10 and relatively prime to 10 are:
  \item 1,3,7,9
  \item The elements of $\Z_{10}^*$ are: $[1], [3], [7], [9]$.
  \item So $\phi(10) = 4$.
  \item Question: If $p$ is prime what is $\phi(p)$?
  \item Answer: $p-1$.
  \item If $n$ is composite then $1<\phi(n)<n-1$.
\end{itemize}

\end{frame}

\begin{frame}{Generalization of Fermat's Little Theorem}

\begin{itemize}
  \item \textbf{Fermat's Little Theorem.} Let $p$ be prime. Let $a$ be relatively prime to $p$.
  \item Then $a^{p-1} \equiv 1 \pmod p$.
  \item Equivalent in terms of $\phi$.
  \item \textbf{Theorem.} Let $p$ be prime. Let $a$ be relatively prime to $p$.
  \item Then $a^{\phi(p)} \equiv 1 \pmod p$.
  \item Equivalent in terms of $\Z_p^*$.
  \item \textbf{Theorem.} Let $p$ be prime. Let $[a]\in\Z_p^*$.
  \item Then $[a]^{\phi(p)} = [1]$.
  \item ``Any element of the group raised to the power of the cardinality of the group equals one.''
  \item Euler's Theorem: Same but drop requirement that modulus is prime.
\end{itemize}

\end{frame}

\begin{frame}{Euler's Theorem}

\begin{itemize}
  \item \textbf{Euler's Theorem.} Let $n>1$ be an integer. Let $a$ be relatively prime to $n$.
  \item Then $a^{\phi(n)} \equiv 1 \pmod n$.
  \item Equivalent in terms of $\Z_n^*$.
  \item \textbf{Theorem.} Let $n>1$ be an integer. Let $[a]\in\Z_n^*$.
  \item Then $[a]^{\phi(n)} = [1]$.
  \item ``Any element of the group raised to the power of the cardinality of the group equals one.''
  \item Let $[a]\in\Z_{10}^*$.
  \item Then $[a]^4 = [1]$.
  \item $[1]^4=[1], [3]^4=[1], [7]^4 = [1], [9]^4 = [1]$.
  \item Suppose $\gcd(a,10)=1$. Then $a^4 \equiv 1 \pmod{10}$.
  \item $21^4=194481 \equiv 1 \pmod {10}$
\end{itemize}

\end{frame}

\begin{frame}{Proof}

\begin{itemize}
  \item \textbf{proof.} Let $[a] \in \Z_n^*$. Consider the set of all multiples of $[a]$ in $\Z_n^*$.
  \item $[1][a], \cdots [i][a], \cdots [j][a], \cdots [n-1][a]$.
  \item \textbf{Claim.} The set of multiples of $[a]$ runs through all of the elements of $\Z_n^*$.
  \item For example, take $n=10$, $[a]=[3]$. The set of multiples of $[3]$ in $\Z_{10}^*$ is:
  \item $[1][3]=[3], [3][3]=[9], [7][3]=[1], [9][3] = [7]$.
  \item \textbf{proof of Claim.} Notice that there are $\phi(n)$ values. So we just need to show they are all different.
  \item Suppose $[i][a]=[j][a]$. Multiplying both sides by $[a]^{-1}$ we get $[i]=[j]$.
  \item In other words, if $[i]\not=[j]$, then $[i][a]\not=[j][a]$.
  \item So all $\phi(n)$ multiples of $[a]$ in $\Z_n^*$ are different.
  \item So the set of multiples of $[a]$ runs through all of the elements of $\Z_n^*$. $\diamondsuit$.
\end{itemize}

\end{frame}

\begin{frame}{Proof continued}

\begin{itemize}
  \item Now consider what happens when we multiply all of the elements of $\Z^*_n$ together.
  \item Let $[x] = [1]\cdots [i]\cdots [j]\cdots[n-1]$.
  \item By the claim
  \item $[x] = [1][a]\cdots[i][a]\cdots[j][a]\cdots[n-1][a]$
  \item  $=[a]^{\phi(n)}[1]\cdots[i]\cdots[j]\cdots[n-1]$
  \item $=[a]^{\phi(n)}[x]$.
  \item Canceling $[x]$, we get $[a]^{\phi(n)}=[1]$.
  \item $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Example}
\begin{itemize}
\item Example: Find the last digit in the decimal expansion of $7^{4000000002}$.
\item Solution: If $n$ is any positive integer, the last digit in the decimal
expansion of $n$ is $n\bmod 10$.
\item So we must find $7^{4000000002} \bmod 10$.
\item 7 is relatively prime to 10 so we can apply Euler's Theorem.
\item $\phi(10) = 4$, so $7^4 \bmod 10 = 1$.
\item So $7^{4000000000} \bmod 10 = 1$.
\item So $7^{4000000002} \equiv 7^2 \pmod{10}$.
\item $7^{4000000002} \bmod 10 = 7^2 \bmod 10 = 49 \bmod 10 = 9$.
\item So the last digit in the decimal expansion of $7^{4000000002}$ is 9.
\end{itemize}
\end{frame}

\begin{frame}{Complete Residue Systems}
\begin{itemize}
\item Recall the definition of a complete residue system.
\item  A \emph{complete system of residues}  or a \emph{complete residue system} mod $n$
\item is a set of integers that
chooses exactly one integer from each of the congruence classes mod $n$.
\item The following are all complete residue systems mod 10.
\item (a) $\singleton{0,1,2,3,4,5,6,7,8,9}$.
\item (b) $\singleton{-4,-3,-2,-1,0,1,2,3,4,5}$.
\item (c) $\singleton{0,11,22,33,44,55,66,77,88,99}$.
\item Equivalently, a complete residue system mod n is a set of n integers that
are pairwise incongruent mod n.
\item Equivalently, a complete residue system mod n is a set of integers $S$ such
that $\setof{[a \mod n]}{a\in S} = \Z_n$.
\end{itemize}
\end{frame}

\begin{frame}{Reduced Residue Systems}
\begin{itemize}
\item \textbf{Definition.} A \emph{reduced residue system} mod $n$
\item is a set of $\phi(n)$ integers that are pairwise incongruent mod n and
relatively prime to $n$.
\item Equivalently, a reduced residue system mod n is a set of integers $S$ such
that $\setof{[a \mod n]}{a\in S} = \Z_n^*$.
\item The following are all reduced residue systems mod 10.
\item (a) $\singleton{1,3,7,9}$.
\item (b) $\singleton{-3,-1,1,3}$.
\item (c) $\singleton{11,33,77,99}$.
\end{itemize}
\end{frame}




\end{document}
