% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 2 \\ Matrices}
\subtitle{Math 325, Linear Algebra \\ Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}



\begin{frame}{Matrices}

\begin{itemize}
\item A (real) \emph{matrix} is a rectangular array of real numbers.
\item We draw matrices, as in the example below, inside of elongated, curved
braces.
\item An $m\times n$ matrix (pronounced ``$m$-by-$n$'') has $m$-rows and $n$-columns.
\item If $A$ is an $m$-by-$n$ matrix, we say its \emph{size} is $m$-by-$n$.
\item We always list the number of rows before the number of columns.
\item The matrix $A$ below is a $4\times 5$ matrix because it has 4 rows
and 5 columns.
\item The size of $A$ is $4\times 5$.
\end{itemize}

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Entries of a matrix}

\begin{itemize}
\item We usually use capital Roman letters (lightface font), like $A$ or $M$, to name matrices.
\item The number in row $i$, column $j$ is called the $(i,j)$-th \emph{entry}.
\item If the matrix is called $A$ then by convention the symbol $a_{ij}$ refers
to the $(i,j)$-th entry of $A$.
We always list the row index first and the column index second.
\item For the matrix $A$ displayed below, the $(2,3)$-entry is 1 and the
$(3,2)$-entry is -3.
\end{itemize}

$$
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Square matrices}

\begin{itemize}
\item If the number of rows equals the number of columns then the matrix is
called a \emph{square} matrix.
\item The matrix $M$ below is a square $2\times 2$ matrix.
\item If a matrix is not square it is called \emph{rectangular.}
\item The $4\times 5$ matrix $A$ from the previous slide is rectangular.
\end{itemize}

$$
M =
\begin{pmatrix}
\frac{1}{3} & \frac{2}{3} \\
0 & -1
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Row vectors and column vectors}

\begin{itemize}
\item A $1\times n$ matrix, with $n>1$, is called a \emph{row vector}.
\item $\ba$ below has size $1\times 4$ so it is a row vector.
\item An $m\times 1$ matrix, with $m>1$, is called a \emph{column vector}.
\item $\bv$ below has size $3\times 1$ so it is a column vector.
\item We often use boldface lower-case letters for row and column vectors,
as below where we used $\ba$ and $\bv$,
instead of names like $A$ and $M$.
\end{itemize}

$$
\begin{matrix}
\ba =
\begin{pmatrix}
2, 2, 0, 1
\end{pmatrix}

&

\bv =
\begin{pmatrix}
1 \\
0 \\
-5
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Vector indexing}

\begin{itemize}
\item We sometimes refer to the entries of a vector as \emph{components}.
\item We use a slightly different notation for the components of a vector
than we do for the entries of a matrix.
\item For a row vector like $\ba$ we will often drop the row
index 1 and write only the column index.
\item So instead of writing $\ba_{1,3} = 0$,
\item we write $a_3 = 0$. Notice also the lightface font.
\item Similarly for column vectors like $\bv$ below we will often
drop the column index 1 and write only the row index.
\item So instead of writing $\bv_{2,1} = 0$,
\item we may write $v_2 = 0$.
\end{itemize}

$$
\begin{matrix}
\ba =
\begin{pmatrix}
2, 2, 0, 1
\end{pmatrix}

&

\bv =
\begin{pmatrix}
1 \\
0 \\
-5
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Transpose of a vector}

\begin{itemize}
\item If $\bv$ is an $n\times 1$ column vector, then $\bv^{\text{T}}$
(pronounced $\bv$-\emph{transpose}) is the corresponding $1\times n$ row vector.
\item If $\ba$ is a $1\times n$ row vector, then $\ba^{\text{T}}$ is the
corresponding $n\times 1$ column vector.
\item $$(1, 0, -5)^{\text{T}} = \begin{pmatrix} 1 \\ 0 \\ -5 \end{pmatrix}$$
\item $$\begin{pmatrix} 1 \\ 0 \\ -5 \end{pmatrix}^{\text{T}} =  (1, 0, -5)$$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Scalars}

\begin{itemize}
\item The things which can be entries in a matrix are called \emph{scalars}.
\item Scalars are just numbers but the word \emph{scalar} emphasizes that it
is a different type of thing than a matrix or a vector.
\item For a real matrix, the scalars are real numbers.
\item Later in the class we will consider matrices in which the scalars
are complex numbers.
\item A $1\times 1$ matrix is identified with a scalar.
\item For example the $1\times 1$ matrix $(42)$ is considered to
be just the scalar 42.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Matrix multiplication}

\begin{itemize}
\item The most important thing we do with matrices is multiply two of
them together.
\item Let $A$ be an $m\times n$ matrix. Let $B$ be a $n\times k$ matrix.
\item Then the product $A B$ is defined.
\item $A B$ is the $m\times k$ matrix defined as follow:
\item Let $C=A B$. Then for $1\leq i \leq m$, and $1\leq j \leq k$,
\item $$c_{ij} = \sum_{t=1}^n a_{it} b_{tj}$$
\item To understand this better, we will look at some special cases first.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Which products are defined?}

\begin{itemize}
\item Suppose $A$ and $B$ are matrices of the sizes given below. Say
whether the product $A B$ is defined and if so say what its size is.
\item $A$ is $3\times 4$ and $B$ is $3\times 5$.
\item  $A B$ is not defined.
\item $A$ is $3 \times 4$ and $B$ is $4 \times 5$.
\item $A B$ is $3 \times 5$.
\item $A$ is $5\times 1$ and $B$ is $6\times 1$.
\item $A B$ is not defined.
\item $A$ is $5\times 1$ and $B$ is $1\times 6$.
\item $A B$ is $5\times 6$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A row vector times a column vector}

\begin{itemize}
\item Let's look at the special case of the previous definition where $n=k=1$.
\item Let $\ba$ be a $1\times m$ row vector. Let $\bv$ be a $m\times 1$ column vector.
\item Then the product $\ba \bv$ is defined.
\item $\ba \bv$ is the $1\times 1$ matrix (i.e. the scalar) defined as follow:
\item $$\ba \bv  = \sum_{t=1}^m a_t v_t$$
\item Let's look at an example to try to understand this.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A row vector times a column vector: Example}

\begin{itemize}
\item Let $\ba$ be the $1\times 3$ row vector below.
\item Let $\bv$ be the $3\times 1$ column vector below.
\item $$\ba \bv  = \sum_{t=1}^3 a_t  v_t = $$
\item $ = 1\cdot 4 + (-1)\cdot (-2) + 2\cdot 5$
\item $=16$
\item The product of a row vector times a column vector
\emph{of the same length} is just the \emph{dot product} of the two vectors.
\end{itemize}

$$
\begin{matrix}

\ba =

\begin{pmatrix}
1 & -1 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
4 \\ -2 \\ 5
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector}

\begin{itemize}
\item Let's look at the special case of the  definition where $k=1$.
\item Let $A$ be an $m\times n$ matrix. Let $\bv$ be a $n\times 1$ column vector.
\item Then the product $A \bv$ is defined.
\item $A \bv$ is the $m\times 1$ column vector defined as follows:
\item Let $\bw=A \bv$. Then for $1\leq i \leq m$,
\item $$w_i = \sum_{t=1}^n a_{it} v_{t}$$
\item Let's look at an example to try to understand what this means.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector: Example}

\begin{itemize}
\item Let $A$ be a $4 \times 5$ matrix.
\item Let $\bv$ be a $5\times 1$ column vector.
\item Then $A \bv$ is a $4 \times 1$ column vector.
\item Let $\bw=A \bv$. Then for $1\leq i \leq 4$,
\item $$w_i = \sum_{t=1}^5 a_{it} v_{t}$$
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
2 \\ 0 \\ -1 \\ 3 \\ 0
\end{pmatrix}
\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{A matrix times a column vector: Example continued}

\begin{itemize}
\item $$w_i = \sum_{t=1}^5 a_{it} v_{t}$$
\item The $i$-th element of $A\bv$ is the $i$-th row of $A$ times $\bv$.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}

&

\bv =

\begin{pmatrix}
2 \\ 0 \\ -1\\ 3 \\ 0
\end{pmatrix}

&

A \bv =

\begin{pmatrix}
-3 \\ 15 \\ 10 \\ -23
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Product of two matrices}

\begin{itemize}
\item Let $A$ be the $3\times 2$ matrix below.
\item Let $B$ be the $2\times 3$ matrix below.
\item Then $A B$ is the $3 \times 3$ matrix $C$ defined by
\item $$c_{ij} = \sum_{t=1}^3 a_{it} b_{tj}$$
\item for $1\leq i \leq 3$, $1\leq j \leq 3$.
\item The $(i,j)$-th entry of $A B$ comes from multiplying row $i$ of $A$
by column $j$ of $B$.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 0 \\
0 & 2 \\
-2 & -3
\end{pmatrix}

&

B =

\begin{pmatrix}
4  & 1 & 0 \\
-1 & 0 & 2
\end{pmatrix}

&

A B =

\begin{pmatrix}
4 & 1 & 0 \\
-2 & 0 & 4 \\
-5 & -2 & -6
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Another way to think of the product of two matrices}

\begin{itemize}
\item You can also think about $A B$ is being $A$ times \emph{each column of} $B$.
\item Let $\bb_j$ be the $j$-th column of $B$.
\item So we can write $B = (\bb_1 \vert \bb_2 \vert \bb_3)$.

\item Then notice that
$$A B = (A\bb_1 \vert A\bb_2 \vert A\bb_3)$$

\item That is, the $j$-th column of $A B$ is gotten by multiplying $A$ times the $j$-th column of $B$.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 0 \\
0 & 2 \\
-2 & -3
\end{pmatrix}

&

B =

\begin{pmatrix}
4  & 1 & 0 \\
-1 & 0 & 2
\end{pmatrix}

&

A B =

\begin{pmatrix}
4 & 1 & 0 \\
-2 & 0 & 4 \\
-5 & -2 & -6
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Multiplication: Summary}

\begin{itemize}
\item Let $\bw$ be a row vector and $\bv$ a column vector of the same length.
\item Then $\bw \bv$ is just the dot product of the two vectors.
\item Let $A$ be an $m\times n$ matrix and $\bv$ an $n\times 1$ column vector.
\item Then $A \bv$ is the $m\times 1$ column vector whose $i$-th element is
the product of row $i$ of $A$ times $\bv$.
\item (Multiply a matrix by a column vector by taking the matrix one row at a time).
\item Let $A$ be an $m\times n$ matrix and $B$ an $n\times k$ matrix.
\item Then the $(i,j)$-th entry of $A B$ comes from multiplying row $i$ of $A$
by column $j$ of $B$.
\item Can also think of $A B$ as the matrix whose $j$-th column is the product
of $A$ times the $j$-th column of $B$.
\item (Multiply two matrices $A B$ by taking $B$ one column at a time.)
\end{itemize}


\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Addition}

\begin{itemize}
\item Matrix addition is defined on two matrices of the \emph{same size}.
\item Let $A$ and $B$ both be $m\times n$ matrices.
\item Then $A + B$ is the $m\times n$ matrix $C$ defined by
\item $c_{ij} = a_{ij} + b_{ij}$
\item We add two matrices by adding their components.
\end{itemize}


\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Addition: Example}

\begin{itemize}
\item Let $A$ and $B$ be the two $3\times 2$ matrices below.
\item Then $A+B$ is also a $3\times 2$ matrix, shown below.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 0 \\
0 & 2 \\
-2 & -3
\end{pmatrix}

&

B =

\begin{pmatrix}
3 & -5 \\
8 & -1 \\
6 & 4
\end{pmatrix}

&

A + B =

\begin{pmatrix}
4 & -5 \\
8 & 1 \\
4 & 1
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Scalar Multiplication}

\begin{itemize}
\item Let $A$ be an $m\times n$ matrix and $c$ a scalar.
\item Then $cA$ is defined to be the $m\times n$ matrix $B$ with
\item $b_{ij} = c a_{ij}$.
\item That is we just multiply each entry in $A$ by $c$.
\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Scalar Multiplication: Example}

\begin{itemize}
\item Let $A$ be the $3\times 2$ matrix below.
\item Let $c = -2$.
\item Then $cA$ is the $3\times 2$ matrix shown below.
\end{itemize}

$$
\begin{matrix}
A =

\begin{pmatrix}
1 & 0 \\
0 & 2 \\
-2 & -3
\end{pmatrix}

&

(-2) A =

\begin{pmatrix}
-2 & 0 \\
0 & -4 \\
4 &  6
\end{pmatrix}

\end{matrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Zero Matrix}

\begin{itemize}
\item $O_{m\times n}$ means the $m\times n$ matrix all of whose entries are zero.
\item
$$O_{2\times 3} =
\begin{pmatrix}
0 & 0 & 0 \\
0 & 0 & 0
\end{pmatrix}
$$
\item If the size of the matrix is clear from context sometimes we just
write $O$ for the zero matrix.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Main diagonal}

\begin{itemize}
\item If $A$ is a square $n\times n$ matrix, then the entries on the
\emph{main diagonal} of $A$ are the entries $a_{11}, a_{22}, \cdots a_{nn}$
whose row and column are equal.
\item Let
$$A =
\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9
\end{pmatrix}
$$
\item Then the entries on the main diagonal of $A$ are 1,5 9.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Diagonal matrix}

\begin{itemize}
\item A \emph{diagonal} matrix is a \emph{square} matrix such that all
of the entries \emph{off of the main diagonal} are zero.
\item Let
$$A =
\begin{pmatrix}
1 & 0 & 0 \\
0 & 5 & 0 \\
0 & 0 & 9
\end{pmatrix}
$$
\item Then $A$ is a diagonal matrix.
\item We sometimes write $\diag(a_1,a_2,\cdots, a_n)$ for the $n\times n$
diagonal matrix whose entries on the main diagonal are $a_1,a_2,\cdots, a_n$.
\item The matrix $A$ above would be written $\diag(1,5,9)$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Identity matrix}

\begin{itemize}
\item The matrix
$$I_n = \diag(1, 1, \cdots 1)$$
is called the $n\times n$ \emph{identity} matrix.
\item
$$
I_3 =
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1
\end{pmatrix}
$$
\item If the size of the matrix is clear from context we will often just
write $I$ instead of $I_n$.
\item So we would write
$$
I=
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1
\end{pmatrix}
$$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix Arithmetic}

\begin{itemize}
\item We have defined three arithmetic operations on matrices.
\item Matrix addition between two matrices of the same size.
\item Matrix scalar multiplication.
\item Matrix multiplication. $A B$ is only defined if the number of
\emph{columns} of $A$ is the same as the number of \emph{rows} of $B$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Properties of matrix addition}

\begin{itemize}
\item Matrix addition is \emph{commutative}.
\item $A+B = B + A$.
\item proof: $a_{ij} + b_{ij} = b_{ij} + a_{ij}$.
\item Matrix addition is \emph{associative.}
\item $A + (B + C) = (A + B) + C$.
\item proof: $a_{ij} + (b_{ij} + c_{ij}) = (a_{ij} + b_{ij}) + c_{ij}$.
\item The zero matrix is an additive identity.
\item $$A+O = O + A = A$$
\item Additive inverses: Let $A$ be a matrix. Then there is another
matrix, $-A$ such that $A+(-A) = O$.
\item proof: Let $-A = (-1) A$.
\item
$$
\begin{matrix}

A =

\begin{pmatrix}
1 & -2 & 3\\
-4 & 5  & -6
\end{pmatrix}

&

-A =

\begin{pmatrix}
-1 & 2 & -3\\
4 & -5  & 6
\end{pmatrix}

\end{matrix}
$$
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Properties of scalar multiplication}

\begin{itemize}
\item Scalar multiplication has these properties:
\item $(c \cdot d) A = c ( d A)$. e.g. $2(3 A) = 6 A$.
\item $(c + d) A = c A + d A$. e.g. $2A + 3A = 5 A$.
\item proof: $(c+d)a_{ij} = c a_{ij} + d a_{ij}$.
\item $c (A + B) = cA + cB$.
\item proof: $c (a_{ij} + b_{ij}) = c a_{ij} + c b_{ij}$.
\item $1 A = A$
\item $0 A = O$.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Properties of matrix multiplication}

\begin{itemize}
\item Matrix multiplication is \emph{not} commutative.
\item $A B = B A$ is often \emph{false}.
\item In fact if $A B$ is defined then $B A$ may not even be defined.
\item Suppose $A$ is $3\times 4$ and $B$ is $4\times 5$. Then $A B$
is $3\times 5$ but $B A$ is not defined.
\item Question: When are both $A B$ and $B A$ defined?
\item When $A$ is $m\times n$ and $B$ is $n\times m$.
\item Then $A B$ is $m \times m$ and $B A$ is $n\times n$.
\item If $n\not= m$ then $A B \not= B A$.
\item So the only time there is even the chance that $A B = B A$ is
if $A$ and $B$ are square matrices of the same size.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix multiplication is not commutative}

\begin{itemize}
\item If $A$ and $B$ are both $n\times n$ matrices then $A B$ and $B A$
are both $n\times n$ matrices.
\item $A B = B A$ is sometimes true and sometimes not true.
\item Let
$$
\begin{matrix}

A =

\begin{pmatrix}
2 & 0 \\
0 & 1
\end{pmatrix}

&

B =

\begin{pmatrix}
1 & 1 \\
0 & 1
\end{pmatrix}

\end{matrix}
$$

\item Then
$$
\begin{matrix}

AB =

\begin{pmatrix}
2 & 2 \\
0 & 1
\end{pmatrix}

&

BA =

\begin{pmatrix}
2 & 1 \\
0 & 1
\end{pmatrix}

\end{matrix}
$$
\item So $AB \not= BA$.

\end{itemize}

\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Sometimes $AB = BA$}

\begin{itemize}
\item Let
$$
\begin{matrix}

A =

\begin{pmatrix}
1 & 2 \\
0 & 1
\end{pmatrix}

&

B =

\begin{pmatrix}
1 & 1 \\
0 & 1
\end{pmatrix}

\end{matrix}
$$

\item Then
$$
\begin{matrix}

AB =

\begin{pmatrix}
1 & 3 \\
0 & 1
\end{pmatrix}

&

BA =

\begin{pmatrix}
1 & 3 \\
0 & 1
\end{pmatrix}

\end{matrix}
$$
\item So $AB = BA$.

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Associative law for matrix multiplication}

\begin{itemize}
\item Matrix multiplication \emph{is} associative.
\item $A (B C) = (A B) C$ whenever the multiplication is defined.
\item Let's first check that the sizes make sense.
\item Let $A$ be $m\times n$, $B$ be $n\times k$, $C$ be $k\times p$.
\item Then $B C$ is $n\times p$ and so $A (B C)$ is $m \times p$.
\item $A B$ is $m\times k$ so $(A B) C$ is $m \times p$.
\item So it is at least possible that $A (B C) = (A B) C$
\item To check that they are equal we could expand out the formula for
the $(i, j)$-th element of $A (B C)$  and the formula for the
$(i, j)$-th element of $(A B) C$ and show that they are equal.
\item This is not hard, but messy.
\item Instead we will postpone the proof that matrix multiplication is
associative to a later lesson where we will be able to do it more easily
and more abstractly.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\beamerdefaultoverlayspecification{}

\begin{frame}{Distributive law for matrix multiplication}

\begin{itemize}
\item<1-> Matrix multiplication distributes over matrix addition
\item<2-> $A (B + C) = A B + A C$
\item<3-> Assuming $A$ is $m\times n$ and $B$ and $C$ are $n\times k$.
\item<4-> $(B + C) A = B A + C A$
\item<5-> Assuming $B$  and $C$ are $m\times n$ and $A$ is $n\times k$.
\item<6-> Proof:
\begin{align*}
&\sum_{t=1}^n a_{it} (b_{tj} + c_{tj}) \\
\uncover<7->{&=  \sum_{t=1}^n a_{it} b_{tj} + a_{it} c_{tj} } \\
\uncover<8->{&=  \sum_{t=1}^n a_{it} b_{tj} + \sum_{t=1}^n a_{it} c_{tj}}
\end{align*}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\beamerdefaultoverlayspecification{<+->}

\begin{frame}{Multiplicative identity}
\begin{itemize}
\item The $n\times n$ identity matrix $I$ is a multiplicative identity.
\item Let $A$ be an $n\times m$ matrix. Then $I A = A$.
\item Let $B$ be an $m \times n$ matrix. Then $B I = B$.
\item Let $C$ be an $n\times n$ matrix. Then $C I = I C = C$.
\item
$$
\begin{pmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1
\end{pmatrix}
\begin{pmatrix}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{pmatrix}
=
\begin{pmatrix}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{pmatrix}
$$
\item
$$
\begin{pmatrix}
a & b \\
c & d \\
e & f
\end{pmatrix}
\begin{pmatrix}
1 & 0\\
0 & 1
\end{pmatrix}
=
\begin{pmatrix}
a & b  \\
c & d  \\
e & f
\end{pmatrix}
$$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Multiplying by the zero matrix}
\begin{itemize}
\item The $m\times n$ zero matrix $O_{m\times n}$ is a multiplicative zero.
\item Let $A$ be an $n\times k$ matrix. Then $O_{m\times n} A = O_{m\times k}$.
\item Let $B$ be an $k \times m$ matrix. Then $B  O_{m\times n} = O_{k\times n}$.
\item Let $O = O_{n\times n}$ be the square $n\times n$ zero matrix. Let $A$ be $n\times n$.
\item Then $O A = A O = O$.
\item
$$
\begin{pmatrix}
0 & 0 & 0 \\
0 & 0 & 0 \\
0 & 0 & 0
\end{pmatrix}
\begin{pmatrix}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{pmatrix}
=
\begin{pmatrix}
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0
\end{pmatrix}
$$
\item
$$
\begin{pmatrix}
a & b \\
c & d \\
e & f
\end{pmatrix}
\begin{pmatrix}
0 & 0\\
0 & 0
\end{pmatrix}
=
\begin{pmatrix}
0 & 0  \\
0 & 0  \\
0 & 0
\end{pmatrix}
$$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\beamerdefaultoverlayspecification{}

\begin{frame}{Matrix multiplication and scalar multiplication}

\begin{itemize}
\item<1-> Matrix multiplication is compatible with scalar multiplication
\item<2-> $c (A B) = (c A) B = A (c B)$
\item<3-> Proof:
\begin{align*}
&c \sum_{t=1}^m a_{it} b_{tj} \\
\uncover<4->{&=  \sum_{t=1}^m (c a_{it}) b_{tj} } \\
\uncover<5->{&=  \sum_{t=1}^m a_{it} (c b_{tj}) }
\end{align*}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}


