\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 6 \\ Matrices and Linear Transformations $\R^n$}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Fall 2018 \\ SFSU }
\date{}

\maketitle

\textbf{Problem 1.}

$$
\begin{pmatrix}
1 & 2 \\
2 & 3 \\
3 & 4 \\
\end{pmatrix}
$$

\medskip

\textbf{Problem 2.}


$$
\begin{pmatrix}
3 & 5 & 1 \\
4 & 6 & 2 \\
\end{pmatrix}
$$

\medskip

\textbf{Problem 3.}

$$
\begin{pmatrix}
0 & 1 \\
1 & 0 \\
\end{pmatrix}
$$

\medskip

\textbf{Problem 4.}

$(7, 4, 1)^{\text{T}}$.

\medskip

\textbf{Problem 5.} This is false. All we know is that the first
two columns of $A$ and $B$ are the same, but the third columns of
the two matrices may be different.

\medskip


\textbf{Problem 6.} There are two ways you might think of this.

\smallskip

(i) $\begin{pmatrix} 3 \\ 4\end{pmatrix} = 3 \be_1 + 4 \be_2$
so
\begin{align*}
&T\begin{pmatrix} 3 \\ 4\end{pmatrix} = 3 T(\be_1) + 4 T(\be_2) \\
&= 3\cdot \begin{pmatrix} 1 \\ 2\end{pmatrix} + 4 \cdot \begin{pmatrix} 2 \\ 3\end{pmatrix}
=\begin{pmatrix} 11 \\ 18\end{pmatrix}
\end{align*}

\medskip

(ii) Let $A$ be the matrix of $T$. From the data given we see that
$$A = \begin{pmatrix}
1 & 2 \\
2 & 3 \\
\end{pmatrix}
.$$
So $T\begin{pmatrix} 3 \\ 4\end{pmatrix} = A\begin{pmatrix} 3 \\ 4\end{pmatrix} = \begin{pmatrix} 11 \\ 18\end{pmatrix}$.

\medskip

\textbf{Problem 7.} 20

\medskip

\textbf{Problem 8.}

\textbf{a.} 4

\smallskip

\textbf{b.} Yes. Covectors are linear by definition. The point is that the component functions
of a linear function are themselves linear.

\smallskip

\textbf{c.} You know that the second row is $(7,8,9)$

\smallskip


\textbf{d.}

$$
\begin{pmatrix}
0 & 0 & 0 \\
7 & 8 & 9 \\
0 & 0 & 0 \\
0 & 0 & 0 \\
\end{pmatrix}
$$

\medskip

\textbf{Problem 9.}

\textbf{a.} 4

\smallskip

\textbf{b.} 2

\medskip

\textbf{Problem 10.} Yes a linear transformation, like any function, is
determined by its component functions. The component covectors of a
linear transformation are just the component functions. Another way of
asking this question is this: If two matrices have the same rows are
they the same matrix. The answer of course is yes.

\end{document}
