% $Header$

\documentclass{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 7 \\ Operations on Linear Transformations}
\subtitle{Math 325, Linear Algebra \\ Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Matrices and Linear Transformations}

\begin{itemize}
\item We now know there is a one-to-one correspondence between $m\times n$ matrices and linear
transformations from $\R^n$ to $\R^m$
\item Every matrix gives us a linear transformation.
\item Every linear transformation is given by a matrix.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Operations on matrices}

\begin{itemize}
\item We also have several operations on matrices:
\item matrix multiplication
\item matrix addition
\item matrix scalar multiplication
\item Question: What operations on linear transformations do these
operations on matrices correspond to?
\item Let's start with matrix multiplication.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Matrix multiplication}

\begin{itemize}
\item Let $A$ be an $m\times n$ matrix and $B$ an $n\times k$ matrix.
\item So $A B$ is an $m\times k$ matrix.
\item Now $T_B:\R^k\map\R^n$ and $T_A:\R^n\map\R^m$.
\item Also $T_{AB}:\R^k\map\R^m$.
\item How could we combine $T_A$ and $T_B$ in order to form $T_{AB}$?
\item Let $\bv\in\R^k$. Because matrix multiplication is \emph{associative}
\item $(A B) \bv = A (B \bv)$.
\item So $T_{AB}(\bv) = T_A\left( T_B(\bv) \right)$.
\item So $T_{AB} = T_A \circ T_B$.
\item Matrix multiplication corresponds to \emph{composition} of linear transformations.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{We used associativity of matrix multiplication}

\begin{itemize}
\item It is true that matrix multiplication corresponds to composition of linear transformations.
\item Unfortunately the argument given on the previous slide doesn't work for us as a proof.
\item The problem is that we never proved that matrix multiplication is associative.
\item We stated the fact in lesson 2, and said we would postpone the proof until later.
\item We will prove that matrix multiplication is associative in this lesson, based on
the fact that matrix multiplication corresponds to composition of linear transformations.
\item But then we can't use the fact that matrix multiplication is associative in order
to prove that matrix multiplication corresponds to composition.
\item We need a different proof.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\beamerdefaultoverlayspecification{}
\begin{frame}{Matrix multiplication corresponds to composition}

\begin{theorem}
Let $A$ be an $m\times n$ matrix and $B$ an $n\times k$ matrix.

\pause

Then $T_{AB} = T_A \circ T_B$.
\end{theorem}

\pause

\begin{proof}
It is enough to show that $T_{AB}$ and $T_A \circ T_B$ agree on the standard
basis vectors.

\pause

Let $\be_j$ be the $j$-th standard basis vector, for $1\leq j\leq k$. We will show that
$(AB)\be_j = A (B \be_j)$.

\pause

Let $\bb_1, \bb_2,\cdots \bb_k$ be the columns of $B$, so that
$B = (\bb_1 \vert \bb_2 \vert \cdots \bb_k)$.

 \pause

 We know that $AB = (A \bb_1 \vert A \bb_2 \vert \cdots A \bb_k)$.

 \pause

 So then $(AB)\be_j = A\bb_j$. (The columns of a matrix equal the images of the standard basis vectors.)

 \pause

 Now $B \be_j = \bb_j$. (Same reason.) So $A (B \be_j) = A\bb_j$.

 \pause

 So $(AB)\be_j = A (B \be_j)$.

\end{proof}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example}

Let
$$
\begin{matrix}
A=
\begin{pmatrix}
1 &  0 & 2 \\
0 & -1 & -3
\end{pmatrix}
&
B =
\begin{pmatrix}
 1 & 1 \\
-1 & 0 \\
 0 & 4
\end{pmatrix}
\end{matrix}
$$

\pause

$T_B:\R^2\map \R^3$ and $T_A:\R^3\map \R^2$. So $T_A \circ T_B : \R^2\map \R^2$.
Find the matrix corresponding to the linear transformation $T_A \circ T_B$.

\pause

Solution. The matrix is

$$ AB =
\begin{pmatrix}
1 & 9 \\
1 & -12
\end{pmatrix}
$$

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix multiplication is associative}

\begin{lemma}
Let $A$ be an $m\times n$ matrix, $B$ an $n\times k$ matrix, and $C$ a $k\times l$ matrix

\pause

Then $A(BC) = (AB)C$.
\end{lemma}

\pause

\textbf{Proof.}

Because there is a one-to-one correspondence between matrices and linear transformations,
it is enough to show that the two corresponding linear transformations are the same.

\pause

\smallskip

We will show that $T_{A(BC)} = T_{(AB)C}$.

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Proof continued}
Let $\bv \in \R^l$. We will show that $T_{A(BC)}(\bv) = T_{(AB)C}(\bv)$.

\uncover<2->{We repeatedly use the fact that matrix multiplication corresponds to composition of linear transformations.}

\begin{align*}
\uncover<3->{ &T_{A(BC)}(\bv) \\}
\uncover<4->{ & = T_A\left( T_{BC} (\bv) \right) \\}
\uncover<5->{ & = T_A\left( T_B \left( T_C(\bv) \right) \right) \\}
\uncover<6->{ & = T_{AB}\left(  T_C(\bv) \right) \\}
\uncover<7->{ & = T_{(AB)C}(\bv)}
\end{align*}

\uncover<8->{ So $T_{A(BC)} = T_{(AB)C}$.}


\uncover<9->{ So $A(BC) = (AB)C \qed$.}

\end{frame}

\beamerdefaultoverlayspecification{<+->}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The identity matrix}

\begin{itemize}
\item Let $I$ be the $n\times n$ identity matrix.
\item Which linear transformation does $I$ correspond to?
\item The \emph{identity} transformation.
\item $T_I:\R^n\map\R^n$.
\item $T_I(\bv) = I \bv = \bv$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The zero matrix}

\begin{itemize}
\item Let $O$ be the $m\times n$ zero matrix.
\item Which linear transformation does $O$ correspond to?
\item The constantly \emph{zero} transformation.
\item $T_O:\R^n\map\R^m$.
\item $T_O(\bv) = O \bv = \mathbf{0}$, for all $\bv$.
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Matrix addition}

\begin{itemize}
\item Which linear transformation does $A+B$ correspond to?
\item $T_A + T_B$.
\item Let $f$ and $g$ be any two functions from $\R^n$ to $\R^m$.
\item Then $f+g$ is defined to be the function
$$(f+g)(\bv) = f(\bv) + g(\bv)$$
\item Now $(A + B)(\bv) = A\bv + B\bv$.
\item So $T_{A+B} = T_A + T_B$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Scalar multiplication}

\begin{itemize}
\item Which linear transformation does $c A$ correspond to?
\item $c T_A$.
\item Let $f$ be any function from $\R^n$ to $\R^m$.
\item Then $c f$ is defined to be the function
$$(c f)(\bv) = c \left( f(\bv) \right)$$
\item Now $(c A)\bv = c(A\bv)$.
\item So $T_{cA} = cT_A$.
\end{itemize}
\end{frame}


\end{document}


