\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,rotating}
\usepackage[shortlabels]{enumitem}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Mouse Pairs and Suslin Cardinals in a Type I Hierarchy}
\author{John R. Steel}

\maketitle

\tableofcontents

We assume $\ADR$ and $\HPC$. Following Jackson \cite{Jackson1}:

Let $\lambda_1$ be a limit of Suslin cardinals, and $\cf(\lambda_1)=\omega$.
Put $S_{<\lambda_1} =\Union{\alpha<\lambda_1} S_{\alpha}$, and
\begin{align*}
\bSigma_0 &= \Union{\omega} S_{<\lambda_1} \\
\bPi_0 &= \Intersection{\omega} S_{<\lambda_1} \
\end{align*}

and

\begin{align*}
\bSigma_{n+1} &= \exists^{\R}\bPi_n \textbf{,\quad}
\bPi_{n+1} = \forall^{\R}\bSigma_n \\
\bdelta_n&=\sup\setof{|R|}{R\text{ is a }\bDelta_n\text{ pwo of } \R}
\end{align*}

for $n\geq 1$. [Rudominer: Should this be $n\geq 0$?]
We have $\Scale(\bSigma_{2n+2})$ and $\Scale(\bPi_{2n+1})$,
for all $n$, with norms into $\bdelta_{2n+1}$. We also have

$$ \bdelta_{2n+1} = \lambda_{2n+1}^+,$$
where $\lambda_{2n+1}$ is a Suslin cardinal of cofinality $\omega$.
(So $\bdelta_1=\lambda_1^+$.) The sequence $\lambda_1,\bdelta_1,\lambda_3,\bdelta_3,\cdots$
enumerates the Suslin cardinals in this hierarchy. For $\kappa=\bdelta_{2n+1}$,
$S_{\kappa}=\bSigma_{2n+2}$. For $\kappa=\lambda_{2n+1}$,
$S_{\kappa}=\bSigma_{2n+1}$.

If $\lambda_1=\omega$, then $\bSigma_n = \bSigma^1_n$ (for $n\geq 1$) in  the
ordinary projective hierarchy. In that case, we have the envelopes

$$
\bGamma^1_{k} = \Union{n<\omega} \Game^{k}(\omega n - \bPi^1_1)
$$

for $k\geq0$. There are self-justifying systems sealing those envelopes. For $k$ even, this gives us an optimal scale on $\bSigma^1_{k+1}$, and for $k$ odd, an optimal scale on $\bPi^1_{k+1}$. By results of
Hjorth, Woodin and Sargsyan (See \cite{Sargsyan1}),
$$
\lambda^1_{2k+1} = \sup\setof{|\leq|}{\leq\text{ is a $\bGamma^1_{2k-1}$ pwo of $\R$}}.
$$

The $\bGamma^1_{k}$ are related to the mice $M_n = $ minimal model with $n$ Woodins, in various ways. For example, let

$$
T^n_{k+1}(\varphi,x) \Ifff M_k \models \varphi\left[ x,\omega_1,\cdots,\omega_n\right],
$$
then each $T^n_{k+1}$ is in $\bGamma^1_{k+1}$, and the  $T^n_{k+1}$ are Wadge cofinal in
$\bGamma^1_{k+1}$.

Our goal is to generalize this theory to the type I hierarchy $\bSigma_n,\bPi_n$ we have
fixed above. The main problem is how to define the analogs of $\bGamma^1_{0}$ and
$M_0=L$. This is a problem even when we are considering the second projective-like hierarchy,
$\bSigma_n= \bSigma_{n+1}^{J_2(\R)}$. (I.e. $\lambda_1=\bdelta^1_{\omega}$.)
Hugh Woodin recently solved this problem in its mouse set incarnation, by showing that the reals in Rudominer's ``ladder mouse'' are precisely the reals that are $\Delta_2^{J_2(\R)}$ in some
$\alpha<\omega_1$. (It turns out that $M^{\text{ld}}$ is somewhat more analogous to $M_1$.)
We show here that Woodin's solution to the Rudominer problem adapts (easily) to the general type I hierarchy above (assuming HPC). We then show that it yields good analogues of the $\bGamma^1_n$'s
and $M_n$'s.

As a corollary we obtain

\begin{theorem} ($\AD^+$) Let $\kappa$ be a Suslin cardinal, and $(P,\Sigma)$ a
mouse pair such that $\kappa<\ords\left( M_{\infty}(P,\Sigma)  \right)$.
Suppose $\cf(\lambda) = \omega$, where $\lambda$ is the largest limit of Suslin
cardinals $\leq \kappa$.

Then $\kappa$ is a cutpoint of $M_{\infty}(P,\Sigma)$.
\end{theorem}


\begin{remark}
It is shown in \cite{Steel1} that if $\kappa$ is a cardinal and a cutpoint of
$M_{\infty}(P,\Sigma)$ for some mouse pair $(P,\Sigma)$, then $\kappa$ is Suslin.
\end{remark}

\begin{corollary}
Assume $\AD_{\R} + \HPC$, and let $\kappa$ be a Suslin cardinal. Suppose
$\cf(\lambda)=\omega$, where $\lambda$ is the largest limit of Suslin cardinals
$\leq \kappa$; then $\kappa$ is a cutpoint of the $\HOD$-sequence.
\end{corollary}

\begin{remark}
It is shown in \cite{Steel1} that the theorem and corollary hold when $\kappa$
is a limit of Suslins, or the next Suslin after a limit of Suslins.
We of course conjecture that they hold for any Suslin cardinal $\kappa$.
\end{remark}

\section{Ladder Mice for $J_2(\R)$}

Woodin's work characterizing $\Delta_2^{J_2(\R)}$-in-$\alpha$ as a mouse set is
unpublished, so we devote this section to describing it.

The ladder mouse $M^{\text{ld}}$ was defined by Rudominer
(\cite{My_Thesis} and \cite{Mouse_Sets} ).

\begin{definition}
$M^{\text{ld}}$ is the minimal iterable pure extender mouse such that for each $k<\omega$,
 there is a $\delta<\ords(M^{\text{ld}})$ such that
\begin{enumerate}[(a)]
\item $M_k(M^{\text{ld}} \vert \delta) \properseg M^{\text{ld}} $, and
\item $M_k(M^{\text{ld}} \vert \delta) \models \delta$ is Woodin, and
\item $M^{\text{ld}} \models \delta$ is a cardinal.
\end{enumerate}

Here $M_k(N)$ is the minimal mouse with $k$ Woodins over $N$. Let $\delta_k$
be the least $\delta$ as in (a) - (c); then $\delta_k<\delta_{k+1}$, and
$\ords(M^{\text{ld}}) = \sup_k \delta_k$. These are the rungs of our ladder.

\end{definition}

Rudominer showed in \cite{My_Thesis} that every real belonging to
$M^{\text{ld}}$ is $\Delta_2^{J_2(\R)}$ in some $\alpha<\omega_1$.

For let $N$ be the $\alpha^{\text{th}}$ level of $M^{\text{ld}}$ that projects to
$\omega$. Say that a premouse $P$ is  $\Pi_1$-iterable iff $\forall \cT$ on $P$,
$\forall k < \omega$,  $\exists$ maximal branch $b$ of
$\cT$ s.t. $ M_k(M(\cT))\initseg M_b^{\cT}$ or $M_b^{\cT}\initseg M_k(M(\cT))$.
This property is $\Pi_1^{J_2(\R)}$.

\begin{aside}
That is what we called $\Pi_0$ in the general setting. It corresponds to infinite recursive
conjunctions of analytical predicates. [Rudominer: Then why bother calling
it $\Pi_1$-iterable here? Why not call it either $\Pi_1^{J_2(\R)}$-iterable
or $\Pi_0$-iterable.]
\end{aside}

One can show that $N$ is the unique $\Pi_1$-iterable premouse that thinks it is
the $\alpha^{\text{th}}$ level projecting to $\omega$. So if $x\in{}^{\omega}\omega$
codes the theory of $N$, the $\singleton{x}$ is $\Pi_1^{J_2(\R)}$ in $\alpha$,
hence $x$ is $\Delta_2^{J_2(\R)}$ in $\alpha$.

\begin{theorem}[Woodin, unpublished] Let $x$ be $\Delta_2^{J_2(\R)}$ in $\alpha<\omega_1$;
then $x\in M^{\text{ld}}$.
\end{theorem}

\begin{proof}
Let $x$ be $\Delta_2^{J_2(\R)}$ in $\alpha$. We assume $\alpha=0$ for simplicity.
(Otherwise, iterate the least measurable of $M^{\text{ld}}$ past $\alpha$, generating
$M^{\text{ld}}\to P$. The argument to come shows $x\in P[g]$ for all $g$
on $\Col(\omega,\alpha)$, hence $x\in P$, hence $x\in M^{\text{ld}}$.)
So we have recursive sequences $\angles{A_n}$ and $\angles{B_n}$ such that
for $n\geq 1$
\begin{enumerate}[(i)]
\item $A_n$ and $B_n$ are $\Sigma^1_{2n}$ subsets of $\R\times\omega$,
uniformly in $n$, and
\item for any $k\in\omega$
\begin{align*}
k\in x &\Iff \exists y \bigwedge_n A_n(y,k) \\
k\notin x &\Iff \exists y \bigwedge_n B_n(y,k). \\
\end{align*}
\end{enumerate}
(We assume $x\subseteq\omega$ now.)

Let $\phibar^n=\sequence{\varphi^n_i}{i\in\omega}$ be a $\Sigma^1_{2n}$
scale on $A_n$, and $\psibar^n=\sequence{\psi^n_i}{i\in\omega}$ be a $\Sigma^1_{2n}$
scale on $B_n$. Let $\leq^A_{n,i}$ be the prewellorder
$\leq_{\varphi^n_i}$, and $\leq^B_{n,i}=\leq_{\psi^n_i}$. Let
\begin{align*}
A_k(y) &\Iff  \bigwedge_n A_n(y,k) \\
B_k(y) &\Iff  \bigwedge_n B_n(y,k).
\end{align*}

We get scales $\thetabar^k$ and $\nubar^k$ on $A_k$ and $B_k$ via

\begin{align*}
\theta^k_{\angles{n,i}}(y) &= \varphi^n_i(y,k) \\
\nu^k_{\angles{n,i}}(y) &= \psi^n_i(y,k).
\end{align*}

Let $S^k\subseteq(\omega \times \bdelta^1_{\omega})^{<\omega}$ and
$T^k\subseteq(\omega \times \bdelta^1_{\omega})^{<\omega}$  be the trees of these scales,
so that

\begin{align*}
k\in x &\Iff \left[ S^k \right] \not= \emptyset \\
k\notin x &\Iff \left[ T^k \right] \not= \emptyset.
\end{align*}

Now we look inside $M^{\text{ld}}$. This model is projectively correct, so it
has the prewellorders $\leq^A_{n,i}$ and $\leq^B_{n,i}$ restricted to its own
reals. Also, for any $y\in M^{\text{ld}}$,

\begin{align*}
\left( M^{\text{ld}} \models A_k(y)  \right) &\Implies A_k(y) \text{, and} \\
\left( M^{\text{ld}} \models B_k(y)  \right) &\Implies B_k(y).
\end{align*}

[Rudominer: Replace $\Implies$ with $\Iff$ above?]

Let $S^k_{M}$ and $T^k_{M}$ be the trees of the scales
$\left( \thetabar^k \right)^{M^{\text{ld}}}$ and
$\left( \nubar^k \right)^{M^{\text{ld}}}$ computed in ${M^{\text{ld}}}$
using the pwo's $\leq^A_{n,i}\intersect M^{\text{ld}}$ and
$\leq^B_{n,i}\intersect M^{\text{ld}}$. These are isomorphic to subtress
of the full $S^k$ and $T^k$. The projections in $M^{\text{ld}}$ are
$A_k\intersect M^{\text{ld}}$ and
$B_k\intersect M^{\text{ld}}$ respectively. For each $k$, at least on of
$S^k_{M}$ and $T^k_{M}$ is wellfounded.

\begin{claim}[Claim 1]
For any $k\in\omega$, $k\in x$ iff $M^{\text{ld}} \models \text{`` }T^k_M$ is
wellfounded, and either $S^k_M$ is illfounded, or $|T^k_M|<|S^k_M|$.''
\end{claim}

Clearly Claim 1 implies $x\in M^{\text{ld}}$, as desired.

\begin{subproof}{(of Claim 1)}
It is enough to show $\Rightarrow$, since then we get $\Leftarrow$ by assuming
$k\notin x$ and using the same argument to show $S^k_M$ is wellfounded and
$|S^k_M| < |T^k_m|$.

So fix $k_0\in x$, and fix $y$ s.t. $A_{k_0}(y)$. Since $B_{k_0} = \emptyset$,
$T^{k_0}_M$ is wellfounded. We must see that $|T^{k_0}_M| < |S^{k_0}_M|$ or
$S^{k_0}_M$ is illfounded.

Let for any $k$

\begin{align*}
\delta_k &=  \text{ least cardinal } \delta \text{ of } M^{\text{ld}} \text{ s.t. }
    M_{2k+5}(M^{\text{ld}}|\delta) \initseg M^{\text{ld}} \text{ and } \\
N_k      &=  M_{2k+5}(M^{\text{ld}}|\delta_k)
\end{align*}
[Rudominer: Two problems with the above text: Firstly we have already defined
$\delta_k$ to be the $k$-th rung of the ladder. Secondly we forgot above to
say that $\delta$ needs to be locally Woodin.]

So $\delta_k$ is Woodin in $N_k$, which has $2k+5$ Woodins above it as well. The
$\delta_k$'s are cofinal in $\lambda=\ords(M^{\text{ld}})$. By iterating,
we may assume that for each $k$, there is an $N_k$-generic $g_k$ on
$\Col(\omega,\delta_k)$ such that

$$
y\in N_k[g_k].
$$

We caution that $g_k$ is only $N_k$-generic, not necessarily $M^{\text{ld}}$-generic,
because $\delta_k$ is only Woodin in $N_k$, not the full $M^{\text{ld}}$.
(But the relevant genericity iterations do move the whold $M^{\text{ld}}$ along,
because $\delta_k$ is a cardinal of $M^{\text{ld}}$. This is what motivated
Rudominer's definition in the first place.)

We now build a generic $G$ for the countable stationary tower $\Q^{M^{\text{ld}}}_{<\lambda}$
such that for all $k$, $G\intersect \Q_{<\delta_k}$ is $N_k$-generic. (Note that
$\Q^{M^{\text{ld}}}_{<\delta_k}=\Q^{N_k}_{<\delta_k}$ because $\delta_k$ is
a cardinal of $M^{\text{ld}}$. But $M^{\text{ld}}$ has more dense subsets of $\Q_{<\delta_k}$
than $N_k$ does.)

\end{subproof}

\end{proof}

\bibliographystyle{amsalpha}

\bibliography{math}

\end{document}
