% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,xcolor}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 17 \\ Determinants and Miscellanesou}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant Theory}

\begin{itemize}
\item \textbf{Theorem.} There exists a unique function, called the \emph{determinant}, and writen $\det$, with the following properties:
\item (1) For every square matrix $A$, $\det(A)$ is a scalar.
\item (2) Performing an elementary row operation of type 1 on a matrix does not change its determinant.
\item (3) Performing an elementary row operation of type 2 on a matrix multiplies the determinant by -1.
\item (4) Performing an elementary row operation of type 3 on a matrix, multiplying one of the rows by the scalar $t$,
multiplies the determinant by the same scalar $t$.
\item (5) The determinant of the identity matrix is 1.
\item We will give a sketch of the proof but we won't go through the proof in detail.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $2\times 2$ matrix.}

\begin{itemize}
\item We have already defined the determinant of $2\times 2$ matrix.
\item
$
\det
\begin{pmatrix}
a & b \\
c & d \\
\end{pmatrix}
= ad - bc
$.
\item Verify each of the properties:
\item (1) For every $2\times 2$ matrix $A$, $\det(A)$ is a scalar.
\item (2) Performing an elementary row operation of type 1 on a matrix does not change its determinant.
\item (3) Performing an elementary row operation of type 2 on a matrix multiplies the determinant by -1.
\item (4) Performing an elementary row operation of type 3 on a matrix, multiplying one of the rows by the scalar $t$,
multiplies the determinant by the same scalar $t$.
\item (5) The determinant of the identity matrix is 1.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Sign of a Permutation}

\begin{itemize}
\item \textbf{Definition} If $\pi$ is a permutation of the numbers $\singleton{1,2,\cdots, n}$
we say that $\pi$ is \emph{even} and that $\sgn(\pi) = 1$ if $\pi$ is obtained via an even number of swaps.
\item Otherwise we say that $\pi$ is odd and $\sgn(\pi) = -1$.
\item For example consider the following permutations of $(1,2,3,4)$.
\item $\pi=(1,2,3,4)$. $\pi$ is even. $\sgn(pi) = 1$. $\pi$ is obtained via zero swaps.
\item $\pi=(3,2,1,4)$. $\pi$ is odd. $\sgn(pi) = -1$. $\pi$ is obtained via one swap.
\item $\pi=(3,2,4,1)$. $\pi$ is even. $\sgn(pi) = 1$. $\pi$ is obtained via two swaps.
\item $\pi=(3,4,2,1)$. $\pi$ is odd. $\sgn(pi) = -1$. $\pi$ is obtained via three swaps.
\item \textbf{Theorem.} $\sgn(\pi)$ is well-defined. (We won't prove this.)
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{General Definition of Determinant}

\begin{itemize}
\item \textbf{Definition} Let $A$ be an $n\times n$ matrix.
\item Then
$$\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}$$
\item The sum is over all permutations of the row indices $(1,2,\cdots, n)$
\item Each term in the sum is the product over the columns of $A$ of one of the entries in that column.
\item In column $i$ we take the entry from row $\pi(i)$.
\item The product is multiplied by $\sgn(\pi)$.
\item i.e. we add the product if $\pi$ is even and subtract the product of $\pi$ is odd.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $2\times 2$ matrix revisited.}

\begin{itemize}
\item Check that the two different definitions coincide for $2\times 2$ matrices.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $3\times 3$ matrix.}

\begin{itemize}
\item
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%







\end{document}


