% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 15 \\ $\Z_n$}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$\Z^*_n$}

\begin{itemize}
  \item Let $n>1$ be an integer.
  \item $\Z*_n$ is pronounced $\Z$-mod-n-star.
  \item It is also written $\Z^*/(n)$.
  \item It is a subset of $\Z_n$. $\Z_n^* \subset \Z_n$.
  \item $\Z^*_n$ is the set of units of $\Z_n$.
  \item Find $\Z_6^*$.
  \item $\Z_6^*=\singleton{[1],[5]}$.
  \item Find $\Z_7^*$.
  \item $\Z_7^*=\singleton{[1],[2],[3],[4],[5],[6]}$.
  \item Find $\Z_8* = \singleton{[1],[3],[5],[7]}$.
\end{itemize}

\end{frame}

\begin{frame}{Other characterizations of $\Z^*_n$}

\begin{itemize}
  \item If $[a]\in\Z_n$ then $[a]\in\z_n^*$
  \item iff $\gcd(a,n)=1$
  \item iff $[a]$ has a multiplicative inverse ($[a]$ is a unit)
  \item iff $a\not 0$ and $[a]$ is not a zero-divisor.
  \item If $p$ is prime, then $\Z_p^*=\setof{[a]\in\Z_p}{a\not=0}$.
  \item $\Z_5^*=\singleton{[1],[2],[3],[4]}$
  \item $\Z_7^*=\singleton{[1],[2],[3],[4],[5],[6]}$.
  \item If $n$ is not prime then this is not the case.
\end{itemize}

\end{frame}

\begin{frame}{Multiplication in $\Z^*_n$}

\begin{itemize}
  \item $\Z_n^*$ is a self-contained arithmetic universe inside of a self-contained
  arithmetic universe.
  \item $\Z_n^*$ is closed under the multiplication of $\Z_n$.
  \item If $[a],[b]\in\Z_n^*$ and $[a]\cdot[b]$ is computed in $\Z_n$, then
  $[a]\cdot[b]\in\Z_n^*$.
  \item For example work in $\Z_8$...
  \item $\Z_8* = \singleton{[1],[3],[5],[7]}$.
  \item $[3]\cdot [3] = [1]$.
  \item $[3]\cdot [5] = [7]$.
  \item $[3] \cdot [7] = [5]$.
\end{itemize}

\end{frame}

\begin{frame}{$\Z^*_n$ is closed under multiplication}

\begin{itemize}
  \item If $[a],[b] in\Z_n^*$ then so is $[a][b]$.
  \item \textbf{Proof} Suppose $[a],[b]\in\Z_n^*$.
  \item So $[a]^{-1}$ and $[b]^{-1}$ exist.
  \item To see $[a][b]\in\Z_n^*$ we must see that $([a][b])^{-1}$ exists.
  \item Notice that $([a][b])\cdot ([a]^{-1}[b]^{-1}) = [1]$.
  \item This means that $([a][b])^{-1} = [a]^{-1}[b]^{-1}$.
  \item So $([a][b])^{-1}$ exists and $[a][b]\in\Z_n^*$. $\qed$.
  \item \textbf{Proof 2}. Suppose $a$ and $b$ are relatively prime to $n$.
  \item Then $a$ and $n$ don't share any prime factors and neither do $b$ and $n$.
  \item But then $ab$ and $n$ don't share any prime factors.
  \item So $ab$ and $n$ are relatively prime. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{$\Z^*_n$ is a commutative group under multiplication}

\begin{itemize}
  \item Multiplication is commutative and associative.
  \item $[1]$ is a multiplicative identity.
  \item Every element has a multiplicative inverse.
  \item Note that $\Z^*_n$ is not closed under addition.
  \item Example: Work in $\Z_{10}$...
  \item $[1]$ and $[3]$ are units but $[4]$ is not.
  \item Note that $\Z^*n$ is closed under additive inverse.
  \item If $[a]\in\Z^*n$ then $-[a]\in\Z^*n$.
  \item Work in $\Z_{10}$...
  \item $[3]\in\Z^*_{10}$. $-[3]=[7]\in\Z^*{10}$.
  \item This is because $-[1]$ is a unit.
\end{itemize}

\end{frame}


\begin{frame}{The Orbit of an element of $\Z_n^*$}

\begin{itemize}
  \item Let $[a]\in\Z^*n$. The \emph{orbit} of $[a]$ is
  the set of powers of $[a]$.
  \item Careful not to confuse powers with multiples.
  \item Example: Work in $\Z^*_7$...
  \item Find the orbit of $[2]$.
  \item $[2]^1 = 2$, $[2]^2 = 4$, $[2]^3=[1]$
  \item $[2]^4 = 2$, $[2]^5 = 4$, $[2]^6=[1]$.
  \item So the orbit of $[2]$ in $\Z^*_7$ is: $[1],[2],[4]$
  \item Find the orbit of $[3]$ in $\Z^*_7$.
  \item $[3]^1=[3]$, $[3]^2=[2]$, $[3]^3=[6]$, $[3]^4=[4]$
  \item $[3]^5=[5]$, $[3]^[6]=[1]$,
  \item $[3]^7=[3]$, $[3]^8=[2]$, $[3]^9=[6]$, $[3]^{10}=[4]$.
  \teim So the orbit of $[3]$ in $\Z^*_7$ is $[1],[2],[3],[4]
\end{itemize}

\end{frame}




\end{document}
