% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 2 \\ Divisibility}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Divisibility}


\begin{itemize}
  \item \textbf{Definition:} Let $a,b\in\Z$. Then we say that $a$ \emph{divides}
  $b$ iff there is a $c\in\Z$ such that $b = a c$.
  \item We write this as $a\divides b$.
  \item If $a$ doesn't divide $b$ we write $a\ndivides b$.
  \item Example: $5 \divides 15$ because $15 = 5 \cdot 3$.
  \item Example: $5 \ndivides 16$ because there is no integer $c$ such that
  $16 = 5  c$.
  \item Be careful not to confuse the notations $a\divides b$ with $a / b$.
\end{itemize}

\end{frame}

\begin{frame}{Divisibility}

\begin{itemize}
  \item True or false: $14 \divides 7$.
  \item True or false: $7 \divides 14$.
  \item True or false: $-7 \divides 14$.
  \item True or false: $7 \divides -14$.
  \item True or false: $7 \divides 0$.
  \item True or false: $0 \divides 7$.
  \item True or false: $0 \divides 0$.
  \item For which values of $a$ does $a \divides 0$?
  \item For which values of $b$ does $0 \divides b$?
  \item The textbook does not allow $a=0$ in the expression $a\divides b$.
  But there is no harm in saying that $0\divides 0$.
\end{itemize}

\end{frame}

\begin{frame}{Divisors and multiples}

\begin{itemize}
  \item Suppose that $a\divides b$, so there is some $c$ such that $b = ac$.
  \item We also say that $a$ is a \emph{divisor} of $b$.
  \item And we also say that $b$ is a \emph{multiple} of $a$.
  \item Example: Because $24 = 6 \times 4$, 24 is a multiple of 6 and 6 is
  a divisor of 24 and $6\divides 24$.
  \item Example: Find all of the positive divisors of 12.
  \item Solution: 1,2,3,4,6,12.
  \item Example: Find all of the divisors of 12.
  \item Solution: $\pm 1, \pm 2, \pm 3, \pm 4, \pm 6, \pm 12$.
\end{itemize}

\end{frame}

\begin{frame}{$a/b \in \Z$}

\begin{itemize}
  \item \textbf{Fact:} Let $a,b\in\Z$ with $a\not=0$.
  \item Then $a\divides b$ iff $b / a \in \Z$.
  \item Example: $4 \divides 12$ because $12/4 \in \Z$.
  \item Example $5 \ndivides 12$ because $12/5 \in \Q \setminus \Z$.
  \item \textbf{proof:} If $a\divides b$ then there is a $c$ such that
  $b = ac$.
  \item So then $b / a = c$ so $b/a \in \Z$.
  \item Conversely, suppose $b/a \in \Z$.
  \item Then, letting $c = b/a$ we have $b = a c$ so $a\divides b$.
  \item $\qed$
  \item Note the logical structure of the proof of an ``iff'' statement.
\end{itemize}

\end{frame}

\begin{frame}{Properties of divisibility}

\begin{itemize}
  \item True or false? For all $a,b,c\in\Z$, if
  $a\divides b$ and $b\divides c$ then $a\divides c$.
  \item How could we decide if a statement like this is true or false?
  \item Since the statement is about all integers, in order to see that it
  is true, it does \emph{not} suffice to give an example.
  \item To see that the statement is true we need to give a proof.
  \item On the other hand, to see that the statement is false it suffices
  to give a single counterexample.
  \item In this case, the statement is true.
  \item proof: If $a\divides b$ then there is a $d$ such that $b=ad$.
  \item If $b\divides c$ there is an $e$ such that $c = b e$.
  \item So $c = (a d) e = a (de)$, so $a\divides c$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Properties of divisibility, 2}

\begin{itemize}
  \item True or false? For all $a,b,c\in\Z$, if $a\divides bc$ then
  $a\divides b$ or $a\divides c$?
  \item This is false. To prove this we only need to give one counterexample.
  \item $6\divides (3\times 4)$ but $6\ndivides 3$ and $6\ndivides 4$.
  \item Notice that I could also give an example where this is true.
  \item $3 \divides 12$ and $12 = 6 \times 2$ and $3\divides 6$.
  \item But one example is not sufficient to prove that a property is true
  for all integers.
\end{itemize}

\end{frame}

\begin{frame}{Integer Linear Combinations}

\begin{itemize}
  \item \textbf{Definition} Let $a,b\in\Z$. Then an
  \emph{integer linear combination} of $a$ and $b$ is any integer of the
  form $xa + yb$ where $x,y\in\Z$.
  \item $-3\times 4 + 2\times 7 = 2$ so 2 is an integer linear combination
  of 4 and 7.
\end{itemize}

\end{frame}

\begin{frame}{Properties of divisibility, 3}

\begin{itemize}
  \item True or false? For all $a,b,c\in\Z$, if $a\divides b$ and $a\divides c$
  then $a$ divides every integer linear combination of $b$ and $c$.
  \item True. Proof: Suppose $a\divides b$ and $a\divides c$.
  \item So there are $n,m\in\Z$ so that $b=na$ and $c=ma$.
  \item Let $r$ be an integer linear combination of $b$ and $c$.
  \item So there are $x,y\in\Z$ such that $r=xb+yc$.
  \item So $r=xna + yma = (xn + ym)a$.
  \item So $r$ is a multiple of $a$, i.e. $a\divides r$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Reading assignment}

\begin{itemize}
  \item Beginning of \emph{Divisibility}, section 5.1 of Rosen. The bottom
  of page 36 and the top of page 37. Stop before Theorem 1.10,
  \emph{The Division Algorithm}.
\end{itemize}

\end{frame}

\end{document}
