% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 13 \\ $\Z_n$}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2021 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$\Z_n$}

\begin{itemize}
  \item Let $n>1$ be an integer.
  \item $\Z_n$ is pronounced $\Z$-mod-n.
  \item It is also written $\Z/(n)$.
  \item It consists of the  set of congruence classes mod $n$.
  \item Example: $\Z_6=\singleton{[0]_6,[1]_6,[2]_6,[3]_6,[4]_6,[5]_6}$.
  \item If $n$ is clear from context, we will drop the subscript and write...
  \item  $\Z_6=\singleton{[0],[1],[2],[3],[4],[5]}$.
  \item You can think of $\Z_n$ as consisting of the integers $0$ to $n-1$,
  \item but in their role as the least non-negative residues modulo $n$.
\end{itemize}
\end{frame}

\begin{frame}{Addition in $\Z_n$}

\begin{itemize}
  \item We define addition in $\Z_n$ as follows:
  \item $[a]_n + [b]_n = [a+b]_n$.
  \item For example, letting $n=6$, addition in $\Z_6$ is defined as:
  \item $[a]_6 + [b]_6 = [a+b]_6 = [(a+b) \bmod 6]_6$.
  \item Working in $\Z_6$...
  \item $[3] + [0] = [3]$.
  \item $[3] + [1] = [4]$.
  \item $[3] + [2] = [5]$.
  \item $[3] + [3] = [0]$.
  \item $[3] + [4] = [1]$.
  \item $[3] + [5] = [2]$.
\end{itemize}

\end{frame}

\begin{frame}{Multiplication in $\Z_n$}

\begin{itemize}
  \item We define multiplication in $\Z_n$ as follows:
  \item $[a]_n \cdot [b]_n = [ab]_n = [(a\cdot b) \bmod n]_n$.
  \item For example, letting $n=6$, multplication in $\Z_6$ is defined as:
  \item $[a]_6 \cdot [b]_6 = [ab]_6 = [(a\cdot b) \bmod 6]_6$.
  \item Working in $\Z_6$...
  \item $[3] \cdot [0] = [0]$.
  \item $[3] \cdot [1] = [3]$.
  \item $[3] \cdot [2] = [0]$.
  \item $[3] \cdot [3] = [3]$.
  \item $[3] \cdot [4] = [0]$.
  \item $[3] \cdot [5] = [3]$.
\end{itemize}

\end{frame}

\begin{frame}{Self-contained arithmetic universe}

\begin{itemize}
  \item $\Z_n$ is a self-contained arithmetic universe.
  \item When you add or multiply two elements of the set you get another element of the set.
  \item The arithmetic is just that of ``wrapping around.''
  \item For example, let $n=6$ and work in $\Z_6$.
  \item  $\Z_6=\singleton{[0],[1],[2],[3],[4],[5]}$.
  \item $[3] + [1] = [4]$.
  \item $[4] + [1] = [5]$.
  \item $[5] + [1] = [0]$.
  \item $[5] + [2] = [1]$.
  \item $[5] + [3] = [2]$.
\end{itemize}

\end{frame}


\begin{frame}{Exponents}
\begin{itemize}
  \item If $[a]\in\Z_n$ and $k$ is a positive integer, then $[a]^k$ means
  \item $[a]\cdot[a]\cdots[a]$, where there are $k$ terms.
  \item Working in $\Z_5$...
  \item $[4]^6=[4]\cdot[4]\cdot[4]\cdot[4]\cdot[4]\cdot[4]$
  \item $=[1]\cdot[4]\cdot[4]\cdot[4]\cdot[4]$
  \item $=[1]\cdot[1]\cdot[4]\cdot[4]$
  \item $=[1]\cdot[1]\cdot[1]$
  \item $=[1]$
  \item Notice the exponent 6 is not in $\Z_5$. It is just a regular integer.
\end{itemize}
\end{frame}

\begin{frame}{Additive Inverses}
\begin{itemize}
  \item If $[a]$ and $[b]$ are in $\Z_n$
  \item and $[a]+[b] = [0]$
  \item then we say that $[a]$ and $[b]$ are additive inverses.
  \item We write $[a] = -[b]$ and $[b] = -[a]$.
  \item For example, working in $\Z_7$, $[3] + [4] = [0]$.
  \item So $[3] = -[4]$ and $[4] = -[3]$.
  \item In general, it is easy to see that, working in $\Z_n$,
  \item $-[a] = [n-a]$.
  \item Notice that $-(-[a]) = [a]$.
  \item Important: $-[a]$ is not an additional element of $\Z_n$. It is another name for one of the elements.
  \item For example, in $\Z_7$, $-[4]$ is another name for $[3]$. It's not an additional element.
\end{itemize}
\end{frame}

\begin{frame}{Additive Inverses and Negatives}
\begin{itemize}
  \item Notice that $-[a]_n = [-a]_n$.
  \item For example in $\Z_5$,
  \item $[2]_5 + [3]_5 = [0]_5$.
  \item So $-[2]_5 = [3]_5$.
  \item Now notice we also have $-2 \equiv 3 \pmod 5$, so $[-2]_5 = [3]_5$.
  \item So $-[2]_5 = [-2]_5 = [3]_5$.
  \item $[-2]_5$ is not an additional element of $\Z_5$. $[-2]_5$ is another name for $[3]_5$.
  \item $\Z_5=\singleton{[0]_5, [1]_5, [2]_5, [3]_5, [4]_5}$.
\end{itemize}
\end{frame}

\begin{frame}{Subtraction}
\begin{itemize}
  \item We can also define subtraction in $\Z_n$.
  \item $[a] - [b] = [a-b] = [(a-b) \bmod n]$.
  \item Equivalently
  \item $[a] - [b] = [a] + (-[b])$.
  \item For example, working in $\Z_9$...
  \item $[5] - [8] = [6]$.
  \item We can see this two different ways:
  \item $[5]-[8] = [5-8] = [-3] = [6]$
  \item $[5] - [8] = [5] + (-[8]) = [5] + [1] = [6]$.
  \item Notice that arithmetic in $\Z_n$ ``wraps around'' the bottom as well as the top.
  \item Working in $\Z_9$, $[0] - [1] = [8]$.
\end{itemize}
\end{frame}

\begin{frame}{$\Z_n$ is a commutative ring}
\begin{itemize}
  \item $[0]$ is an additive identity: $[a] + [0] = [a]$.
  \item Every element has an additive inverse.
  \item Addition is commutative: $[a]+[b]=[b]+[a]$.
  \item Addition is associative: $([a]+[b])+[c]=[a]+([b]+[c])$.
  \item $[1]$ is a multiplicative identity.
  \item Multiplication is commutative: $[a]\cdot[b]=[b]\cdot[a]$.
  \item Multiplication is associative: $([a]\cdot[b])\cdot[c]=[a]\cdot([b]\cdot[c])$.
  \item Distributive law: $[a]\cdot([b]+[c]) = ([a]\cdot[b])+([a]\cdot [c])$.
\end{itemize}
\end{frame}

\begin{frame}{Multiplying with negatives}
\begin{itemize}
  \item Multiplying with negatives works the same way it does with integers.
  \item $-[a] = -[1] \cdot [a]$.
  \item $(-[a])\cdot [b] = -([a]\cdot [b])$
  \item For example, working in $\Z_5$...
  \item Compute $-[2] \cdot [3]$.
  \item $-[2] \cdot [3] = -([2]\cdot[3]) = -[6] = -[1] = [4]$
\end{itemize}
\end{frame}

\begin{frame}{Negative one}
\begin{itemize}
  \item In any $\Z_n$, $-[1] = [n-1]$.
  \item For example, in $\Z_{100}$, $-[1] = [99]$.
  \item Example. Find $99^2 \bmod 100$.
  \item Work in $\Z_{100}$
  \item $[99] = -[1]$.
  \item $[99]^2 = (-[1])^2 = [1]$.
  \item So $99^2 \equiv 1 \pmod {100}$.
  \item i.e. $99^2 \bmod 100 = 1$.
\end{itemize}
\end{frame}

\begin{frame}{Multiplicative Inverses}
\begin{itemize}
  \item If $[a]$ and $[b]$ are in $\Z_n$
  \item and $[a]\cdot [b] = [1]$
  \item then $[a]$ and $[b]$ are called multiplicative inverses.
  \item We write $[a] = [b]^{-1}$ and $[b] = [a]^{-1}$.
  \item For example, working in $\Z_7$...
  \item $[2]\cdot[4] = [1]$.
  \item So $[2]^{-1} = [4]$ and $[4]^{-1} = [2]$.
  \item It's always true in any $\Z_n$ that $[1]\cdot[1] = [1]$,
  \item and that $-[1]\cdot-[1] = [1]$.
  \item So $[1]^{-1} = [1]$ and $(-[1])^{-1} = -[1]$.
  \item Working in $\Z_{100}$, $[99]^{-1}=[99]$.
\end{itemize}
\end{frame}

\begin{frame}{Units}
\begin{itemize}
  \item Sometimes $[a]^{-1}$ doesn't exist.
  \item If $[a]$ has a multiplicative inverse it is called a \emph{unit}.
  \item For example, working in $\Z_9$...
  \item $[4]\cdot [7] = [1]$ so $[4]^{-1} = [7]$ and $[7]^{-1} = [4]$.
  \item So in $\Z_9$, $[4]$ and $[7]$ are units.
  \item But $[3]$ doesn't have a multiplicative inverse in $\Z_9$.
  \item So $[3]$ is not a unit.
\end{itemize}
\end{frame}

\begin{frame}{Which are the units?}
\begin{itemize}
  \item Question: Which elements of $\Z_9$ are units?
  \item Notice that $[a]\cdot[x] = [1]$ in $\Z_9$ is the same as
  \item $ax \equiv 1 \pmod 9$.
  \item We know when there is a solution to this congruence.
  \item Answer: When $a$ is relatively prime to $9$.
  \item So the units of $\Z_9$ are: $[1],[2],[4],[5],[7],[8]$.
  \item In general:
  \item \textbf{Theorem.} $[a]$ is a unit of $\Z_n$ iff $\gcd(a, n) = 1$.
\end{itemize}
\end{frame}

\begin{frame}{Zero Divisors}
\begin{itemize}
  \item We work in $\Z_9$...
  \item Notice that $[3] \cdot [6] = [0]$.
  \item \textbf{Definition.} In any $\Z_n$ suppose $[a]$ and $[b]$ are non-zero
  \item and $[a]\cdot [b] = [0]$. Then $[a]$ and $[b]$ are called \emph{zero-divisors}.
  \item Which are the zero-divisors of $\Z_n$?
  \item $[a]$ is a zero-divisor of $\Z_n$ iff $a\not=0$ and
  \item there is a non-zero solution to $ax \equiv 0 \pmod n$.
  \item Let $d=\gcd(a,n)$. $d\divides 0$. So there are solutions.
  \item $x=0$ is one solution. $[a]$ is a zero-divisor iff there is a second solution.
  \item There are $d$ solutions. $[a]$ is a zero-divisor iff $d\not=1$.
  \item \textbf{Theorem.} $[a]$ is a zero-divisor of $\Z_n$ iff $a\not=0$ and $\gcd(a, n) \not= 1$.
\end{itemize}
\end{frame}

\begin{frame}{Zero Divisors and Units}
\begin{itemize}
  \item So every non-zero element of $\Z_n$ is either a unit or a zero-divisor.
  \item We work in $\Z_9$...
  \item $[1]$ is a unit. ($[1]$ is a unit in $\Z_n$ for every $n$.)
  \item $[2]$ is a unit. $[2]\cdot [5] = [1]$.
  \item $[3]$ is a zero-divisor. $[3]\cdot[6] = [0]$.
  \item $[4]$ is a unit. $[4]\cdot [7] = [1]$
  \item $[5]$ is a unit. $[5]\cdot [2] = [1]$
  \item $[6]$ is a zero-divisor. $[6]\cdot [3] = [0]$
  \item $[7]$ is a unit. $[7]\cdot [4] = [1]$
  \item $[8]$ is a unit. $[8]\cdot [8] = [1]$
  \item $[8] = -[1]$ and $-[1]$ is a unit in every $\Z_n$.
\end{itemize}
\end{frame}


\end{document}
