% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 7 \\ Fundamental Theorem of Arithmetic}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Prime Numbers}

\begin{itemize}
  \item \textbf{Definition.} Let $n>1$ be an integer. Then $n$ is \emph{prime} iff its only positive divisors are 1 and itself.
  \item If $n>1$ is not prime it is called \emph{composite}.
  \item Example: 11 is prime. 12 is composite.
  \item Notice that 1 and 0 are special cases. They are not considered prime or composite.
  \item We could define prime and composite for negative integers less than -1 but we will not.
\end{itemize}

\end{frame}


\begin{frame}{The Fundamental Theorem of Arithmetic}

\begin{itemize}
  \item \textbf{Theorem.} Every integer greater than 1 can be expressed as a product of primes.
  \item This \emph{prime decomposition} or \emph{prime factorization} is unique up to the order of the primes.
  \item Example: $12 = 2\cdot 2 \cdot 3$ is the unique prime factorization of 12 up to the order of the primes.
  \item What does up to the order of the primes mean?
  \item $12 = 2 \cdot 3 \cdot 2$.
\end{itemize}

\end{frame}

\begin{frame}{The Fundamental Theorem of Arithmetic}

\begin{itemize}
  \item We can extend the theorem to all non-zero integers by allowing an initial factor of $\pm 1$.
  \item If we say the primes have to be in increasing ordering we can get rid of ``up to the order of the primes.''
  \item \textbf{Theorem.} Let $n$ be a non-zero integer. Then $n$ has a unique prime decomposition:
  \item $n=\pm1 \cdot p_1^{e_1} \cdot p_2^{e_2} \cdots p_k^{e_k}$
  \item where each $p_i$ is prime and $p_1<p_2<\cdots < p_k$
  \item and each $e_i$ is a positive integer.
  \item Example: $12 = 1 \times 2^2 \times 3^1$.
  \item We will prove the Fundamental Theorem of Arithmetic over the course of the next few slides.
  \item This is an existence and uniqueness proof so there will be two directions. First we
  will prove that every integer has some prime decomposition.
\end{itemize}

\end{frame}

\begin{frame}{Fundamental Theorem of Arithmetic: Existence}

\begin{itemize}
  \item Every integer greater than 1 has some prime decomposition.
  \item \textbf{proof.} Towards a contradiction, suppose there are some integers greater than 1 that do not have a prime decomposition.
  \item Let $n$ be the least such integer. (There is a least such integer by the wellordering property.)
  \item Now $n$ cannot be prime because then $n=n$ is already a prime decomposition. So $n$ is composite.
  \item So $n = ab$ with $1<a<n$ and $1<b<n$.
  \item Since $a,b < n$, each of $a$ and $b$ have a prime decomposition.
  \item But $n=ab$ so the prime decomposition of $a$ followed by the prime decomposition of $b$ is a prime decomposition of $n$.
  \item Contradiction. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Fundamental Theorem of Arithmetic: Towards Uniqueness}
\begin{itemize}
  \item Towards being able to prove the uniqueness part of the Fundamental Theorem
  there are a few things we need to understand first.
  \item Recall that the following statement is \emph{not} true.
  \item Let $a,b,c$ be integers. Suppose $a\divides bc$. Then $a\divides b$ or $a\divides c$.
  \item Counter-example: Let $a=6, b=2, c=3$.
  \item However it turns out that the statement \emph{is} true if $a$ is prime.
  \item In fact this turns out to be a defining characteristic of primes.
  \item First we examine the case where $a$ is relatively prime to $b$.
\end{itemize}

\end{frame}

\begin{frame}{Relative Prime Divisor}

\begin{itemize}
  \item \textbf{Lemma.} Let $a,b,c$ be integers. Suppose $a\divides bc$ and $\gcd(a,b)=1$. Then $a\divides c$.
  \item Example: Let $a=6$, $b=5$, $c=12$.
  \item $6\divides 30 = 5\cdot 12$.
  \item $\gcd(6,5) = 1$
  \item The lemma tells us that $6\divides 12$.
  \item \textbf{proof.} Since $\gcd(a,b)=1$ there are integers $x$ and $y$ such that $xa + yb = 1$.
  \item Multiplying by $c$ we get that $xac + ybc = c$.
  \item Since $a\divides xac$ and $a\divides ybc$, $a\divides c$. $\qed$.
\end{itemize}

\end{frame}


\begin{frame}{Prime vs Relatively Prime}

\begin{itemize}
  \item Let $p$ be a prime and $b$ an integer greater than 1.
  \item Then exactly one of the following two things holds:
  \item (i) $p \divides b$, or
  \item (ii) $\gcd(p,b) = 1$ ($p$ and $b$ are relatively prime.)
  \item \textbf{proof.} If $p\divides b$ then $\gcd(p,b) = p$.
  \item If $p\ndivides b$
  then, since the only positive divisors of $p$ are 1 and $p$, the only common
  divisor of $p$ and $b$ is 1.
  \item So $\gcd(p,b) = 1. \qed$.
\end{itemize}

\end{frame}

\begin{frame}{Corollaries}

\begin{itemize}
  \item \textbf{Corollary.} Every two different primes are relatively prime.
  \item \textbf{Corollary.} If $p$ and $q$ are primes and $p\divides q$ then $p=q$.
  \item \textbf{Corollary} If $p$ is prime and $p\divides bc$ then $p\divides b$ or $p\divides c$.
  \item \textbf{proof} If $p\ndivides b$ then $p$ and $b$ are relatively prime, so $p\divides c$. $\qed$.
\end{itemize}
\end{frame}

\begin{frame}{Prime Dividing a Product}

\begin{itemize}
  \item \textbf{Theorem.} Suppose $p$ is prime and $b_1,b_2,\cdots b_k$ are positive integers.
  \item Suppose $p \divides (b_1 b_2 \cdots b_k)$.
  \item Then $p \divides b_1$ or $p\divides b_2$ or $\cdots$ $p\divides b_k$.
  \item \textbf{proof:} If $k=2$ this is the last Corollary from the previous slide.
  \item Let $k$ be least so that the Theorem fails to be true. So $k\geq 3$.
  \item $p\divides b_1(b_2\cdots b_k)$. So $p\divides b_1$ or $p\divides (b_2\cdots b_k)$.
  \item If $p\divides b_1$ we are done.
  \item Suppose $p\divides (b_2\cdots b_k)$.
  \item Since there are $k-1$ terms on the right and $k$ was
  least so that the Theorem failed, $p\divides b_2$ or $\cdots$ $p\divides b_k$. $\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Prime Dividing Product of Primes}

\begin{itemize}
  \item \textbf{Corollary.} Suppose $p, q_1, q_2, \cdots q_k$ are all primes.
  \item Suppose $p \divides (q_1 q_2 \cdots q_k)$.
  \item Then $p = q_1$ or $p=q_2$ or $\cdots$ $p=q_k$.
\end{itemize}
\end{frame}


\begin{frame}{Fundamental Theorem of Arithmetic: Proof of Uniqueness}
\begin{itemize}
  \item Let $n$ be an integer greater than 1. Suppose $n$ has two different
  prime decompositions:
  \item $n=p_1 p_2 \cdots p_s = q_1 q_2 \cdots q_t$.
  \item We will show that $s=t$ and the $p_i$ and $q_i$ are the same up to reordering.
  \item Suppose towards a contradiction that this were not true.
  \item Let $n$ be the least positive integer which has two inequivalent prime decompositions.
  \item (I'm calling two prime decompositions equivalent if one is just a rearrangement of the other.)
\end{itemize}
\end{frame}

\begin{frame}{Proof, continued}
\begin{itemize}
  \item $n=p_1 p_2 \cdots p_s = q_1 q_2 \cdots q_t$.
  \item Notice that $p_1 \divides (q_1 q_2 \cdots q_t)$.
  \item By the Corollary on the previous slide, $p_1 = q_i$ for some $i$ with $1\leq i \leq t$.
  \item For notational simplicity suppose $p_1=q_1$.
  \item Dividing both sides of the equation by $p_1 = q_1$ we get
  \item $\nprime = p_2 \cdots p_s = q_2 \cdots q_t$.
  \item Since $\nprime < n$, $\nprime$ does not have two inequivalent decompositions.
  \item But this contradicts the fact that we started with two inequivalent decompositions and canceled a common term from both sides. $\qed$.
\end{itemize}
\end{frame}

\end{document}
