\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,numprint}
\input{mathdefs}
\input{theoremstyles}

\npthousandsep{,}
\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Homework for Lesson 4 \\ Linear Transformations on $\R^n$}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Spring 2019 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.} Suppose $f:\R^n\map\R^m$ and $f$ is linear.
True or false? Explain your answer:
It is necessarily true that $f(\mathbf{0}_n) = \mathbf{0}_m$.
In other words $f$ maps the origin of $\R^n$ to the origin of $\R^m$.
(Hint: Consider what it means for $f$ to preserve scalar multiplication when
the scalar is zero.)

\bigskip
\bigskip

\textbf{Problem 2.} Let $f:\R^1\map\R^1$ be defined by
$$f(x) = 2x+1$$
Is $f$ linear? Explain why or why not. (Note that in $\R^1$ the vectors are
just the numbers so vector addition means just ordinary addition of numbers.)

\bigskip
\bigskip

\textbf{Problem 3.} Are constant functions linear?

\smallskip

\textbf{(a)} Let $f:\R^{2}\map\R^2$ be defined by $f(\bv) = \textbf{0}$ for
all $\bv$. Is $f$ linear? Why or why not?

\bigskip
\bigskip

\textbf{(b)} Let $f:\R^{2}\map\R^2$ be defined by $f(\bv) = (1,0)^{\text{T}}$ for
all $\bv$. Is $f$ linear? Why or why not?

\bigskip
\bigskip

\textbf{(c)} Let $c$ be a real number and
$f:\R^{2}\map\R^1$ be defined by $f(\bv) = c$ for
all $\bv$. For which values of $c$ is $f$ linear? Explain

\bigskip
\bigskip

\textbf{Problem 4.} Equivalent definitions of linear.

\smallskip

We defined
that a function $f$ is linear if it preserves vector addition and
preserves scalar multiplication.

\smallskip

Say that $f$ \emph{preserves linear combinations of length 2}
if for all vectors $\bv, \bw$ and all scalars $c, d$,
$$f(c\bv + d\bw) = c f(\bv) + d f (\bw)$$

\bigskip

\textbf{(a)} Suppose that $f$ preserves linear combinations of length 2.
Does it follow that $f$ is linear? Explain why or why not.

\bigskip
\bigskip

\textbf{(b)} Suppose that $f$ is linear. Does it follow that $f$ preserves
linear combinations of length 2? Explain why or why not.

\bigskip
\bigskip

Say that $f$ \emph{preserves all linear combinations} if
for all vectors $\bv_1,\bv_2 \cdots \bv_k$, and all scalars
$c_1, c_2, \cdots c_k$,

$$f(c_1 \bv_1 + c_2 \bv_2 + \cdots + c_k \bv_k)
= c_1 f(\bv_1) + c_2 f(\bv_2) + \cdots + c_k f(\bv_k)$$

\bigskip

\textbf{(c)} Suppose that $f$ is linear. Does it follow that $f$ preserves
all linear combinations? Explain why or why not.

\bigskip
\bigskip

\textbf{Problem 5.} Composition of linear functions.

\smallskip

Suppose $f:\R^n\map\R^m$ and $g:\R^m\map\R^k$ are both linear.
Let $h=g\circ f$ so that $h:\R^n\map\R^k$.

\smallskip

Is $h$ necessarily linear? Explain why or why not?

\bigskip
\bigskip

\textbf{Problem 6.}  Suppose $A$ is a matrix such that $T_A:\R^3\map\R^2$ and
$$
\begin{matrix}
T_A \begin{pmatrix} 0 \\ 0 \\ 1 \end{pmatrix} = \begin{pmatrix} 1 \\ 2 \end{pmatrix}\text{; and }
&
T_A \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} = \begin{pmatrix} 3 \\ 4 \end{pmatrix}\text{; and }
&
T_A \begin{pmatrix} 0 \\ 1 \\ 0 \end{pmatrix} = \begin{pmatrix} 5 \\ 6 \end{pmatrix}
\end{matrix}
$$

Give the matrix $A$.

\bigskip
\bigskip

\textbf{Problem 7.}  Suppose $T:\R^2\map\R^2$ is a linear transformation
that swaps the two standard basis vectors. That is, it
maps $\be_1$ to $\be_2$ and maps $\be_2$ to $\be_1$. Write down the matrix
for $T$.

\bigskip
\bigskip

\textbf{Problem 8.}  Let
$$
A = \begin{pmatrix}
9 & 8 & 7 \\
6 & 5 & 4 \\
3 & 2 & 1
\end{pmatrix}
$$

Where does $T_A$ send the third standard basis vector of $\R^3$?

\bigskip
\bigskip


\textbf{Problem 9.} Suppose $T:\R^2\map\R^2$ is a linear transformation and
$T(\be_1) = \begin{pmatrix} 1 \\ 2\end{pmatrix}$
and $T(\be_2) = \begin{pmatrix} 2 \\ 3\end{pmatrix}$. Then find
$T\begin{pmatrix} 3 \\ 4\end{pmatrix}$.


\bigskip
\bigskip


\end{document}
