\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Review of Beginning Inner Model Theory by Bill Mitchell}

\maketitle

The program of inner model theory is to define and study the properties of
canonical models of large cardinal axioms. The minimal and historically first
inner model is $L$, G\"{o}del's constructible universe, and this model serves
as the touchstone for what a ``canonical" model should be like. The smallest
large cardinals, for example weakly compact cardinals, can consistently exist
in $L$ and so $L$ may be considered the canonical inner model for these small
large cardinal axioms. But stronger large cardinal axioms are inconsistent
with ``$V=L$" and hence the program to find larger inner models is pursued.

We mention some of the properties of $L$ that contribute to it being considered
canonical. $L$ is a model of ZFC + GCH and the reason that AC and GCH are true
in $L$ can be explained by \emph{structural} properties of $L$. The axiom of choice
is true in $L$ because $L$ has a natural simply-definable global wellordering.
The GCH is true in $L$ because, working in $L$, if $\kappa\geq\omega$ is a cardinal
and $A\subseteq\kappa$  then $\exists \alpha<\kappa^{+}$ s.t.
$A\in L_{\alpha}$ (and $|L_{\alpha}| \leq \kappa$).
We will refer to this property as \emph{condensation} (though usually that term refers
to a different structural property that implies this property.)
$L$ is unique: If $M$ is a transitive proper class and $M\models$``$V=L$" then $M=L$.

The chapter under review begins in section 1.1 with a review of the theory of $L$
and goes on in later sections to describe, in varying levels of details, inner
models for progressively stronger large cardinal axioms: one measurable cardinal,
many measurable cardinals of higher Mitchell order, a strong cardinal. The theory
of the the last two models is given in some detail. Besides this main thread,
other important themes in inner model theory are intertwined in the narrative
of the chapter but in less detail: sharps, the covering lemma and the core model.
Besides describing the ideas mentioned above another stated goal of the chapter
is to prepare the reader for the next three chapters
in the Handbook which extend the conversation to include more
modern variants of the theory with stronger large cardinal axioms (Woodin cardinals
and beyond) and which describe the core model and covering theorems in more detail.
Bill Mitchell, the author of the chapter, is recognized as one of the founding
fathers of inner model theory. For this reason alone it is a pleasure to read
this chapter.

In the remainder of this review we will give some more details about the type of
material described in the chapter and we will point out a small error in one
of the definitions and give a correction.

Historically the next inner model to be studied after $L$, and the subject
of section 1.2 of the chapter, was $L[U]$, the canonical inner model
for a measurable cardinal. Here $U$ is a normal measure in $L[U]$ on some
$\kappa$ that is a measurable cardinal in $L[U]$. All of the canonical
properties of $L$ mentioned above generalize to $L[U]$ with some small provisos.
If $U^{\prime}$ is a normal measure in $L[U^{\prime}]$ on \emph{the same} $\kappa$
then $U^{\prime}=U$. Also $U$ is the only normal measure in $L[U]$ and $\kappa$ is the
only measurable cardinal in $L[U]$. $L[U]$ is not quite as unique as $L$ in that there
are other $L[U^{\prime}]$ models at other $\kappa^{\prime}$'s. But given any two such models one is
elementarily embeddable into the other. $L[U]$ is a model of GCH but the condensation
property is not true below $\kappa$ if we use the $L_{\alpha}[U]$ as levels.
This is becuase for $\alpha\leq\kappa$, $U\intersect L_{\alpha}[U]=\emptyset$ and
so $L_{\alpha}[U]=L_{\alpha}$ and so it is not reasonable to expect that
$L_{\alpha}[U]$ contains enough sets for condensation to hold. But there is a more
modern organization of the model in which partial information about $U$ is
used as a predicate in constructing the earlier levels of the model and using
this presentation we recover condensation. $L[U]$ has a definable wellorder but its
definition is more complex than the one for $L$. Restricted to reals the wellorder
for $L$ is $\Delta^1_2$ and the one for $L[U]$ is $\Delta^1_3$.

Historically the next step in the progression we are outlining, and the subject
of section 2, is the models of the form
$L[\vec{U}]$ where $\vec{U}$ is a coherent sequence of normal measures in
$L[\vec{U}]$. These models were introduced and studied by Bill Mitchell, the
author of the chapter under review. These models capture a class of large cardinal properties
stronger than a measurable limit of measurable cardinals, namely the property that
$\kappa$ is a measurable cardinal of higher Mitchell order.

If $U_1$ and $U_2$ are normal measures on $\kappa$ then $U_1 \vartriangleleft U_2$ iff
$U_1 \in \text{Ult}(V, U_2)$. $\vartriangleleft$ is wellfounded and $o(U)$ denotes the
rank of $U$ under this partial order. $o(\kappa) = \text{sup}
\setof{o(U) + 1}{U \text{ is a normal measure on } \kappa}$  is called
the Mitchell order of $\kappa$. So  $o(U) = 0$ if $U$
concentrates on non-measurable cardinals, $o(\kappa) = 1$ if $\kappa$ is measurable
but doesn't have a $U$ that concentrates on measurables, $o(U) = 1$ if $U$
concentrates on cardinals $\alpha$ with $o(\alpha) = 1$ and $o(\kappa) = 2$ if
$\kappa$ admits a $U$ with $o(U) = 1$ but not a $U$ that concentrates on cardinals
$\alpha$ that admit a measure of order 1. A sequence of measures $\vec{U}$ is
called \emph{coherent} if the measures are organized in a certain precise way
implying, among other things, that the measures on a fixed $\kappa$ are ordered
increasingly under $\vartriangleleft$.

Models of the form $L[\vec{U}]$ can satisfy
that there exists many measurable cardinals with many normal measures and
with Mitchell order up to $o(\kappa)=\kappa^{++}$. All of the canonical properties
of $L$ generalize to $L[\vec{U}]$ with some more provisos.
$L[\vec{U}]$ is a model of GCH but the situation with condensation is similar
to what was discussed above for $L[U]$. $L[\vec{U}]$ has a definable wellorder
which, when restricted to reals, is $\Delta^1_3$. (Many measurable cardinals is
not enough to increase the complexity of the wellorder beyond $\Delta^1_3$. That
does not occur until we get past the model with one Woodin cardinal.) As for uniqueness,
the only normal measures and the only measurable cardinals in $L[\vec{U}]$ are
those on the sequence $\vec{U}$. If $\vec{U_1}$ and $\vec{U_2}$ have the same
domain (meaning the same set of measurable cardinals and the same number
of measures on each cardinal), we would like to conclude that
$L[\vec{U_1}] = L[\vec{U_2}]$. This is true if the lengths of the sequences
are definable but it turns out that the hypothesis
$L[\vec{U_i}]\models \text{``}\vec{U_i} \text{ is coherent"}$
is not sufficient for the result in general. It is sufficient to assume that the $\vec{U_i}$ consist
of measures in $V$ (not just in the inner model) and that in $V$ the sequences satisfy a property called
\emph{weak coherence}. With this assumption it is true that $L[\vec{U_1}] = L[\vec{U_2}]$
and the restrictions of the two sequences to the inner model are equal.

But the definition of weakly coherent given in the chapter is not correct.
In Definition 2.6 on page 1466 the definition of a \emph{coherent} sequence $\vec{U}$
is given as
\begin{enumerate}
\item $\dom(\vec{U}) = \setof{(\kappa, \beta)}{\kappa < \length(\vec{U}) \AND \beta < o^{\vec{U}}(\kappa)}$,
where $\length(\vec{U})$ and  $o^{\vec{U}}(\kappa)$ are some ordinals.
\item If $(\kappa, \beta) \in \dom(\vec{U})$ then $\vec{U}(\kappa, \beta)$ is a normal measure on $\kappa$.
\item If $U=\vec{U}(\kappa, \beta)$ then $o^{i^{U}(\vec{U})}(\kappa) = \beta$ and $i^{U}(\vec{U})(\kappa, \betaprime)=
\vec{U}(\kappa, \betaprime)$ for all $\betaprime<\beta$, where $i^{U}$ is the ultrapower embedding by $U$.
\end{enumerate}


At the bottom of page 1469 a sequence $\vec{U}$ is defined to be \emph{weakly coherent}
if condition 3 above is replaced by the condition: If $U=\vec{U}(\kappa, \beta)$  then $o^{V}(U) = \beta$.

This condition is not sufficient for the purposes that the author will use it. For example Lemma 2.10 on
page 1470 states: Suppose that $\vec{U}$ and $\vec{W}$ are weakly coherent sequences of measures in $V$ with
the same domain. Then $L[\vec{U}] = L[\vec{W}]$ and $\vec{U}(\kappa,\beta)\intersect L[\vec{U}] =
\vec{W}(\kappa,\beta)\intersect L[\vec{W}]$ for every $(\kappa,\beta)$ in their common domain.

A simple example suffices to see that this fails using the definition of weakly coherent given above. Suppose
$o(\kappa) = 2$ and there exists two different normal measures on $\kappa$ of order 1,
$U_1\not=W_1$. Let $A\subset \kappa$ be in $U_1$ but
not in $W_1$ and consist only of measurable cardinals. There is such an $A$ since both $U_1$ and $W_1$ concentrate
on measurable cardinals.
 Let $\vec{U}$ and $\vec{W}$ be sequences of measures with common domain
$\setof{(\alpha, 0)}{\alpha \in A} \union \singleton{(\kappa, 0), (\kappa, 1)}$. Let
$\vec{U}(\alpha, 0) = \vec{W}(\alpha, 0)$ be any measure of order zero on $\alpha$ for $\alpha \in A$.
Let $\vec{U}(\kappa, 0) = \vec{W}(\kappa, 0)$ be any measure of order zero on $\kappa$. Finally
let $\vec{U}(\kappa, 1) = U_1$ and $\vec{W}(\kappa, 1) = W_1$. Then $\vec{U}$ and $\vec{W}$ satisfy
the definition of weakly coherent given in the chapter but, since $A\in L[\vec{U}]\intersect L[\vec{W}]$,
$U_1\intersect L[\vec{U}] \not= W_1\intersect L[\vec{W}]$.

A correct definition of weakly coherent may be found in Mitchell's own earlier \cite{Mitchell-Revisited}.
There a sequence $\vec{U}$ is defined to be weakly coherent
if condition 3 above is replaced by just the first half of condition 3:
If $U=\vec{U}(\kappa, \beta)$ then $o^{i^{U}(\vec{U})}(\kappa) = \beta$.
It is shown in \cite{Mitchell-Revisited} that this
condition suffices for the proof of Lemma 2.10.
Some authors use the notation
$o^{\vec{U}}(U) = \beta$ for this condition (but the chapter under review does not introduce this notation.)
Thus the correction may be construed as consisting of changing a single symbol:
For weak coherence, instead of $o^V(U) = \beta$ we want $o^{\vec{U}}(U) = \beta$. The reason our
counterexample above fails to satisfy this correct definition of weakly coherent is that although
$o^{V}(W_1) = 1$, $o^{\vec{W}}(W_1) = 0$.

In \cite{Mitchell-Revisited} it is pointed out that an example of a weakly coherent sequence
$\vec{U}$ (in the correct sense) is given by letting, for every measurable cardinal $\alpha$ and every $\beta<o(\alpha)$,
$\vec{U}(\alpha,\beta)$ be any normal measure on $\alpha$ of order $\beta$. Notice that in
this example $o^{\vec{U}}(\alpha) = o^V(\alpha)$ for every $\alpha$ and so this example does
actually satisfy the definition of weak coherence given in the chapter. The reason that our
counterexample above differs from this example is that the set $A$ did not contain all measurable
cardinals below $\kappa$.

Although this point is not explained in detail in the chapter,
it is explained in \cite{Mitchell-Revisited}
that the motivation for considering weakly coherent sequences is the following problem:
Given a measurable cardinal $\kappa$ of high Mitchell order in $V$, construct a sequence $\vec{U}$ such that
in $L[\vec{U}]$ $U$ is coherent and $\kappa$ has high Mitchell order. In the case of
a single measurable cardinal the analogous problem was easy. If $\kappa$ is measureable let $U$ be a normal
measure on $\kappa$ in $V$. Then $U\intersect L[U]$ is a normal measure on $\kappa$ in
$L[U]$. But for $\beta>=\omega$ it is unknown whether the hypothesis $o(\kappa) = \beta$
implies that there is a coherent $\vec{U}$ such $o^{\vec{U}}(\kappa) = \beta$.
But the example from the previous paragraph shows that it is possible to construct a weakly coherent
sequence $\vec{U}$ with $o^{\vec{U}}(\kappa) = \beta$.
This explains the motivation behind Corollary 2.11 of the chapter which says that either
this $\vec{U}$ is coherent in $L[\vec{U}]$ or else there is an inner model
of $\exists \kappa (o(\kappa) = \kappa^{++})$.


\bibliographystyle{amsalpha}

\bibliography{math}

\end{document}
