% $Header$

\documentclass{beamer}
% \documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 0 \\ Introduction \\ Interplay of Concrete and Abstract}
\subtitle{Math 325, Linear Algebra \\ Fall 2018 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}


\begin{frame}{Preview}
\begin{itemize}
\item This intro is a course preview
\item Not tested on this preview.
\item Not expected to understand all of the math in this preview.
\item Just to give you a sense of where we are headed.
\item Will whip through some ideas very quickly.
\item Don't worry, we'll go through it more slowly later.
\end{itemize}
\end{frame}

\begin{frame}{Concrete problems}

  \begin{itemize}
  \item  Main theme: interplay between
  the \emph{concrete} and the \emph{abstract} in mathematics.
  \item What do I mean by concrete?
  \item A concrete problem focuses on a specific example.
  \item It doesn't try to understand a more general situation.
  \item It often uses actual numbers and the answer is often a number.
  \end{itemize}

\end{frame}

\begin{frame}{Abstract theory}

  \begin{itemize}
  \item What do I mean by abstract?
  \item An abstract theory tries to analyze a general class of things rather than a single specific example.
  \item The goal is to understand how systems work at a more fundamental level.
  \item In abstract mathematics we are more interested in proving theorems than in finding the
  answer to any particular numeric problem.
  \item The style of abstract mathematics is definition-theorem-proof.
  \end{itemize}

\end{frame}

\begin{frame}{Example of a concrete problem}

Consider the following system of equations:

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

\begin{itemize}
\item There are 5 \emph{unknowns}: $x, y, z, u, v$.
\item There are 4 \emph{equations}.
\item The numbers 2, 8, 13, -3 constitute the \emph{right-hand-side}
\item The goal is to find values for the unknowns in order to make all of the
equations simultaneously true.
\end{itemize}

\end{frame}

\begin{frame}{Many details}

This is an example of a concrete problem.

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

\begin{itemize}
\item Sometimes concrete problems are hard to think about because they make you focus on too many details.
\item The details make the problem feel complicated.
\item The idea of abstraction is: Instead of trying to solve this particular problem right away,
let's try to understand a whole class of problems similar to this one.
\item Does this move make our job harder or easier?
\end{itemize}

\end{frame}

\begin{frame}{Moving towards abstraction}

\begin{align*}
 x + 3y + 2z - u  \qquad &= 2 \\
2x + 6y + z + 4u + 3v  &= 8 \\
-x -3y  -3z + 3u + v  &= 13 \\
3x + 9y + 8z  -7u + 2v  &= -3 \\
\end{align*}

In this course we will learn how to express this system as a \emph{matrix equation}: $A \bx = \bb$.

$$
\begin{matrix}
A =
\begin{pmatrix}
1 & 3 & 2 & -1 & 0 \\
2 & 6 & 1 & 4 & 3\\
-1 & -3 & -3 & 3 & 1 \\
3 & 9 & 8 & -7 & 2
\end{pmatrix}
&
\bx =
\begin{pmatrix}
x \\ y \\ z \\ u \\ v
\end{pmatrix}
&
\bb =
\begin{pmatrix}
2 \\ 8 \\ 13 \\ -3
\end{pmatrix}
\end{matrix}
$$

\end{frame}

\begin{frame} {Less detail}
This version of the problem has less detail:

\pause

Find a \emph{vector} $\bx$ such that

$$A \bx = \bb.$$
\end{frame}

\begin{frame} {More abstract}
This version of the problem is more abstract:

\pause

Given \emph{a} matrix $A$ and \emph{a} vector $\bb$ find \emph{all} vectors $\bx$ such that

$$A \bx = \bb.$$

\pause

By making this move to a more general problem we have entered into matrix \emph{theory}.
\end{frame}

\begin{frame} {Why make the move to abstraction?}
\begin{itemize}
\item The original concrete problem is not that hard.
\item If our only goal was to learn how to solve that problem, maybe it would
not be worth it to make the move to abstraction, the move to matrix
theory.
\item It would be more efficient to just learn a technique for solving that problem.
\item By making the move to abstraction, we hope to get a deeper understanding
of what is going on.
\item This deeper understanding will then help us understand how to solve
other concrete problems.
\end{itemize}
\end{frame}

\begin{frame} {Example}
\begin{itemize}
\item Suppose it turns out that the original system of 4 equations in 5 unknowns
does not have any solutions?
\item What should we do next?  Should we give up?
\item If we didn't study abstract theory we might
not have any idea what to do next.
\item Studying abstract theory gives a suggestion for what to do.
\item To see this, we have to go even more abstract...
\end{itemize}
\end{frame}


\begin{frame} {Even more abstract}
\begin{itemize}
\item We are going to learn that associated with every matrix $A$ there is
a \emph{function} called $T_A$.
\item $T_A$ is called the \emph{linear transformation} associated with the
matrix $A$.
\item For the matrix that came from our earlier system of 4 equations and 5 unknowns,
$T_A$ is a function that takes input vectors in $\R^5$ and gives output vectors in $\R^4$.
\item We are going to learn that solving the original concrete problem is equivalent
to looking for an input $\bx$ in $\R^5$ such that $T_A(\bx) = \bb$.
\item If there is no solution to the original system of equations that is
equivalent to saying that there is no input vector $\bx$ in $\R^5$ such that $T_A(\bx) = \bb$.
\end{itemize}
\end{frame}


\begin{frame} {Abstraction leads to a way to proceed}
\begin{itemize}
\item By thinking of the problem in terms of the function $T_A$ we are lead
to pose this follow-up problem:
\item Find an input vector $\bx$ in $\R^5$ so that $T_A(\bx)$ is as \emph{close} as possible
in $\R^4$ to $\bb$?
\item We will learn that finding such an input vector $\bx$ in $\R^5$ is equivalent to finding the
best \emph{approximate} solution to the original system of equations.
\item The move to abstraction gave us the idea of finding the best approximate solution to the
original system of equations and suggested a technique for finding it.
\item We might not have thought of this if we stuck with the concrete original problem.
\end{itemize}
\end{frame}

\begin{frame}{The interplay of the abstract and the concrete}
\begin{itemize}
\item Concrete problems suggest new areas where we can develop abstract theory.
\item Abstract theory gives us deeper insight into solving concrete problems.
\item Some people are more excited by learning how to solve concrete problems.
\item Some people are more excited by learning abstract theory.
\item The subject of Linear Algebra has something exciting for everyone!
\end{itemize}
\end{frame}



\end{document}


