% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 13 \\ Vector Spaces}
\subtitle{Math 325, Linear Algebra \\Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Going more abstract}

\begin{itemize}
\item In this lesson we will define the notion of a \emph{vector space}.
\item This is a generalization of the notion of vectors in $\R^n$.
\item So we are going  more abstract.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Definition of Real Vector Space}

\begin{itemize}
\item \textbf{Definition.} A \emph{vector space} over $\R$, also called
a \emph{real vector space}, consists of the following:
\item A \emph{set of objects} $V$ (that we will call \emph{vectors}),
\item Two \emph{operations} on the vectors:
\item \emph{addition}. If $\bv$ and $\bw$ are vectors in $V$ then
we know how to form their sum $\bv+\bw$.
\item \emph{scalar multiplication}. If $\bv$ is a vector in $V$ and
$c\in\R$ then we know how to form the product $c\bv$.
\item The two operations satisfy the following properties:
\item Vector addition is commutative and associative.
\item There is a zero vector $\bzero$ that is an additive identity.
\item For each vector $\bv$ there is an additive inverse vector $-\bv$.
\item Scalar multiplication is associative and distributive over vector addition.
\item For any vector $\bv$, $1\bv = \bv$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: $\R^n$ is a real vector space.}

\begin{itemize}
\item This is not surprising because we intentionally defined \emph{vector space}
to be a generalization of $\R^n$.
\item We already know that if $\bv, \bw, \bu$ are vectors in $\R^n$ and
$c,d\in\R$ then:
\item $\bv+\bw = \bw + \bv$
\item $(\bv + \bw) + \bu = \bv + (\bw + \bu)$
\item $\bv + \textbf{0} = \textbf{0} + \bv = \bv$.
\item $\bv + (-\bv) = \textbf{0}$
\item $c(\bv + \bw) = c\bv + c\bw$
\item $(c+d)(\bv) = c\bv + d\bv$
\item $c(d\bv) = (c d) \bv$.
\item $1 \bv = \bv$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The zero vector and the zero scalar}

\begin{itemize}
\item \textbf{Lemma.} Let $V$ be a vector space and let $\bv \in V$ be
any vector and let $c$ be any scalar. Then
\item \textbf{(a)} $0 \bv = \bzero$.
\item \textbf{(b)} $c \bzero = \bzero$.
\item \textbf{proof.} \textbf{(a)}. Let $\bw = 0\bv$. We must show $\bw=\bzero$.
\item $\bw = 0\bv = (0+0)\bv = 0\bv + 0\bv = \bw+\bw$.
\item So $\bw = \bw + \bw$.
\item Subtracting $\bw$ from both sides gives $\bzero = \bw$.
\item \textbf{(b)}. $c\bzero = c(\bzero + \bzero) = c\bzero + c\bzero$.
\item So as above, $c\bzero = \bzero$. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Subspaces}

\begin{itemize}
\item \textbf{Definition.} Let $V$ be a vector space. Then a \emph{subspace}
of $V$ is
\item  a subset $W\subseteq V$ that contains $\bzero$ and is closed
under vector addition and scalar multiplication:
\item $W\subseteq V$
\item $\bzero \in W$.
\item If $\bv,\bw\in W$ then $\bv+\bw\in W$.
\item If $\bv\in W$ and $c\in \R$ then $c\bv\in W$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Subspaces are vector spaces}

\begin{itemize}
\item \textbf{Lemma.} $W$ is also a vector space, under the same operations
as $\bv$.
\item \textbf{proof.}
\item For example, to see that vector addition is commutative, let
$\bv,\bw\in W$.
\item Well then $\bv,\bw\in V$ and $V$ is a vector space so $\bv+\bw=\bw+\bv$
when considered as vectors in $V$.
\item Because $W$ is a subspace $\bv + \bw\in W$, and $\bw + \bv\in W$,
and vector addition in $W$ is the same as vector addition in $V$, so
$\bv+\bw=\bw+\bv$ in $W$ too.
\item The argument for all of the other properties of being a vector space
is similar. $\qed$.
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: the $x$-,$y$-plane in $\R^3$}

\begin{itemize}
\item Consider the vector space $V = \R^3$.
\item Let $W $ be the $x$-,$y$-plane in $\R^3$.
\item $W = \setof{(x,y,z)^{\text{T}}}{z = 0}$.
\item \textbf{Claim.} $W$ is a subspace of $V$.
\item \textbf{proof.} We need to see that $\textbf{0}\in W$ and that
$W$ is closed under vector
addition and scalar multiplication.
\item $\textbf{0} \in W$ because the origin is in the $x$-,$y$-plane.
\item Let $\bv,\bw\in W$, we need to see that $\bv+\bw\in W$.
\item Write $\bv=(x_0, y_0, 0)^{\text{T}}$, $\bw = (x_1, y_1,0)^{\text{T}}$.
\item So $\bv+\bw = (x_0 + x_1, y_0 + y_1, 0)^{\text{T}}$.
\item So $\bv+\bw \in W$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof of claim, continued.}

\begin{itemize}
\item Let $c\in \R$. We need to see that $c \bv\in W$.
\item $c\bv = c(x_0, y_0, 0)^{\text{T}}$
\item $=(cx_0, cy_0, 0)^{\text{T}}$
\item So $c\bv\in W$.
\item So $W$ is closed under both vector addition and scalar multiplication.
\item So $W$ is a subspace. $\qed$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: The line $y=x$ in $\R^2$}

\begin{itemize}
\item Consider the vector space $V = \R^2$.
\item Let $W$ be the set of points on the line $y=x$
\item $W=\setof{(x,y)^{\text{T}}}{x=y}$.
\item \textbf{Claim.} $W$ is a subspace of $V$.
\item \textbf{proof.} We need to see that $\textbf{0}\in W$ and that
$W$ is closed under vector
addition and scalar multiplication.
\item $\textbf{0} \in W$ because the origin is on the line $y=x$.
\item Let $\bv,\bw\in W$, we need to see that $\bv+\bw\in W$.
\item Write $\bv=(a,a)^{\text{T}}$, $\bw = (b,b)^{\text{T}}$.
\item So $\bv+\bw = (a+b,a+b)^{\text{T}}$.
\item So $\bv+\bw \in W$.
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Proof of claim, continued.}

\begin{itemize}
\item Let $c\in \R$. We need to see that $c \bv\in W$.
\item $c\bv = c(a,a)^{\text{T}}$
\item $=(ca,ca)^{\text{T}}$
\item So $c\bv\in W$.
\item So $W$ is closed under both vector addition and scalar multiplication.
\item So $W$ is a subspace. $\qed$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: The Zero Subspace}

\begin{itemize}
\item Let $V$ be any vector space.
\item Let $W=\singleton{\bzero}$.
\item \textbf{Claim.} $W$ is a subspace of $v$, called the zero subspace.
\item \textbf{proof.} Clearly $W\subseteq V$ and $\bzero\in W$.
\item $W$ is closed under addition since $\bzero+\bzero=\bzero$.
\item $W$ is closed under scalar multiplication since $c\bzero=\bzero$.
\item So $W$ is a subspace. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Example: Not a sub-space}

\begin{itemize}
\item Let $V$ be any vector space.
\item Let $\bv\in V$, $\bv\not=\bzero$.
\item Let $W=\singleton{\bv}$.
\item Then $W$ is \emph{not} a subspace.
\item Why?
\end{itemize}
\end{frame}


\end{document}


