\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Homework Solutions, Week 5 \\ Math 310, Elementary Number Theory \\ Fall 2019}
\author{Mitch Rudominer}

\maketitle

\textbf{Solution to Exercise 1.12 from Shoup} Show that two integers are relatively prime iff there is no one prime that divides both of them.
\begin{proof}
Let $a,b\in \Z$. First assume that $a$ and $b$ are relatively prime.
Then the only positive common divisor of $a$ and $b$ is 1, so in particular $a$ and $b$ have no common prime divisor.

Conversely, suppose that $a$ and $b$ are not relatively prime. Then they have some positive common divisor $d$. Let $p$ be any prime that divides $d$. Then $d$ divides both $a$ and $b$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 1.15 from Shoup} An integer $a$ is called \textbf{square-free} if it is not divisible by the square of any integer greater than 1.

(a) $a$ is square-free iff $a=\pm p_1 \cdots p_r$, where the $p_i$'s are distinct primes.

\begin{proof}
Let $a=\pm p_1^{e_1} \cdots p_r^{e_r}$ be the prime factorization of $a$.
The condition on the right-hand-side is equivalent to $e_1=\cdots=e_r=1$.

Suppose $e_i\geq 2$ for some $i$. Then $p_i^2\divides a$ so $a$ is not square-free.

Conversely, suppose $d$ is positive and $d^2\divides a$. Let $p$ be a prime
divisor of $d$. Then $p$ is a prime divisor of $a$ so $p=p_i$ for some $i$.
Since $d^2\divides a$, $p^2\divides a$,so $e_i\geq 2$.
\end{proof}

\medskip

(b) Every positive integer $n$ can be expressed uniquely as $n=ab^2$, where $a$ and $b$ are positive integers, and $a$ is square-free.

\begin{proof}
Let $n$ be a positive integer and let $n=p_1^{e_1}\cdots p_r^{e_r}$
be the prime factorization of $n$, with the $p_i$ distinct. For each $i$,
if $e_i$ is even let $f_i$ be such that $e_i=2f_i$ and if $e_i$ is odd
let $f_i$ be such that $e_i=2f_i + 1$.

Let $a$ be the product of all of those $p_i$ such that $e_i$ is odd. Since
$a$ is a product of distinct primes, $a$ is square free.

Let $b=p_1^{f_1}\cdots p_r^{f_r}$.

Then $b^2=p_1^{2f_1}\cdots p_r^{2f_r}$ and 
$ab^2=p_1^{e_1}\cdots p_r^{e_r} = n$.

To show uniqueness, suppose that $n=ab^2$ where $a$ and $b$ are positive integers and $a$ is square-free. We will show that $a$ and $b$ must be as we just constructed above. Let $p_i$ be a prime factor of $n$ and let $e_i$ be the exponent of $p_i$ in the prime factorization of $n$. Let $f_i$ be the exponent of $p_i$ in the prime factorization of $b$ and let $g_i$ be the exponent of $p_i$ in the prime factorization of $a$. Here $f_i$ and $g_i$ are both allowed to be 0. Since $a$ is square free, $g_i$ is either zero or one.
We see that $e_i = 2f_i + g_i$.  Thus $g_i=1$ iff $e_i$ is odd. Thus $a$ is
the product of the prime factors of $n$ whose exponents are odd. In
other words, $a$ is the same $a$ that we constructed above. But then $b^2=n/a$ so $b$ is the same $b$ that we constructed above. This proves uniqueness.
\end{proof}

\bigskip

\textbf{Solution to Exercise 6.6}

\begin{enumerate}
\item[(a)] $\nu_5(100) = 2$ because $100=2^2 \cdot 5^2$.
\item[(b)] $\nu_5(-100) = 2$ because $-100=-1 \cdot 2^2 \cdot 5^2$.
\item[(c)] We are asked to find the smallest positive integers $a,b,c$ such that 
$\nu_2(a)=b$ and $\nu_3(b)=c$ and $\nu_2(c) > \nu_3(c)$. The smallest positive $c$ such
that $\nu_2(c) > \nu_3(c)$ is $c=2$. For this $c$ we have 
$\nu_2(c) = 1, \nu_3(c)=0$. The smallest positive integer $b$ such that 
$\nu_3(b)=2$ is $b=3^2=9$. The smallest positive integer $a$ such that
$\nu_2(a)=9$  is $a=2^9=512$.
\end{enumerate}

\bigskip

\textbf{Solution to Exercise 1.20 from Shoup} Let $a,b,c\in\Z$.
\begin{enumerate}
\item[(a)] $\lcm(a,b)=\lcm(b,a)$
\begin{subproof}
This is trivial because being a ``common multiple of $a$ and $b$'' is the same thing
as being a ``common multiple of $b$ and $a$.''
\end{subproof}
\item[(b)] $\lcm(a,b)=\av{a} \Iff b\divides a$.
\begin{subproof}
First consider the case $a\geq 0$ so we can ignore the absolute value signs. 

We will give two proofs of the fact that $\lcm(a,b) = a \Iff b\divides a$.

For the first proof we use the $\nu$ function.

\begin{align*}
\lcm(a,b) = a &\Iff \prod_p p^{\max\left(\nu_p(a)\nu_p(b)\right)} = \prod_p p^{\nu_p(a)} \\
              &\Iff \forall p \, \nu_p(a)\geq \nu_p(b) \\
              &\Iff b\divides a
\end{align*}


For the second proof of the fact $\lcm(a,b) = a \Iff b\divides a$ we directly use
divisibility.

If $a=\lcm(a,b)$ then in particular $a$ is a multiple of $b$ so $b\divides a$. Conversely if
$b\divides a$ then $a$ is a common multiple of $a$ and $b$ and $a$ divides every common
multiple of $a$ and $b$ so $a=\lcm(a,b)$.
 
Now suppose $a<0$ so $\av{a}=-a$. Notice $a$ and $-a$ have the same multiples. 
So $\lcm(a,b) = \av{a} \Iff \lcm(\av{a},b) = \av{a} \Iff b\divides\av{a} \Iff b\divides a$.
\end{subproof}

\item[(c)] $\lcm(a,a)=\lcm(a,1)=\av{a}$.
\begin{subproof}
The first equality is trivial becase a ``common multiple of $a$ and $a$'' is the same
thing as a multiple of $a$ which is the same thing as a ``common multiple of $a$ and 1'' 
since every integer is a multiple of 1.

For the second equality, first take the case that $a\geq 0$ so we can ignore the absolute
value signs. Then clearly $a$ itself is
the least non-negative multiple of $a$ so $a = \lcm(a,a)$.

Now take the case that $a$ is negative. Then clearly $\av{a}$ is the least positive
multiple of $a$ and so the least common multiple of $a$ and 1.
\end{subproof}

\item[(d)] $\lcm(ca,cb) = \av{c}\lcm(a,b)$.
\begin{subproof}
First lets consider the case $c\geq 0$ so we can ignore the absolute value signs.

We give two proofs of the fact that $\lcm(ca,cb) = c\lcm(a,b)$.

For the first proof we use the $\nu$ function.

\begin{align*}
\lcm(ca,cb)  &=  \prod_p p^{\max\left(\nu_p(ca),\nu_p(cb)\right)} \\
             &=  \prod_p p^{\max\left(\nu_p(c)+\nu_p(a),\nu_p(c) + \nu_p(b)\right)} \\
             &=  \prod_p p^{\nu_p(c) + \max(\nu_p(a),\nu_p(b))} \\
             &=  \prod_p p^{\nu_p(c)} \cdot \prod_p p^{\max(\nu_p(a),\nu_p(b))}\\
             &=  c \lcm(a,b) \\
\end{align*}

For the second proof of the fact that $\lcm(ca,cb) = c\lcm(a,b)$ we use straight divisibility.

Let $m=\lcm(a,b)$. So $ca\divides cm$ and $cb\divides cm$ and so $cm$ is a common multiple of $ca$ and $cb$.

Let $k$ be any common multiple of $ca$ and $cb$. So $k=cas=cbt$, for some integers $s,t$. 
Cancelling $c$ we get $as=bt$ is a common multiple of $a$ and $b$ and so a multiple of $m$: $as=bt=mr$
for some $r$. So $k=cmr$ so $cm\divides k$. We have shown that $m$ is a common multiple of $a$ and $b$ and that every common multiple of 
$a$ and $b$ divides m, so $m=\lcm(a,b)$.

Now suppose $c$ is negative. Then $\lcm(ca,cb) = \lcm(-ca,-cb) = -c \lcm(a,b) = \av{c} \lcm(a,b)$.
\end{subproof}

\end{enumerate}


\bigskip

\textbf{Solution to Exercise 1.21 from Shoup} Show that for all $a,b\in\Z$ we have:
\begin{enumerate}
\item[(a)] $\gcd(a,b) \cdot \lcm(a,b) = \av{ab}$.

If $a$ or $b$ are zero then both sides of the equation are zero. So now assume that $a$ and $b$ are non-zero.

First assume that $a$ and $b$ are positive so we can ignore the absolute value sign.

We will give two proofs of the fact that $\gcd(a,b) \cdot \lcm(a,b) = ab$.

For the first proof we use the $\nu$ function.

\begin{align*}
\gcd(a,c) \cdot \lcm(a,b)  &=  \prod_p p^{\min\left(\nu_p(a),\nu_p(b)\right)}\cdot \prod_p p^{\max\left(\nu_p(a),\nu_p(b)\right)} \\
           &= \prod_p p^{\min\left(\nu_p(a),\nu_p(b)\right)+\max\left(\nu_p(a),\nu_p(b)\right)} \\
           &= \prod_p p^{\nu_p(a) + \nu_p(b)} \\
           &= ab \\
\end{align*}


For the second proof of the fact that $\gcd(a,b) \cdot \lcm(a,b) = ab$ when $a,b>0$ we will use straight divisibility.
First we handle the special case that $a$ and $b$ are relatively prime.

\textbf{Lemma} Suppose $a$ and $b$ are relatively prime and positive. Then $\lcm(a,b) = ab$.
\begin{subproof}
Clearly $ab$ is a common multiple of $a$ and $b$. To see that it is the least common multiple, let
$k$ be any common multiple of $a$ and $b$. So there are $s$ and $t$ such that $k=sa=tb$.
Since $a\divides (sa)$, $a\divides (tb)$. Since $a$ is relatively prime to $b$, $a\divides t$. 
So write $t=ar$. Then $k=arb$. So $ab\divides k$. We have shown that $ab$ divides every common multiple of 
$a$ and $b$ so $ab=\lcm(a,b)$.
\end{subproof}

Now we consider the general case where $a$ and $b$ are not necessarily relatively prime. We want to show that
$\gcd(a,b) \cdot \lcm(a,b) = ab$ when $a,b>0$.

Let $d=\gcd(a,b)$. Since $a,b>0$, $d>0$.


Since $d\divides a$ and $d\divides b$ let $\abar, \bbar$ be such that
$a=d\abar$, $b=d\bbar$. 

By Exercise 1.10 from Shoup, $\abar$ and $\bbar$ are relatively prime.
Let $m=d\abar\bbar = ab/d$. We must show that $m=\lcm(a,b)$.

By the Lemma we just proved $\lcm(\abar,\bbar) = \abar \bbar$.

By part (d) of the previous exercise $\lcm(d\abar,d\bbar) = d\lcm(\abar,\bbar)$.
So $\lcm(a,b) = d\abar\bbar=m$.

If $a<0$ and $b\geq 0$ then $\lcm(a,b) = \lcm(-a,b) = \gcd(-a,b) (-ab) = \gcd(a,b) \av{ab}$.

Similarly if $b<0$ and $a\geq 0$.

If $a<0$ and $b<0$ then $\lcm(a,b) = \lcm(-a,-b) = \gcd(-a,-b) (-a\cdot -b) = \gcd(a,b) \av{ab}$.

\item[(b)] This part follows immediately from part (a).
\end{enumerate}

\bigskip

\textbf{Solution to Exercise 1.27 from Shoup.} For all $a,b\in\Z$,
$$\gcd(a+b,\lcm(a,b)) = \gcd(a,b).$$
\begin{proof}
First we handle the case that $a$ and $b$ are relatively prime. In this case we must show that
$a+b$ and $ab$ are relativley prime. This is true because if $p$ is a prime and $p$ divides
both $ab$ and $a+b$ then $p$ divides one of $a$ or $b$, but then since $p$ divides $a+b$ it
divides both $a$ and $b$, which is impossible since $a$ and $b$ are relatively prime.

Now let $a,b$ be arbitrary integers. Let $d=\gcd(a,b)$ and let $\abar,\bbar$ be such that
$a=d\abar,b=d\bbar$. Then $\abar,\bbar$ are relatively prime and letting $m=d\abar\bbar$ we have
that $m=\lcm(a,b)$. By what we have just shown,
$\gcd(\abar+\bbar,\abar\bbar) = 1$. Multiplying by $d$ and using a fact from an exercise we proved earlier,
$\gcd(d\abar+d\bbar,d\abar\bbar)=d$. So $\gcd(a+b,m)=d$ as was to be shown.
\end{proof}

\bigskip

\textbf{Solution to Exercise 1.28 from Shoup.} Show that for every positive integer $k$, there exists $k$
consecutive composite integers. Thus, there are arbitrarily loarge gaps
between primes.

\begin{note}
Do not feel bad if you were not able to solve this problem on your own--it required a flash of insight and
was not at all obvious. I expected you would ask for a hint. However it is important that you do understand
the proof now.
\end{note}

\begin{proof}
For every integer $k\geq 3$, consider the interval of integers ${[k!+2, k!+k]}$. This interval has length $k-1$.
We claim that every integer in that interval is composite. In fact ever integer in that interval is greater than $k$,
and we will show that every integer in that interval is divisible by an integer less than $k$ and greater than 1.
Every integer in that interval is of the form $k!+n$ for some $n$ with $1< n \leq k$. Since $n\divides (k!)$
and $n\divides n$, $n\divides (k!+n)$. Thus $(k!+n)$ is composite.
\end{proof}

\bigskip

\textbf{Solution to Exercise 1.29 from Shoup.} Let $p$ be prime. Show that for all $a,b\in\Z$ we have
$\nu_p(a+b)\geq\min\left\{\nu_p(a),\nu_p(b)\right\}$ and $\nu_p(a+b)=\nu_p(a)$ if $\nu_p(a)<\nu_p(b)$.

\begin{proof}
Let $N=\nu_p(a)$ and $M=\nu_p(b)$ and suppose without loss of generality that $N\leq M$.
Let $\abar,\bbar$ be such that $a=p^N\abar$, $b=p^M\bbar$, $p\ndivides\abar$, $p\ndivides\bbar$.
Then 
$$a+b=p^N\left(\abar+p^{M-N}\bbar\right).$$

 So $p^N\divides(a+b)$. So $\nu_p(a+b)\geq N = \min\left\{\nu_p(a),\nu_p(b)\right\}$.

Now suppose that $N<M$. This implies that $p\divides(p^{M-N}\bbar)$. But $p\ndivides\abar$.
Therefore $p\ndivides(\abar+p^{M-N}\bbar)$. This means that $\nu_p(a+b)=N=\nu_p(a)$.
Notice the point here is that if $N=M$ then we would have that $a+b=p^N(\abar+\bbar)$
and we would not know whether or not $p\divides(\abar+\bbar)$ (it might or it might not).
If $p\divides(\abar+\bbar)$ then $\nu_p(a+b) > N$.

\bigskip

\textbf{Solution to Exercise 1.34 from Shoup.}

\textbf{(a)} That the intersection of two ideals is an ideal was a homework problem form an earlier week.

\bigskip

\textbf{(b)}. Let $m=\lcm(a,b)$. We are supposed to show that $(m) = (a) \intersect (b)$.
$(a) \intersect (b)$ is precisely the set of common multiples of $a$ and $b$. So $m$ is, by definition
of $\lcm$, the least non-negative element of $(a)\intersect(b)$. But from the proof that
$\Z$ is a PID, this tells us that $m$ generates the ideal $(a)\intersect(b)$, i.e.
$(m)=(a)\intersect(b)$.

\end{proof}



\end{document}

