\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Homework Solutions, Week 4 \\ Math 310, Elementary Number Theory \\ Fall 2019}
\author{Mitch Rudominer}

\maketitle

\textbf{Solution to Exercise 4.22} In each case below, exprecss $\gcd(a,b)$ as a linear combination of $a$ and $b$.
\begin{enumerate}
\item $a=8,b=12$. $\gcd(8,12)=4$. $4=12-8$.
\item $a=20,b=12$. $\gcd(20,12)=4$. $4=2\cdot 12 - 20$.
\item $a=10,b=7$. $\gcd(10,7)=1$. $1=3\cdot 7 - 2 \cdot 10$.
\item $a=20,b=14$. $\gcd(20,14)=2$. $2=3 \cdot 14-2\cdot20$.
\end{enumerate}

\bigskip

\textbf{Solution to Shoup Exercise 1.8} Let $I$ be a non-empty set of integers that is closed under addition. Then $I$ is an ideal iff $-a\in I$ for all $a\in I$.
\begin{proof}
If $I$ is an ideal and $a\in I$ then $-a\in I$ because $-a = -1 \cdot a$ and $I$ is closed under scalar multiplication.

Conversely, suppose that for all $a\in I$, $-a\in I$. To see that $I$ is an ideal we need to see that $I$ is closed under scalar multiplication.
Let $a\in I$ and let $n\in \Z$. We need to see that $na\in I$.

First consider the case $n=0$. Then $na=0$ so we need to show that $0\in I$. This is true because $a$ and $-a$ are in $I$ and $I$ is closed under addition.

Next consider the case $n>0$. We prove by induction on $n$ that $na\in I$. For $n=1$ this is true. For the inductive step, assume $na\in I$ and we will show that $(n+1)a\in I$. This is true because $(n+1)a= na +a$ and both $na$ and $a$ are in $I$.

Finally consider the case $n<0$. We need to see that $na\in I$. Since $-n$ is positive, we now know that $-na\in I$. But then our hypothesis implies $-(-na)=na$ is also in $I$.
\end{proof}

\bigskip

\textbf{Solution to Shoup Exercise 1.9} Show that for all integers $a,b,c$ we have:
\begin{enumerate}
\item[(a)] $\gcd(a,b)=\gcd(b,a)$.
\begin{proof}
This follows immediately from several of our different characterizations of $\gcd$ which are easily seen to be symmetric in $a$ and $b$. For example we know that $\gcd$ is the greatest of the common divisors and the common divisors of $a$ and $b$ are the same as the common divisors of $b$ and $a$.
\end{proof}

\item[(b)] $\gcd(a,b) = \av{a} \Iff a\divides b$.
\begin{proof}
If $\gcd(a,b) = \av{a}$ then $\av{a}$ is a common divisor of $a$ and $b$
so in particular it is divisor of $b$. So $a\divides b$ or $-a\divides b$. But if $-a\divides b$ then also $a\divides b$.

Conversely, suppose $a\divides b$. Then also $-a\divides b$ so $\av{a} \divides b$. So $\av{a}$ is a non-negative common divisor of $a$ and $b$. To see it is the greatest common divisor, it suffices to see that every common
divisor divides it. Suppose $c$ is a common divisor of $a$ and $b$. In 
particular $c\divides a$ so $c$ divides $\av{a}$.
\end{proof}

\item[(c)] $\gcd(a,0)=\gcd(a,a) = \av{a}$ and $\gcd(a,1) = 1$.
\begin{proof}
That $\gcd(a,0)=\gcd(a,a) = \av{a}$ is easy to see if we think about ideals.
The principal ideal generated by $a$ and $0$ is the same as the principal
ideal generated by $a$ (because 0 is in that ideal anyway) which is the
same as the principal ideal generated by $-a$. In other words
$(a,0)=(a)=(-a) = (\av{a})$.

That $\gcd(a,1)=1$ may also be seen by thinking about ideals.
$(a,1) = (1)$ because $1\in (a,1)$ and $1\in (1)$ so $(a,1)=(1)=\Z$.
\end{proof}

\item[(d)] $\gcd(ca,cb) = \av{c} \gcd(a,b)$.
\begin{proof}
Let $d=\gcd(a,b)$ and let $\bar{c}=\av{c}$. We want to see that $\bar{c}d=\gcd(ca,cb)$. Since $d\divides a$, $cd \divides ca$, so also $-cd \divides ca$ so $\bar{c}d\divides ca$. Similarly $\bar{c}d \divides cb$. So $\bar{c}d$ is a non-negative
common divisor of $a$ and $b$. To see that it is the greatest common divisor
it suffices to see that $\bar{c}d$ can be written as a linear combination
of $ca$ and $cb$. Let $x$ and $y$ be integers such that $d=xa+yb$.
Then $cd=x(ca) + y(cb)$. Then $-cd=-x(ca) -y(cb)$. So $\bar{c}d$ can
be written as an integer linear combination of $ca$ and $cb$.
\end{proof}
\end{enumerate}

\bigskip

\textbf{Solution to Shoup Exercise 1.10} Show that for al integers $a,b$ with $d:= \gcd(a,b) \not= 0$, we have $\gcd(a/d,b/d)=1$.
\begin{proof}
Let $a,b,d\in \Z$ with $d=\gcd(a,b)$. Suppose $d\not=0$.
Since $d\divides a$ and $d\divides b$ and $d\not=0$, $a/d$ and $b/d$ are
integers. Let $\bar{a}=a/d$ and $\bar{b}=b/d$. We must show that $\bar{a}$
and $\bar{b}$ are relatively prime.
Since $d=\gcd(a,b)$ there are integers $x,y$ such that $d=xa+yb$. Dividing
both sides by $d$ we get $1=x\bar{a}+y\bar{b}$. So $\bar{a}$ and $\bar{b}$ are relatively prime.
\end{proof}

\bigskip

\textbf{Solution to Shoup Exercise 1.11} Let $n$ be an integer. Show that if $a,b$ are relatively prime integers, each of which divides $n$, then $ab$ divides $n$.
\begin{proof}
Suppose $a,b$ are relatively prime and $a\divides n$ and $b\divides n$.
We must show that $ab\divides n$.

Let $x,y$ be integers such that $xa+yb=1$. Then $xan + ybn = n$.
To see that $ab\divides n$, it suffices to see that $ab\divides xan$ and
$ab\divides ybn$. Since $b\divides n$ there is an integer 
$c$ such that $n=bc$.
Thus $xan=xabc$. Thus $ab\divides xan$. Similarly since $a\divides n$ there
is an integer $d$ such that $n=ad$. Thus $ybn=ybad$ so $ab\divides ybn$.
So $ab\divides n$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 5.1} Let $p$ be a prime and let $a$ be any integer. Then either
\begin{itemize}
\item $p\divides a$ and $\gcd(p,a)=p$ or
\item $p\ndivides a$ and $\gcd(p,a)=1$.
\end{itemize}
\begin{proof}
Let $d=\gcd(p,a)$. Then in particular $d\divides p$ and $d$ is positive (since $p$ is positive $d\not=0$) so $d=1$ or $d=p$.

If $p\divides a$ then $p$ is a common divisor of $p$ and $a$ so we must have $d=p$.

If $p\ndivides a$ then $p$ is not a common divisor of $p$ and $a$ so
$d$ must be 1.
\end{proof}

\bigskip

\textbf{Solution to Exercise 1.2 from Shoup} Let $n$ be a composite integer. Show that there exists a prime $p$ dividing $n$ with $p\leq n^{1/2}$.
\begin{proof}
Since $n$ is composite there are $a,b$ with $1<a,b<n$ such that $n=ab$. Without loss of generality, assume that $a\leq b$. Since $a>1$, there is some prime $p$ such that $p\divides a$. Then $p\divides n$. Also $1<p\leq a$.
Thus $p^2\leq a^2 \leq ab=n$. So $p\leq n^{1/2}$.
\end{proof}

\end{document}

