\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm}

\input{mathdefs}
\input{theoremstyles}

\pagestyle{plain}

\begin{document}

\title{Homework Solutions, Week 3 \\ Math 310, Elementary Number Theory \\ Fall 2019}
\author{Mitch Rudominer}

\maketitle

\textbf{Solution to Exercise 3.14}
\begin{itemize}
\item $17 \mod 100 = 17$
\item $100 \mod 17 = 15$
\item $-17 \bmod 100 = 83$
\item $-100 \bmod 17 = 2$
\item $27 \bmod 3 = 0$
\item $4 \bmod 3 = 1$
\item $3 \bmod 3 = 0$
\item $2 \bmod 3 = 2$
\item $1 \bmod 3 = 1$
\item $0 \bmod  = 0$
\item $-1 \bmod 3 = 2$
\item $-2 \bmod 3 = 1$
\item $-3 \bmod 3 = 0$
\item $-4 \bmod 3 = 2$
\end{itemize}

\bigskip

\textbf{Solution to Exercise 1.1 on page 4 of \emph{Shoup}}.
Let $a,b,d\in\Z$ with $d\not=0$. Show that $a\divides b$ if and only if
$da\divides db$.
\begin{proof}
Suppose $a\divides b$ and let $c$ be such that $b=ac$. Then $bd=adc$ so
$ad \divides bd$. Conversely suppose $ad \divides bd$ and let $c$ be
such that $bd = adc$. Since $d\not=0$ we can cancel it to yield
$b=ac$. So $a\divides b$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.5} Suppose $I_1$ and $I_2$ are ideals. Show that $I_1 \intersect I_2$ is an ideal.
\begin{proof}
We must show that  $I_1 \intersect I_2$  satisfies the the three clauses in the definition of an ideal:
\begin{itemize}
\item It is non-empty.
\item It is closed under addition.
\item It is closed under scalar multiplication.
\end{itemize}
Since 0 is in both $I_1$ and $I_2$, 0 is in  $I_1 \intersect I_2$ and so  $I_1 \intersect I_2$  is non-empty.
To see that $I_1 \intersect I_2$ is closed under addition, let $a, b\in I_1 \intersect I_2$ and we will show that $a+b\in  I_1 \intersect I_2$.
Since  $a, b\in I_1 \intersect I_2$,  $a, b\in I_1$. Since $I_1$ is an ideal, $a+b\in I_1$. By the same argument $a+b\in I_2$.
So $a+b\in I_1\intersect I_2$.

To see that $I_1 \intersect I_2$ is closed under scalar multiplication, let $a \in I_1 \intersect I_2$ and let $x\in\Z$ and we will show that $xa\in  I_1 \intersect I_2$.
Since  $a\in I_1 \intersect I_2$,  $a\in I_1$. Since $I_1$ is an ideal, $xa\in I_1$. By the same argument $xa\in I_2$.
So $xa\in I_1\intersect I_2$.
\end{proof}

Now suppose that $\cS$ is a non-empty collection of ideals. Prove that the intersection
of all ideals in $\cS$ is also an ideal.

\begin{proof}
This proof is very similar to the previous one except it is a little more abstract because rather than having two ideals $I_1$ and $I_2$ we now have
a set $\cS$ of ideals. In order to follow this proof, think about the fact that if $\cS = \singleton{I_1, I_2}$ then this proof reduces to the previous proof.

Let $I$ be the intersection of all ideals in $\cS$. To see that $I$ is an ideal we must show that it is non-empty, closed under addition, and closed under scalar multiplication. It is non-empty because $0\in I$ because 0 is in every ideal in $\cS$. Let $a, b \in I$. Then $a$ and $b$ are in every ideal in $\cS$ so $a+b$ is in every ideal in $\cS$ so $a+b\in I$. Let $a\in I$ and $x\in\Z$. Then $a$ is in every ideal in $\cS$ so $xa$ is in every ideal in $\cS$ so $xa\in I$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.8} Show that $a\Z$ is an ideal of $\Z$ that contains $a$.
\begin{proof}
$a\Z$ contains $a$ because $a=1\cdot a$ is a multiple of $a$. Thus $a\Z$ is not empty. To see that $a\Z$ is closed under addition, note that
$xa + ya = (x+y)a$ so that the sum of two multiples of $a$ is again a multiple of $a$. To see that $a\Z$ is closed under scalar multiplication,
note that $y(xa) = (yx)a$ so that a scalar multiple of a multiple of $a$ is also a multiple of $a$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.10}  Note that we will use Lemma 4.9 so we know that $(a) = a\Z$.

$b\in (a)$ iff $a\divides b$.
\begin{proof}
This is easy because $b\in (a)$ iff $b\in a\Z$ iff $b$ is a multiple of $a$ iff $a\divides b$.
\end{proof}

For any ideal $I$, $b\in I$ iff $(b)\subseteq I$.
\begin{proof}
If $(b)\subseteq I$ then obviously $b\in I$. For the other direction, suppose $b\in I$. The reason we get $(b)\subseteq I$ is that $I$ is closed under scalar multiplication and so every multiple of $b$ is also in $I$.
\end{proof}

$(b)\subseteq (a)$ iff $a\divides b$.
\begin{proof}
If $(b)\subseteq (a)$ then $b\in (a)$ so $a\divides b$ by the first part of this exercise. If $a\divides b$ then $b\in (a)$ so by the second part of
this exercise $(b)\subseteq (a)$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.12}  Show that $I_1+I_2$ is an ideal.
\begin{proof}
Let $I=I_1+I_2$. $0\in I$ because $0=0+0$. So $I$ is not empty. To see that $I$ is closed under addition,
let $a,b\in I$. Then there are $a_1,b_1\in I_1$ and $a_2,b_2\in I_2$ such that $a=a_1+a_2$ and $b=b_1+b_2$.
Then $a+b = (a_1+b_1) + (a_2 + b_2)$ and $(a_1+b_1)\in I_1$ and $(a_2+b_2)\in I_2$ so $a+b\in I$.
To see that $I$ is closed under scalar multiplication, let $a=a_1+a_2$ be as above and let $x\in\Z$.
Then $xa=xa_1+xa_2\in I$.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.13}  Let $(a)$ and $(b)$ be two principal ideals. Show that $(a)+(b)$ is equal to the set of all integer linear combinations of $a$ and $b$.
\begin{proof}
This follows just by unravelling definitions. An element of $(a)+(b)$ looks like $xa+yb$ for some $x,y\in\Z$.
But this is exactly an integer linear combination of $a$ and $b$ by definition.
\end{proof}

\bigskip

\textbf{Solution to Exercise 4.15}  Show that $(a,b)=(a)+(b)$.
\begin{proof}
$(a)+(b)$ is an ideal that contains $a$ and $b$ so $(a,b)\subseteq(a)+(b)$. By Lemma 4.3 every ideal
is closed under integer linear combinations, so $(a,b)$ contains all integer linear combinations of $a$ and $b$
so by the previous exercise $(a)+(b)\subseteq (a,b)$.
\end{proof}

\end{document}

