\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Solutions to Homework for Lesson 10 \\ Linear transformations on vector spaces}
\author{Math 325, Linear Algebra \\ Mitch Rudominer \\ Spring 2019 \\ SFSU }
\date{}

\maketitle


\textbf{Problem 1.} $T$ is linear for all values of $x_0$. To see this we
must see that $T$ preserves vector addition and scalar multiplication.
Notice that the codomain of $T$ is $\R^1$ so in the codomain vector addition
means just ordinary addition of real numbers. $T$ preserves vector addition
because $T(f + g) = (f+g)(x_0) = f(x_0) + g(x_0) = T(f) + T(g)$.
$T$ preserves scalar multiplication because $T(cf) = (cf)(x_0)
= c(f(x_0)) = cT(f)$.

\bigskip

\textbf{Problem 2.}

\textbf{(a)} The kernel is the set of all functions that assign zero to $x_0$,
i.e. $\ker(T) = \setof{f}{f(x_0) = 0}$. To see that this is a subspace of $\cF$
we must see that it contains the zero function $Z$ and is closed under vector addition
and scalar multiplication. The set contains the zero function $Z$ becuase
$Z(x_0) = 0$. The set is closed under vector addition because if $f(x_0) = 0$
and $g(x_0) = 0$ then $(f+g)(x_0) = 0$. The set is closed under scalar
multiplication because if $f(x_0) = 0$ the $cf(x_0) = 0$.

\bigskip

\textbf{(b)} The kernel is the set of constant functions
$\cC = \setof{f\in\cF}{f \text{ is a constant function} }$. This is because
the zero vector of the codomain is $Z$, the zero function, and the kernel
is the set of functions whose derivative is $Z$. The derivative of $f$ is
$Z$ iff $f$ is a constant function. To see that $\cC$ is a subspace notice
that $\cC$ contains $Z$ and is closed under vector addition (because the sum of
two constant functions is a constant function) and scalar multiplication
(because a scalar multiple of a constant function is a constant function.)

\bigskip

\textbf{(c)} Let $O$ be the $2\times 2$ zero matrix. Then $T=T_O$. So $T$
is linear. $\ker(T) = \R^2$. That is because $T$ sends every vector in
$\R^2$ to $\bzero$. It is obvious that the kernel is a subspace of $\R^2$ because
the kernel is all of $\R^2$.

\bigskip

\textbf{(d)} Let $I$ be the $2\times 2$ identity matrix. Then $T=T_I$. So $T$
is linear. $\ker(T) = \singleton{\bzero}$. That is because $T$ sends every vector in
$\R^2$ to itself so it only sends $\bzero$ to $\bzero$. The kernel in this
case is the zero subspace which we showed in class is a subspace.

\bigskip

\textbf{(e)} Let $A=
\begin{pmatrix}
1 & 0 \\
0 & 0
\end{pmatrix}
$.
Then $T=T_A$ so $T$ is linear. The kernel of $T$ is the $y$-axis of $\R^2$, i.e.
$\ker(T) = \setof{(x,y)^{\text{T}}}{x=0}$. This contains the origin and
is closed under vector addition and scalar multiplication, so it is a subspace.

\bigskip

\textbf{Problem 3.} By looking at the kernels from problem 2 we can tell
that only the transformation in part (d) is one-to-one. This is because
a linear transformation is one-to-one iff its kernel is trivial, i.e.
its kernel is the zero subspace.

\bigskip

\textbf{Problem 4.}

\textbf{(a)} The image of $T$ is all of $\R$. That is because for any $c\in\R$
there is a function $f$ such that $f(x_0) = c$: for example the constant
$c$ function. For such an $f$ we have $T(f) = c$, so $c\in\ran(T)$.
In this case it is obvious that the range of the transformation
is a subspace of the codomain because the range is all of the codomain, i.e.
$T$ is onto.

\bigskip

\textbf{(b)} The image of $T$ is $\singleton{\bzero}$ and this is the zero
subspace.

\bigskip

\textbf{(c)} The image of $T$ is all of $\R^2$ so it is a subspace of $\R^2$,

\bigskip

\textbf{(d)} Let $A=
\begin{pmatrix}
1 & 0 \\
1 & 0
\end{pmatrix}
$.
Then $T=T_A$ so $T$ is linear.
The image of $T$ is the line $y=x$ in $\R^2$, i.e.
$\setof{(x,x)^{\text{T}}}{x\in\R}$. This is a subspace of $\R^2$ because
it contains the origin and is closed under vector addition and scalar
multiplication.

\bigskip

\textbf{Problem 5.} Let

\textbf{(a)} $W$ is a subspace of $V$ because
\begin{itemize}
\item $W$ contains the zero vector because that is the zero function $Z$ and
$Z$ is a polynomial of degree $\leq 3$.
\item $W$ is closed under vector addition because the sum of two polynomials
of degree at most 3 is itself a polynomial of degree at most 3.
\item $W$ is closed under scalar multiplication because if you multiply
a polynomial of degree at most 3 by a scalar the result is also a polynomial
of degree at most 3.
\end{itemize}

\bigskip

\textbf{(b)} The image of $T$ is the set of polynomials of degree at most 2.
That is because if you differentiate a polynomial of degree 3 you get a
polynomial of degree 2. The set of polynomials of degree at most 2 is a
subspace of the set of all polynomials using exactly the same argument
that was used in (a) above to show the same thing about the set of
polynomials of degree at most 3.





\end{document}
