% $Header$

%\documentclass{beamer}
\documentclass[handout]{beamer}
\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,hyperref}
\input{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 17 \\ Wilson's Theorem}
\subtitle{Math 310, Elementary Number Theory \\ Fall 2020 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{$(n-1)! \bmod n$}

\begin{itemize}
  \item In this lesson we are going to study
  \item $(n-1)! \bmod n$
  \item for $n$ a positive integer.
  \item For example, $10! \bmod 11$
  \item $1\cdot 2\cdot 3 \cdot 4 \cdot 5 \cdot 6 \cdot 7 \cdot 8 \cdot 9 \cdot 10 \bmod 11$
  \item or $99! \bmod 100$
  \item $1\cdot 2 \cdot 3 \cdots 98 \cdot 99 \bmod 100$.
\end{itemize}

\end{frame}

\begin{frame}{Examples}

\begin{itemize}
  \item Consider this table of the values of $(n-1)! \bmod n$.
  \item https://en.wikipedia.org/wiki/Wilson\%27s\_theorem\#Example
  \item From these examples, let's try to devise some theorems.
\end{itemize}

\end{frame}

\begin{frame}{Wilson's Theorem}

\begin{itemize}
  \item \textbf{Theorem.} Let $p$ be prime. Then $(p-1)! \bmod p = p-1$.
  \item Example: $10! \bmod 11 = 10$.
  \item Example: $22! \bmod 23 = 22$.
  \item Example: $100! \bmod 101 = 100$.
  \item Another way of looking at this:
  \item \textbf{Theorem.} Let $p$ be prime. Then $p \divides ((p-1)! + 1)$.
  \item Example: $11 \divides (10!+1)$.
  \item Example: $23 \divides (22!+1)$.
  \item Example: $101 \divides (100! + 1)$.
  \item Another way of looking at this:
  \item \textbf{Wilson's Theorem.} Let $p$ be prime. Then $(p-1)! \equiv -1 \pmod p$.
  \item Example: $10! \equiv -1  \pmod {11}$.
  \item Example: $22! \equiv -1  \pmod {23}$.
  \item Example: $100! \equiv -1 \pmod {101}$.
\end{itemize}

\end{frame}

\begin{frame}{When $n$ is not prime}

\begin{itemize}
  \item \textbf{Theorem.} Let $n>1$ and suppose $(n-1)! \equiv -1 \pmod n$.
  \item Then $n$ is prime.
  \item Equivalently:
  \item \textbf{Theorem.} Let $n>1$ be composite. Then $(n-1)! \not\equiv -1 \pmod n$.
  \item $3! \bmod 4 = 2$.
  \item If $n \geq 6$ and $n$ is composite...
  \item then $(n-1)! \bmod n = 0$.
  \item i.e. $n \divides (n-1)!$
  \item For example $100\divides 99!$.
  \item $99! \bmod 100 = 0$.
\end{itemize}

\end{frame}

\begin{frame}{Proof of composite case}

\begin{itemize}
  \item \textbf{Theorem.} Let $n \geq 6$ be composite.
  \item then $n \divides (n-1)!$
  \item For example $100\divides 99!$.
  \item \textbf{proof.} Since $n$ is composite, $n=ab$ with $1< a \leq b < n$.
  \item For example $100 = 5\cdot 20$.
  \item If $a<b$ then $(n-1)! = 1\cdot 2 \cdots a \cdots b \cdots (n-1)$.
  \item So $n=ab \divides (n-1)!$.
  \item For example $100 = 5\cdot 20 \divides 99!$.
  \item If $n$ cannot be written as $ab$ with $a<b$ then $n=p^2$ for some odd prime $p$.
  \item Then $2p < n$ so $(n-1)! = 1\cdot 2 \cdots p \cdots 2p \cdots n-1$.
  \item So $n=p^2 \divides (n-1)!$ $\qed$
\end{itemize}

\end{frame}

\begin{frame}{Proof of Wilson's Theorem}

\begin{itemize}
  \item \textbf{Wilson's Theorem.} Let $p$ be prime. Then $(p-1)! \equiv -1 \pmod p$.
  \item \textbf{proof}. Consider what this means in the language of $\Z_p^*$.
  \item It means that the product of all of the elements of $\Z_p^*$ is $-[1]$.
  \item For example, let $p=7$. Wilson's Theorem says that in $\Z_7^*$,
  \item $[1][2][3][4][5][6] = -[1] = [6]$.
  \item Every element of $\Z_p^*$ has a multiplicative inverse in $\Z_p^*$.
  \item For example in $\Z_7^*$,
  \item $[1][1]=[1], [2][4]=[1],[3][5]=[1]$ and $[6][6] = [1]$.
  \item So in $\Z_7^*$, $[1][2][3][4][5][6] = [1][6] = [6]$.
  \item In general, when we take the product of all of the elements of $\Z_p^*$,
  \item every element will cancel with its multiplicative inverse
  \item except for $[1]$ and $-[1]$.
  \item These two are their own multiplicative inverses.
\end{itemize}

\end{frame}

\begin{frame}{Proof of Wilson's Theorem, page 2}
\begin{itemize}
  \item So to prove Wilson's theorem we need the following
  \item \textbf{Lemma.} Let $p$ be prime. In $\Z_p^*$,
  \item $[1]$ and $-[1]$ are the only two elements that are their own multiplicative inverse.
  \item In other words, these are the only two square-roots of $[1]$.
  \item \textbf{proof.} Let $1<a<(p-1)$. We need to prove that $[a]^2 \neq [1]$ in $\Z_p^*$.
  \item i.e. we need to prove that $a^2 \not\equiv 1 \pmod p$.
  \item We need to prove that $p\ndivides (a^2 - 1)$.
  \item But $a^2 - 1 = (a-1)(a+1)$ and $(a-1)<(a+1)<p$.
  \item So $p\ndivides (a-1)$ and $p\ndivides (a+1)$.
  \item Since $p$ is prime, $p\ndivides (a-1)(a+1)$.$\qed$.
\end{itemize}

\end{frame}

\begin{frame}{Square Roots in $\Z_p^*$}
\begin{itemize}
  \item \textbf{Corollary.} Let $p$ be prime. In $\Z_p^*$ every element has at most two square roots.
  \item If $[b]^2 = [a]$ and $[c]^2 = [a]$ then $[c]=\pm [b]$.
  \item \textbf{proof.} Suppose $[b]^2=[c]^2$. Multiply both sides by $([c]^{-1})^2$.
  \item $([b][c]^{-1})^2 = [1]$. So $[b][c]^{-1}$ is a square-root of $[1]$.
  \item By the previous lemma, $[b][c]^{-1}=\pm[1]$.
  \item Multiply both sides by $[c]$.
  \item $[b]=\pm[c]$. $\qed$.
  \item Example. In $\Z_{101}^*$ find all square roots of $[20]$.
  \item $11^2=121\equiv 20 \pmod{101}$.
  \item So $[11]$ and $-[11]=[90]$ are two square roots of $[20]$ in $\Z_{101}^*$.
  \item $121$ is prime so these are the only two square roots.
\end{itemize}
\end{frame}

\begin{frame}{Examples}
\begin{itemize}
  \item \textbf{Example:} Find $42! \bmod 43$.
  \item Solution: 42.
  \item \textbf{Example:} Find $41! \bmod 43$.
  \item Solution: $41! = 42!/42$. Working in $\Z_{43}^* \cdots$
  \item $[1][2]\cdots[41] = -[1][42]^{-1}=-[1](-[1])^{-1} = [1]$.
  \item So $41! \bmod 43 = 1$.
  \item \textbf{Corollary.} Let $p$ be prime. Then $(p-2)! \equiv 1 \pmod{p}$.
  \item \textbf{Example:} Find $40!  \bmod 43$.
  \item Solution: $40! = 42!/(41\cdot 42)$. Working in $\Z_{43}^* \cdots$
  \item $[1][2]\cdots[40] = -[1][41]^{-1}[42]^{-1}=-[1](-[2])^{-1}(-[1])^{-1} = (-[2])^{-1}$
  \item $= -[2]^{-1} = -[22] = [21]$.
  \item So $40! \bmod 43=21$.
\end{itemize}
\end{frame}

\begin{frame}{More Complicated Example}
\begin{itemize}
  \item \textbf{Example:} Find $58! \bmod 3599$.
  \item Solution: Find the prime factorization of 3,599.
  \item Since the problem contains $58!$ and 59 is prime and
  \item this lesson is about Wilson's theorem, guess that $59\divides 3,599$.
  \item $3,599 = 59 \cdot 61$.
  \item Idea: Find $58! \bmod 59$ and $58! \bmod 61$ and then use the Chinese Remainder Theorem.
  \item $58! \bmod 59 = -1$. $58!\bmod 61$ is found by working in $\Z_{61}^* \cdots$
  \item $-[2]^{-1}=-[31]=[30]$.
  \item Let $x=58! \bmod 3599$. Then $x$ satisfies:
  \item $x \equiv -1 \pmod {59}$
  \item $x \equiv 30 \pmod {61}$.
\end{itemize}
\end{frame}

\begin{frame}{More Complicated Example, Continued}
\begin{itemize}
  \item \textbf{Example:} Find $58! \bmod 3599$.
  \item Solve
  \item $x \equiv -1 \pmod {59}$
  \item $x \equiv 30 \pmod {61}$.
  \item $M_1 = 61$. Find $y_1$, the multiplicative inverse of 61 mod 59.
  \item $61 \bmod 59 = 2$ and $2\cdot 30 \equiv 1 \pmod {59}$ so $y_1=30$.
  \item $M_2 = 59$. Find $y_2$, the multiplicative inverse of 59 mod 61.
  \item $59 \equiv -2 \pmod{61}$ and $(-2)\cdot(-31) \equiv 1 \pmod {61}$.
  \item $y_2 = (-31) \bmod 61 = 30$.
  \item $x = \left( (-1)(61)(30) + (30)(59)(30) \right) \bmod 3599=$
  \item $=884$.
  \item So $58! \bmod 3599 = 884$
\end{itemize}
\end{frame}


\end{document}
