% $Header$

\documentclass{beamer}
%\documentclass[handout]{beamer}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx,xcolor}
\include{mathdefs}

\graphicspath{{images/}}

\newtheorem*{claim}{claim}
\newtheorem*{observation}{Observation}
\newtheorem*{warning}{Warning}
\newtheorem*{question}{Question}
\newtheorem{remark}[theorem]{Remark}

\newenvironment*{subproof}[1][Proof]
{\begin{proof}[#1]}{\renewcommand{\qedsymbol}{$\diamondsuit$} \end{proof}}

\mode<presentation>
{
  \usetheme{Singapore}
  % or ...

  \setbeamercovered{invisible}
  % or whatever (possibly just delete it)
}


\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title{Lesson 17 \\ Determinants and Transpose}
\subtitle{Math 325, Linear Algebra \\ Spring 2019 \\ SFSU}
\author{Mitch Rudominer}
\date{}



% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant Theory}

\begin{itemize}
\item \textbf{Theorem.} There exists a unique function, called the \emph{determinant}, and writen $\det$, with the following properties:
\item (1) For every square matrix $A$, $\det(A)$ is a scalar.
\item (2) Performing an elementary row operation of type 1 on a matrix does not change its determinant.
\item (3) Performing an elementary row operation of type 2 on a matrix multiplies the determinant by -1.
\item (4) Performing an elementary row operation of type 3 on a matrix, multiplying one of the rows by the scalar $t$,
multiplies the determinant by the same scalar $t$.
\item (5) The determinant of the identity matrix is 1.
\item We will give a sketch of the proof but we won't go through the proof in detail.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $2\times 2$ matrix.}

\begin{itemize}
\item We have already defined the determinant of $2\times 2$ matrix.
\item
$
\det
\begin{pmatrix}
a & b \\
c & d \\
\end{pmatrix}
= ad - bc
$.
\item Verify each of the properties:
\item (1) For every $2\times 2$ matrix $A$, $\det(A)$ is a scalar.
\item (2) Performing an elementary row operation of type 1 on a matrix does not change its determinant.
\item (3) Performing an elementary row operation of type 2 on a matrix multiplies the determinant by -1.
\item (4) Performing an elementary row operation of type 3 on a matrix, multiplying one of the rows by the scalar $t$,
multiplies the determinant by the same scalar $t$.
\item (5) The determinant of the identity matrix is 1.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{The Sign of a Permutation}

\begin{itemize}
\item \textbf{Definition} If $\pi$ is a permutation of the numbers $\singleton{1,2,\cdots, n}$
we say that $\pi$ is \emph{even} and that $\sgn(\pi) = 1$ if $\pi$ is obtained via an even number of swaps.
\item Otherwise we say that $\pi$ is odd and $\sgn(\pi) = -1$.
\item For example consider the following permutations of $(1,2,3,4)$.
\item $\pi=(1,2,3,4)$. $\pi$ is even. $\sgn(pi) = 1$. $\pi$ is obtained via zero swaps.
\item $\pi=(3,2,1,4)$. $\pi$ is odd. $\sgn(pi) = -1$. $\pi$ is obtained via one swap.
\item $\pi=(3,2,4,1)$. $\pi$ is even. $\sgn(pi) = 1$. $\pi$ is obtained via two swaps.
\item $\pi=(3,4,2,1)$. $\pi$ is odd. $\sgn(pi) = -1$. $\pi$ is obtained via three swaps.
\item \textbf{Theorem.} $\sgn(\pi)$ is well-defined. (We won't prove this.)
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{General Definition of Determinant}

\begin{itemize}
\item \textbf{Definition} Let $A$ be an $n\times n$ matrix.
\item Then
$$\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}$$
\item The sum is over all permutations of the row indices $(1,2,\cdots, n)$
\item Each term in the sum is the product over the columns of $A$ of one of the entries in that column.
\item In column $i$ we take the entry from row $\pi(i)$.
\item The product is multiplied by $\sgn(\pi)$.
\item i.e. we add the product if $\pi$ is even and subtract the product of $\pi$ is odd.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $2\times 2$ matrix revisited}

\begin{itemize}
\item Check that the two different definitions coincide for $2\times 2$ matrices.
\item
$
A=
\begin{pmatrix}
a & b \\
c & d \\
\end{pmatrix}
=
\begin{pmatrix}
a_{1,1} & a_{1,2} \\
a_{2,1} & a_{2,2} \\
\end{pmatrix}
$
\item
$
\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}
$
\item There are two permutations of $(1,2)$:
\item $(1,2)$ is even, $(2,1)$ is odd.
\item
$
\det(A) = a_{1,1}a_{2,2} - a_{2,1}a_{2,2}
$
\item $= ad-bc$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a $3\times 3$ matrix.}

\begin{itemize}
\item
$
A=
\begin{pmatrix}
a & b & c\\
d & e & f \\
g & h & i \\
\end{pmatrix}
=
\begin{pmatrix}
a_{1,1} & a_{1,2}  & a_{1,3}\\
a_{2,1} & a_{2,2}  & a_{2,3} \\
a_{3,1} & a_{3,2}  & a_{3,3} \\
\end{pmatrix}
$
\item
$
\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}
$
\item There are six permutations of $(1,2,3)$:
\item $(1,2,3), (2,3,1), (3,1,2)$ are even
\item $(1,3,2), (3,2,1), (2,1,3)$ are odd.
\item
$
\det(A) = a_{1,1}a_{2,2}a_{3,3}  + a_{2,1}a_{3,2}a_{1,3} + a_{3,1}a_{1,2}a_{2,3}
-a_{1,1}a_{3,2}a_{2,3} -a_{3,1}a_{2,2}a_{1,3} - a_{2,1}a_{1,2}a_{3,1}
$
\item $= aei + dhc + gbf - ahf - gec - dbi$
\item $= a(ei - hf) - d(bi-hc) + g(bf -ec)$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant existence and uniqueness theorem}

\begin{itemize}
\item \textbf{sketch of proof of first theorem} We claim that the determinant, as defined by the formula we have
been discussing, satisfies the five properties from the first theorem:
\item (1) For every square matrix $A$, $\det(A)$ is a scalar.
\item (5) The determinant of the identity matrix is 1.
\item This follows easily from the formula for $\det$ applied to the identity matrix because the only non-zero term in the
sum is the product of the diagonals, which is 1.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Proof sketch, continued}

\begin{itemize}
\item (3) Performing an elementary row operation of type 2 on a matrix multiplies the determinant by -1.
\item This follows from the formula for $\det$ because the sign of a single swap is -1.
\item (4) Performing an elementary row operation of type 3 on a matrix, multiplying one of the rows by the scalar $t$,
multiplies the determinant by the same scalar $t$.
\item This follows from the formula for $\det$ because each term in the sum is multiplied by $t$ so the sum is multiplied by $t$.
\item (2) Performing an elementary row operation of type 1 on a matrix does not change its determinant.
\item This is not as easy to see immediately but you end up adding and subtracting the same amount to the sum as in
th $2\times 2$ case. $\qed$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of an upper triangular matrix}

\begin{itemize}
\item \textbf{Theorem} The determinant of an upper-triangular matrix
is the product of it's diagonals.
\item \textbf{proof} In the formula
$$
\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}
$$
all terms except the product of the diagonal entries is zero. $\qed$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Example with a $3\times 3$ matrix.}

\begin{itemize}
\item
$
A=
\begin{pmatrix}
a & b & c\\
d & e & f \\
g & h & i \\
\end{pmatrix}
=
\begin{pmatrix}
a & b & c\\
0 & e & f \\
0 & 0 & i \\
\end{pmatrix}
$
\item
$\det(A) = aei + dhc + gbf - ahf - gec - dbi$
\item
$= aei + 0 + 0 - 0 - 0 -0$
\item $=aei$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinants and Gaussian Elimination}
\begin{itemize}
\item \textbf{Theorem.} Let $A$ be a square matrix and let $U$ be
the row-echelon form matrix resulting from performing Gaussian elimination
on $A$.
\item $\det(A) = \det(U)$ if the number of swaps used in Gaussian elimination
is event.
\item $\det(A) = -\det(U)$ if the number of swaps used in Gaussian elimination is odd.
\item \textbf{proof} This is easy because row operations of type 1 do not
change the determinant and row operations of type 2 multiply the determinant by -1. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Computing determinants using Gaussian elimination}
\begin{itemize}
\item Compute the determinant of the matrix
$A=
\begin{pmatrix}
-1 & 2 & 0 \\
3 & -6 & 5 \\
2 & 0 & 2
\end{pmatrix}
$
\item Perform Gaussian elimination
\item
$
\SkipImplies
\begin{pmatrix}
-1 & 2 & 0 \\
0 & 0 & 5 \\
0 & 4 & 2
\end{pmatrix}
\SkipImplies
\begin{pmatrix}
-1 & 2 & 0 \\
0 & 4 & 2 \\
0 & 0 & 5 \\
\end{pmatrix}
= U
$
\item $U$ is upper-triangular so $\det(U) = -1 \cdot 4 \cdot 5 = -20$.
\item Gaussian elimination involved one swap, so $\det(A) = 20$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a singular matrix}
\begin{itemize}
\item Compute the determinant of the matrix
$A=
\begin{pmatrix}
-1 & 2 & 0 \\
3 & -6 & 0 \\
2 & 0 & 4
\end{pmatrix}
$
\item Perform Gaussian elimination
\item
$
\SkipImplies
\begin{pmatrix}
-1 & 2 & 0 \\
0 & 0 & 0 \\
0 & 4 & 4
\end{pmatrix}
\SkipImplies
\begin{pmatrix}
-1 & 2 & 0 \\
0 & 4 & 4 \\
0 & 0 & 0 \\
\end{pmatrix}
= U
$
\item $U$ is upper-triangular so $\det(U) = -1 \cdot 4 \cdot 0 = 0$.
\item So $\det(A) = 0$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinants and invertibility}
\begin{itemize}
\item \textbf{Theorem.} Let $A$ be a square matrix.
\item Then $A$ is singular iff $\det(A) = 0$.
\item i.e. $A$ is invertible iff $\det(A) \not= 0$.
\item \textbf{proof} Use Gaussian elimination to put $A$ into row-echelon
form $U$.
\item Then $U$ is actually in upper-triangular form.
\item If $A$ is singular then the bottom row of $U$ is all zeros and
so the product of the diagonal entries is zero and so $\det(A)=0$.
\item if $A$ is nonsingular then all of the diagonal entries of $U$ are
nonzero and so $\det(A)\not=0$. $\qed$.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Transpose of a matrix}
\begin{itemize}
\item \textbf{Definition} Let $A$ be an $m\times n$ matrix.
\item Then $\transpose{A}$ is the $n\times m$ matrix obtained by
converting the rows of $A$ into columns.
\item More precisely if $B=\transpose{A}$ then $b_{ij} = a_{ji}$.
\item Example. Let
$
A=
\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6
\end{pmatrix}
$
\item Then
$
\transpose{A} =
\begin{pmatrix}
1 & 4 \\
2 & 5 \\
3 & 6 \\
\end{pmatrix}
$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Transpose of a square matrix}
\begin{itemize}
\item If $A$ is a square matrix then $\transpose{A}$ is obtained from $A$
by reflecting across the diagonal.
\item Example. Let
$
A=
\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9
\end{pmatrix}
$
\item Then
$
\transpose{A} =
\begin{pmatrix}
1 & 4  & 7 \\
2 & 5  & 8\\
3 & 6  & 9\\
\end{pmatrix}
$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of transpose}
\begin{itemize}
\item \textbf{Theorem.} Let $A$ be a square matrix. Then $\det(A) = \det(\transpose{A})$.
\item \textbf{sketch of proof.}  In the formula
$$
\det(A) = \sum_{\pi} (\sgn(pi)) a_{\pi(1),1} a_{\pi(2), 2} \cdots, a_{\pi(n), n}
$$
each one of the terms in the sum is a product consisting of exactly one entry from each row and exactly one entry from each column.
\item This is symmetric with respect to rows and columns. $\qed$.
\item Example. Consider a $2\times 2$ matrix.
\item
$
A =
\begin{pmatrix}
a & b \\
c & d \\
\end{pmatrix}
\text{, }
\det(A) = ad - bc
$
\item
$
\transpose{A} =
\begin{pmatrix}
a & c \\
b & d \\
\end{pmatrix}
\text{, }
\det(\transpose{A}) = ad - cb = \det(A)
$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Determinant of a product}
\begin{itemize}
\item \textbf{Theorem.} Let $A$ and $B$ be square matrices of the same size. Then $\det(AB) = \det(A) \det(B)$.
\item We will skip the proof.
\item Example.
\item Let
$
A =
\begin{pmatrix}
1 & -2 \\
-1 & 4 \\
\end{pmatrix}
\text{, }
\det(A) = 4-2=2
$
\item
$ Let
B =
\begin{pmatrix}
1 & 5 \\
0 & 3 \\
\end{pmatrix}
\text{, }
\det(B) = 3 - 0 = 3,
$
\item
$
AB =
\begin{pmatrix}
1 & -2 \\
-1 & 4 \\
\end{pmatrix}
\begin{pmatrix}
1 & 5 \\
0 & 3 \\
\end{pmatrix}
=
\begin{pmatrix}
1 & -1 \\
-1 & 7 \\
\end{pmatrix}
$
\item $\det(AB) = 7 - 1 = 6$
\item So $\det(AB) = \det(A) \det(B).
$
\end{itemize}
\end{frame}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%






\end{document}


