\documentclass[oneside,12pt]{amsart}

\usepackage{amsmath,amssymb,latexsym,eucal,amsthm,graphicx}

\input{mathdefs}
\input{theoremstyles}

\graphicspath{{images/}}

\pagestyle{plain}

\begin{document}

\title{Syllabus}
\author{Math 310, Elementary Number Theory  \\ Fall 2021 \\ SFSU }
\date{}

\maketitle

\textbf{Basics:}
\begin{itemize}
\item \textbf{Instructor:} Mitch Rudominer
\item \textbf{e-mail:} rudominer@sfsu.edu
\item \textbf{Zoom:} All lectures and office hours are remote on Zoom. The Zoom info can be found in the \emph{General Course Info} section at the top of the iLearn
page for the class.
\item \textbf{Lecture:}  T,Th 8:00 - 9:15 am \emph{On Zoom}.
\item \textbf{Office Hours:} I'll stay for a few minutes after each class for quick questions. Otherwise, office hours are by appointment.
Please send me an email to arrange it.
\end{itemize}

\bigskip

\textbf{Main topics of this course:}
The theory of the integers: divisibility, congruence,
primitive roots, quadratic residues, quadratic reciprocity.

\bigskip

\textbf{Learning outcomes:}
When a student has successfully completed this course they will be able to:
\begin{itemize}
\item understand number-theoretic proofs
\item explain and be able to use the theorems of Fermat, Wilson and Euler
\item given an integer $n$, determine whether or not there are any primitive
roots modulo $n$ and if so, how many
\item perform the extended Euclidean algorithm and the Chinese remainder theorem algorithm
\item express the greatest common divisor of two integers $a$ and $b$ as an integer
linear combination of $a$ and $b$
\item solve systems of simultaneous linear congruences
\end{itemize}

\bigskip

\textbf{How to get help:}
\begin{itemize}
\item Ask a question during class.
\item Ask a question in the class discussion forum on iLearn.
\item Send me an e-mail.
\item Arrange office hours.
\end{itemize}

\bigskip

\textbf{iLearn:}
\begin{itemize}
\item There is a lot of information on the class iLearn page. Use it frequently.
\item This document
\item The class discussion forum. Register to receive e-mail notifications.
\item Lecture Notes
\item Recordings of lectures
\item Homework (As iLearn practice quizzes)
\item Quizzes (As iLearn quizzes)
\end{itemize}

\bigskip

\textbf{Lectures}
\begin{itemize}
\item This will be a traditional lecture format course.
\item I will use slides and a Google Jamboard to present the lectures.
\item We will also use Google Colab for numerical calculations.
\item The slides, Jamboard and Colab will be available to you via links on the iLearn site.
\item The lectures will be recorded on Zoom with links to the recordings on the iLearn site.
\item My lectures and the slides will be the primary source.
\end{itemize}

\bigskip

\textbf{Colab}
\begin{itemize}
\item We will use Colab in this class.
\item Please go to https://colab.research.google.com and make sure you have access to Colab.
\item You will need a Google account. If you don't have one already you can make one for free.
\end{itemize}

\bigskip

\textbf{Text}
\begin{itemize}
\item There is no required textbook for this course.
\item You can get all the information you need to do well from my lectures and the slides.
\item But if you want an additional source, I have recommended two free online books on the iLearn page.
\item For each class, I will mention optional reading material from those online sources.
\end{itemize}

\bigskip

\textbf{Exams:}
\begin{itemize}
\item \textbf{Midterm:} Thursday October 28, 2021, in class.
\item \textbf{Final Exam:} Tuesday, December 14, 8:00 am- 10:00 am
\end{itemize}

\bigskip

\textbf{Homework:}
\begin{itemize}
\item Homework in the form of iLearn practice quizzes will be assigned in every class:
  \begin{itemize}
    \item These are graded automatically by iLearn.
    \item You can retake the iLearn practice quiz as many times as you want and you are given the correct answers when you finish.
    \item Each time you take the quiz the question order is randomized.
    \item Your best score will be recorded.
    \item There is no deadline or time limit. You have until the end of the semester to finish any of the practice quizzes.
    \item But you will really want to get each of these practice quizzes done before the next class because there will
    be an in-class quiz based very closely on the practice quizzes.
    \item The sum of all of the practice quizzes will count as 10\% of your final grade.
    \item You may work together with classmates on your homework and you may ask questions about the homework on the
    class discussion forum. But in the end you should be able to answer all of the questions on the practice quiz
    by yourself because that is the only way you will do well on the real quiz.
  \end{itemize}
\end{itemize}

\bigskip

\textbf{Quizzes}
\begin{itemize}
\item There will be a quiz during the last 15 minutes of every class (except the first class) covering the material
learned in the previous class.
\item The quiz will be given in the form of an iLearn quiz. It will be automatically graded by iLearn and you will
have your grade and the corrected quiz as soon as the quiz closes at 9:20.
\item The quiz will be based very closely on the practice quiz. So if, after taking the practice quiz several times,
you can get a 100\% on the practice quiz, you are likely to do well on the real quiz.
\item I will drop your five lowest quiz grades.
\item The grades on the remaining quizzes together will comprise 50\% of your final grade.
\item You are not allowed to collaborate with anyone else while taking the quiz. Doing so will be considered cheating.
\item The quizzes are ``open book". You are allowed to consult your notes, the class slides, even the Internet (though that is unlikely to help you.)
\item You are allowed to use Colab or any other calculator you wish to use during the quiz. For some questions you will be required to use Colab.
\end{itemize}

\bigskip

\textbf{Grades:} Percentages are flexible so as to give you the highest possible grade.
\begin{itemize}
\item Homework 10\%
\item Quizzes  50\% ($\pm5\%$) (I will drop you 5 lowest quizzes.)
\item Midterm 15\% ($\pm5\%$)
\item Final: 25\% ($\pm5\%$)
\item Letter grades will be assigned based on straight numerical grades
without curving and using the usual mapping of number grades to letter grades:

\begin{center}
  \begin{tabular}{ | l | l |}
    \hline
    Number Grade & Letter Grade  \\ \hline \hline
    93 -- 100 & A\\ \hline
    90 -- 92 & A- \\ \hline
    87 -- 89 & B+ \\ \hline
    83 -- 86 & B \\ \hline
    80 -- 82 & B- \\ \hline
    etc & etc \\ \hline
  \end{tabular}
\end{center}

\end{itemize}

\textbf{Policy on WU grades and International Students:}

WU (Withdrawal Unauthorized) indicates that an enrolled student did not withdraw from the course and also failed to complete course requirements. It is used when, in the opinion of the instructor, completed assignments or course activities or both were insufficient to make normal evaluation of academic performance possible. For purposes of grade point average and progress point computation, this symbol is equivalent to an F.

International students who have enrollment unit requirements for their stay in the US need to be aware that a course with a WU grade is not viewed as a course attempted whereas a course with an F grade is considered to be a course attempted. The Mathematics Department will not approve petitions to retroactively change a grade of WU to a grade of F.

\textbf{Disability access:}
Students with disabilities who need reasonable accommodations are encouraged to contact
the instructor.  The Disability Programs and Resource Center (DPRC) is available to facilitate
the reasonable accommodations process. The DPRC is located in the Student Service Building
and can be reached by telephone (voice/415-338-2472, video phone/415-335-7210) or by
email (dprc@sfsu.edu).

\textbf{Student disclosures of sexual violence}
SF State fosters a campus free of sexual violence including sexual harassment, domestic
violence, dating violence, stalking, and/or any form of sex or gender discrimination.  If you
disclose a personal experience as an SF State student, the course instructor is required to
notify the Title IX Coordinator by completing the report form available
at http://titleix.sfsu.edu, emailing vpsaem@sfsu.edu or calling 338-2032.

\textbf{To disclose any such violence confidentially, contact:}
\begin{itemize}
\item The SAFE Place - (415) 338-2208; http://www.sfsu.edu/\~safe\_plc/
\item Counseling and Psychological Services Center - (415) 338-2208; http://psyservs.sfsu.edu/
\item For more information on your rights and available resources:  http://titleix.sfsu.edu
\end{itemize}

\end{document}
